### 2013-07-12 Direction.tex
Initial commit. --Lindley

%
%%%%%%%%%%%%%%%%%%%%%
% Special Instructions
%%%%%%%%%%%%%%%%%%%%%
%
% Please follow these instructions to use line numbering package (lineno.sty):
%
%   http://fourforces.wordpress.com/2008/04/24/add-line-numbers-using-linenosty-in-revtex-4/
%

\RequirePackage{lineno}

\documentclass[aps,prc,twocolumn,groupedaddress,showpacs,amsmath,amssymb,floatfix,superscriptaddress]{revtex4}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{times}
\usepackage{textcomp}

\bibliographystyle{apsrev}

\usepackage{color}

\newcommand{\nuebar}{$\overline{\nu}_{e}$}
\newcommand{\PerTonDay}{(ton$\cdot$day)$^{-1}$}
\hyphenation{KamLAND}

\begin{document}

\bibliographystyle{h-physrev3}

\title{Directionality in Kilo-ton Scale Scintillating Neutrino Detectors}

% All university affiliations addresses go here:
\newcommand{\ucla}{\affiliation{University of California Los Angeles, Los Angeles, CA 90049, USA}}
\newcommand{\chicago}{\affiliation{University of Chicago, Chicago, IL, XXXX, USA}}

%
\author{C.~Aberle}\ucla
\author{A.~Elagin}\chicago
\author{L.~Winslow}\ucla

\date{\today}

\begin{abstract}
Liquid scintillator detectors are some of the most prevalent detectors due to their good energy resolution and scalability to large volumes. This has made them particularly attractive for neutrino measurements, and this technology is at the heart of many of the most important measurements in this field. Their weakness is that they only provide information on the energy of the particle. The extraction of a second signal like the particle direction would greatly enhance the scientific reach of these detectors especially for searches for neutrinoless double-beta decay. In this paper, we develop a technique for extracting particle direction and evaluate different detector advances that could be used to make direction reconstruction a reality in a kilo-ton scale detector.
\end{abstract}

\pacs{23.40.$-$s, 21.10.Tg, 14.60.Pq, 27.60.$+$j}

\maketitle

\section{Introduction}
Liquid scintillator based detectors are responsible for several of the critical measurements that have cemented our understanding of three-neutrino oscillations. These measurements include KamLAND's measurement reactor antineutrino oscillation at a distance of $\sim$200~km, Borexino's measurement of $^{7}$Be solar neutrino oscillation and most recently the short baseline reactor antineutrino experiments that measured oscillations due to $\theta_{13}$ at a distance of 1~km: Daya Bay, Double Chooz, and RENO.  Scintillating neutrino detectors will continue to be important  for the next set of neutrino measurements from the determination of the neutrino mass hierarchy to elastic scattering measurements and sterile neutrino searches and for non-proliferation applications. Their scalability to large volumes makes them particularly attractive for neutrinoless double-beta ($0\nu\beta\beta$) decay searches, and currently one of the best limits for the $0\nu\beta\beta$ half-life comes from KamLAND-Zen.

The advantage of liquid scintillators for measurements in the $\sim$1~MeV range is their simple scalability from  1~ton to 1~kiloton while providing energy resolutions of $\sim$5\%. This is roughly a factor of two better than water Cerenkov detectors, the other technology that is easily scaled to these  large masses. While the energy resolution is good due to the abundant scintillation light, this light is isotropic and does not retain the directional information of the primary particle.   In contrast, the direction of the particle can be reconstructed from the Cerenkov cones in water-based detectors, but the energy resolution rapidly degrades below $\sim$5~MeV.

In a scintillating detector, Cerenkov light is produced although most is absorbed and reemitted as part of the scintillation processes; however, some fraction retains its directional information. If this directional Cerenkov light can be isolated from the copious isotropic scintillation light then it may be possible to reconstruct the direction of the primary particle or at least determine something about the topography of the event. The addition of directionality is a powerful tool for background rejection and could be used to search for new physics especially in the case that $0\nu\beta\beta$ is observed. In this paper, we develop a technique for separating the Cerenkov and scintillation using the photon arrival times and evaluate different detector technologies that would allow the realization direction reconstruction in kilo-ton scale scintillating neutrino detectors.

\section{Light Production in Liquid Scintillators}
Molecules vibrate...index of refraction = Cerenkov.

\section{Geant4 Simulation}
In order to study the effects relevant to directionality reconstruction in liquid scintillators, a Geant4 \cite{geant4one,geant4two} simulation has been set up. The presented simulation application has been developed using standard Geant4 functionality for the optical model of the liquid scintillator. Version Geant4.9.6 is used and the simulation appliction has been built from a customized official application examples. The simulation is kept simple to provide generally applicable information about the main factors in directionality reconstruction.

The implemented detector geometry is a sphere of 6.5~m (or 0.65~m) diameter filled with liquid scintillator. The scintillator properties have been chosen to match the KamLAND scintillator as the default simulation settings \cite{tbd}: 80~\% n-Dodecane, 20~\% Pseudocumene (1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole). The implemented scintillator properties include the atomic composition and density ($\rho$ = 0.7752~g/ml), the wavelength-dependent attenuation length and refractive index, the scintillation emission spectrum, emission rise time ($\tau_r$ = 1.0~ns) and emission decay time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 0.13), scintillator light yield (LY, 9030.5 photons/MeV) and the Birks constant ($kB$ = 0.106~mm/MeV) \cite{tbd}. Variations from the baseline KamLAND case are discussed below when applicable. The sphere itself is the photodetector and the following detector properties have been varied: Transit time spread (TTS, default $\sigma = 0.1~ns$) and wavelength-dependent quantum efficiency (QE) for photoelectron production. The default is the bialkali photocathode of the Double Chooz photomultiplier tubes (PMTs) \cite{tbd} which has been verified to be the same type as the KamLAND photocathode. We used the Double Chooz QE because the data was available with finer spacing in wavelength. The sphere surface is treated as fully absorbing (no reflections) with a coverage of 100~\%. Reemission of absorbed photons in the scintillator bulk volume and scattering has not been included. Reflections, reemission and scattering are expected to be corrections to the main effects studied in this paper and are subject of future work.

In sections \ref{detector_timing_sec} to \ref{edep_size_sec}, we effectively include vertex reconstruction uncertainties. To this end, we use the KamLAND vertex resolution of 12.0~cm \cite{tbd} ($\sigma$ in one dimension) to randomly draw a vertex around the known true vertex for each event. This vertex is subsequently used as a reconstructed vertex and consequently, a time of flight (TOF) correction relative to the center of the sphere is applied for each photon hit depending on the distance between the reconstructed vertex and the hit position on the sphere. For the TOF correction a single effective speed of light is used. This effective speed was extracted at the mean wavelength of the photons which create PEs (408~nm). The 12.0~cm vertex resolution corresponds to a time resolution broadening of about 0.6~ns ($\sigma$). Note that the vertex resolution itself is dependent on scintillator and detector properties. For example, the PMTs used in KamLAND have a TTS of 1.28~ns ($\sigma$) and significant improvement of the vertex resolution could be possible with faster photodetectors. In section \ref{reconstruction_sec}, we discuss first results on more realistic reconstruction.

Before we discuss the simulation results for different simulation settings in the following sections, we highlight the effects which contribute to the timing of the scintillator detector system. First, the simulated travel time of the initial 5~MeV electron is between 0.10 and 0.15~ns, while the travel distance is about 3~cm. Second, the scintillation light emission follows a scintillator-specific distribution characterized by rise and decay time(s). Past neutrino experiments were not as interested in the effect of scintillation rise time which is the reason why there is a lack of accurate numbers. Before the solutes in liquid scintillator can emit optical photons, the energy has to be transfered from the solvent to the solute. The time constant of this energy transfer accounts for a rise time in scintillation. We assume a rise time of 1~ns, more detailed studies are needed in the future. The two time constants used to describe the falling edge of the scintillator emission time distribution (quoted above) are values specific to the KamLAND scintillator. Third, chromatic dispersion turns out to be an important effect in a 6.5~m diameter detector. Due to the wavelength-dependence of the refractive index the speed of light in the scintillator (the group velocity is used in Geant4 for normal dispersion) is also wavelength dependent. In order to study the size of this effect, we extracted results from a simplified simulation of 5~MeV electrons at the center of the sphere where we used instantaneous scintillation and did apply neither TTS nor TOF correction. The true time distribution of photoelectrons were analyzed for scintillation light and Cerenkov light separately. Photoelectrons coming from Cerenkov light are created about 0.5~ns earlier than PEs from scintillation light on average. The RMS values for both the Cerenkov and scintillation light true PE time distributions are both about 0.5~ns. Note that these numbers include the effect of the finite electron travel time. On the detector side, the time information of single photoelectrons is affected by the TTS of the photodetectors, a number which can be different by orders of magnitude depending on the detector type. The default TTS of 0.1~ns ($\sigma$) is a value which can be achieved with LAPPDs \cite{tbd}. In fact, even significantly lower TTS numbers are realistic with this novel technology. Finally, the accuracy of the vertex reconstruction is reflected in broadening and distortions of the time spectrum (after the TOF correction has been applied). Another effect which is to be addressed in future work is the granularity of the photodetectors which adds to the uncertainty in the light path and thus the TOF correction.

%% time relative to e- start, i.e. still use some truth information!!
%%statement that rise time is not super-crucial (see April 11 talk)
%% include wavelength spectra, emission spectra, absorption (?), refractive index 
%% draw time cut into plots 
%% Other settings: physics list, theCerenkovProcess->SetMaxNumPhotonsPerStep(20);theCerenkovProcess->SetMaxBetaChangePerStep(10.0); theCerenkovProcess->SetTrackSecondariesFirst(true); theScintillationProcess->SetScintillationYieldFactor(1.); theScintillationProcess->SetTrackSecondariesFirst(true); theScintillationProcess->SetFiniteRiseTime(true); particle gun, 5MeV e- in x direction from center, random polarization
%% discussion: can chromatic dispersion be disentangled? With very high statistics maybe yes (see Ben Monreal, Matt Wetstein).
%% discussion: many scintillator properties depend on each other. Optimization of scintillator properties could lead to better results for directionality reconstruction. i.e. slower scintillation (but vertex reco worse!, LY tends to be lower) , lower wavelength, 
%% principle: Cerenkov light carries direction, time cut to separate from Scintillation
%% cite LAPPDs
%% idea: filter for part of the PMTs or two kinds of PMTs: separation by wavelength
%% coverage is just a reduction by factor 0.22 (flat over wavelengths etc.), not studied in the paper?  
%% Idea: see if there is asymmetry in light even with much more abundant Scintillation light. Early detection of Cerenkov light. 
%% Cross check of calculations in MIT paper with simulation. 
%% idea of fake vertex results: show where we can gain the most. Just used the 12cm from KamLAND. This is pessimistic! Also just used a ToF correction which uses the speed of light of 
%% at the 0.1 ns level in a 6.5 m detector one has to be careful with the input of nref. If the slope is not smooth enough the group velocity calcuation is creating wavelength-dependent distortions. 
%% Reflections and absorption/reemission and scattering will shift events to higher times were they will likely not pass the time cut. They will effectively change the scintillation/Cerenkov ratio. Scattering might distort the angular variables and is subject of future work. 
%% In small det. the chromatic dispersion doesn't help too much. (In the small detector the detector properties are the driving factor.)
%% Buffer sim. This will look like a 0.65m + 0.5m scintillator det. A little more chromatic dispersion and more absorbance. 
%% show simple center-of-gravity-reco? depends on the reco effort this week. 
%% discussion: some scintillator and detector properties have complicated effects on the directionality, also other requirements have to be considered (high LY, high resolutionc). Multiparameter optimization problem.  
%% nuebar scattering stuff
%% cos_theta plot instead of rings?

\section{Detector Timing}
\label{detector_timing_sec}
\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}
        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with default settings: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$). PEs from Cerenkov light only (blue) and scintillation light only (red) are compared. The number of PEs per event after a 33.0~ns time cut is 37 from scintillation and 64 from Cerenkov. \label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME}}
        \end{center}
\end{figure}
\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{graphs/theta_phi.pdf}
        \caption[]{Photoelectron hit positions for 1000 events (5~MeV electrons in x direction) in the default simulation after the effective position reconstruction effect was included and the 33.0~ns time cut has been applied. The azimuthal angle $\Phi$ is zero in x direction and $\pi/2$ in y direction, while $\cos(\Theta)$ ranges from -1 (-z direction) to 1 (z direction). \label{hitpattern_default}}
        \end{center}
\end{figure}
\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}
        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with TTS = 1.28~ns: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 1.28~ns ($\sigma$). PEs from Cerenkov light only (blue) and scintillation light only (red) are compared. The number of PEs per event after a 33.0~ns time cut is 153 from scintillation and 60 from Cerenkov. \label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME}}
        \end{center}
\end{figure}

The settings of the default simulation are described in the previous section. Figure \ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME} shows the TOF corrected photoelectron detection time for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x axis. The photoelectrons induced by Cerenkov light arrive earlier, as expected due to the instantaneous emission and the differences in the photon speed. There is however overlap with the PE time distribution coming from scintillation light. In order to compare simulations with different parameters to each other a fixed time cut $t \leq$ 33.0~ns is done (using the truth information of the electron starting time) instead of a more realistic event-by-event time reconstruction. For the default simulation case, the number of PEs coming from Cerenkov light after the time cut (64) is about 58~\% of the total number of PEs from Cerenkov light (110). For scintillation the number of PEs after the time cut (37) is only 0.68~\% (5445).

Figure \ref{hitpattern_default} displays the PE hit pattern after the 33.0~ns time cut has been applied (1000 events plotted together). Although the time cut is an oversimplification of actual time reconstruction effects, we can use it to indicate the spatial distribution of hits after timing information has been used to separate Cerenkov and scintillation light. The Cerenkov ring structure can be clearly seen (opening angle between 40\textdegree and 50\textdegree) which demonstrates that the directional signal conveyed by the Cerenkov photons is not erased by scattering of the initial 5~MeV electrons.

If the 17 inch KamLAND PMTs \cite{tbd} (TTS = 1.28~ns) are used in the simulation, the broadening of the time distributions leads to a strongly decreased ratio of Cerenkov over Scintillation light after the time cut (see FIG. \ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME}). This demonstrates that the TTS of the photodetectors is critical for directionality reconstruction. Note that the same KamLAND vertex resolution of 12.0~cm has been used in both cases. For detectors with TTS = 0.1~ns ($\sigma$) this number is pessimistic since faster detectors can improve the vertex resolution accuracy.

\section{Detector Wavelength Response}
\label{detector_wavelength_response_sec} 
In the previous section we saw that low photodetector TTS is highly benefitial for discrimination between Cerenkov and scintillation light. The second important photodetector parameter is the wavelength-dependent QE. Since Cerenkov photons which passed 6.5~m of scintillator have higher average wavelengths than scintillation photons, a photodetector which is more sensitive at high wavelengths increases not only the absolute number of PEs but also the ratio between Cerenkov- and scintillation-induced PEs. Therefore, a simulation has been run with the QE of an extended red-sensitive GaAsP photocathode (the data has been digitized using Hamamatsu R3809U-63 QE data).

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}
        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with red-sensitive photocathode: Detector diameter = 6.5~m, red-sensitive GaAsP photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$). PEs from Cerenkov light only (blue) and scintillation light only (red) are compared. The number of PEs per event after a 33.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov. \label{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME}}
        \end{center}
\end{figure}

Figure \ref{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME} shows the results for the modified simulation with high QE in the red spectral region. The higher absolute number of PE coming from Cerenkov light and the increased Cerenkov/scintillation ratio after the time cut (almost doubled compared to the default case) would significantly enhance the directionality reconstruction capabilities.

\section{Scintillator Emission Spectrum}
\label{scintillator_emission_sec}
One important contributing factor to the separation in time between Cerenkov and scintillation photon hits is the higher average light speed for the photons coming from the Cerenkov effect due to their higher average wavelength. In this section we present two more sets of simulations where the scintillator emission spectrum was changed.

Recently, the use of quantum dots (QDs) in liquid scintillators has been studied as a possibility to improve future large scale neutrino experiments \cite{tbd}. One important motivation is the control of the emission spectra by tuning the size or composition of the quantum dots. The emission spectrum of alloyed core/shell CdS$_x$Se$_{1-x}$/ZnS Trilite450 \cite{tbd} quantum dots was measured. This spectrum is a symmetric peak centered around 461~nm with FWHM = 29~nm.

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_QD461nmScintSpec_TIME.pdf}
        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with QD emission spectrum at 461~nm: Detector diameter = 6.5~m, bialkali photocathode, QD emission spectrum at 461~nm, TTS = 0.1~ns ($\sigma$). PEs from Cerenkov light only (blue) and scintillation light only (red) are compared. The number of PEs per event after a 33.0~ns time cut is 180 from scintillation and 64 from Cerenkov. \label{6p5Meter_5MeVElectrons_Bialkali_QD461nmScintSpec_TIME}}
        \end{center}
\end{figure}

In order to isolate the effect of a different emission spectrum, the other simulation settings, including the KamLAND absorption spectrum are kept unchanged. In FIG. \ref{6p5Meter_5MeVElectrons_Bialkali_QD461nmScintSpec_TIME} the result for this simulation is presented. Compared to the default case shown in FIG. \ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME} the separation is worse because the scintillation light wavelengths are higher than in the KamLAND emission spectrum. However, advances in the production of commercial quantum dot samples could yield quantum dots which have similar, single peak emission shapes at lower wavelengths. This case has been simulated using the same spectral shape of the measured Trilite450 emission but shifted to lower wavelengths such that the emission peak is centered at 384~nm. This peak emission value has been measured for other types of QDs, however with a much more pronounced tail \cite{tbd}. The resulting PE time distribution shows (as expected) a clearer separation of Cerenkov and Scintillation light. After the simple 33.0~ns time cut we obtain slightly better numbers compared to the default simulation: The number of Cerenkov-induced PE after the time cut is unchanged (64) while the number of PEs coming from scintillation light is decreased to 25 (compared to 37 for the default simulation).

\section{Energy Dependence and Detector Size}
\label{edep_size_sec}
include this?

The number of Cerenkov photons decreases disproportionally with decreasing electron energy. The number of PEs after the 33~ns time cut for 1~MeV electrons is 8.8 PE from scintillation and 7.7 PE from Cerenkov light. This illustrates that directionality reconstruction benefits from higher initial energies but even at 1~MeV there might be accessible directionality information.

An additional set of simulations of a 0.65~m diameter ($\approx$ 1~m$^3$) detector with otherwise default parameters was performed. In the smaller detector the chromatic dispersion is not as important and absorption in the scintillator bulk is reduced. The latter effect increases the absolute number of photoelectrons. The former effect, on one hand, reduces the separation in time between Cerenkov and Scintillation light. On the other hand, it reduces the broadening of both time distributions which helps more sophisticated vertex and track reconstruction with the help of fast photodetectors like for example the LAPPDs. However, with the simplified effective vertex reconstruction and a time cut of 3.265~ns we get 87 PEs from scintillation and 87 PEs from Cerenkov. The time cut was adjusted such that the fraction of selected PEs from Cerenkov light matches the fraction in the large detector ($\approx$ 58\%). The QD spectrum discussed in section \ref{scintillator_emission_sec} (both shifted and unshifted) did not change the PE time distribution significantly. We also simulated the red-sensitive photocathode discussed in section \ref{detector_wavelength_response_sec} in the small detector. The amount of PE increases again significantly and with the 3.265~ns time cut the number of PE coming from Cerenkov light is 182 while the number of PE from scintillation light is 96.

In summary, the photodetector properties (TTS and QE) become more important relative to the scintillator emission and absorption spectra when the detector size is decreased from 6.5~m to 0.65~m.

%%idea: use (some) PMTs with high QE for high lambda (and/or low QE for normal lambda). 
%% add some more info why we are interested in a smaller detector
%% motivation: Daedalus/ISODAR (?), 0nuBB 
\section{Reconstruction}
\label{reconstruction_sec}
Hopefully results

\section{Conclusion}
This is going to work.

\section{Acknowledgements}
Matt, Henry, Janet, Katsushi. Need Andrey's DOE numbers.

\bibliography{DirectionBibliography}

%%20 inch PMTs: 
%%H.~Kume, S.~Sawaki, M.~Ito, K.~Arisaka, T.~Kajita, A.~Nishimura, and A.~Suzuki, Nucl. Instrum. Methods Phys. Res. \bf{205}, 443 (1983).
%%17 inch PMTs:

\end{document}

### 2013-07-15 Direction.tex
More files. --Lindley

@@
 % All university affiliations addresses go here:
-\newcommand{\ucla}{\affiliation{University of California Los Angeles, Los Angeles, CA 90049, USA}}
-\newcommand{\chicago}{\affiliation{University of Chicago, Chicago, IL, XXXX, USA}}
+\newcommand{\ucla}{\affiliation{University of California Los Angeles, Los Angeles, CA 90095, USA}}
+\newcommand{\chicago}{\affiliation{University of Chicago, Chicago, IL, 60637, USA}}
 

### 2013-07-15 Direction.tex

@@
 \author{A.~Elagin}\chicago
+\author{M.~Wetstein}\chicago
 \author{L.~Winslow}\ucla
@@
 \section{Light Production in Liquid Scintillators}
-Molecules vibrate...index of refraction = Cerenkov.
+Liquid scintillators are cocktails of aromatic hydrocarbons.
+When charged particles move through a scintillator, the molecules are excited, predominantly the the non-localized electrons in the $\pi$-bonds of the phenyl groups.
+Vibrational and rotational modes of the molecules are turned into heat within picoseconds through collisions with other molecules.
+Within $\sim$10 picoseconds, the $\pi$-electrons de-excite to the first excited state from higher levels through radiationless transitions.
+The first excited state de-excites through photon emission.
+There are two characteristic times for this de-excitation depending if the singlet state or the triplet state was excited.
+The singlet state will de-excite within nanoseconds while the triplet state de-excites on the order of 10's or 100's of nanoseconds.
+These two processes are fluorescence and phosphorescence respectively.
+The exact time constants for these processes are determined by the composition of the scintillator.
+
+The molecules in liquid scintillators are not isolated.
+Radiationless processes transfer energy between the molecules.
+The probability of energy transfer between molecules increases as the overlap between the molecule's absorption and emission spectra increase.
+The absorption and emission spectra overlap at some level for all molecules.
+Therefore, if there is only one type of molecule in the scintillator cocktail, the light output would be reduced due to inefficiencies in the energy transfer through multiple absorption and reemission processes.
+Aromatic solutes or fluors are added to the primary solvent to shift the wavelength of the photons to longer wavelengths where the scintillator is more transparent.
+This wavelength-shifting is also used to match the quantum efficiency as a function of wavelength for the photodetectors being used.
+One typical scintillator mixture uses pseudo-cumene as the solvent with 1-5~g/L of PPO as the fluor.
+This mixture has a peak emission at about 400~nm where photomultiplier tubes (PMTs) are most sensitive and the pseudo-cumene is relatively transparent.
+
+A good liquid scintillator will produce $\sim$10,000 photons isotropically per MeV of deposited energy.
+Although less abundant, if a particle is moving faster than the speed of light in the medium then Cerenkov light will be produced as well.
+This light is emitted in a cone pointed in the direction of the particles trajectory and with a continuous spectrum wavelengths weighted toward shorter wavelengths but extending well into the red,
+\begin{equation}
+\label{eqCerenkov}
+\frac{dN}{d\lambda dx} = \frac{2 \pi \alpha Z^2}{\lambda^2} ( 1 - \frac{1}{\beta^2 n(\lambda)^2} )
+\end{equation}
+where $n(\lambda)$ is the wavelength-dependent index of refraction and $\beta$ is the velocity of the incoming particle.
+The Cerenkov light produced at wavelengths shorter than the absorption cutoff of the scintillator will be absorbed and re-emitted as isotropic light, but wavelengths longer than this cutoff will propagate across the detector retaining their directional information, roughly 60 photons per MeV assuming a 400~nm absorption cutoff.
+These undisturbed Cerenkov photons will have timing determined by the group velocity in the liquid,
+\begin{equation}
+\label{eqGroup}
+v_{g}(\lambda) = \frac{c_{vacuum}}{n(\lambda) + dn(\lambda)/dlog(\lambda)}.
+\end{equation}
+These longer wavelengths Cerenkov photons arrive in principle before the scintillation light which is slowed by both the scintillation processes and the shorter wavelengths involved.
+Thus, with sufficient timing resolution and sensitivity to longer wavelength, it should be possible to separate the directional Cerenkov light and the isotropic scintillation light in time and then reconstruct the direction of the initial particle.
 

### 2013-07-16 Direction.tex
Halfway done with first references. --Lindley

@@
 Liquid scintillator based detectors are responsible for several of the critical measurements that have cemented our understanding of three-neutrino oscillations.
-These measurements include KamLAND's measurement reactor antineutrino oscillation at a distance of $\sim$200~km, Borexino's measurement of $^{7}$Be solar neutrino oscillation and most recently the short baseline reactor antineutrino experiments that measured oscillations due to $\theta_{13}$ at a distance of 1~km: Daya Bay, Double Chooz, and RENO.
+These measurements include KamLAND's measurement reactor antineutrino oscillation at a distance of $\sim$200~km\cite{kam2013}, Borexino's measurement of $^{7}$Be solar neutrino oscillation\cite{borexino} and most recently the short baseline reactor antineutrino experiments that measured oscillations due to $\theta_{13}$ at a distance of 1~km: Daya Bay\cite{dbtwo}, Double Chooz\cite{dctwo, dchydrogen}, and RENO\cite{reno}.
 Scintillating neutrino detectors will continue to be important  for the next set of neutrino measurements from the determination of the neutrino mass hierarchy to elastic scattering measurements and sterile neutrino searches and for non-proliferation applications.

### 2013-07-16 Direction.tex
Add most general references. --Lindley

@@
 \usepackage{textcomp}
+\usepackage{hyperref}
 
@@
 These measurements include KamLAND's measurement reactor antineutrino oscillation at a distance of $\sim$200~km\cite{kam2013}, Borexino's measurement of $^{7}$Be solar neutrino oscillation\cite{borexino} and most recently the short baseline reactor antineutrino experiments that measured oscillations due to $\theta_{13}$ at a distance of 1~km: Daya Bay\cite{dbtwo}, Double Chooz\cite{dctwo, dchydrogen}, and RENO\cite{reno}.
-Scintillating neutrino detectors will continue to be important  for the next set of neutrino measurements from the determination of the neutrino mass hierarchy to elastic scattering measurements and sterile neutrino searches and for non-proliferation applications.
-Their scalability to large volumes makes them particularly attractive for neutrinoless double-beta ($0\nu\beta\beta$) decay searches, and currently one of the best limits for the $0\nu\beta\beta$ half-life comes from KamLAND-Zen.
+Scintillating neutrino detectors will continue to be important  for the next set of neutrino measurements from the determination of the neutrino mass hierarchy\cite{juno,reno50} to elastic scattering measurements\cite{isodarscatt} and sterile neutrino searches\cite{isodar,nist} and for non-proliferation applications\cite{nucifer, songs}.
+Their scalability to large volumes makes them particularly attractive for neutrinoless double-beta ($0\nu\beta\beta$) decay searches, and currently one of the best limits for the $0\nu\beta\beta$ half-life comes from KamLAND-Zen\cite{KZ0nu}.
 
@@
 where $n(\lambda)$ is the wavelength-dependent index of refraction and $\beta$ is the velocity of the incoming particle.
-The Cerenkov light produced at wavelengths shorter than the absorption cutoff of the scintillator will be absorbed and re-emitted as isotropic light, but wavelengths longer than this cutoff will propagate across the detector retaining their directional information, roughly 60 photons per MeV assuming a 400~nm absorption cutoff.
+The Cerenkov light produced at wavelengths shorter than the absorption cutoff of the scintillator will be absorbed and re-emitted as isotropic light, but wavelengths longer than this cutoff will propagate across the detector retaining their directional information, roughly 60 photons per MeV assuming a 400~nm absorption cutoff\cite{qdot}.
 These undisturbed Cerenkov photons will have timing determined by the group velocity in the liquid,
@@
 On the detector side, the time information of single photoelectrons is affected by the TTS of the photodetectors, a number which can be different by orders of magnitude depending on the detector type.
-The default TTS of 0.1~ns ($\sigma$) is a value which can be achieved with LAPPDs \cite{tbd}.
-In fact, even significantly lower TTS numbers are realistic with this novel technology.
+The default TTS of 0.1~ns ($\sigma$) is a value which can be achieved with the large area picosecond photodetectors (LAPPDs)\cite{LAPPDSum,LAPPDTDR} and possibly hybrid photodetectors (HPDs)\cite{hpdThesis}.
+In fact, even significantly lower TTS numbers are realistic with the LAPPD.
 Finally, the accuracy of the vertex reconstruction is reflected in broadening and distortions of the time spectrum (after the TOF correction has been applied).

### 2013-07-16 Direction.tex
Add acknowledgements. --Lindley

@@
 \section{Acknowledgements}
-Matt, Henry, Janet, Katsushi. Need Andrey's DOE numbers.
+The authors would like to thank H.
+Frisch for useful discussion and comments in preparing this work.
+The authors would like to thank K.
+Arisaka for discussion on the capabilities of traditional PMTs in addition to HPDs.
+Finally, L.
+Winslow would like to thank J.M.
+Conrad for suggesting that the LAPPD may have sufficient timing resolution to allow the reconstruction of the electron direction in $0\nu\beta\beta$ and that a collaboration with the University of Chicago should be pursued.
+A.
+Elagin and M.
+Wetstein are supported by DOE XXXX.
+C.
+Aberle and L.
+Winslow are supported by funds from University of California Los Angeles.
 

### 2013-07-17 Direction.tex
One-line fix in section 3 concerning effective speed of light. --Christoph
Added new plots for the MC sections. Removed the effective vertex reconstruction since we now have real reconstruction results. Changed the text accordingly. Updated the PE numbers throughout the document and used 34.0 ns cut because we work with this number in the reconstruction. Some polishing of the text in sections 3 through 7. --Christoph

@@
 The presented simulation application has been developed using standard Geant4 functionality for the optical model of the liquid scintillator.
-Version Geant4.9.6 is used and the simulation appliction has been built from a customized official application examples.
+Version Geant4.9.6 is used and the simulation application has been built from a customized official application example.
 The simulation is kept simple to provide generally applicable information about the main factors in directionality reconstruction.
@@
 The implemented detector geometry is a sphere of 6.5~m (or 0.65~m) diameter filled with liquid scintillator.
-The scintillator properties have been chosen to match the KamLAND scintillator as the default simulation settings \cite{tbd}: 80~\% n-Dodecane, 20~\% Pseudocumene (1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole).
+Default scintillator properties have been chosen to match the KamLAND scintillator \cite{tbd}: 80~\% n-Dodecane, 20~\% Pseudocumene (1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole).
 The implemented scintillator properties include the atomic composition and density ($\rho$ = 0.7752~g/ml), the wavelength-dependent attenuation length and refractive index, the scintillation emission spectrum, emission rise time ($\tau_r$ = 1.0~ns) and emission decay time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 0.13), scintillator light yield (LY, 9030.5 photons/MeV) and the Birks constant ($kB$ = 0.106~mm/MeV) \cite{tbd}.
 Variations from the baseline KamLAND case are discussed below when applicable.
-The sphere itself is the photodetector and the following detector properties have been varied: Transit time spread (TTS, default $\sigma = 0.1~ns$) and wavelength-dependent quantum efficiency (QE) for photoelectron production.
-The default is the bialkali photocathode of the Double Chooz photomultiplier tubes (PMTs) \cite{tbd} which has been verified to be the same type as the KamLAND photocathode.
-We used the Double Chooz QE because the data was available with finer spacing in wavelength.
-The sphere surface is treated as fully absorbing (no reflections) with a coverage of 100~\%.
 Reemission of absorbed photons in the scintillator bulk volume and scattering has not been included.
@@
 
-In sections \ref{detector_timing_sec} to \ref{edep_size_sec}, we effectively include vertex reconstruction uncertainties.
-To this end, we use the KamLAND vertex resolution of 12.0~cm \cite{tbd} ($\sigma$ in one dimension) to randomly draw a vertex around the known true vertex for each event.
+The sphere surface is used as the photodetector.
+It is treated as fully absorbing (no reflections) with a coverage of 100~\%.
+Two important photodetector properties have also been varied: Transit time spread (TTS, default $\sigma$ = 0.1~ns) and wavelength-dependent quantum efficiency (QE) for photoelectron production.
+The default is the bialkali photocathode of the Double Chooz photomultiplier tubes (PMTs) \cite{tbd}.
+It has been verified to be the same type as the KamLAND photocathode.
+We used the Double Chooz QE because the data was available with finer spacing in wavelength.
+
+%deprecated text: fake vertex reconstruction not needed anymore. 
+%To this end, we use the KamLAND vertex resolution of 12.0~cm \cite{tbd} ($\sigma$ in one dimension) to randomly draw a vertex around the known true vertex for each event.
 This vertex is subsequently used as a reconstructed vertex and consequently, a time of flight (TOF) correction relative to the center of the sphere is applied for each photon hit depending on the distance between the reconstructed vertex and the hit position on the sphere.
 For the TOF correction a single effective speed of light is used.
-This effective speed was extracted at the mean wavelength of the photons which create PEs (408~nm).
+This effective speed was extracted at the mean wavelength of both Scintillation and Cerenkov photons (408~nm).
 The 12.0~cm vertex resolution corresponds to a time resolution broadening of about 0.6~ns ($\sigma$).
@@
 In section \ref{reconstruction_sec}, we discuss first results on more realistic reconstruction.
-
 Before we discuss the simulation results for different simulation settings in the following sections, we highlight the effects which contribute to the timing of the scintillator detector system.
@@
 Second, the scintillation light emission follows a scintillator-specific distribution characterized by rise and decay time(s).
-Past neutrino experiments were not as interested in the effect of scintillation rise time which is the reason why there is a lack of accurate numbers.
 Before the solutes in liquid scintillator can emit optical photons, the energy has to be transfered from the solvent to the solute.
-The time constant of this energy transfer accounts for a rise time in scintillation.
+The time constant of this energy transfer accounts for a rise time in scintillation light emission.
+Past neutrino experiments were not highly interested in the effect of scintillation rise time which is the reason why there is a lack of accurate numbers.
 We assume a rise time of 1~ns, more detailed studies are needed in the future.
 The two time constants used to describe the falling edge of the scintillator emission time distribution (quoted above) are values specific to the KamLAND scintillator.
-Third, chromatic dispersion turns out to be an important effect in a 6.5~m diameter detector.
+Third, chromatic dispersion turns out to be an important effect in a 6.5~m diameter detector at the level of precision needed for direction reconstruction.
 Due to the wavelength-dependence of the refractive index the speed of light in the scintillator (the group velocity is used in Geant4 for normal dispersion) is also wavelength dependent.
 In order to study the size of this effect, we extracted results from a simplified simulation of 5~MeV electrons at the center of the sphere where we used instantaneous scintillation and did apply neither TTS nor TOF correction.
-The true time distribution of photoelectrons were analyzed for scintillation light and Cerenkov light separately.
+The true time distributions of photoelectrons were analyzed for scintillation light and Cerenkov light separately.
 Photoelectrons coming from Cerenkov light are created about 0.5~ns earlier than PEs from scintillation light on average.
@@
 Note that these numbers include the effect of the finite electron travel time.
+
 On the detector side, the time information of single photoelectrons is affected by the TTS of the photodetectors, a number which can be different by orders of magnitude depending on the detector type.
@@
 
-%% time relative to e- start, i.e. still use some truth information!!
+In Sections \ref{detector_timing_sec} to \ref{edep_size_sec}, we comparatively study the photoelectron timing for different detector properties.
+We follow the idea to increase discrimination between Cerenkov and scintillation light by detector timing.
+The main quantities provided by the Geant4 simulation are the photoelectron hit positions and the detection times after TTS has been applied.
+In section \ref{reconstruction_sec} these quantities are used for event reconstruction.
 %%statement that rise time is not super-crucial (see April 11 talk)
-%% include wavelength spectra, emission spectra, absorption (?), refractive index 
-%% draw time cut into plots 
+%% include wavelength spectra, emission spectra, absorption (?), refractive index  ?
 %% Other settings: physics list, theCerenkovProcess->SetMaxNumPhotonsPerStep(20);theCerenkovProcess->SetMaxBetaChangePerStep(10.0); theCerenkovProcess->SetTrackSecondariesFirst(true); theScintillationProcess->SetScintillationYieldFactor(1.); theScintillationProcess->SetTrackSecondariesFirst(true); theScintillationProcess->SetFiniteRiseTime(true); particle gun, 5MeV e- in x direction from center, random polarization
@@
 slower scintillation (but vertex reco worse!, LY tends to be lower) , lower wavelength,
-%% principle: Cerenkov light carries direction, time cut to separate from Scintillation
-%% cite LAPPDs
 %% idea: filter for part of the PMTs or two kinds of PMTs: separation by wavelength
 %% coverage is just a reduction by factor 0.22 (flat over wavelengths etc.), not studied in the paper?  
-%% Idea: see if there is asymmetry in light even with much more abundant Scintillation light. Early detection of Cerenkov light. 
-%% Cross check of calculations in MIT paper with simulation. 
-%% idea of fake vertex results: show where we can gain the most. Just used the 12cm from KamLAND. This is pessimistic! Also just used a ToF correction which uses the speed of light of 
 %% at the 0.1 ns level in a 6.5 m detector one has to be careful with the input of nref. If the slope is not smooth enough the group velocity calcuation is creating wavelength-dependent distortions. 
@@
 %% In small det. the chromatic dispersion doesn't help too much. (In the small detector the detector properties are the driving factor.)
-%% Buffer sim. This will look like a 0.65m + 0.5m scintillator det. A little more chromatic dispersion and more absorbance. 
-%% show simple center-of-gravity-reco? depends on the reco effort this week. 
 %% discussion: some scintillator and detector properties have complicated effects on the directionality, also other requirements have to be considered (high LY, high resolutionc).
 Multiparameter optimization problem.
-%% nuebar scattering stuff
-%% cos_theta plot instead of rings?
+%% nuebar scattering --> one of next papers
 
@@
         \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}
-        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with default settings: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
-PEs from Cerenkov light only (blue) and scintillation light only (red) are compared.
-The number of PEs per event after a 33.0~ns time cut is 37 from scintillation and 64 from Cerenkov.
+        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with default settings: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
+PEs from Cerenkov light (black) and scintillation light (red) are compared.
+The number of PEs per event after a 34.0~ns time cut is 171 from scintillation and 108 from Cerenkov.
 \label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME}}
@@
         \begin{center}
-        \includegraphics[scale=0.40]{graphs/theta_phi.pdf}
-        \caption[]{Photoelectron hit positions for 1000 events (5~MeV electrons in x direction) in the default simulation after the effective position reconstruction effect was included and the 33.0~ns time cut has been applied.
-The azimuthal angle $\Phi$ is zero in x direction and $\pi/2$ in y direction, while $\cos(\Theta)$ ranges from -1 (-z direction) to 1 (z direction).
-\label{hitpattern_default}}
+        \includegraphics[scale=0.40]{graphs/cos_psi_comp_h.pdf}
+        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction for 1000 events (5~MeV electrons).
+Default simulation settings were used.
+A 33.0~ns time cut (black) is compared with a 34.0~ns time cut (red).
+\label{cerenkov_cone}}
         \end{center}
@@
         \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}
-        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with TTS = 1.28~ns: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 1.28~ns ($\sigma$).
-PEs from Cerenkov light only (blue) and scintillation light only (red) are compared.
-The number of PEs per event after a 33.0~ns time cut is 153 from scintillation and 60 from Cerenkov.
+        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with TTS = 1.28~ns: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 1.28~ns ($\sigma$).
+PEs from Cerenkov light (black) and scintillation light (red) are compared.
+The number of PEs per event after a 34.0~ns time cut is 349 from scintillation and 88 from Cerenkov.
 \label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME}}
@@
 
-The settings of the default simulation are described in the previous section.
-Figure \ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME} shows the TOF corrected photoelectron detection time for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x axis.
+First we discuss the default simulation settings which are described in the previous section.
+Figure \ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME} shows the TTS-smeared photoelectron detection time for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x-axis.
 The photoelectrons induced by Cerenkov light arrive earlier, as expected due to the instantaneous emission and the differences in the photon speed.
 There is however overlap with the PE time distribution coming from scintillation light.
-In order to compare simulations with different parameters to each other a fixed time cut $t \leq$ 33.0~ns is done (using the truth information of the electron starting time) instead of a more realistic event-by-event time reconstruction.
-For the default simulation case, the number of PEs coming from Cerenkov light after the time cut (64) is about 58~\% of the total number of PEs from Cerenkov light (110).
-For scintillation the number of PEs after the time cut (37) is only 0.68~\% (5445).
+In order to compare simulations with different parameters to each other a fixed time cut $t \leq$ 34.0~ns (or 33.0~ns?) is applied (using the truth information of the electron starting time) instead of an event-by-event time reconstruction (see Section \ref{reconstruction_sec} for reconstruction results).
+For the default simulation case, the number of PEs per event coming from Cerenkov light after the 34.0~ns time cut (108) is 98~\% of the total number of PEs from Cerenkov light (110).
+For scintillation the number of PEs after the time cut (171) is only 3.2~\% of the total scintillation-induced PEs (5445).
 
-Figure \ref{hitpattern_default} displays the PE hit pattern after the 33.0~ns time cut has been applied (1000 events plotted together).
-Although the time cut is an oversimplification of actual time reconstruction effects, we can use it to indicate the spatial distribution of hits after timing information has been used to separate Cerenkov and scintillation light.
+Figure \ref{cerenkov_cone} displays the PE hit pattern after the 34.0~ns time cut has been applied (1000 events plotted together).
+Although this time cut is an oversimplification of actual time reconstruction effects, we can use it to indicate the spatial distribution of hits after timing information has been used to separate Cerenkov and scintillation light.
 The Cerenkov ring structure can be clearly seen (opening angle between 40\textdegree and 50\textdegree) which demonstrates that the directional signal conveyed by the Cerenkov photons is not erased by scattering of the initial 5~MeV electrons.
@@
 \ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME}).
-This demonstrates that the TTS of the photodetectors is critical for directionality reconstruction.
-Note that the same KamLAND vertex resolution of 12.0~cm has been used in both cases.
-For detectors with TTS = 0.1~ns ($\sigma$) this number is pessimistic since faster detectors can improve the vertex resolution accuracy.
+This demonstrates that low photodetector TTS is critical for directionality reconstruction and motivates use of novel detector types.
 
@@
 \label{detector_wavelength_response_sec} 
-In the previous section we saw that low photodetector TTS is highly benefitial for discrimination between Cerenkov and scintillation light.
-The second important photodetector parameter is the wavelength-dependent QE.
+Comlementary to lowering the photodetector TTS, optimization of the wavelength-dependent QE is promising.
 Since Cerenkov photons which passed 6.5~m of scintillator have higher average wavelengths than scintillation photons, a photodetector which is more sensitive at high wavelengths increases not only the absolute number of PEs but also the ratio between Cerenkov- and scintillation-induced PEs.
-Therefore, a simulation has been run with the QE of an extended red-sensitive GaAsP photocathode (the data has been digitized using Hamamatsu R3809U-63 QE data).
+Therefore, a simulation has been run with the QE of an extended red-sensitive GaAsP photocathode (digitized Hamamatsu R3809U-63 QE data was used).
 
@@
         \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with red-sensitive photocathode: Detector diameter = 6.5~m, red-sensitive GaAsP photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
-PEs from Cerenkov light only (blue) and scintillation light only (red) are compared.
-The number of PEs per event after a 33.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov.
+PEs from Cerenkov light (black) and scintillation light (red) are compared.
+The number of PEs per event after a 34.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov.
 \label{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME}}
@@
 Figure \ref{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME} shows the results for the modified simulation with high QE in the red spectral region.
-The higher absolute number of PE coming from Cerenkov light and the increased Cerenkov/scintillation ratio after the time cut (almost doubled compared to the default case) would significantly enhance the directionality reconstruction capabilities.
+The higher absolute number of PE coming from Cerenkov light (factor of $\approx$ 2) and the increased Cerenkov/scintillation ratio (factor of $\approx$ 1.6) after the time cut would significantly enhance the directionality reconstruction capabilities.
 
@@
         \begin{center}
-        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_QD461nmScintSpec_TIME.pdf}
-        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with QD emission spectrum at 461~nm: Detector diameter = 6.5~m, bialkali photocathode, QD emission spectrum at 461~nm, TTS = 0.1~ns ($\sigma$).
-PEs from Cerenkov light only (blue) and scintillation light only (red) are compared.
-The number of PEs per event after a 33.0~ns time cut is 180 from scintillation and 64 from Cerenkov.
-\label{6p5Meter_5MeVElectrons_Bialkali_QD461nmScintSpec_TIME}}
+        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME.pdf}
+        \caption[]{PE times for the simulation of 1000 electrons (5~MeV) with a quantum dot emission spectrum centered at 384~nm: Detector diameter = 6.5~m, bialkali photocathode, QD emission spectrum at 384~nm, TTS = 0.1~ns ($\sigma$).
+PEs from Cerenkov light only (black) and scintillation light only (red) are compared.
+The number of PEs per event after a 34.0~ns time cut is 124 from scintillation and 107 from Cerenkov.
+\label{6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME}}
         \end{center}
@@
 In order to isolate the effect of a different emission spectrum, the other simulation settings, including the KamLAND absorption spectrum are kept unchanged.
-In FIG.
-\ref{6p5Meter_5MeVElectrons_Bialkali_QD461nmScintSpec_TIME} the result for this simulation is presented.
 Compared to the default case shown in FIG.
@@
 This peak emission value has been measured for other types of QDs, however with a much more pronounced tail \cite{tbd}.
-The resulting PE time distribution shows (as expected) a clearer separation of Cerenkov and Scintillation light.
-After the simple 33.0~ns time cut we obtain slightly better numbers compared to the default simulation: The number of Cerenkov-induced PE after the time cut is unchanged (64) while the number of PEs coming from scintillation light is decreased to 25 (compared to 37 for the default simulation).
+The resulting PE time distribution shows a clearer separation of Cerenkov and Scintillation light (see FIG.
+\ref{6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME}).
+After the 34.0~ns cut on the TTS-smeared PE time we obtain slightly better numbers compared to the default simulation: The number of Cerenkov-induced PE after the time cut is unchanged while the number of PEs coming from scintillation light is decreased due to the higher average photon travel times.
 
@@
 The number of Cerenkov photons decreases disproportionally with decreasing electron energy.
-The number of PEs after the 33~ns time cut for 1~MeV electrons is 8.8 PE from scintillation and 7.7 PE from Cerenkov light.
+The number of PEs after the 34.0~ns time cut for 1~MeV electrons is 37 PE from scintillation and 13 PE from Cerenkov light.
 This illustrates that directionality reconstruction benefits from higher initial energies but even at 1~MeV there might be accessible directionality information.
@@
 On the other hand, it reduces the broadening of both time distributions which helps more sophisticated vertex and track reconstruction with the help of fast photodetectors like for example the LAPPDs.
-However, with the simplified effective vertex reconstruction and a time cut of 3.265~ns we get 87 PEs from scintillation and 87 PEs from Cerenkov.
+
+%tbd (calculate numbers): However, with the simplified effective vertex reconstruction and a time cut of 3.4~ns we get  PEs from scintillation and 87 PEs from Cerenkov.
 The time cut was adjusted such that the fraction of selected PEs from Cerenkov light matches the fraction in the large detector ($\approx$ 58\%).

### 2013-07-17 Direction.tex
Additions from Henry. --Lindley

@@
 
-\title{Directionality in Kilo-ton Scale Scintillating Neutrino Detectors}
+\title{Measuring Directionality in Double Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors}
 
@@
 \author{M.~Wetstein}\chicago
+\author{H.~J.~Frisch}\chicago
 \author{L.~Winslow}\ucla
@@
 \begin{abstract}
-Liquid scintillator detectors are some of the most prevalent detectors due to their good energy resolution and scalability to large volumes.
-This has made them particularly attractive for neutrino measurements, and this technology is at the heart of many of the most important measurements in this field.
-Their weakness is that they only provide information on the energy of the particle.
-The extraction of a second signal like the particle direction would greatly enhance the scientific reach of these detectors especially for searches for neutrinoless double-beta decay.
-In this paper, we develop a technique for extracting particle direction and evaluate different detector advances that could be used to make direction reconstruction a reality in a kilo-ton scale detector.
+Large liquid-scintillator-based detectors have proven to be
+exceptionally effective for low energy neutrino measurements due to
+their good energy resolution and scalability to large volumes. The
+addition of directional information using Cherenkov light and fast
+timing would greatly enhance the scientific reach of these detectors,
+especially for searches for neutrinoless double-beta decay. In this
+paper, we develop a technique for extracting particle direction and
+evaluate different detector advances that could be used to make
+direction reconstruction a reality in a kiloton-scale detector.
 \end{abstract}
@@
 \section{Introduction}
-Liquid scintillator based detectors are responsible for several of the critical measurements that have cemented our understanding of three-neutrino oscillations.
-These measurements include KamLAND's measurement reactor antineutrino oscillation at a distance of $\sim$200~km\cite{kam2013}, Borexino's measurement of $^{7}$Be solar neutrino oscillation\cite{borexino} and most recently the short baseline reactor antineutrino experiments that measured oscillations due to $\theta_{13}$ at a distance of 1~km: Daya Bay\cite{dbtwo}, Double Chooz\cite{dctwo, dchydrogen}, and RENO\cite{reno}.
-Scintillating neutrino detectors will continue to be important  for the next set of neutrino measurements from the determination of the neutrino mass hierarchy\cite{juno,reno50} to elastic scattering measurements\cite{isodarscatt} and sterile neutrino searches\cite{isodar,nist} and for non-proliferation applications\cite{nucifer, songs}.
-Their scalability to large volumes makes them particularly attractive for neutrinoless double-beta ($0\nu\beta\beta$) decay searches, and currently one of the best limits for the $0\nu\beta\beta$ half-life comes from KamLAND-Zen\cite{KZ0nu}.
+Liquid scintillator based detectors are responsible for several of the
+critical measurements that have determined our present understanding
+of neutrino masses and mixings. These measurements include KamLAND's
+measurement of reactor antineutrino oscillation at a distance of
+$\sim$200~km\cite{kam2013}, Borexino's measurement of $^{7}$Be solar
+neutrino oscillation\cite{borexino}, and most recently the short
+baseline reactor antineutrino experiments that measured oscillations
+due to $\theta_{13}$ at a distance of 1~km: Daya Bay\cite{dbtwo},
+Double Chooz\cite{dctwo, dchydrogen}, and RENO\cite{reno}.
+Scintillator-based neutrino detectors will continue to be important for the
+next set of neutrino measurements, from the determination of the
+neutrino mass hierarchy\cite{juno,reno50} to elastic scattering
+measurements\cite{isodarscatt} and sterile neutrino
+searches\cite{isodar,nist} and for non-proliferation
+applications\cite{nucifer, songs}.
 
-The advantage of liquid scintillators for measurements in the $\sim$1~MeV range is their simple scalability from  1~ton to 1~kiloton while providing energy resolutions of $\sim$5\%.
-This is roughly a factor of two better than water Cerenkov detectors, the other technology that is easily scaled to these  large masses.
-While the energy resolution is good due to the abundant scintillation light, this light is isotropic and does not retain the directional information of the primary particle.
-In contrast, the direction of the particle can be reconstructed from the Cerenkov cones in water-based detectors, but the energy resolution rapidly degrades below $\sim$5~MeV.
+The scalability of these detectors to large volumes also makes them
+highly competitive for neutrinoless double-beta ($0\nu\beta\beta$)
+decay searches in which the final state consists of a
+positron-electron pair with energies in the few MeV range.  Currently
+one of the best limits for the $0\nu\beta\beta$ half-life comes from
+KamLAND-Zen\cite{KZ0nu}.
 
-In a scintillating detector, Cerenkov light is produced although most is absorbed and reemitted as part of the scintillation processes; however, some fraction retains its directional information.
-If this directional Cerenkov light can be isolated from the copious isotropic scintillation light then it may be possible to reconstruct the direction of the primary particle or at least determine something about the topography of the event.
-The addition of directionality is a powerful tool for background rejection and could be used to search for new physics especially in the case that $0\nu\beta\beta$ is observed.
-In this paper, we develop a technique for separating the Cerenkov and scintillation using the photon arrival times and evaluate different detector technologies that would allow the realization direction reconstruction in kilo-ton scale scintillating neutrino detectors.
+The advantage of liquid scintillators for measurements in the
+$\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
+providing energy resolutions of $\sim$5\%. This is roughly a factor of
+two better than water Cerenkov detectors, the other developed
+technology that can be economically scaled to these large
+masses. However, for scintillator-based detectors, while the energy
+resolution is good due to the abundance of light, the light is
+isotropic and does not retain the directional information of the
+primary particle.  In contrast, the direction of the particle can be
+reconstructed from the Cerenkov cones in water-based detectors,
+although the energy resolution rapidly degrades below $\sim$5~MeV. For
+double-beta decay in particular, but also for neutrino interactions,
+the directional information can be a strong suppressant of
+backgrounds.
+
+In a liquid-scintillation-based detector, Cerenkov light is also
+produced, although most is absorbed and re-emitted as part of the
+scintillation processes.  However, some fraction retains its
+directional information. If this directional Cerenkov light can be
+isolated from the copious isotropic scintillation light, it may be
+possible to reconstruct the direction of the primary particle or, in
+the case of double beta decay, to determine the existence and topology
+of the pair.  The addition of directionality is thus a powerful tool
+for background rejection.  In this paper, we develop a technique for
+separating the Cerenkov and scintillation using the photon arrival
+times and evaluate different detector technologies that would allow
+the realization direction reconstruction in kilo-ton scale
+scintillating neutrino detectors.
 
 \section{Light Production in Liquid Scintillators}
-Liquid scintillators are cocktails of aromatic hydrocarbons.
-When charged particles move through a scintillator, the molecules are excited, predominantly the the non-localized electrons in the $\pi$-bonds of the phenyl groups.
-Vibrational and rotational modes of the molecules are turned into heat within picoseconds through collisions with other molecules.
-Within $\sim$10 picoseconds, the $\pi$-electrons de-excite to the first excited state from higher levels through radiationless transitions.
-The first excited state de-excites through photon emission.
-There are two characteristic times for this de-excitation depending if the singlet state or the triplet state was excited.
-The singlet state will de-excite within nanoseconds while the triplet state de-excites on the order of 10's or 100's of nanoseconds.
-These two processes are fluorescence and phosphorescence respectively.
-The exact time constants for these processes are determined by the composition of the scintillator.
+Liquid scintillators are cocktails of aromatic hydrocarbons. When
+charged particles move through a scintillator, the molecules are
+excited, predominantly the the non-localized electrons in the
+$\pi$-bonds of the phenyl groups~\cite{scintillator_ref}. Vibrational
+and rotational modes of the molecules are turned into heat within
+picoseconds through collisions with other molecules.  Within $\sim$10
+picoseconds, the $\pi$-electrons de-excite to the first excited state
+from higher levels through radiationless transitions. The first
+excited state de-excites through photon emission. There are two
+characteristic times for this de-excitation, depending if the singlet
+state or the triplet state was excited.  The singlet state will
+de-excite within nanoseconds while the triplet state de-excites on the
+order of 10's or 100's of nanoseconds. These two processes are
+fluorescence and phosphorescence respectively. The exact time
+constants for these processes are determined by the composition of the
+scintillator.
 
-The molecules in liquid scintillators are not isolated.
-Radiationless processes transfer energy between the molecules.
-The probability of energy transfer between molecules increases as the overlap between the molecule's absorption and emission spectra increase.
-The absorption and emission spectra overlap at some level for all molecules.
-Therefore, if there is only one type of molecule in the scintillator cocktail, the light output would be reduced due to inefficiencies in the energy transfer through multiple absorption and reemission processes.
-Aromatic solutes or fluors are added to the primary solvent to shift the wavelength of the photons to longer wavelengths where the scintillator is more transparent.
-This wavelength-shifting is also used to match the quantum efficiency as a function of wavelength for the photodetectors being used.
-One typical scintillator mixture uses pseudo-cumene as the solvent with 1-5~g/L of PPO as the fluor.
-This mixture has a peak emission at about 400~nm where photomultiplier tubes (PMTs) are most sensitive and the pseudo-cumene is relatively transparent.
+The molecules in liquid scintillators are not isolated. Radiationless
+processes transfer energy between the molecules. The probability of
+energy transfer between molecules increases as the overlap between the
+molecule's absorption and emission spectra increase. The absorption
+and emission spectra overlap at some level for all
+molecules. Consequently if there is only one type of molecule in the
+scintillator cocktail, the light output would be reduced due to
+inefficiencies in the energy transfer through multiple absorption and
+reemission processes. Aromatic solutes or fluors are added to the
+primary solvent to shift the wavelength of the photons to longer
+wavelengths where the scintillator is more transparent. This
+wavelength-shifting is also used to match the quantum efficiency as a
+function of wavelength for the photodetectors being used. One typical
+scintillator mixture uses pseudo-cumene as the solvent with 1-5~g/L of
+PPO as the fluor. This mixture has a peak emission at about 400~nm
+where bialkali photomultiplier tubes (PMTs) are most sensitive and the
+pseudo-cumene is relatively transparent.
 
-A good liquid scintillator will produce $\sim$10,000 photons isotropically per MeV of deposited energy.
-Although less abundant, if a particle is moving faster than the speed of light in the medium then Cerenkov light will be produced as well.
-This light is emitted in a cone pointed in the direction of the particles trajectory and with a continuous spectrum wavelengths weighted toward shorter wavelengths but extending well into the red,
+A good liquid scintillator will produce $\sim$10,000 photons
+isotropically per MeV of deposited energy. Although less abundant,
+Cerenkov light will be produced as well if a particle is moving faster
+than the speed of light in the medium.  This light is emitted in a
+cone pointed in the direction of the particles trajectory and with a
+continuous spectrum weighted toward shorter wavelengths
+but extending well into the red. The spectrum is described by:
 \begin{equation}
@@
 \end{equation}
-where $n(\lambda)$ is the wavelength-dependent index of refraction and $\beta$ is the velocity of the incoming particle.
-The Cerenkov light produced at wavelengths shorter than the absorption cutoff of the scintillator will be absorbed and re-emitted as isotropic light, but wavelengths longer than this cutoff will propagate across the detector retaining their directional information, roughly 60 photons per MeV assuming a 400~nm absorption cutoff\cite{qdot}.
-These undisturbed Cerenkov photons will have timing determined by the group velocity in the liquid,
+where $n(\lambda)$ is the wavelength-dependent index of refraction and
+$\beta$ is the velocity of the incoming particle. The Cerenkov light
+produced at wavelengths shorter than the absorption cutoff of the
+scintillator will be absorbed and re-emitted as isotropic light, but
+wavelengths longer than this cutoff will propagate across the
+detector, retaining their directional information. The yield is
+roughly 60 photons per MeV, assuming a 400~nm absorption
+cutoff\cite{qdot}. These undisturbed Cerenkov photons will have timing
+determined by the group velocity in the liquid,
 \begin{equation}
@@
 \end{equation}
-These longer wavelengths Cerenkov photons arrive in principle before the scintillation light which is slowed by both the scintillation processes and the shorter wavelengths involved.
-Thus, with sufficient timing resolution and sensitivity to longer wavelength, it should be possible to separate the directional Cerenkov light and the isotropic scintillation light in time and then reconstruct the direction of the initial particle.
+The longer wavelengths Cerenkov photons typically arrive before
+the scintillation light, which is slowed by both the scintillation
+processes and the shorter wavelengths involved. Thus, with sufficient
+timing resolution and sensitivity to longer wavelength, it should be
+possible to separate the directional Cerenkov light and the isotropic
+scintillation light in time and then reconstruct the direction of the
+initial particle.
 

### 2013-07-17 Direction.tex
Text from Matt and update the acknowledgements. --Lindley

@@
 \label{reconstruction_sec}
-Hopefully results
+WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration).
+It provides a framework for generic event cleaning, track reconstruction, and particle id, and it comes equipped with variety of pre-built algorithms.
+A collaboration between Iowa State, the University of Chicago, and Argonne National Laboratory has continued adding to the code, developing new track-fitting techniques for Water Cherenkov detectors built around the unique capabilities of microchannel plate photodetectors (MCPs): timing resolutions below 100 picoseconds and imaging resolutions below a centimeter.
+In this paper, we use only the simple, low-level reconstruction tools native to the package, though in future works we hope to implement more sophisticated techniques.
+
+The simplest approach to vertex reconstruction, used in this paper, is commonly referred to as a “point fit”.
+It assumes that, to first order, the light was emitted a single point in space-time (x0,y0,z0,t0).
+In reality, the light is emitted along the extended, multi-scattered track of the electron.
+However, at the energies discussed in this paper, the extent of the electron track is small (a few cm) compared to the scale of the detector, which has a radius of 6 meters.
+It is also small compared with the vertex resolutions typical of experiments with this scale.
+As starting point, the assumption works well.
+
+The first algorithm we use to find a seed vertex relies on exact numerical calculations of possible vertex candidates.
+Given a single point source, we need four constraints to solve for the four unknowns of the vertex.
+A collection of N random quadruplets of hits is selected and used to solve for N corresponding vertex candidates.
+Some of these solutions will produce nonsensical results due to optical scattering of the emitted light, slow decay constant of the scintillation, finite detector resolution, and the limitations of our assumption of a point source.
+Nonetheless, over the N unique quadruplets, we expect some of the solutions to lie near the true event vertex.
+In this paper we chose N to be 80, determined by trial and error.
+However, given many hundreds of hits, the number of possible 4-hit combinations is much larger and one can choose a different N to better suit one’s particular experimental context.
+
+Once the 80 vertex candidates have been solved for, we test the goodness of each candidate and select the one that best fits the full ensemble of hits.
+Distances are calculated between each hit and the candidate point-source.
+From these distances, we calculate the transit time, given the speed of light in our scintillator medium.
+This transit time plus the t0 of the 4-vertex provides a predicted arrival time for each hit.
+The “time residual” is the calculated as difference between the predicted and the measured arrival time, given the vertex candidate.
+Since we assume a point-source in this paper, we use what we call the “simple time residual”, as distinct from the “extended time residual”, which assumes an extended, track-like geometry.
+The distribution of time residuals for each hit with respect to a vertex candidate has a width that is minimized when the hypothesis vertex is close to the true vertex.
+The t0 of the vertex candidate is left to float, and is adjusted to maximize the likelihood comparison with a Gaussian test function.
+The width of that gaussian test function is hand optimized to correspond with the width of the “best fit”.
+The vertex candidate with the largest maximum is selected to serve as the best candidate seed vertex.
+Once a single best seed vertex has been chosen, the same time-residual figure-of-merit is used in a MINUIT-based fit over a continuum of possible vertices in the neighborhood of the seed.
+
+Once the vertex fitting is completed, we determine the direction by taking the centroid of all vectors pointing from the fitted event vertex and the hits on the detector.
+Since our initial time cut greatly enhances the purity of the Cherenkov light in this sample of hits and, since the Cherenkov light is directional (compared to the isotropic scintillation light), this fit provides a good measure of the track direction.
+This approach does not rely on the highly constrained Cherenkov geometry of the event, which contains further information.
+Future work will focus on better algorithms.
+Nonetheless, for a proof-of-principles, this method works sufficiently well..
 
@@
 \section{Acknowledgements}
-The authors would like to thank H.
-Frisch for useful discussion and comments in preparing this work.
-The authors would like to thank K.
-Arisaka for discussion on the capabilities of traditional PMTs in addition to HPDs.
-Finally, L.
-Winslow would like to thank J.M.
-Conrad for suggesting that the LAPPD may have sufficient timing resolution to allow the reconstruction of the electron direction in $0\nu\beta\beta$ and that a collaboration with the University of Chicago should be pursued.
+L.
+Winslow would like to thank for useful discussion on the topic.
+L.
+Winslow would also like to thank K.
+Arisaka for discussions on the possible reach of traditional PMTs and the characteristics of HPDs.
 A.
-Elagin and M.
+Elagin, H.J.
+Frisch and M.
 Wetstein are supported by DOE XXXX.

### 2013-07-17 Direction.tex
adding plots from Andrey to text. --Lindley
Pulled the time distribution plots together in one two-column figure. Removed the 4th time distribution plot with the QD emsission spectrum. Changed caption text and updated the references to the plots. Other minor edits of text in sections 3 to 7. --Christoph

@@
 \usepackage{multirow}
-\usepackage{graphicx}
+\usepackage{graphicx,subfigure}
 \usepackage{times}
@@
 \label{detector_timing_sec}
-\begin{figure}
+\begin{figure*}
         \begin{center}
-        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}
-        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with default settings: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
+        \subfigure[Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        \subfigure[Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        \subfigure[Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with different settings.
+PEs from Cerenkov light (black) and scintillation light (grey) are compared.
+The dashed vertical line illustrates a time cut at 34~ns.
+(a) Default simulation including bialkali photocathode and TTS = 0.1~ns ($\sigma$).
+The number of PEs per event after the 34.0~ns time cut is 171 from scintillation and 108 from Cerenkov.
+(b) Default simulation settings but for TTS = 1.28~ns (KamLAND 17 in.
+PMTs).
+The number of PEs per event after the 34.0~ns time cut is 349 from scintillation and 88 from Cerenkov.
+(c) Default simulation settings but for GaAsP photocathode.
+The number of PEs per event after the 34.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov.
+\label{time_plots_comparison}}
+        \end{center}
+\end{figure*}
+%%\begin{figure}
+%%        \begin{center}
+%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}
+%%        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with default settings: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
 PEs from Cerenkov light (black) and scintillation light (red) are compared.
@@
 \label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME}}
-        \end{center}
-\end{figure}
+%%        \end{center}
+%%\end{figure}
 \begin{figure}
@@
 \end{figure}
-\begin{figure}
-        \begin{center}
-        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}
-        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with TTS = 1.28~ns: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 1.28~ns ($\sigma$).
+%%\begin{figure}
+%%        \begin{center}
+%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}
+%%        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with TTS = 1.28~ns: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 1.28~ns ($\sigma$).
 PEs from Cerenkov light (black) and scintillation light (red) are compared.
@@
 \label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME}}
-        \end{center}
-\end{figure}
+%%        \end{center}
+%%\end{figure}
 
 First we discuss the default simulation settings which are described in the previous section.
-Figure \ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME} shows the TTS-smeared photoelectron detection time for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x-axis.
+Figure \ref{time_plot_comparison} (a) shows the TTS-smeared photoelectron detection time for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x-axis.
 The photoelectrons induced by Cerenkov light arrive earlier, as expected due to the instantaneous emission and the differences in the photon speed.
@@
 If the 17 inch KamLAND PMTs \cite{tbd} (TTS = 1.28~ns) are used in the simulation, the broadening of the time distributions leads to a strongly decreased ratio of Cerenkov over Scintillation light after the time cut (see FIG.
-\ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME}).
+\ref{time_plots_comparison} (b)).
 This demonstrates that low photodetector TTS is critical for directionality reconstruction and motivates use of novel detector types.
@@
 Therefore, a simulation has been run with the QE of an extended red-sensitive GaAsP photocathode (digitized Hamamatsu R3809U-63 QE data was used).
-
-\begin{figure}
-        \begin{center}
-        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}
-        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with red-sensitive photocathode: Detector diameter = 6.5~m, red-sensitive GaAsP photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
+%%\begin{figure}
+%%        \begin{center}
+%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}
+%%        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with red-sensitive photocathode: Detector diameter = 6.5~m, red-sensitive GaAsP photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
 PEs from Cerenkov light (black) and scintillation light (red) are compared.
@@
 \label{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME}}
-        \end{center}
-\end{figure}
-
-Figure \ref{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME} shows the results for the modified simulation with high QE in the red spectral region.
+%%        \end{center}
+%%\end{figure}
+Figure \ref{time_plots_comparison} (c) shows the results for the modified simulation with high QE in the red spectral region.
 The higher absolute number of PE coming from Cerenkov light (factor of $\approx$ 2) and the increased Cerenkov/scintillation ratio (factor of $\approx$ 1.6) after the time cut would significantly enhance the directionality reconstruction capabilities.
@@
 
-\begin{figure}
-        \begin{center}
-        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME.pdf}
-        \caption[]{PE times for the simulation of 1000 electrons (5~MeV) with a quantum dot emission spectrum centered at 384~nm: Detector diameter = 6.5~m, bialkali photocathode, QD emission spectrum at 384~nm, TTS = 0.1~ns ($\sigma$).
+%%\begin{figure}
+%%        \begin{center}
+%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME.pdf}
+%%        \caption[]{PE times for the simulation of 1000 electrons (5~MeV) with a quantum dot emission spectrum centered at 384~nm: Detector diameter = 6.5~m, bialkali photocathode, QD emission spectrum at 384~nm, TTS = 0.1~ns ($\sigma$).
 PEs from Cerenkov light only (black) and scintillation light only (red) are compared.
@@
 \label{6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME}}
-        \end{center}
-\end{figure}
+%%        \end{center}
+%%\end{figure}
 
@@
 Compared to the default case shown in FIG.
-\ref{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME} the separation is worse because the scintillation light wavelengths are higher than in the KamLAND emission spectrum.
+\ref{time_plot_comparison} (a) the separation is worse because the scintillation light wavelengths are higher than in the KamLAND emission spectrum.
 However, advances in the production of commercial quantum dot samples could yield quantum dots which have similar, single peak emission shapes at lower wavelengths.
@@
 This peak emission value has been measured for other types of QDs, however with a much more pronounced tail \cite{tbd}.
-The resulting PE time distribution shows a clearer separation of Cerenkov and Scintillation light (see FIG.
-\ref{6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME}).
-After the 34.0~ns cut on the TTS-smeared PE time we obtain slightly better numbers compared to the default simulation: The number of Cerenkov-induced PE after the time cut is unchanged while the number of PEs coming from scintillation light is decreased due to the higher average photon travel times.
+The resulting PE time distribution shows a slightly clearer separation of Cerenkov and Scintillation light.
+After the 34.0~ns cut on the TTS-smeared PE time we obtain a higher Cerenkov/scintillation ratio compared to the default simulation: The number of Cerenkov-induced PE after the time cut is unchanged while the number of PEs coming from scintillation light is decreased due to the higher average photon travel times.
 
@@
 
+\begin{figure}
+        \begin{center}
+        \includegraphics[scale=0.20]{graphs/hDirX.pdf}
+        \includegraphics[scale=0.20]{graphs/hDirY.pdf}
+        \includegraphics[scale=0.20]{graphs/hDirZ.pdf}	
+         \includegraphics[scale=0.20]{graphs/hVtxX.pdf}
+        \includegraphics[scale=0.20]{graphs/hVtxY.pdf}
+        \includegraphics[scale=0.20]{graphs/hVtxZ.pdf}	
+        \caption[]{ We need a caption.}
+        \end{center}
+\end{figure}
+
 \section{Conclusion}

### 2013-07-18 Direction.tex
Added Reference for speed of light in glass and added sentence below the corresponding equation. --Christoph

@@
 cutoff\cite{qdot}. These undisturbed Cerenkov photons will have timing
-determined by the group velocity in the liquid,
+determined by the group velocity \cite{group_velocity_article} in the liquid,
 \begin{equation}
@@
 \end{equation}
+In Geant4 optical photons are assigned the group velocity in the wavelength 
+region of normal dispersion ($dn(\lambda)/dlog(\lambda)$>0) which is relevant to 
+our study. 
 The longer wavelengths Cerenkov photons typically arrive before
@@
 In section \ref{reconstruction_sec}, we discuss first results on more realistic reconstruction.
+\begin{figure*}[tbh]
+        \begin{center}
+        \subfigure[Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        \subfigure[Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        \subfigure[Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with different settings.
+PEs from Cerenkov light (black) and scintillation light (grey) are compared.
+The dashed vertical line illustrates a time cut at 34~ns.
+(a) Default simulation including bialkali photocathode and TTS = 0.1~ns ($\sigma$).
+The number of PEs per event after the 34.0~ns time cut is 171 from scintillation and 108 from Cerenkov.
+(b) Default simulation settings but for TTS = 1.28~ns (KamLAND 17 in.
+PMTs).
+The number of PEs per event after the 34.0~ns time cut is 349 from scintillation and 88 from Cerenkov.
+(c) Default simulation settings but for GaAsP photocathode.
+The number of PEs per event after the 34.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov.
+\label{time_plots_comparison}}
+        \end{center}
+\end{figure*}
 Before we discuss the simulation results for different simulation settings in the following sections, we highlight the effects which contribute to the timing of the scintillator detector system.
@@
 Multiparameter optimization problem.
-%% nuebar scattering --> one of next papers
+%% nuebar scattering --> one of next paper
 
@@
 \label{detector_timing_sec}
-\begin{figure*}
-        \begin{center}
-        \subfigure[Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
-        \subfigure[Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
-        \subfigure[Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
-        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with different settings.
-PEs from Cerenkov light (black) and scintillation light (grey) are compared.
-The dashed vertical line illustrates a time cut at 34~ns.
-(a) Default simulation including bialkali photocathode and TTS = 0.1~ns ($\sigma$).
-The number of PEs per event after the 34.0~ns time cut is 171 from scintillation and 108 from Cerenkov.
-(b) Default simulation settings but for TTS = 1.28~ns (KamLAND 17 in.
-PMTs).
-The number of PEs per event after the 34.0~ns time cut is 349 from scintillation and 88 from Cerenkov.
-(c) Default simulation settings but for GaAsP photocathode.
-The number of PEs per event after the 34.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov.
-\label{time_plots_comparison}}
-        \end{center}
-\end{figure*}
 %%\begin{figure}
@@
 %%\end{figure}
-\begin{figure}
-        \begin{center}
-        \includegraphics[scale=0.40]{graphs/cos_psi_comp_h.pdf}
-        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction for 1000 events (5~MeV electrons).
-Default simulation settings were used.
-A 33.0~ns time cut (black) is compared with a 34.0~ns time cut (red).
-\label{cerenkov_cone}}
-        \end{center}
-\end{figure}
 %%\begin{figure}
@@
 For scintillation the number of PEs after the time cut (171) is only 3.2~\% of the total scintillation-induced PEs (5445).
+\begin{figure}[bt]
+        \begin{center}
+        \includegraphics[scale=0.40]{graphs/cos_psi_comp_h.pdf}
+        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction for 1000 events (5~MeV electrons).
+Default simulation settings were used.
+A 33.0~ns time cut (black) is compared with a 34.0~ns time cut (red).
+\label{cerenkov_cone}}
+        \end{center}
+\end{figure}
 

### 2013-07-18 Direction.tex
Improved captions. --Lindley

@@
         \begin{center}
-        \subfigure[Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
-        \subfigure[Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
-        \subfigure[Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        \subfigure[ Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        \subfigure[ Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        \subfigure[ Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
         \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with different settings.
@@
 The dashed vertical line illustrates a time cut at 34~ns.
-(a) Default simulation including bialkali photocathode and TTS = 0.1~ns ($\sigma$).
+(a) Default simulation: bialkali photocathode and TTS = 0.1~ns ($\sigma$).
 The number of PEs per event after the 34.0~ns time cut is 171 from scintillation and 108 from Cerenkov.
@@
         \includegraphics[scale=0.40]{graphs/cos_psi_comp_h.pdf}
-        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction for 1000 events (5~MeV electrons).
-Default simulation settings were used.
-A 33.0~ns time cut (black) is compared with a 34.0~ns time cut (red).
+        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction for 1000 events (5~MeV electrons at the center in x direction): $\cos(\Psi) = x_{hit}/\vec{r}_{hit}$.
+Default simulation settings were used and both Cerenkov and Scintillation light are included.
+A 33.0~ns time cut (black) is compared with a 34.0~ns time cut (grey).
 \label{cerenkov_cone}}

### 2013-07-18 Direction.tex
Add J.M. Conrad to acknowledgements. --Lindley
Add Andrey Figures and put them in a nice full page figure. --Lindley

@@
 
-\section{Energy Dependence and Detector Size}
-\label{edep_size_sec}
-include this?
-
-The number of Cerenkov photons decreases disproportionally with decreasing electron energy.
-The number of PEs after the 34.0~ns time cut for 1~MeV electrons is 37 PE from scintillation and 13 PE from Cerenkov light.
-This illustrates that directionality reconstruction benefits from higher initial energies but even at 1~MeV there might be accessible directionality information.
-
-An additional set of simulations of a 0.65~m diameter ($\approx$ 1~m$^3$) detector with otherwise default parameters was performed.
-In the smaller detector the chromatic dispersion is not as important and absorption in the scintillator bulk is reduced.
-The latter effect increases the absolute number of photoelectrons.
-The former effect, on one hand, reduces the separation in time between Cerenkov and Scintillation light.
-On the other hand, it reduces the broadening of both time distributions which helps more sophisticated vertex and track reconstruction with the help of fast photodetectors like for example the LAPPDs.
-
-%tbd (calculate numbers): However, with the simplified effective vertex reconstruction and a time cut of 3.4~ns we get  PEs from scintillation and 87 PEs from Cerenkov.
-The time cut was adjusted such that the fraction of selected PEs from Cerenkov light matches the fraction in the large detector ($\approx$ 58\%).
-The QD spectrum discussed in section \ref{scintillator_emission_sec} (both shifted and unshifted) did not change the PE time distribution significantly.
-We also simulated the red-sensitive photocathode discussed in section \ref{detector_wavelength_response_sec} in the small detector.
-The amount of PE increases again significantly and with the 3.265~ns time cut the number of PE coming from Cerenkov light is 182 while the number of PE from scintillation light is 96.
-
-In summary, the photodetector properties (TTS and QE) become more important relative to the scintillator emission and absorption spectra when the detector size is decreased from 6.5~m to 0.65~m.
-
 %%idea: use (some) PMTs with high QE for high lambda (and/or low QE for normal lambda). 
@@
 \label{reconstruction_sec}
+
+\begin{figure*}[tbh]
+        \begin{center}
+        \subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
+        \subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
+        \subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
+        \subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
+        \subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
+        \subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
+        \caption[]{We need a caption}
+        \end{center}
+\end{figure*}
+
 WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration).
@@
 
-\begin{figure}
-        \begin{center}
-        \includegraphics[scale=0.20]{graphs/hDirX.pdf}
-        \includegraphics[scale=0.20]{graphs/hDirY.pdf}
-        \includegraphics[scale=0.20]{graphs/hDirZ.pdf}	
-         \includegraphics[scale=0.20]{graphs/hVtxX.pdf}
-        \includegraphics[scale=0.20]{graphs/hVtxY.pdf}
-        \includegraphics[scale=0.20]{graphs/hVtxZ.pdf}	
-        \caption[]{ We need a caption.}
-        \end{center}
-\end{figure}
-
 \section{Conclusion}
@@
 L.
-Winslow would like to thank for useful discussion on the topic.
+Winslow would like to thank J.M.
+Conrad for useful discussion on the topic.
 L.

### 2013-07-18 Direction.tex
Add new text from Matt and figure caption. --Lindley

@@
         \subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \caption[]{We need a caption}
+        \caption[]{\label{fig:reco} (Top) Reconstructed direction, $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|}, \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons (5MeV).
+In the simulation the electrons are produced along the X-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate from the center of the detector, $\vec{r}$ = (0,0,0).
+Detector diameter, $R$ = 6.5~m.
+Only photons with arrival time of $T<$ 34~ns are used in the reconstruction.
+The quantum efficiency of the bialkali photocathode is taken into account.
+(Bottom) Reconstructed vertex position, $(x,y,z)$, for the simulation of 1000 electrons (5~MeV).
+In the simulation the electrons are produced along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate from the center of the detector, $\vec{r}$ = (0,0,0).
+Detector diameter, $R$ = 6.5~m.
+Only photons with arrival time of $T<$ 34~ns are used in the reconstruction.
+The quantum efficiency of the bialkali photocathode is taken into account.
+}
         \end{center}
@@
 
-WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration).
+WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration) \cite{Andy}.
+It provides a framework for generic event cleaning, track reconstruction, and particle id, and it comes equipped with variety of pre-built algorithms.
+A collaboration between Iowa State, the University of Chicago, and Argonne National Laboratory has continued adding to the code, developing new track-fitting techniques for Water Cherenkov detectors based on advanced photosensors with sub-cm imaging capabilities and timing resolutions below 100 picoseconds \cite{fast_group_and_LAPPD}.
+In this paper, we use only the simple, low-level reconstruction tools native to the package, though in future works we hope to implement more sophisticated algorithms.
+
+The results presented in this paper rely on a simple vertex reconstruction algorithm, commonly known as a "point fit" \cite{reference_Super_K}.
+It assumes that all of the scintillation and Cherenkov light is emitted from a single point in space-time (x0,y0,z0,t0).
+In actuality, the light is emitted along an extended, multi-scattered electron track.
+However, at the energies discussed in this paper, the extent of this track is small (a few cm) compared to the scale of the detector (R=6 meters) and thus typical photon transit distances.
+
+The first step of the reconstruction process relies on exact numerical calculations of vertex candidates from quadruplets of hits.
+Given a single point source, we need four constraints to solve for the four unknowns of the vertex \cite{Michael Smy} This approach would provide an exact solution in the case of four prompt, unscattered photons originating from a common point.
+However, many of these randomly chosen quadruplets will produce anomalous solutions due to real world effects such as optical scattering, delayed emission, and deviations from the point-like geometry.
+Nonetheless, we found that any chosen subset of 80 quadruplets was a sufficiently large ensemble to assure that some solutions will be close to the true vertex.
+
+Once a small set of vertex candidates have been solved for, we test the goodness of each candidate and select the one that best fits the full ensemble of photon hits.
+The goodness of fit is determined based on the distribution of an observable known as the "point time residual" [super K].
+The point time residual is calculated by taking the difference between the measured time of a photon hit, and the predicted time of the hit, given its distance from the vertex hypothesis, the speed of light in the scintillator, and the hypothesized t0 of the event.
+The width of the time residual distribution over all hits is minimized when the hypothesized vertex is near the true vertex.
+Based on this figure of merit, we select the vertex with the narrowest time residual distribution from among the 80 candidates.
+This best candidate is used as the seed value for a MINUIT-based optimization [MINUIT], where the vertex is allowed to vary over a continuum of locations until the global optimum is found.
+
+The direction of the electron track is determined by taking the centroid of all vectors pointing from the fitted vertex to the hits on the detector.
+Since the Cherenkov light his highly directional, and since our timing cut enhances the purity of the Cherenkov light in the sample, this calculation provides a good measure of the track direction.
+Future work will focus on more sophisticated algorithms, taking better advantage of the highly constrained Cherenkov geometry.
+For the purpose of this paper, the current method provides a very good proof-of-principle.
+
+For the purpose of testing the reconstruction algorithm described above we use 1000 simulated electrons with the energy of 5~MeV.
+The electrons are similated at the center of the detector, $\vec{r}$ = (0,0,0), along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0).
+
+Figure~\ref{fig:reco} show vertex reconstruction.
+The vertex is reasonably well reconstructed around center of the detector, $\vec{r}$ = (0,0,0).
+The RMS of the distributions for all three reconstructed coordinates are smaller than 3.5~cm.
+The shift along x-axis is attributed to the fact that electron is moving towards that direction and has a finite track length while the reconstruction algorithm treats all photons as they were emitted from the same point.
+
+The reconstruction of the direction is shown in Fig.~\ref{fig:reco}. For majority of the events we are able to reconsruct the direction along x-axis.
+
+%
+%This is Matt's text from July 17.
+%
+%WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration).
 It provides a framework for generic event cleaning, track reconstruction, and particle id, and it comes equipped with variety of pre-built algorithms.
@@
 
-The simplest approach to vertex reconstruction, used in this paper, is commonly referred to as a “point fit”.
+%The simplest approach to vertex reconstruction, used in this paper, is commonly referred to as a Òpoint fitÓ.
 It assumes that, to first order, the light was emitted a single point in space-time (x0,y0,z0,t0).
@@
 
-The first algorithm we use to find a seed vertex relies on exact numerical calculations of possible vertex candidates.
+%The first algorithm we use to find a seed vertex relies on exact numerical calculations of possible vertex candidates.
 Given a single point source, we need four constraints to solve for the four unknowns of the vertex.
@@
 In this paper we chose N to be 80, determined by trial and error.
-However, given many hundreds of hits, the number of possible 4-hit combinations is much larger and one can choose a different N to better suit one’s particular experimental context.
+However, given many hundreds of hits, the number of possible 4-hit combinations is much larger and one can choose a different N to better suit oneÕs particular experimental context.
 
-Once the 80 vertex candidates have been solved for, we test the goodness of each candidate and select the one that best fits the full ensemble of hits.
+%Once the 80 vertex candidates have been solved for, we test the goodness of each candidate and select the one that best fits the full ensemble of hits.
 Distances are calculated between each hit and the candidate point-source.
@@
 This transit time plus the t0 of the 4-vertex provides a predicted arrival time for each hit.
-The “time residual” is the calculated as difference between the predicted and the measured arrival time, given the vertex candidate.
-Since we assume a point-source in this paper, we use what we call the “simple time residual”, as distinct from the “extended time residual”, which assumes an extended, track-like geometry.
+The Òtime residualÓ is the calculated as difference between the predicted and the measured arrival time, given the vertex candidate.
+Since we assume a point-source in this paper, we use what we call the Òsimple time residualÓ, as distinct from the Òextended time residualÓ, which assumes an extended, track-like geometry.
 The distribution of time residuals for each hit with respect to a vertex candidate has a width that is minimized when the hypothesis vertex is close to the true vertex.
 The t0 of the vertex candidate is left to float, and is adjusted to maximize the likelihood comparison with a Gaussian test function.
-The width of that gaussian test function is hand optimized to correspond with the width of the “best fit”.
+The width of that gaussian test function is hand optimized to correspond with the width of the Òbest fitÓ.
 The vertex candidate with the largest maximum is selected to serve as the best candidate seed vertex.
@@
 
-Once the vertex fitting is completed, we determine the direction by taking the centroid of all vectors pointing from the fitted event vertex and the hits on the detector.
+%Once the vertex fitting is completed, we determine the direction by taking the centroid of all vectors pointing from the fitted event vertex and the hits on the detector.
 Since our initial time cut greatly enhances the purity of the Cherenkov light in this sample of hits and, since the Cherenkov light is directional (compared to the isotropic scintillation light), this fit provides a good measure of the track direction.

### 2013-07-19 Direction.tex
Changed DC QE sentence. --Christoph

@@
 Two important photodetector properties have also been varied: Transit time spread (TTS, default $\sigma$ = 0.1~ns) and wavelength-dependent quantum efficiency (QE) for photoelectron production.
-The default is the bialkali photocathode of the Double Chooz photomultiplier tubes (PMTs) \cite{tbd}.
-It has been verified to be the same type as the KamLAND photocathode.
-We used the Double Chooz QE because the data was available with finer spacing in wavelength.
+Default is the QE of a bialkali photocathode (Hamamatsu R7081 PMT), digitized values come from the Double Chooz \cite{dctwo} Monte Carlo simulation.
+The KamLAND 17 inch PMTs use the same photocathode type (bialkali) with similar quantum efficiency.
 

### 2013-07-19 Direction.tex
Changes by Matt on the reconstruction section and Acknowledgments. Changes by Christoph and Lindley to sections 3 through 6. Addition of detector view with 0nuBB event (version 0). New references. Clean up of deprecated text and comments. July 19 (version 1 of draft). --Christoph

@@
 excited, predominantly the the non-localized electrons in the
-$\pi$-bonds of the phenyl groups~\cite{scintillator_ref}. Vibrational
+$\pi$-bonds of the phenyl groups~\cite{birks_book}. Vibrational
 and rotational modes of the molecules are turned into heat within
@@
 \label{eqGroup}
-v_{g}(\lambda) = \frac{c_{vacuum}}{n(\lambda) + dn(\lambda)/dlog(\lambda)}.
+v_{g}(\lambda) = \frac{c_{vacuum}}{n(\lambda) - dn(\lambda)/d\textnormal{log}(\lambda)}.
 \end{equation}
 In Geant4 optical photons are assigned the group velocity in the wavelength 
-region of normal dispersion ($dn(\lambda)/dlog(\lambda)$>0) which is relevant to 
-our study. 
+region of normal dispersion which is relevant to our study. 
 The longer wavelengths Cerenkov photons typically arrive before
@@
 In order to study the effects relevant to directionality reconstruction in liquid scintillators, a Geant4 \cite{geant4one,geant4two} simulation has been set up.
-The presented simulation application has been developed using standard Geant4 functionality for the optical model of the liquid scintillator.
-Version Geant4.9.6 is used and the simulation application has been built from a customized official application example.
-The simulation is kept simple to provide generally applicable information about the main factors in directionality reconstruction.
+The presented simulation has been developed using standard Geant4 (version 4.9.6) functionality for the liquid scintillator optical model.
+The simulation has been built from a customized official Geant4 application example and it is kept rather generic to provide generally applicable information about the main factors in directionality reconstruction.
 
-The implemented detector geometry is a sphere of 6.5~m (or 0.65~m) diameter filled with liquid scintillator.
-Default scintillator properties have been chosen to match the KamLAND scintillator \cite{tbd}: 80~\% n-Dodecane, 20~\% Pseudocumene (1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole).
-The implemented scintillator properties include the atomic composition and density ($\rho$ = 0.7752~g/ml), the wavelength-dependent attenuation length and refractive index, the scintillation emission spectrum, emission rise time ($\tau_r$ = 1.0~ns) and emission decay time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 0.13), scintillator light yield (LY, 9030.5 photons/MeV) and the Birks constant ($kB$ = 0.106~mm/MeV) \cite{tbd}.
+\begin{figure}
+        \begin{center}
+        \includegraphics[scale=0.25]{graphs/detector_view.png}
+        \caption[]{Two back-to-back electrons with 1.41~MeV each (equally divided energy of $^116$Cd $0\nu\beta\beta$ decay) at the center of the sphere with initial directions along the x and -x-axis.
+Only Cerenkov photons are drawn to illustrate the directionality of the event.
+\label{detector_view}}
+        \end{center}
+\end{figure}
+
+The implemented detector geometry is a sphere of 6.5~m diameter filled with liquid scintillator (see Figure \ref{detector_view}).
+Default scintillator properties have been chosen to match the KamLAND scintillator \cite{kamland2003}: 80~\% n-Dodecane, 20~\% Pseudocumene (1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole).
+The implemented scintillator properties include the atomic composition and density ($\rho$ = 0.78~g/ml), the wavelength-dependent attenuation length \cite{tajimaMaster} and refractive index \cite{OlegThesis}, the scintillation emission spectrum, emission rise time ($\tau_r$ = 1.0~ns) and emission decay time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 0.13) \cite{tajimaThesis}, scintillator light yield (LY, 9030 photons/MeV) and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV) \cite{ChrisThesis}.
 Variations from the baseline KamLAND case are discussed below when applicable.
-Reemission of absorbed photons in the scintillator bulk volume and scattering has not been included.
-Reflections, reemission and scattering are expected to be corrections to the main effects studied in this paper and are subject of future work.
+Reflections, reemission of absorbed photons in the scintillator bulk volume and scattering has not been included.
+These effects are expected to be corrections to the main effects studied in this paper and are subject of future work.
 
-The sphere surface is used as the photodetector.
-It is treated as fully absorbing (no reflections) with a coverage of 100~\%.
-Two important photodetector properties have also been varied: Transit time spread (TTS, default $\sigma$ = 0.1~ns) and wavelength-dependent quantum efficiency (QE) for photoelectron production.
+The inner sphere surface is used as the photodetector.
+It is treated as fully absorbing (no reflections) with a photodetector coverage of 100~\%.
+Two important photodetector properties have been varied: Transit time spread (TTS, default $\sigma$ = 0.1~ns) and wavelength-dependent quantum efficiency (QE) for photoelectron production.
 Default is the QE of a bialkali photocathode (Hamamatsu R7081 PMT), digitized values come from the Double Chooz \cite{dctwo} Monte Carlo simulation.
-The KamLAND 17 inch PMTs use the same photocathode type (bialkali) with similar quantum efficiency.
+We note that the KamLAND 17 inch PMTs use the same photocathode type (bialkali) with similar quantum efficiency.
 
-%deprecated text: fake vertex reconstruction not needed anymore. 
-%To this end, we use the KamLAND vertex resolution of 12.0~cm \cite{tbd} ($\sigma$ in one dimension) to randomly draw a vertex around the known true vertex for each event.
-This vertex is subsequently used as a reconstructed vertex and consequently, a time of flight (TOF) correction relative to the center of the sphere is applied for each photon hit depending on the distance between the reconstructed vertex and the hit position on the sphere.
-For the TOF correction a single effective speed of light is used.
-This effective speed was extracted at the mean wavelength of both Scintillation and Cerenkov photons (408~nm).
-The 12.0~cm vertex resolution corresponds to a time resolution broadening of about 0.6~ns ($\sigma$).
-Note that the vertex resolution itself is dependent on scintillator and detector properties.
-For example, the PMTs used in KamLAND have a TTS of 1.28~ns ($\sigma$) and significant improvement of the vertex resolution could be possible with faster photodetectors.
-In section \ref{reconstruction_sec}, we discuss first results on more realistic reconstruction.
 \begin{figure*}[tbh]
 \begin{center}
-        \subfigure[ Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
-        \subfigure[ Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
-        \subfigure[ Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        \subfigure[ ~Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        \subfigure[ ~Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        \subfigure[ ~Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
         \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with different settings.
@@
 (a) Default simulation: bialkali photocathode and TTS = 0.1~ns ($\sigma$).
-The number of PEs per event after the 34.0~ns time cut is 171 from scintillation and 108 from Cerenkov.
-(b) Default simulation settings but for TTS = 1.28~ns (KamLAND 17 in.
+After the 34.0~ns time cut we get 171~PEs from scintillation and 108~PEs from Cerenkov light.
+(b) Default simulation settings except for TTS = 1.28~ns (KamLAND 17 in.
 PMTs).
-The number of PEs per event after the 34.0~ns time cut is 349 from scintillation and 88 from Cerenkov.
-(c) Default simulation settings but for GaAsP photocathode.
-The number of PEs per event after the 34.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov.
+After the 34.0~ns time cut we get 349~PEs from scintillation and 88~PEs from Cerenkov light.
+(c) Default simulation settings except for a GaAsP photocathode.
+After the 34.0~ns time cut we get 226~PEs from scintillation and 229~PEs from Cerenkov light.
 \label{time_plots_comparison}}
@@
 \end{figure*}
-Before we discuss the simulation results for different simulation settings in the following sections, we highlight the effects which contribute to the timing of the scintillator detector system.
+Before the simulation results for different simulation settings are discussed in the following sections, we highlight the effects which contribute to the timing of the scintillator detector system.
 First, the simulated travel time of the initial 5~MeV electron is between 0.10 and 0.15~ns, while the travel distance is about 3~cm.
@@
 Past neutrino experiments were not highly interested in the effect of scintillation rise time which is the reason why there is a lack of accurate numbers.
-We assume a rise time of 1~ns, more detailed studies are needed in the future.
+We assume a rise time of 1.0~ns, more detailed studies are needed in the future.
 The two time constants used to describe the falling edge of the scintillator emission time distribution (quoted above) are values specific to the KamLAND scintillator.
 Third, chromatic dispersion turns out to be an important effect in a 6.5~m diameter detector at the level of precision needed for direction reconstruction.
-Due to the wavelength-dependence of the refractive index the speed of light in the scintillator (the group velocity is used in Geant4 for normal dispersion) is also wavelength dependent.
-In order to study the size of this effect, we extracted results from a simplified simulation of 5~MeV electrons at the center of the sphere where we used instantaneous scintillation and did apply neither TTS nor TOF correction.
-The true time distributions of photoelectrons were analyzed for scintillation light and Cerenkov light separately.
-Photoelectrons coming from Cerenkov light are created about 0.5~ns earlier than PEs from scintillation light on average.
+Due to the wavelength-dependence of the refractive index the speed of light in the scintillator (see Equation (\ref{eqGroup})) increases with increasing photon wavelengths for normal dispersion.
+In order to study the size of this effect, we extracted results from a simplified simulation of 5~MeV electrons at the center of the sphere where we used instantaneous scintillation and applied QE but no TTS.
+The true hit time distributions of photoelectrons were analyzed for scintillation light and Cerenkov light separately.
+Photoelectrons coming from Cerenkov light are on average created about 0.5~ns earlier than PEs from scintillation light.
 The RMS values for both the Cerenkov and scintillation light true PE time distributions are both about 0.5~ns.
@@
 In fact, even significantly lower TTS numbers are realistic with the LAPPD.
-Finally, the accuracy of the vertex reconstruction is reflected in broadening and distortions of the time spectrum (after the TOF correction has been applied).
-Another effect which is to be addressed in future work is the granularity of the photodetectors which adds to the uncertainty in the light path and thus the TOF correction.
+Finally, the accuracy of the vertex reconstruction is reflected in broadening and distortions of the time spectrum (after time-of-flight correction has been applied).
+Another effect which is to be addressed in future work is the granularity of the photodetectors which adds to the uncertainty in the light path and thus the time-of-flight correction.
 
-In Sections \ref{detector_timing_sec} to \ref{edep_size_sec}, we comparatively study the photoelectron timing for different detector properties.
-We follow the idea to increase discrimination between Cerenkov and scintillation light by detector timing.
-The main quantities provided by the Geant4 simulation are the photoelectron hit positions and the detection times after TTS has been applied.
+In Sections \ref{detector_timing_sec} to \ref{scintillator_emission_sec}, we comparatively study the photoelectron timing for different detector configurations.
+We follow the idea to increase discrimination between Cerenkov and scintillation light by accurate detector timing.
+The main relevant quantities provided by the Geant4 simulation are the photoelectron hit positions and the detection times after TTS has been applied.
 In section \ref{reconstruction_sec} these quantities are used for event reconstruction.
-%%statement that rise time is not super-crucial (see April 11 talk)
-%% include wavelength spectra, emission spectra, absorption (?), refractive index  ?
-%% Other settings: physics list, theCerenkovProcess->SetMaxNumPhotonsPerStep(20);theCerenkovProcess->SetMaxBetaChangePerStep(10.0); theCerenkovProcess->SetTrackSecondariesFirst(true); theScintillationProcess->SetScintillationYieldFactor(1.); theScintillationProcess->SetTrackSecondariesFirst(true); theScintillationProcess->SetFiniteRiseTime(true); particle gun, 5MeV e- in x direction from center, random polarization
-%% discussion: can chromatic dispersion be disentangled? With very high statistics maybe yes (see Ben Monreal, Matt Wetstein).
-%% discussion: many scintillator properties depend on each other.
-Optimization of scintillator properties could lead to better results for directionality reconstruction.
-i.e.
-slower scintillation (but vertex reco worse!, LY tends to be lower) , lower wavelength,
-%% idea: filter for part of the PMTs or two kinds of PMTs: separation by wavelength
-%% coverage is just a reduction by factor 0.22 (flat over wavelengths etc.), not studied in the paper?  
-%% at the 0.1 ns level in a 6.5 m detector one has to be careful with the input of nref. If the slope is not smooth enough the group velocity calcuation is creating wavelength-dependent distortions. 
-%% Reflections and absorption/reemission and scattering will shift events to higher times were they will likely not pass the time cut.
-They will effectively change the scintillation/Cerenkov ratio.
-Scattering might distort the angular variables and is subject of future work.
-%% In small det. the chromatic dispersion doesn't help too much. (In the small detector the detector properties are the driving factor.)
-%% discussion: some scintillator and detector properties have complicated effects on the directionality, also other requirements have to be considered (high LY, high resolutionc).
-Multiparameter optimization problem.
-%% nuebar scattering --> one of next paper
 
@@
 \label{detector_timing_sec}
-%%\begin{figure}
-%%        \begin{center}
-%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}
-%%        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with default settings: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
-PEs from Cerenkov light (black) and scintillation light (red) are compared.
-The number of PEs per event after a 34.0~ns time cut is 171 from scintillation and 108 from Cerenkov.
-\label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME}}
-%%        \end{center}
-%%\end{figure}
-%%\begin{figure}
-%%        \begin{center}
-%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}
-%%        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with TTS = 1.28~ns: Detector diameter = 6.5~m, bialkali photocathode, KamLAND scintillator emission spectrum, TTS = 1.28~ns ($\sigma$).
-PEs from Cerenkov light (black) and scintillation light (red) are compared.
-The number of PEs per event after a 34.0~ns time cut is 349 from scintillation and 88 from Cerenkov.
-\label{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME}}
-%%        \end{center}
-%%\end{figure}
 
-First we discuss the default simulation settings which are described in the previous section.
-Figure \ref{time_plot_comparison} (a) shows the TTS-smeared photoelectron detection time for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x-axis.
-The photoelectrons induced by Cerenkov light arrive earlier, as expected due to the instantaneous emission and the differences in the photon speed.
-There is however overlap with the PE time distribution coming from scintillation light.
-In order to compare simulations with different parameters to each other a fixed time cut $t \leq$ 34.0~ns (or 33.0~ns?) is applied (using the truth information of the electron starting time) instead of an event-by-event time reconstruction (see Section \ref{reconstruction_sec} for reconstruction results).
+First we discuss results for the default simulation settings which are described in the previous section.
+Figure \ref{time_plots_comparison} (a) shows the TTS-smeared photoelectron detection times for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x-axis.
+The photoelectrons induced by Cerenkov light arrive earlier, as expected due to the instantaneous emission and the higher average photon speed (higher wavelengths) compared to scintillation after QE weighting.
+There is however significant overlap with the PE time distribution coming from scintillation light.
+In order to compare simulations with different parameters to each other a fixed time cut $t \leq$ 34.0~ns (or 33.0~ns) is applied using the truth information of the electron starting time instead of an event-by-event time reconstruction (see Section \ref{reconstruction_sec} for reconstruction results).
 For the default simulation case, the number of PEs per event coming from Cerenkov light after the 34.0~ns time cut (108) is 98~\% of the total number of PEs from Cerenkov light (110).
 For scintillation the number of PEs after the time cut (171) is only 3.2~\% of the total scintillation-induced PEs (5445).
-\begin{figure}[bt]
+This indicates the effectiveness of a time cut to separate Cerenkov light from scintillation light.
+An important figure of merit when comparing different simulation settings is the ratio of Cerenkov/scintillation induced PE after the 34.0~ns time cut since a higher ratio means more directional information per PE.
+The ratio is 0.63 for the default settings.
+Note that this ratio depends on the time cut, which will be optimized in future work together with the absolute number of PEs per event passing the cut.
+\begin{figure}
         \begin{center}
         \includegraphics[scale=0.40]{graphs/cos_psi_comp_h.pdf}
-        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction for 1000 events (5~MeV electrons at the center in x direction): $\cos(\Psi) = x_{hit}/\vec{r}_{hit}$.
-Default simulation settings were used and both Cerenkov and Scintillation light are included.
+        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction along the x-axis: $\cos(\Psi) = x_{hit}/\vec{r}_{hit}$.
+1000 electron events with 5~MeV at the center are summed.
+Default simulation settings are used and both Cerenkov and scintillation light are included.
 A 33.0~ns time cut (black) is compared with a 34.0~ns time cut (grey).
@@
 
-Figure \ref{cerenkov_cone} displays the PE hit pattern after the 34.0~ns time cut has been applied (1000 events plotted together).
-Although this time cut is an oversimplification of actual time reconstruction effects, we can use it to indicate the spatial distribution of hits after timing information has been used to separate Cerenkov and scintillation light.
-The Cerenkov ring structure can be clearly seen (opening angle between 40\textdegree and 50\textdegree) which demonstrates that the directional signal conveyed by the Cerenkov photons is not erased by scattering of the initial 5~MeV electrons.
+Figure \ref{cerenkov_cone} displays the PE hit pattern after cutting on time.
+Although this time cut is a simplification of actual time reconstruction effects, we can use it to indicate the spatial distribution of hits after timing information has been used to separate Cerenkov and scintillation light.
+The Cerenkov ring structure can be clearly seen (peak at about 46\textdegree) which demonstrates that the directional signal conveyed by the Cerenkov photons is not erased by scattering of the initial 5~MeV electrons.
 
-If the 17 inch KamLAND PMTs \cite{tbd} (TTS = 1.28~ns) are used in the simulation, the broadening of the time distributions leads to a strongly decreased ratio of Cerenkov over Scintillation light after the time cut (see FIG.
+If the 17 inch KamLAND PMTs \cite{tajimaMaster,kume_1983} (TTS = 1.28~ns) are used in the simulation, the broadening of the time distributions leads to a strongly decreased ratio of Cerenkov over scintillation light (0.25) after the 34~ns time cut (see FIG.
 \ref{time_plots_comparison} (b)).
-This demonstrates that low photodetector TTS is critical for directionality reconstruction and motivates use of novel detector types.
+This shows that low photodetector TTS is critical for directionality reconstruction and motivates the use of novel detector types.
 
@@
 \label{detector_wavelength_response_sec} 
-Comlementary to lowering the photodetector TTS, optimization of the wavelength-dependent QE is promising.
+Complementary to lowering the photodetector TTS, optimization of the wavelength-dependent QE is promising.
 Since Cerenkov photons which passed 6.5~m of scintillator have higher average wavelengths than scintillation photons, a photodetector which is more sensitive at high wavelengths increases not only the absolute number of PEs but also the ratio between Cerenkov- and scintillation-induced PEs.
-Therefore, a simulation has been run with the QE of an extended red-sensitive GaAsP photocathode (digitized Hamamatsu R3809U-63 QE data was used).
-%%\begin{figure}
-%%        \begin{center}
-%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}
-%%        \caption[]{Effectively reconstructed PE times for the simulation of 1000 electrons (5~MeV) with red-sensitive photocathode: Detector diameter = 6.5~m, red-sensitive GaAsP photocathode, KamLAND scintillator emission spectrum, TTS = 0.1~ns ($\sigma$).
-PEs from Cerenkov light (black) and scintillation light (red) are compared.
-The number of PEs per event after a 34.0~ns time cut is 50.8 from scintillation and 171 from Cerenkov.
-\label{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME}}
-%%        \end{center}
-%%\end{figure}
+Therefore, a simulation has been run with the QE of an extended red-sensitive GaAsP photocathode (Hamamatsu R3809U-63 QE data was used).
+
 Figure \ref{time_plots_comparison} (c) shows the results for the modified simulation with high QE in the red spectral region.
-The higher absolute number of PE coming from Cerenkov light (factor of $\approx$ 2) and the increased Cerenkov/scintillation ratio (factor of $\approx$ 1.6) after the time cut would significantly enhance the directionality reconstruction capabilities.
+The higher absolute number of PE coming from Cerenkov light (factor of $\approx$ 2) and the increased Cerenkov/scintillation ratio (1.6) after the time cut would significantly enhance the directionality reconstruction capabilities given that high coverage can be achieved with red-sensitive photocathodes.
 
@@
 \label{scintillator_emission_sec}
-One important contributing factor to the separation in time between Cerenkov and scintillation photon hits is the higher average light speed for the photons coming from the Cerenkov effect due to their higher average wavelength.
-In this section we present two more sets of simulations where the scintillator emission spectrum was changed.
-
-Recently, the use of quantum dots (QDs) in liquid scintillators has been studied as a possibility to improve future large scale neutrino experiments \cite{tbd}.
-One important motivation is the control of the emission spectra by tuning the size or composition of the quantum dots.
-The emission spectrum of alloyed core/shell CdS$_x$Se$_{1-x}$/ZnS Trilite450 \cite{tbd} quantum dots was measured.
-This spectrum is a symmetric peak centered around 461~nm with FWHM = 29~nm.
-
-%%\begin{figure}
-%%        \begin{center}
-%%        \includegraphics[scale=0.40]{graphs/6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME.pdf}
-%%        \caption[]{PE times for the simulation of 1000 electrons (5~MeV) with a quantum dot emission spectrum centered at 384~nm: Detector diameter = 6.5~m, bialkali photocathode, QD emission spectrum at 384~nm, TTS = 0.1~ns ($\sigma$).
-PEs from Cerenkov light only (black) and scintillation light only (red) are compared.
-The number of PEs per event after a 34.0~ns time cut is 124 from scintillation and 107 from Cerenkov.
-\label{6p5Meter_5MeVElectrons_Bialkali_QD384nmScintSpec_TIME}}
-%%        \end{center}
-%%\end{figure}
+An alternative route towards increasing the separation in time between Cerenkov and scintillation photon hits is the tuning of the scintillator emission spectrum.
+Recently, the use of quantum dots (QDs) in liquid scintillators has been studied as a possibility to improve future large scale neutrino experiments \cite{qdot2}.
+One major motivation for quantum-dot-doped scintillator is control of the emission spectra by tuning the size or composition of the quantum dots.
+The emission spectrum of commercial alloyed core/shell CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured.
+This spectrum shows a symmetric peak centered around 461~nm with FWHM = 29~nm.
 
@@
 Compared to the default case shown in FIG.
-\ref{time_plot_comparison} (a) the separation is worse because the scintillation light wavelengths are higher than in the KamLAND emission spectrum.
+\ref{time_plots_comparison} (a) the separation is worse (as expected) because the scintillation light wavelengths are higher than in the KamLAND emission spectrum.
 However, advances in the production of commercial quantum dot samples could yield quantum dots which have similar, single peak emission shapes at lower wavelengths.
-This case has been simulated using the same spectral shape of the measured Trilite450 emission but shifted to lower wavelengths such that the emission peak is centered at 384~nm.
-This peak emission value has been measured for other types of QDs, however with a much more pronounced tail \cite{tbd}.
-The resulting PE time distribution shows a slightly clearer separation of Cerenkov and Scintillation light.
-After the 34.0~ns cut on the TTS-smeared PE time we obtain a higher Cerenkov/scintillation ratio compared to the default simulation: The number of Cerenkov-induced PE after the time cut is unchanged while the number of PEs coming from scintillation light is decreased due to the higher average photon travel times.
+This case has been simulated using the same spectral shape of the measured core-shell quantum dot emission but shifted to lower wavelengths such that the emission peak is centered at 384~nm.
+This peak emission value has been measured for other types of QDs, however with a much more pronounced tail \cite{qdot2}.
+The resulting PE time distribution shows improved separation of Cerenkov and scintillation light compared to the default simulation.
+After the 34.0~ns cut on the TTS-smeared PE time we obtain a higher Cerenkov/scintillation ratio of 0.86 (107 PE from Cerenkov light and 124 PE from scintillation).
+The number of Cerenkov-induced PE after the time cut is unchanged while the number of PEs coming from scintillation light is decreased due to the higher average photon travel times.
 
-%%idea: use (some) PMTs with high QE for high lambda (and/or low QE for normal lambda). 
-%% add some more info why we are interested in a smaller detector
-%% motivation: Daedalus/ISODAR (?), 0nuBB 
 \section{Reconstruction}
@@
 
-WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration) \cite{Andy}.
-It provides a framework for generic event cleaning, track reconstruction, and particle id, and it comes equipped with variety of pre-built algorithms.
-A collaboration between Iowa State, the University of Chicago, and Argonne National Laboratory has continued adding to the code, developing new track-fitting techniques for Water Cherenkov detectors based on advanced photosensors with sub-cm imaging capabilities and timing resolutions below 100 picoseconds \cite{fast_group_and_LAPPD}.
+WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration)~\cite{Blake}.
+It provides a framework for generic event cleaning, track reconstruction, and particle identification, and it comes equipped with variety of pre-built algorithms.
+A collaboration between Iowa State, the University of Chicago, and Argonne National Laboratory has continued adding to the code, developing new track-fitting techniques for Water Cherenkov detectors based on advanced photosensors with sub-cm imaging capabilities and timing resolutions below 100 picoseconds~\cite{{fasttiming},{LAPPDcollab}}.
 In this paper, we use only the simple, low-level reconstruction tools native to the package, though in future works we hope to implement more sophisticated algorithms.
 
-The results presented in this paper rely on a simple vertex reconstruction algorithm, commonly known as a "point fit" \cite{reference_Super_K}.
-It assumes that all of the scintillation and Cherenkov light is emitted from a single point in space-time (x0,y0,z0,t0).
+The results presented in this paper rely on a simple vertex reconstruction algorithm, commonly known as a ``point fit''~\cite{SuperKalgo}.
+It assumes that all of the scintillation and Cherenkov light is emitted from a single point in space-time $(x_0,y_0,z_0,t_0)$.
 In actuality, the light is emitted along an extended, multi-scattered electron track.
-However, at the energies discussed in this paper, the extent of this track is small (a few cm) compared to the scale of the detector (R=6 meters) and thus typical photon transit distances.
+However, at the energies discussed in this paper, the extent of this track is small (a few cm) compared to the scale of the detector (R=6.5 meters) and thus typical photon transit distances.
 
 The first step of the reconstruction process relies on exact numerical calculations of vertex candidates from quadruplets of hits.
-Given a single point source, we need four constraints to solve for the four unknowns of the vertex \cite{Michael Smy} This approach would provide an exact solution in the case of four prompt, unscattered photons originating from a common point.
-However, many of these randomly chosen quadruplets will produce anomalous solutions due to real world effects such as optical scattering, delayed emission, and deviations from the point-like geometry.
-Nonetheless, we found that any chosen subset of 80 quadruplets was a sufficiently large ensemble to assure that some solutions will be close to the true vertex.
+Given a single point source, we need four constraints to solve for the four unknowns of the vertex ~\cite{Smy}.
+This approach would provide an exact solution in the case of four prompt, unscattered photons originating from a common point.
+However, many of these randomly chosen quadruplets will produce anomalous solutions due to real world effects such as delayed emission and deviations from the point-like geometry.
+Nonetheless, we found that any chosen subset of 400 quadruplets was a sufficiently large ensemble to assure that some solutions will be close to the true vertex.
 
 Once a small set of vertex candidates have been solved for, we test the goodness of each candidate and select the one that best fits the full ensemble of photon hits.
-The goodness of fit is determined based on the distribution of an observable known as the "point time residual" [super K].
-The point time residual is calculated by taking the difference between the measured time of a photon hit, and the predicted time of the hit, given its distance from the vertex hypothesis, the speed of light in the scintillator, and the hypothesized t0 of the event.
+The goodness of fit is determined based on the distribution of an observable known as the ``point time residual''~\cite{SuperKalgo}.
+The point time residual is calculated by taking the difference between the measured time of a photon hit, and the predicted time of the hit, given its distance from the vertex hypothesis, the speed of light in the scintillator, and the hypothesized $t_0$ of the event.
 The width of the time residual distribution over all hits is minimized when the hypothesized vertex is near the true vertex.
-Based on this figure of merit, we select the vertex with the narrowest time residual distribution from among the 80 candidates.
-This best candidate is used as the seed value for a MINUIT-based optimization [MINUIT], where the vertex is allowed to vary over a continuum of locations until the global optimum is found.
+Based on this figure of merit, we select the vertex with the narrowest time residual distribution from among the 400 candidates.
+This best candidate is used as the seed value for a MINUIT-based optimization~\cite{MINUIT}, where the vertex is allowed to vary over a continuum of locations until the global optimum is found.
 
@@
 
-%
-%This is Matt's text from July 17.
-%
-%WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration).
-It provides a framework for generic event cleaning, track reconstruction, and particle id, and it comes equipped with variety of pre-built algorithms.
-A collaboration between Iowa State, the University of Chicago, and Argonne National Laboratory has continued adding to the code, developing new track-fitting techniques for Water Cherenkov detectors built around the unique capabilities of microchannel plate photodetectors (MCPs): timing resolutions below 100 picoseconds and imaging resolutions below a centimeter.
-In this paper, we use only the simple, low-level reconstruction tools native to the package, though in future works we hope to implement more sophisticated techniques.
-
-%The simplest approach to vertex reconstruction, used in this paper, is commonly referred to as a Òpoint fitÓ.
-It assumes that, to first order, the light was emitted a single point in space-time (x0,y0,z0,t0).
-In reality, the light is emitted along the extended, multi-scattered track of the electron.
-However, at the energies discussed in this paper, the extent of the electron track is small (a few cm) compared to the scale of the detector, which has a radius of 6 meters.
-It is also small compared with the vertex resolutions typical of experiments with this scale.
-As starting point, the assumption works well.
-
-%The first algorithm we use to find a seed vertex relies on exact numerical calculations of possible vertex candidates.
-Given a single point source, we need four constraints to solve for the four unknowns of the vertex.
-A collection of N random quadruplets of hits is selected and used to solve for N corresponding vertex candidates.
-Some of these solutions will produce nonsensical results due to optical scattering of the emitted light, slow decay constant of the scintillation, finite detector resolution, and the limitations of our assumption of a point source.
-Nonetheless, over the N unique quadruplets, we expect some of the solutions to lie near the true event vertex.
-In this paper we chose N to be 80, determined by trial and error.
-However, given many hundreds of hits, the number of possible 4-hit combinations is much larger and one can choose a different N to better suit oneÕs particular experimental context.
-
-%Once the 80 vertex candidates have been solved for, we test the goodness of each candidate and select the one that best fits the full ensemble of hits.
-Distances are calculated between each hit and the candidate point-source.
-From these distances, we calculate the transit time, given the speed of light in our scintillator medium.
-This transit time plus the t0 of the 4-vertex provides a predicted arrival time for each hit.
-The Òtime residualÓ is the calculated as difference between the predicted and the measured arrival time, given the vertex candidate.
-Since we assume a point-source in this paper, we use what we call the Òsimple time residualÓ, as distinct from the Òextended time residualÓ, which assumes an extended, track-like geometry.
-The distribution of time residuals for each hit with respect to a vertex candidate has a width that is minimized when the hypothesis vertex is close to the true vertex.
-The t0 of the vertex candidate is left to float, and is adjusted to maximize the likelihood comparison with a Gaussian test function.
-The width of that gaussian test function is hand optimized to correspond with the width of the Òbest fitÓ.
-The vertex candidate with the largest maximum is selected to serve as the best candidate seed vertex.
-Once a single best seed vertex has been chosen, the same time-residual figure-of-merit is used in a MINUIT-based fit over a continuum of possible vertices in the neighborhood of the seed.
-
-%Once the vertex fitting is completed, we determine the direction by taking the centroid of all vectors pointing from the fitted event vertex and the hits on the detector.
-Since our initial time cut greatly enhances the purity of the Cherenkov light in this sample of hits and, since the Cherenkov light is directional (compared to the isotropic scintillation light), this fit provides a good measure of the track direction.
-This approach does not rely on the highly constrained Cherenkov geometry of the event, which contains further information.
-Future work will focus on better algorithms.
-Nonetheless, for a proof-of-principles, this method works sufficiently well..
-
 \section{Conclusion}
-This is going to work.
+The ability to reconstruct direction in kiloton-scale scintillating detectors would be a major technological advance.
+It would find immediate application in neutrino experiments especially those searching for $0\nu\beta\beta$.
+More generally, this technique could be applied wherever scintillation-based detectors are used.
+A Geant4 simulation of a simple spherical detector corresponding to a kiloton of scintillator shows that timing on the order of 0.1~ns is required to separate the directional Cerenkov light from the more abundant scintillation light.
+This separation can be improved using photodetectors with more red sensitivity and liquid scintillators with a more narrow emission spectrum shifted to shorter wavelengths.
+Furthermore, simple reconstruction algorithms adapted from those for water Cerenkov detectors are able to reliable reconstruct the position and direction of 5~MeV electrons.
+More detailed simulation and advanced reconstruction algorithms will need to be developed to move to lower energies and more complicated event topography, like those in $0\nu\beta\beta$, but the technique already appears very promising.
 
 \section{Acknowledgements}
+We would like to thank and acknowledge Andrew Blake at University of Cambridge for his work authoring the WCSimAnalysis code used to perform the reconstructions shown in this paper.
+Special thanks to the neutrino reconstruction group at Iowa State, particularly Mayly Sanchez, Ioana Anghel, and Tian Xin for their continued work in developing the WCSimAnalaysis algorithms and for their insights and expertise regarding issues related to Cherenkov reconstruction with fast-timing.
+Thanks also to Micheal Smy for his development of the quadruplet-based, vertex-finding method.
 L.
@@
 
-%%20 inch PMTs: 
-%%H.~Kume, S.~Sawaki, M.~Ito, K.~Arisaka, T.~Kajita, A.~Nishimura, and A.~Suzuki, Nucl. Instrum. Methods Phys. Res. \bf{205}, 443 (1983).
-%%17 inch PMTs:
-
 \end{document}

### 2013-07-20 Direction_v2.tex
The updates from Henry. --Lindley

--- 2013-07-19 Direction.tex
@@
 \usepackage{hyperref}
+\usepackage{lineno}
 
@@
 \newcommand{\PerTonDay}{(ton$\cdot$day)$^{-1}$}
+\newcommand{\GEANT}{GEANT4}
 \hyphenation{KamLAND}
+
+\linenumbers
 
@@
 addition of directional information using Cherenkov light and fast
-timing would greatly enhance the scientific reach of these detectors,
-especially for searches for neutrinoless double-beta decay. In this
+timing would enhance the scientific reach of these detectors,
+especially for searches for neutrino-less double-beta decay. In this
 paper, we develop a technique for extracting particle direction and
@@
 of neutrino masses and mixings. These measurements include KamLAND's
-measurement of reactor antineutrino oscillation at a distance of
+measurement of reactor anti-neutrino oscillation at a distance of
 $\sim$200~km\cite{kam2013}, Borexino's measurement of $^{7}$Be solar
 neutrino oscillation\cite{borexino}, and most recently the short
-baseline reactor antineutrino experiments that measured oscillations
+baseline reactor anti-neutrino experiments that measured oscillations
 due to $\theta_{13}$ at a distance of 1~km: Daya Bay\cite{dbtwo},
@@
 measurements\cite{isodarscatt} and sterile neutrino
-searches\cite{isodar,nist} and for non-proliferation
+searches\cite{isodar,nist}, and for non-proliferation
 applications\cite{nucifer, songs}.
@@
 The scalability of these detectors to large volumes also makes them
-highly competitive for neutrinoless double-beta ($0\nu\beta\beta$)
-decay searches in which the final state consists of a
-positron-electron pair with energies in the few MeV range.  Currently
-one of the best limits for the $0\nu\beta\beta$ half-life comes from
+highly competitive for neutrino-less double-beta ($0\nu\beta\beta$)
+decay searches in which the final state consists of a pair of
+electrons with energies in the few MeV range.  Currently one of the
+best limits for the $0\nu\beta\beta$ half-life comes from
 KamLAND-Zen\cite{KZ0nu}.
@@
 charged particles move through a scintillator, the molecules are
-excited, predominantly the the non-localized electrons in the
+excited, predominantly via the non-localized electrons in the
 $\pi$-bonds of the phenyl groups~\cite{birks_book}. Vibrational
@@
 picoseconds, the $\pi$-electrons de-excite to the first excited state
-from higher levels through radiationless transitions. The first
+from higher levels through radiation-less transitions. The first
 excited state de-excites through photon emission. There are two
@@
 
-The molecules in liquid scintillators are not isolated. Radiationless
+The molecules in liquid scintillators are not isolated. Radiation-less
 processes transfer energy between the molecules. The probability of
@@
 inefficiencies in the energy transfer through multiple absorption and
-reemission processes. Aromatic solutes or fluors are added to the
+re-emission processes. Aromatic solutes or fluors are added to the
 primary solvent to shift the wavelength of the photons to longer
@@
 than the speed of light in the medium.  This light is emitted in a
-cone pointed in the direction of the particles trajectory and with a
+cone centered on the direction of the particle trajectory, and with a
 continuous spectrum weighted toward shorter wavelengths
-but extending well into the red. The spectrum is described by:
+but extending well into the red. The spectrum is described by~\cite{Cherenkov_spectrum}:
 \begin{equation}
@@
 \end{equation}
-In Geant4 optical photons are assigned the group velocity in the wavelength 
-region of normal dispersion which is relevant to our study. 
-The longer wavelengths Cerenkov photons typically arrive before
+The longer wavelength Cerenkov photons typically arrive before
 the scintillation light, which is slowed by both the scintillation
 processes and the shorter wavelengths involved. Thus, with sufficient
-timing resolution and sensitivity to longer wavelength, it should be
+timing resolution and sensitivity to longer wavelengths it should be
 possible to separate the directional Cerenkov light and the isotropic
-scintillation light in time and then reconstruct the direction of the
+scintillation light, and then to reconstruct the direction of the
 initial particle.
@@
 \section{Geant4 Simulation}
-In order to study the effects relevant to directionality reconstruction in liquid scintillators, a Geant4 \cite{geant4one,geant4two} simulation has been set up.
-The presented simulation has been developed using standard Geant4 (version 4.9.6) functionality for the liquid scintillator optical model.
-The simulation has been built from a customized official Geant4 application example and it is kept rather generic to provide generally applicable information about the main factors in directionality reconstruction.
+In order to study the effects relevant to directional
+reconstruction in liquid scintillators, a 
+\GEANT~\cite{geant4one,geant4two} simulation has been set up. The 
+simulation has been developed using \GEANT Version 4.9.6 for the
+liquid scintillator optical model, in which optical photons are
+assigned the group velocity in the wavelength region of normal
+dispersion.
 
@@
         \includegraphics[scale=0.25]{graphs/detector_view.png}
-        \caption[]{Two back-to-back electrons with 1.41~MeV each (equally divided energy of $^116$Cd $0\nu\beta\beta$ decay) at the center of the sphere with initial directions along the x and -x-axis.
-Only Cerenkov photons are drawn to illustrate the directionality of the event.
-\label{detector_view}}
+        \caption[]{Two back-to-back electrons with 1.41~MeV each
+        (equally divided energy of $^{116}$Cd $0\nu\beta\beta$ decay) at
+        the center of the sphere with initial directions along the x
+        and -x-axis. Only Cerenkov photons are drawn to illustrate the
+        directionality of the event. \label{detector_view}}
         \end{center}
@@
 
-The implemented detector geometry is a sphere of 6.5~m diameter filled with liquid scintillator (see Figure \ref{detector_view}).
-Default scintillator properties have been chosen to match the KamLAND scintillator \cite{kamland2003}: 80~\% n-Dodecane, 20~\% Pseudocumene (1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole).
-The implemented scintillator properties include the atomic composition and density ($\rho$ = 0.78~g/ml), the wavelength-dependent attenuation length \cite{tajimaMaster} and refractive index \cite{OlegThesis}, the scintillation emission spectrum, emission rise time ($\tau_r$ = 1.0~ns) and emission decay time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 0.13) \cite{tajimaThesis}, scintillator light yield (LY, 9030 photons/MeV) and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV) \cite{ChrisThesis}.
-Variations from the baseline KamLAND case are discussed below when applicable.
-Reflections, reemission of absorbed photons in the scintillator bulk volume and scattering has not been included.
-These effects are expected to be corrections to the main effects studied in this paper and are subject of future work.
+The detector geometry is a sphere of 6.5~m diameter filled
+with liquid scintillator doped with $^{116}$Cd (see Figure
+\ref{detector_view}). The default
+scintillator properties have been chosen to match the KamLAND
+scintillator \cite{kamland2003}: 80~\% n-Dodecane, 20~\% Pseudocumene
+(1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole). The
+scintillator properties implemented in the simulation include the
+atomic composition and density ($\rho$ = 0.78~g/ml), the
+wavelength-dependent attenuation length \cite{tajimaMaster} and
+refractive index \cite{OlegThesis}, the scintillation emission
+spectrum, emission rise time ($\tau_r$ = 1.0~ns) and emission decay
+time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with
+relative weights of 0.87 and 0.13)
+\cite{tajimaThesis}, scintillator light yield (LY, 9030 photons/MeV)
+and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)
+\cite{ChrisThesis}. Variations from the baseline KamLAND case are
+discussed below. Re-emission of absorbed photons in the scintillator
+bulk volume and scattering have not yet been included, but are not
+expected to change the conclusions here.
 
-The inner sphere surface is used as the photodetector.
-It is treated as fully absorbing (no reflections) with a photodetector coverage of 100~\%.
-Two important photodetector properties have been varied: Transit time spread (TTS, default $\sigma$ = 0.1~ns) and wavelength-dependent quantum efficiency (QE) for photoelectron production.
-Default is the QE of a bialkali photocathode (Hamamatsu R7081 PMT), digitized values come from the Double Chooz \cite{dctwo} Monte Carlo simulation.
-We note that the KamLAND 17 inch PMTs use the same photocathode type (bialkali) with similar quantum efficiency.
+The inner sphere surface is used as the photodetector. It is treated
+as fully absorbing (no reflections), with a photodetector coverage of
+100~\%. Two important photodetector properties have been varied: 1)
+the transit time spread (TTS, default $\sigma$ = 0.1~ns) and 2) the
+wavelength-dependent quantum efficiency (QE) for photoelectron
+production. The default is the QE of a bialkali photocathode (Hamamatsu
+R7081 PMT), for which digitized values come from the Double Chooz \cite{dctwo}
+Monte Carlo simulation. We note that the KamLAND 17-inch PMTs use the
+same photocathode type with similar quantum efficiency.
 
@@
 \begin{center}
-        \subfigure[ ~Default simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
-        \subfigure[ ~Increased TTS (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
-        \subfigure[ ~Red-sensitive photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
-        \caption[]{PE times after TTS application for the simulation of 1000 electrons (5~MeV) with different settings.
-PEs from Cerenkov light (black) and scintillation light (grey) are compared.
-The dashed vertical line illustrates a time cut at 34~ns.
-(a) Default simulation: bialkali photocathode and TTS = 0.1~ns ($\sigma$).
-After the 34.0~ns time cut we get 171~PEs from scintillation and 108~PEs from Cerenkov light.
-(b) Default simulation settings except for TTS = 1.28~ns (KamLAND 17 in.
-PMTs).
-After the 34.0~ns time cut we get 349~PEs from scintillation and 88~PEs from Cerenkov light.
-(c) Default simulation settings except for a GaAsP photocathode.
-After the 34.0~ns time cut we get 226~PEs from scintillation and 229~PEs from Cerenkov light.
-\label{time_plots_comparison}}
+        \subfigure[ ~Default
+        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        \subfigure[ ~Increased TTS
+        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        \subfigure[ ~Red-sensitive
+        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        \caption[]{Photo-electron (PE) arrival times after application
+        of the transit-time spread (TTS) for the simulation
+        of 1000 electrons (5~MeV) with different values of the TTS and
+        wavelength response. PEs from
+        Cerenkov light (black, solid line) and scintillation light
+        (red, dotted line) are
+        compared. The dashed vertical line illustrates a time cut at
+        34~ns. (a) Default simulation: bialkali photocathode and TTS =
+        0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
+        from scintillation and 108~PEs from Cerenkov light. (b)
+        Default simulation settings except for TTS = 1.28~ns (KamLAND
+        17 in. PMTs). After the 34.0~ns time cut we get 349~PEs from
+        scintillation and 88~PEs from Cerenkov light. (c) Default
+        simulation settings except for a GaAsP photocathode. After the
+        34.0~ns time cut we get 226~PEs from scintillation and 229~PEs
+        from Cerenkov light. \label{time_plots_comparison}}
         \end{center}
 \end{figure*}
-Before the simulation results for different simulation settings are discussed in the following sections, we highlight the effects which contribute to the timing of the scintillator detector system.
-First, the simulated travel time of the initial 5~MeV electron is between 0.10 and 0.15~ns, while the travel distance is about 3~cm.
-Second, the scintillation light emission follows a scintillator-specific distribution characterized by rise and decay time(s).
-Before the solutes in liquid scintillator can emit optical photons, the energy has to be transfered from the solvent to the solute.
-The time constant of this energy transfer accounts for a rise time in scintillation light emission.
-Past neutrino experiments were not highly interested in the effect of scintillation rise time which is the reason why there is a lack of accurate numbers.
-We assume a rise time of 1.0~ns, more detailed studies are needed in the future.
-The two time constants used to describe the falling edge of the scintillator emission time distribution (quoted above) are values specific to the KamLAND scintillator.
-Third, chromatic dispersion turns out to be an important effect in a 6.5~m diameter detector at the level of precision needed for direction reconstruction.
-Due to the wavelength-dependence of the refractive index the speed of light in the scintillator (see Equation (\ref{eqGroup})) increases with increasing photon wavelengths for normal dispersion.
-In order to study the size of this effect, we extracted results from a simplified simulation of 5~MeV electrons at the center of the sphere where we used instantaneous scintillation and applied QE but no TTS.
-The true hit time distributions of photoelectrons were analyzed for scintillation light and Cerenkov light separately.
-Photoelectrons coming from Cerenkov light are on average created about 0.5~ns earlier than PEs from scintillation light.
-The RMS values for both the Cerenkov and scintillation light true PE time distributions are both about 0.5~ns.
-Note that these numbers include the effect of the finite electron travel time.
 
-On the detector side, the time information of single photoelectrons is affected by the TTS of the photodetectors, a number which can be different by orders of magnitude depending on the detector type.
-The default TTS of 0.1~ns ($\sigma$) is a value which can be achieved with the large area picosecond photodetectors (LAPPDs)\cite{LAPPDSum,LAPPDTDR} and possibly hybrid photodetectors (HPDs)\cite{hpdThesis}.
-In fact, even significantly lower TTS numbers are realistic with the LAPPD.
-Finally, the accuracy of the vertex reconstruction is reflected in broadening and distortions of the time spectrum (after time-of-flight correction has been applied).
-Another effect which is to be addressed in future work is the granularity of the photodetectors which adds to the uncertainty in the light path and thus the time-of-flight correction.
+Before the simulation results for different simulation settings are
+discussed in the following sections, we highlight the effects which
+contribute to the timing of the scintillator detector system. First,
+the simulated travel time of the initial 5~MeV electron is between
+0.10 and 0.15~ns, while the travel distance is about 3~cm. Second, the
+scintillation light emission follows a distribution characterized by
+scintillator-specific rise and decay times. Before the solutes in
+liquid scintillator can emit optical photons, the energy has to be
+transferred from the solvent to the solute. The time constant of this
+energy transfer accounts for a rise time in scintillation light
+emission. Past neutrino experiments were not highly sensitive to the
+effect of the scintillation rise time, which is the reason why there
+is a lack of accurate numbers. We assume a rise time of 1.0~ns, more
+detailed studies are needed in the future. The two time constants used
+to describe the falling edge of the scintillator emission time
+distribution (quoted above) are values specific to the KamLAND
+scintillator. Third, chromatic dispersion turns out to be an important
+effect in a 6.5~m diameter detector at the level of precision needed
+for direction reconstruction. Due to the wavelength-dependence of the
+refractive index the speed of light in the scintillator (see Equation
+(\ref{eqGroup})) increases with increasing photon wavelengths for
+normal dispersion. In order to study the size of this effect, we
+extracted results from a simplified simulation of 5~MeV electrons at
+the center of the sphere where we used instantaneous scintillation
+emission with the quantum-efficiency applied, but not including
+the transit-time spread. The true hit time distributions of
+photoelectrons were analyzed for scintillation light and Cerenkov
+light separately. Photoelectrons coming from Cerenkov light are on
+average created about 0.5~ns earlier than PEs from scintillation
+light. The RMS values from PE time distributions for Cerenkov and
+scintillation light are both about 0.5~ns. Note that these numbers
+include the effect of the finite electron travel time.
 
-In Sections \ref{detector_timing_sec} to \ref{scintillator_emission_sec}, we comparatively study the photoelectron timing for different detector configurations.
-We follow the idea to increase discrimination between Cerenkov and scintillation light by accurate detector timing.
-The main relevant quantities provided by the Geant4 simulation are the photoelectron hit positions and the detection times after TTS has been applied.
-In section \ref{reconstruction_sec} these quantities are used for event reconstruction.
+The measurement of the arrival times of single photoelectrons is
+affected by the TTS of the photodetectors, a number which can be
+different by orders of magnitude depending on the detector type. The
+default TTS of 0.1~ns ($\sigma$) is a value which can be achieved with
+the large area picosecond photodetectors
+(LAPPDs)\cite{LAPPDSum,LAPPDTDR} and possibly hybrid photodetectors
+(HPDs)\cite{hpdThesis}; even significantly lower TTS numbers are
+realistic with the LAPPD. The uncertainties in the vertex
+reconstruction are reflected in broadening and distortions of the time
+spectra.
+
+In Sections \ref{detector_timing_sec} to
+\ref{scintillator_emission_sec}, we study the
+photoelectron timing for different detector configurations. We follow
+the idea to increase discrimination between Cerenkov and scintillation
+light by accurate detector timing. The main relevant quantities
+provided by the \GEANT simulation are the photoelectron hit positions
+and the detection times after the TTS resolution has been applied. In
+Section~\ref{reconstruction_sec} these quantities are used for event
+reconstruction.
 
@@
 
-First we discuss results for the default simulation settings which are described in the previous section.
-Figure \ref{time_plots_comparison} (a) shows the TTS-smeared photoelectron detection times for 1000 simulated electrons with 5~MeV energy in the center of the detector with initial momentum direction coinciding with the x-axis.
-The photoelectrons induced by Cerenkov light arrive earlier, as expected due to the instantaneous emission and the higher average photon speed (higher wavelengths) compared to scintillation after QE weighting.
-There is however significant overlap with the PE time distribution coming from scintillation light.
-In order to compare simulations with different parameters to each other a fixed time cut $t \leq$ 34.0~ns (or 33.0~ns) is applied using the truth information of the electron starting time instead of an event-by-event time reconstruction (see Section \ref{reconstruction_sec} for reconstruction results).
-For the default simulation case, the number of PEs per event coming from Cerenkov light after the 34.0~ns time cut (108) is 98~\% of the total number of PEs from Cerenkov light (110).
-For scintillation the number of PEs after the time cut (171) is only 3.2~\% of the total scintillation-induced PEs (5445).
-This indicates the effectiveness of a time cut to separate Cerenkov light from scintillation light.
-An important figure of merit when comparing different simulation settings is the ratio of Cerenkov/scintillation induced PE after the 34.0~ns time cut since a higher ratio means more directional information per PE.
-The ratio is 0.63 for the default settings.
-Note that this ratio depends on the time cut, which will be optimized in future work together with the absolute number of PEs per event passing the cut.
+We first discuss results for the default simulation settings described
+in the previous section. Figure \ref{time_plots_comparison} (a) shows
+the TTS-smeared photoelectron detection times for 1000 simulated
+electrons with 5~MeV energy in the center of the detector with initial
+momentum directions coinciding with the x-axis. The photoelectrons
+induced by Cerenkov light arrive earlier, as expected due to the
+instantaneous emission and the higher average photon speed longer
+wavelengths) compared to scintillation light. There is however
+significant overlap of the two arrival time distributions.
+
+In order
+to compare simulations with different parameters, a fixed time cut $t
+\leq$ 34.0~ns  is applied using the truth information 
+instead of an event-by-event time reconstruction (see Section
+\ref{reconstruction_sec} for reconstruction results). For the default
+simulation case, the number of PEs per event coming from Cerenkov
+light after the 34.0~ns time cut (108) is 98~\% of the total number of
+PEs from Cerenkov light (110). For scintillation the number of PEs
+after the time cut (171) is only 3.2~\% of the total
+scintillation-induced PEs (5445). This indicates the effectiveness of
+a time cut to separate Cerenkov light from scintillation light. An
+important figure of merit when comparing different simulation settings
+is the ratio of Cerenkov/scintillation induced PE after the 34.0~ns
+time cut since a higher ratio means more directional information per
+PE. The ratio is 0.63 for the default settings. Note that this ratio
+depends on the time cut, which will be optimized in future work
+together with the absolute number of PEs per event passing the cut.
+
 \begin{figure}
@@
         \includegraphics[scale=0.40]{graphs/cos_psi_comp_h.pdf}
-        \caption[]{Angular distribution of photoelectron hits relative to the original electron direction along the x-axis: $\cos(\Psi) = x_{hit}/\vec{r}_{hit}$.
-1000 electron events with 5~MeV at the center are summed.
-Default simulation settings are used and both Cerenkov and scintillation light are included.
-A 33.0~ns time cut (black) is compared with a 34.0~ns time cut (grey).
-\label{cerenkov_cone}}
+        \caption[]{The angular distribution of photoelectron hits
+        relative to the original electron direction $\cos(\Psi) =
+        x_{hit}/\vec{r}_{hit}$. The sample consists of 1000 events
+        with a 5~MeV electron produced at the detector center. Default
+        simulation settings are used and both Cerenkov and
+        scintillation light are included. } 
+        \label{cerenkov_cone}
         \end{center}
@@
 
-Figure \ref{cerenkov_cone} displays the PE hit pattern after cutting on time.
-Although this time cut is a simplification of actual time reconstruction effects, we can use it to indicate the spatial distribution of hits after timing information has been used to separate Cerenkov and scintillation light.
-The Cerenkov ring structure can be clearly seen (peak at about 46\textdegree) which demonstrates that the directional signal conveyed by the Cerenkov photons is not erased by scattering of the initial 5~MeV electrons.
+Figure \ref{cerenkov_cone} displays the angular distribution of PE
+hits after the time cut. Although this time cut is a simplification of actual time
+reconstruction effects, we can use it to indicate the spatial
+distribution of hits after timing information has been used to
+separate Cerenkov and scintillation light. The Cerenkov ring structure
+can be clearly seen in the peak at about 46\textdegree, demonstrating
+that the directional signal conveyed by the Cerenkov photons is not
+erased by scattering of the initial 5~MeV electrons.
 
-If the 17 inch KamLAND PMTs \cite{tajimaMaster,kume_1983} (TTS = 1.28~ns) are used in the simulation, the broadening of the time distributions leads to a strongly decreased ratio of Cerenkov over scintillation light (0.25) after the 34~ns time cut (see FIG.
-\ref{time_plots_comparison} (b)).
-This shows that low photodetector TTS is critical for directionality reconstruction and motivates the use of novel detector types.
+When the 17 inch KamLAND PMTs \cite{tajimaMaster,kume_1983} (TTS =
+1.28~ns) are used in the simulation, the broadening of the time
+distributions leads to a strongly decreased ratio of Cerenkov over
+scintillation light (0.25) after the 34~ns time cut (see
+FIG. \ref{time_plots_comparison} (b)). This shows that a low
+photodetector TTS is critical for directionality reconstruction and
+motivates the use of novel detector types.
 
@@
 \label{detector_wavelength_response_sec} 
-Complementary to lowering the photodetector TTS, optimization of the wavelength-dependent QE is promising.
-Since Cerenkov photons which passed 6.5~m of scintillator have higher average wavelengths than scintillation photons, a photodetector which is more sensitive at high wavelengths increases not only the absolute number of PEs but also the ratio between Cerenkov- and scintillation-induced PEs.
-Therefore, a simulation has been run with the QE of an extended red-sensitive GaAsP photocathode (Hamamatsu R3809U-63 QE data was used).
+In addition to improving the photodetector TTS, optimization of the
+wavelength-dependent QE is promising. Since Cerenkov photons which
+passed 6.5~m of scintillator have higher average wavelengths than
+scintillation photons, a photodetector which is more sensitive at high
+wavelengths increases not only the absolute number of PEs but also the
+ratio between Cerenkov- and scintillation-induced PEs.
 
-Figure \ref{time_plots_comparison} (c) shows the results for the modified simulation with high QE in the red spectral region.
-The higher absolute number of PE coming from Cerenkov light (factor of $\approx$ 2) and the increased Cerenkov/scintillation ratio (1.6) after the time cut would significantly enhance the directionality reconstruction capabilities given that high coverage can be achieved with red-sensitive photocathodes.
+We have run the simulation with the QE of an extended red-sensitive
+GaAsP photocathode (Hamamatsu R3809U-63 QE data was used).  Figure
+\ref{time_plots_comparison} (c) shows the results for the modified
+simulation with high QE in the red spectral region. The higher
+absolute number of photo-electrons coming from Cerenkov light (factor
+of $\approx$ 2) and the increased Cerenkov/scintillation ratio (1.6)
+after the time cut would significantly enhance the directionality
+reconstruction.
 
@@
 \label{scintillator_emission_sec}
-An alternative route towards increasing the separation in time between Cerenkov and scintillation photon hits is the tuning of the scintillator emission spectrum.
-Recently, the use of quantum dots (QDs) in liquid scintillators has been studied as a possibility to improve future large scale neutrino experiments \cite{qdot2}.
-One major motivation for quantum-dot-doped scintillator is control of the emission spectra by tuning the size or composition of the quantum dots.
-The emission spectrum of commercial alloyed core/shell CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured.
-This spectrum shows a symmetric peak centered around 461~nm with FWHM = 29~nm.
+An alternative route towards increasing the separation in time between
+Cerenkov and scintillation photon hits is the tuning of the
+scintillator emission spectrum. Recently, the use of quantum dots
+(QDs) in liquid scintillators has been studied as a possibility to
+improve future large scale neutrino experiments \cite{qdot2}. One
+major motivation for quantum-dot-doped scintillator is control of the
+emission spectra by tuning the size or composition of the quantum
+dots. The emission spectrum of commercial alloyed core/shell
+CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured. This spectrum shows a
+symmetric peak centered around 461~nm with FWHM = 29~nm.
 
-In order to isolate the effect of a different emission spectrum, the other simulation settings, including the KamLAND absorption spectrum are kept unchanged.
-Compared to the default case shown in FIG.
-\ref{time_plots_comparison} (a) the separation is worse (as expected) because the scintillation light wavelengths are higher than in the KamLAND emission spectrum.
-However, advances in the production of commercial quantum dot samples could yield quantum dots which have similar, single peak emission shapes at lower wavelengths.
-This case has been simulated using the same spectral shape of the measured core-shell quantum dot emission but shifted to lower wavelengths such that the emission peak is centered at 384~nm.
-This peak emission value has been measured for other types of QDs, however with a much more pronounced tail \cite{qdot2}.
-The resulting PE time distribution shows improved separation of Cerenkov and scintillation light compared to the default simulation.
-After the 34.0~ns cut on the TTS-smeared PE time we obtain a higher Cerenkov/scintillation ratio of 0.86 (107 PE from Cerenkov light and 124 PE from scintillation).
-The number of Cerenkov-induced PE after the time cut is unchanged while the number of PEs coming from scintillation light is decreased due to the higher average photon travel times.
+In order to isolate the effect of a different emission spectrum, the
+other simulation settings, including the KamLAND absorption spectrum,
+are kept unchanged. Compared to the default case shown in
+FIG. \ref{time_plots_comparison} (a) the separation is worse (as
+expected) because the scintillation light wavelengths are higher than
+in the KamLAND emission spectrum. However, advances in the production
+of commercial quantum dot samples could yield quantum dots which have
+similar, single peak emission shapes at lower wavelengths. This case
+has been simulated using the same spectral shape of the measured
+core-shell quantum dot emission but shifted to lower wavelengths such
+that the emission peak is centered at 384~nm. This peak emission value
+has been measured for other types of QDs, however with a much more
+pronounced tail \cite{qdot2}. The resulting PE time distribution shows
+improved separation of Cerenkov and scintillation light compared to
+the default simulation. After the 34.0~ns cut on the TTS-smeared PE
+time we obtain a higher Cerenkov/scintillation ratio of 0.86 (107 PE
+from Cerenkov light and 124 PE from scintillation). The number of
+Cerenkov-induced PEs after the time cut is unchanged while the number
+of PEs coming from scintillation light is decreased due to the higher
+average photon travel times.
 
@@
         \subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \caption[]{\label{fig:reco} (Top) Reconstructed direction, $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|}, \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons (5MeV).
-In the simulation the electrons are produced along the X-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate from the center of the detector, $\vec{r}$ = (0,0,0).
-Detector diameter, $R$ = 6.5~m.
-Only photons with arrival time of $T<$ 34~ns are used in the reconstruction.
-The quantum efficiency of the bialkali photocathode is taken into account.
-(Bottom) Reconstructed vertex position, $(x,y,z)$, for the simulation of 1000 electrons (5~MeV).
-In the simulation the electrons are produced along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate from the center of the detector, $\vec{r}$ = (0,0,0).
-Detector diameter, $R$ = 6.5~m.
-Only photons with arrival time of $T<$ 34~ns are used in the reconstruction.
-The quantum efficiency of the bialkali photocathode is taken into account.
-}
+        \caption[]{\label{fig:reco} (Top) The reconstructed direction,
+        $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|},
+        \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons
+        (5MeV). In the simulation the electrons are produced along the
+        X-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
+        from the center of the 6.5m-diameter detector, $\vec{r}$ =
+        (0,0,0). Only photons with arrival time of $T<$ 34~ns are used
+        in the reconstruction. The quantum efficiency of the bialkali
+        photocathode is taken into account. (Bottom) The reconstructed
+        vertex position, $(x,y,z)$, for the same simulation.}
+
 \end{center}
@@
 
-WCSimAnalysis is a Water Cherenkov reconstruction package developed by Andrew Blake for the Long Baseline Neutrino Experiment (LBNE collaboration)~\cite{Blake}.
-It provides a framework for generic event cleaning, track reconstruction, and particle identification, and it comes equipped with variety of pre-built algorithms.
-A collaboration between Iowa State, the University of Chicago, and Argonne National Laboratory has continued adding to the code, developing new track-fitting techniques for Water Cherenkov detectors based on advanced photosensors with sub-cm imaging capabilities and timing resolutions below 100 picoseconds~\cite{{fasttiming},{LAPPDcollab}}.
-In this paper, we use only the simple, low-level reconstruction tools native to the package, though in future works we hope to implement more sophisticated algorithms.
+WCSimAnalysis is a Water Cherenkov reconstruction package developed by
+Andrew Blake for the Long Baseline Neutrino Experiment (LBNE
+collaboration)~\cite{Blake}. It provides a framework for generic event
+cleaning, track reconstruction, and particle identification, and 
+comes equipped with variety of pre-built algorithms. A collaboration
+between Iowa State, the University of Chicago, and Argonne National
+Laboratory has continued adding to the code, developing new
+track-fitting techniques for Water Cherenkov detectors based on
+advanced photosensors with sub-cm imaging capabilities and timing
+resolutions below 100
+picoseconds~\cite{{fasttiming},{LAPPDcollab}}. In this paper, we use
+only the reconstruction tools native to the package,
+though in future work we hope to implement more sophisticated
+algorithms.
 
-The results presented in this paper rely on a simple vertex reconstruction algorithm, commonly known as a ``point fit''~\cite{SuperKalgo}.
-It assumes that all of the scintillation and Cherenkov light is emitted from a single point in space-time $(x_0,y_0,z_0,t_0)$.
-In actuality, the light is emitted along an extended, multi-scattered electron track.
-However, at the energies discussed in this paper, the extent of this track is small (a few cm) compared to the scale of the detector (R=6.5 meters) and thus typical photon transit distances.
+The results presented in this paper rely on a simple vertex
+reconstruction algorithm, commonly known as a ``point
+fit''~\cite{SuperKalgo}. It assumes that all of the scintillation and
+Cherenkov light is emitted from a single point in space-time
+$(x_0,y_0,z_0,t_0)$. In actuality, the light is emitted along an
+extended, multi-scattered electron track. However, at the energies
+discussed in this paper, the extent of this track is small (a few cm)
+compared to the scale of the detector (R=6.5 meters) and thus typical
+photon transit distances.
 
-The first step of the reconstruction process relies on exact numerical calculations of vertex candidates from quadruplets of hits.
-Given a single point source, we need four constraints to solve for the four unknowns of the vertex ~\cite{Smy}.
-This approach would provide an exact solution in the case of four prompt, unscattered photons originating from a common point.
-However, many of these randomly chosen quadruplets will produce anomalous solutions due to real world effects such as delayed emission and deviations from the point-like geometry.
-Nonetheless, we found that any chosen subset of 400 quadruplets was a sufficiently large ensemble to assure that some solutions will be close to the true vertex.
+The first step of the reconstruction process relies on exact numerical
+calculations of vertex candidates from quadruplets of hits. Given a
+single point source, we need four constraints to solve for the four
+unknowns of the vertex ($x,y,z, and t_0$)~\cite{Smy}. This approach
+would provide an exact solution in the case of four prompt,
+unscattered photons originating from a common point. However, many of
+these randomly chosen quadruplets will produce anomalous solutions due
+to real world effects such as delayed emission and deviations from the
+point-like geometry. Nonetheless, we found that any chosen subset of
+400 quadruplets was a sufficiently large ensemble to assure that some
+solutions will be close to the true vertex.
 
-Once a small set of vertex candidates have been solved for, we test the goodness of each candidate and select the one that best fits the full ensemble of photon hits.
-The goodness of fit is determined based on the distribution of an observable known as the ``point time residual''~\cite{SuperKalgo}.
-The point time residual is calculated by taking the difference between the measured time of a photon hit, and the predicted time of the hit, given its distance from the vertex hypothesis, the speed of light in the scintillator, and the hypothesized $t_0$ of the event.
-The width of the time residual distribution over all hits is minimized when the hypothesized vertex is near the true vertex.
-Based on this figure of merit, we select the vertex with the narrowest time residual distribution from among the 400 candidates.
-This best candidate is used as the seed value for a MINUIT-based optimization~\cite{MINUIT}, where the vertex is allowed to vary over a continuum of locations until the global optimum is found.
+Once a small set of vertex candidates has been found, we test the
+goodness of each vertex and select the one that best fits the full
+ensemble of photon hits. The goodness of fit is determined based on
+the distribution of an observable known as the ``point time
+residual''~\cite{SuperKalgo}. The point time residual is calculated by
+taking the difference between the measured time of a photon hit, and
+the predicted time of the hit, given its distance from the vertex
+hypothesis, the speed of light in the scintillator (see
+Equ.\ref{eqGroup} , and the hypothesized $t_0$ of the event. The width
+of the time residual distribution over all hits is minimized when the
+hypothesized vertex is near the true vertex. Based on this figure of
+merit, we select the vertex with the narrowest time residual
+distribution from among the 400 candidates. This best candidate is
+used as the seed value for a MINUIT-based optimization~\cite{MINUIT},
+where the vertex is allowed to vary over a continuum of locations
+until the global optimum is found.
 
-The direction of the electron track is determined by taking the centroid of all vectors pointing from the fitted vertex to the hits on the detector.
-Since the Cherenkov light his highly directional, and since our timing cut enhances the purity of the Cherenkov light in the sample, this calculation provides a good measure of the track direction.
-Future work will focus on more sophisticated algorithms, taking better advantage of the highly constrained Cherenkov geometry.
-For the purpose of this paper, the current method provides a very good proof-of-principle.
+The direction of the electron track is determined by taking the
+centroid of all vectors pointing from the fitted vertex to the hits on
+the detector. Since the Cherenkov light is highly directional, and
+since the timing cut enhances the purity of the Cherenkov light in the
+sample, this calculation provides a good measure of the track
+direction.
 
-For the purpose of testing the reconstruction algorithm described above we use 1000 simulated electrons with the energy of 5~MeV.
-The electrons are similated at the center of the detector, $\vec{r}$ = (0,0,0), along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0).
+For the purpose of testing the reconstruction algorithm described
+above we use 1000 simulated electrons with an energy of 5~MeV. The
+electrons are simulated at the center of the detector, $\vec{r}$ =
+(0,0,0), along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0).
 
-Figure~\ref{fig:reco} show vertex reconstruction.
-The vertex is reasonably well reconstructed around center of the detector, $\vec{r}$ = (0,0,0).
-The RMS of the distributions for all three reconstructed coordinates are smaller than 3.5~cm.
-The shift along x-axis is attributed to the fact that electron is moving towards that direction and has a finite track length while the reconstruction algorithm treats all photons as they were emitted from the same point.
+Figure~\ref{fig:reco} shows the vertex reconstruction. The vertex is
+reasonably well reconstructed around the center of the detector, $\vec{r}$
+= (0,0,0).\\\ 
+XXXXX(ACTUALLY IT'S NOT- IT'S BADLY PULLED, NO?)XXX\\\
+The RMS of the distributions for all three reconstructed
+coordinates are smaller than 3.5~cm. The shift along the x-axis is
+partially attributed to the fact that electron is moving towards that direction
+and has a finite track length while the reconstruction algorithm
+treats all photons as they were emitted from the same point.\\
+XXX SO WHAT'S THE REST?\\
+The reconstruction of the direction also is shown in
+Fig.~\ref{fig:reco}. For the majority of the events we are able to
+reconstruct the direction of the electron direction.
 
-The reconstruction of the direction is shown in Fig.~\ref{fig:reco}. For majority of the events we are able to reconsruct the direction along x-axis.
+\section{Conclusions}
+The ability to reconstruct direction in kiloton-scale scintillation
+detectors would be a major technological advance for neutrino
+experiments, especially those also
+searching for $0\nu\beta\beta$. More generally, this technique could
+be applied wherever scintillation-based detectors are used. A \GEANT
+simulation of a simple spherical detector corresponding to a kiloton
+of scintillator shows that timing on the order of 0.1~ns is required
+to separate the directional Cerenkov light from the more abundant
+scintillation light. This separation can be improved using
+photodetectors with more red sensitivity and liquid scintillators with
+a more narrow emission spectrum shifted to shorter
+wavelengths. Furthermore, simple reconstruction algorithms adapted
+from those for water Cerenkov detectors are able to reliably
+reconstruct the position and direction of 5~MeV electrons. More
+detailed simulation and advanced reconstruction algorithms will need
+to be developed to move to lower energies and more complicated event
+topography, such as those in $0\nu\beta\beta$, but the technique already
+appears very promising.
 
-\section{Conclusion}
-The ability to reconstruct direction in kiloton-scale scintillating detectors would be a major technological advance.
-It would find immediate application in neutrino experiments especially those searching for $0\nu\beta\beta$.
-More generally, this technique could be applied wherever scintillation-based detectors are used.
-A Geant4 simulation of a simple spherical detector corresponding to a kiloton of scintillator shows that timing on the order of 0.1~ns is required to separate the directional Cerenkov light from the more abundant scintillation light.
-This separation can be improved using photodetectors with more red sensitivity and liquid scintillators with a more narrow emission spectrum shifted to shorter wavelengths.
-Furthermore, simple reconstruction algorithms adapted from those for water Cerenkov detectors are able to reliable reconstruct the position and direction of 5~MeV electrons.
-More detailed simulation and advanced reconstruction algorithms will need to be developed to move to lower energies and more complicated event topography, like those in $0\nu\beta\beta$, but the technique already appears very promising.
-
-\section{Acknowledgements}
-We would like to thank and acknowledge Andrew Blake at University of Cambridge for his work authoring the WCSimAnalysis code used to perform the reconstructions shown in this paper.
-Special thanks to the neutrino reconstruction group at Iowa State, particularly Mayly Sanchez, Ioana Anghel, and Tian Xin for their continued work in developing the WCSimAnalaysis algorithms and for their insights and expertise regarding issues related to Cherenkov reconstruction with fast-timing.
-Thanks also to Micheal Smy for his development of the quadruplet-based, vertex-finding method.
-L.
-Winslow would like to thank J.M.
-Conrad for useful discussion on the topic.
-L.
-Winslow would also like to thank K.
-Arisaka for discussions on the possible reach of traditional PMTs and the characteristics of HPDs.
-A.
-Elagin, H.J.
-Frisch and M.
-Wetstein are supported by DOE XXXX.
-C.
-Aberle and L.
-Winslow are supported by funds from University of California Los Angeles.
+\section{Acknowledgments}
+We thank Andrew Blake at University of Cambridge for his work
+authoring the WCSimAnalysis code used to perform the reconstructions
+shown in this paper. Special thanks are due to the neutrino
+reconstruction group at Iowa State, particularly Mayly Sanchez, Ioana
+Anghel, and Tian Xin, for their continued work in developing the
+WCSimAnalaysis algorithms and for their insights and expertise
+regarding issues related to Cherenkov reconstruction with
+fast-timing. Thanks also to Micheal Smy for his development of the
+quadruplet-based, vertex-finding method. L. Winslow would like to
+thank J.M. Conrad for useful discussion on the topic, and K. Arisaka
+for discussions on the possible reach of traditional PMTs and the
+characteristics of HPDs. C. Aberle and L. Winslow are supported by
+funds from University of California Los Angeles. The work at UC is
+partially supported by DOE DE-SC0008172, NSF PHY-1066014, and the
+Grainger Foundation.
 

### 2013-07-20 Direction_v2.tex
complete read through including references. --Lindley

@@
 %   http://fourforces.wordpress.com/2008/04/24/add-line-numbers-using-linenosty-in-revtex-4/
-%
+%Cherenkov
 
@@
 
-\linenumbers
+%\linenumbers
 
@@
 electrons with energies in the few MeV range.  Currently one of the
-best limits for the $0\nu\beta\beta$ half-life comes from
+best limits for the $0\nu\beta\beta$ half-life comes from the scintillating detector
 KamLAND-Zen\cite{KZ0nu}.
@@
 providing energy resolutions of $\sim$5\%. This is roughly a factor of
-two better than water Cerenkov detectors, the other developed
+two better than water Cherenkov detectors, the other developed
 technology that can be economically scaled to these large
@@
 primary particle.  In contrast, the direction of the particle can be
-reconstructed from the Cerenkov cones in water-based detectors,
+reconstructed from the Cherenkov cones in water-based detectors,
 although the energy resolution rapidly degrades below $\sim$5~MeV. For
@@
 
-In a liquid-scintillation-based detector, Cerenkov light is also
+In a liquid-scintillation-based detector, Cherenkov light is also
 produced, although most is absorbed and re-emitted as part of the
 scintillation processes.  However, some fraction retains its
-directional information. If this directional Cerenkov light can be
+directional information. If this directional Cherenkov light can be
 isolated from the copious isotropic scintillation light, it may be
@@
 for background rejection.  In this paper, we develop a technique for
-separating the Cerenkov and scintillation using the photon arrival
+separating the Cherenkov and scintillation light using the photon arrival
 times and evaluate different detector technologies that would allow
-the realization direction reconstruction in kilo-ton scale
+the realization of direction reconstruction in kilo-ton scale
 scintillating neutrino detectors.
@@
 isotropically per MeV of deposited energy. Although less abundant,
-Cerenkov light will be produced as well if a particle is moving faster
+Cherenkov light will be produced as well if a particle is moving faster
 than the speed of light in the medium.  This light is emitted in a
@@
 continuous spectrum weighted toward shorter wavelengths
-but extending well into the red. The spectrum is described by~\cite{Cherenkov_spectrum}:
+but extending well into the red. The spectrum is described by~\cite{Cherenkov34}:
 \begin{equation}
-\label{eqCerenkov}
-\frac{dN}{d\lambda dx} = \frac{2 \pi \alpha Z^2}{\lambda^2} ( 1 - \frac{1}{\beta^2 n(\lambda)^2} )
+\label{eqCherenkov}
+\frac{dN}{d\lambda dx} = \frac{2 \pi \alpha Z^2}{\lambda^2} \left [ 1 - \frac{1}{\beta^2 n(\lambda)^2} \right ]
 \end{equation}
 where $n(\lambda)$ is the wavelength-dependent index of refraction and
-$\beta$ is the velocity of the incoming particle. The Cerenkov light
+$\beta$ is the velocity of the incoming particle. The Cherenkov light
 produced at wavelengths shorter than the absorption cutoff of the
@@
 roughly 60 photons per MeV, assuming a 400~nm absorption
-cutoff\cite{qdot}. These undisturbed Cerenkov photons will have timing
+cutoff\cite{qdot}. These undisturbed Cherenkov photons will have timing
 determined by the group velocity \cite{group_velocity_article} in the liquid,
@@
 \end{equation}
-The longer wavelength Cerenkov photons typically arrive before
+The longer wavelength Cherenkov photons typically arrive before
 the scintillation light, which is slowed by both the scintillation
@@
 timing resolution and sensitivity to longer wavelengths it should be
-possible to separate the directional Cerenkov light and the isotropic
+possible to separate the directional Cherenkov light and the isotropic
 scintillation light, and then to reconstruct the direction of the
@@
 reconstruction in liquid scintillators, a 
-\GEANT~\cite{geant4one,geant4two} simulation has been set up. The 
-simulation has been developed using \GEANT Version 4.9.6 for the
-liquid scintillator optical model, in which optical photons are
+\GEANT~\cite{geant4one,geant4two} simulation has been constructed. The 
+simulation uses \GEANT~Version 4.9.6 with the default liquid scintillator optical model, in which optical photons are
 assigned the group velocity in the wavelength region of normal
@@
         \caption[]{Two back-to-back electrons with 1.41~MeV each
-        (equally divided energy of $^{116}$Cd $0\nu\beta\beta$ decay) at
+        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay) at
         the center of the sphere with initial directions along the x
-        and -x-axis. Only Cerenkov photons are drawn to illustrate the
+        and -x-axis. Only Cherenkov photons are drawn to illustrate the
         directionality of the event. \label{detector_view}}
@@
 
-The detector geometry is a sphere of 6.5~m diameter filled
-with liquid scintillator doped with $^{116}$Cd (see Figure
-\ref{detector_view}). The default
+The detector geometry is a sphere of 6.5~m diameter filled with scintillator.
+FIG.~\ref{detector_view} shows the geometry and the Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event.
+The default
 scintillator properties have been chosen to match the KamLAND
@@
 refractive index \cite{OlegThesis}, the scintillation emission
-spectrum, emission rise time ($\tau_r$ = 1.0~ns) and emission decay
+spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns) and emission decay
 time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with
@@
         wavelength response. PEs from
-        Cerenkov light (black, solid line) and scintillation light
+        Cherenkov light (black, solid line) and scintillation light
         (red, dotted line) are
@@
         0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
-        from scintillation and 108~PEs from Cerenkov light. (b)
+        from scintillation and 108~PEs from Cherenkov light. (b)
         Default simulation settings except for TTS = 1.28~ns (KamLAND
         17 in. PMTs). After the 34.0~ns time cut we get 349~PEs from
-        scintillation and 88~PEs from Cerenkov light. (c) Default
+        scintillation and 88~PEs from Cherenkov light. (c) Default
         simulation settings except for a GaAsP photocathode. After the
         34.0~ns time cut we get 226~PEs from scintillation and 229~PEs
-        from Cerenkov light. \label{time_plots_comparison}}
+        from Cherenkov light. \label{time_plots_comparison}}
         \end{center}
@@
 Before the simulation results for different simulation settings are
-discussed in the following sections, we highlight the effects which
+discussed, we note that  we concentrate on 5~MeV single electrons since this is the energy where scintillating detectors become more practical than water Cherenkov detectors. 
+We now highlight the effects which
 contribute to the timing of the scintillator detector system.  First,
@@
 the transit-time spread. The true hit time distributions of
-photoelectrons were analyzed for scintillation light and Cerenkov
-light separately. Photoelectrons coming from Cerenkov light are on
+photoelectrons were analyzed for scintillation light and Cherenkov
+light separately. Photoelectrons coming from Cherenkov light are on
 average created about 0.5~ns earlier than PEs from scintillation
-light. The RMS values from PE time distributions for Cerenkov and
+light. The RMS values from PE time distributions for Cherenkov and
 scintillation light are both about 0.5~ns. Note that these numbers
@@
 (HPDs)\cite{hpdThesis}; even significantly lower TTS numbers are
-realistic with the LAPPD. The uncertainties in the vertex
-reconstruction are reflected in broadening and distortions of the time
-spectra.
+realistic with the LAPPD. A similar effect to the smearing due to the TTS, is the effect of the uncertainties in the vertex
+reconstruction.
 
@@
 \ref{scintillator_emission_sec}, we study the
-photoelectron timing for different detector configurations. We follow
-the idea to increase discrimination between Cerenkov and scintillation
-light by accurate detector timing. The main relevant quantities
-provided by the \GEANT simulation are the photoelectron hit positions
+photoelectron timing for different detector configurations. We focus on
+the idea to increase discrimination between Cherenkov and scintillation
+light by using improved detector timing. The main relevant quantities
+provided by the \GEANT~simulation are the photoelectron hit positions
 and the detection times after the TTS resolution has been applied. In
-Section~\ref{reconstruction_sec} these quantities are used for event
+Section~\ref{reconstruction_sec} these quantities are then used for event
 reconstruction.
@@
 We first discuss results for the default simulation settings described
-in the previous section. Figure \ref{time_plots_comparison} (a) shows
+in the previous section. FIG.~\ref{time_plots_comparison} (a) shows
 the TTS-smeared photoelectron detection times for 1000 simulated
@@
 momentum directions coinciding with the x-axis. The photoelectrons
-induced by Cerenkov light arrive earlier, as expected due to the
-instantaneous emission and the higher average photon speed longer
-wavelengths) compared to scintillation light. There is however
+induced by Cherenkov light arrive earlier, as expected due to the
+instantaneous emission and the higher average photon speed compared to scintillation light. There is however
 significant overlap of the two arrival time distributions.
 
-In order
-to compare simulations with different parameters, a fixed time cut $t
-\leq$ 34.0~ns  is applied using the truth information 
-instead of an event-by-event time reconstruction (see Section
-\ref{reconstruction_sec} for reconstruction results). For the default
-simulation case, the number of PEs per event coming from Cerenkov
-light after the 34.0~ns time cut (108) is 98~\% of the total number of
-PEs from Cerenkov light (110). For scintillation the number of PEs
-after the time cut (171) is only 3.2~\% of the total
+In order to compare simulations with different parameters, a fixed time cut $t
+\leq$ 34.0~ns  is applied using the truth information to isolate the Cherenkov light in this early time window. For the default
+simulation case, the number of PEs per event coming from Cherenkov
+light in the early time window (108) is 98~\% of the total number of
+PEs from Cherenkov light (110). For scintillation, the number of PEs (171) is only 3.2~\% of the total
 scintillation-induced PEs (5445). This indicates the effectiveness of
-a time cut to separate Cerenkov light from scintillation light. An
+a time cut to separate Cherenkov light from scintillation light. An
 important figure of merit when comparing different simulation settings
-is the ratio of Cerenkov/scintillation induced PE after the 34.0~ns
-time cut since a higher ratio means more directional information per
-PE. The ratio is 0.63 for the default settings. Note that this ratio
-depends on the time cut, which will be optimized in future work
-together with the absolute number of PEs per event passing the cut.
+is the ratio of Cherenkov/scintillation induced PE in the early time window since a higher ratio means more directional information per
+PE. The ratio is 0.63 for the default settings.
 
@@
         \begin{center}
-        \includegraphics[scale=0.40]{graphs/cos_psi_comp_h.pdf}
+        \includegraphics[scale=0.40]{graphs/cos_psi_34_h.pdf}
         \caption[]{The angular distribution of photoelectron hits
@@
         with a 5~MeV electron produced at the detector center. Default
-        simulation settings are used and both Cerenkov and
+        simulation settings are used and both Cherenkov and
         scintillation light are included. } 
-        \label{cerenkov_cone}
+        \label{Cherenkov_cone}
         \end{center}
@@
 
-Figure \ref{cerenkov_cone} displays the angular distribution of PE
+FIG.~\ref{Cherenkov_cone} displays the angular distribution of PE
 hits after the time cut. Although this time cut is a simplification of actual time
 reconstruction effects, we can use it to indicate the spatial
-distribution of hits after timing information has been used to
-separate Cerenkov and scintillation light. The Cerenkov ring structure
+distribution of hits in this early time window. The Cherenkov ring structure
 can be clearly seen in the peak at about 46\textdegree, demonstrating
-that the directional signal conveyed by the Cerenkov photons is not
+that the directional signal conveyed by the Cherenkov photons is not
 erased by scattering of the initial 5~MeV electrons.
@@
 1.28~ns) are used in the simulation, the broadening of the time
-distributions leads to a strongly decreased ratio of Cerenkov over
-scintillation light (0.25) after the 34~ns time cut (see
-FIG. \ref{time_plots_comparison} (b)). This shows that a low
+distributions leads to a strongly decreased ratio of Cherenkov over
+scintillation light (0.25) for $t<34$~ns (see
+FIG.~\ref{time_plots_comparison} (b)). This shows that a low
 photodetector TTS is critical for directionality reconstruction and
@@
 In addition to improving the photodetector TTS, optimization of the
-wavelength-dependent QE is promising. Since Cerenkov photons which
+wavelength-dependent QE is promising. Since Cherenkov photons which
 passed 6.5~m of scintillator have higher average wavelengths than
@@
 wavelengths increases not only the absolute number of PEs but also the
-ratio between Cerenkov- and scintillation-induced PEs.
+ratio between Cherenkov- and scintillation-induced PEs.
 
 We have run the simulation with the QE of an extended red-sensitive
-GaAsP photocathode (Hamamatsu R3809U-63 QE data was used).  Figure
-\ref{time_plots_comparison} (c) shows the results for the modified
+GaAsP photocathode (Hamamatsu R3809U-63 QE data was used).  FIG.~\ref{time_plots_comparison} (c) shows the results for the modified
 simulation with high QE in the red spectral region. The higher
-absolute number of photo-electrons coming from Cerenkov light (factor
-of $\approx$ 2) and the increased Cerenkov/scintillation ratio (1.6)
-after the time cut would significantly enhance the directionality
+absolute number of photo-electrons coming from Cherenkov light (factor
+of $\approx$ 2) and the increased Cherenkov/scintillation ratio (1.6) in the early time window would significantly improve the directionality
 reconstruction.
@@
 An alternative route towards increasing the separation in time between
-Cerenkov and scintillation photon hits is the tuning of the
+Cherenkov and scintillation photon hits is the tuning of the
 scintillator emission spectrum. Recently, the use of quantum dots
@@
 are kept unchanged. Compared to the default case shown in
-FIG. \ref{time_plots_comparison} (a) the separation is worse (as
+FIG.~\ref{time_plots_comparison} (a) the separation is worse (as
 expected) because the scintillation light wavelengths are higher than
@@
 pronounced tail \cite{qdot2}. The resulting PE time distribution shows
-improved separation of Cerenkov and scintillation light compared to
+improved separation of Cherenkov and scintillation light compared to
 the default simulation. After the 34.0~ns cut on the TTS-smeared PE
-time we obtain a higher Cerenkov/scintillation ratio of 0.86 (107 PE
-from Cerenkov light and 124 PE from scintillation). The number of
-Cerenkov-induced PEs after the time cut is unchanged while the number
+time we obtain a higher Cherenkov/scintillation ratio of 0.86 (107 PE
+from Cherenkov light and 124 PE from scintillation). The number of
+Cherenkov-induced PEs after the time cut is unchanged while the number
 of PEs coming from scintillation light is decreased due to the higher
@@
         from the center of the 6.5m-diameter detector, $\vec{r}$ =
-        (0,0,0). Only photons with arrival time of $T<$ 34~ns are used
+        (0,0,0). Only photons with arrival time of $t<$ 34~ns are used
         in the reconstruction. The quantum efficiency of the bialkali
@@
 
-WCSimAnalysis is a Water Cherenkov reconstruction package developed by
-Andrew Blake for the Long Baseline Neutrino Experiment (LBNE
-collaboration)~\cite{Blake}. It provides a framework for generic event
+The timing studies show that in the early time window, $t\leq$ 34.0~ns, Cherenkov light dominates.
+In this paper, we apply reconstruction tools for a water Cherenkov detector, WCSimAnalysis, with only small modifications to remove water specific properties, to the problem of reconstructing the position and direction of 5~MeV electrons from this early light.
+WCSimAnalysis is a Water Cherenkov reconstruction package developed for the Long Baseline Neutrino Experiment (LBNE
+collaboration)\cite{Blake}. It provides a framework for generic event
 cleaning, track reconstruction, and particle identification, and 
-comes equipped with variety of pre-built algorithms. A collaboration
-between Iowa State, the University of Chicago, and Argonne National
-Laboratory has continued adding to the code, developing new
-track-fitting techniques for Water Cherenkov detectors based on
+comes equipped with variety of pre-built algorithms. It is continuing to be expanded using new
+track-fitting techniques for Water Cherenkov detectors\cite{Sanchez2012525} based on
 advanced photosensors with sub-cm imaging capabilities and timing
-resolutions below 100
-picoseconds~\cite{{fasttiming},{LAPPDcollab}}. In this paper, we use
-only the reconstruction tools native to the package,
-though in future work we hope to implement more sophisticated
-algorithms.
+resolutions below 100 picoseconds\cite{LAPPDSum,LAPPDTDR}.
 
@@
 reconstruction algorithm, commonly known as a ``point
-fit''~\cite{SuperKalgo}. It assumes that all of the scintillation and
+fit''\cite{SuperKalgo}. It assumes that all of the scintillation and
 Cherenkov light is emitted from a single point in space-time
@@
 single point source, we need four constraints to solve for the four
-unknowns of the vertex ($x,y,z, and t_0$)~\cite{Smy}. This approach
+unknowns of the vertex $(x,y,z,t_0)$\cite{Smy}. This approach
 would provide an exact solution in the case of four prompt,
-unscattered photons originating from a common point. However, many of
+un-scattered photons originating from a common point. However, many of
 these randomly chosen quadruplets will produce anomalous solutions due
@@
 
-Once a small set of vertex candidates has been found, we test the
+Once a set of vertex candidates has been found, we test the
 goodness of each vertex and select the one that best fits the full
@@
 the distribution of an observable known as the ``point time
-residual''~\cite{SuperKalgo}. The point time residual is calculated by
+residual''\cite{SuperKalgo}. The point time residual is calculated by
 taking the difference between the measured time of a photon hit, and
 the predicted time of the hit, given its distance from the vertex
-hypothesis, the speed of light in the scintillator (see
-Equ.\ref{eqGroup} , and the hypothesized $t_0$ of the event. The width
+hypothesis, a single effective speed of light in the scintillator, and the hypothesized $t_0$ of the event. The width
 of the time residual distribution over all hits is minimized when the
@@
 merit, we select the vertex with the narrowest time residual
-distribution from among the 400 candidates. This best candidate is
-used as the seed value for a MINUIT-based optimization~\cite{MINUIT},
-where the vertex is allowed to vary over a continuum of locations
-until the global optimum is found.
+distribution from among the 400 candidates.
 
-The direction of the electron track is determined by taking the
+The direction of the electron track is then determined by taking the
 centroid of all vectors pointing from the fitted vertex to the hits on
@@
 electrons are simulated at the center of the detector, $\vec{r}$ =
-(0,0,0), along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0).
-
-Figure~\ref{fig:reco} shows the vertex reconstruction. The vertex is
+(0,0,0), along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). FIG.~\ref{fig:reco} shows the vertex reconstruction. The vertex is
 reasonably well reconstructed around the center of the detector, $\vec{r}$
-= (0,0,0).\\\ 
-XXXXX(ACTUALLY IT'S NOT- IT'S BADLY PULLED, NO?)XXX\\\
-The RMS of the distributions for all three reconstructed
+= (0,0,0), except along the x-axis. The RMS of the distributions for all three reconstructed
 coordinates are smaller than 3.5~cm. The shift along the x-axis is
-partially attributed to the fact that electron is moving towards that direction
+partially attributed to the untuned effective speed of light and partially to the fact that electron is moving towards that direction
 and has a finite track length while the reconstruction algorithm
-treats all photons as they were emitted from the same point.\\
-XXX SO WHAT'S THE REST?\\
+treats all photons as they were emitted from the same point.
 The reconstruction of the direction also is shown in
-Fig.~\ref{fig:reco}. For the majority of the events we are able to
-reconstruct the direction of the electron direction.
+Fig.~\ref{fig:reco}. It shows that the majority of the events are well reconstructed in the correct direction. This is a very promising result given the simplicity of the algorithms used.
 
@@
 searching for $0\nu\beta\beta$. More generally, this technique could
-be applied wherever scintillation-based detectors are used. A \GEANT
-simulation of a simple spherical detector corresponding to a kiloton
+be applied wherever scintillation-based detectors are used. A \GEANT~simulation of a simple spherical detector corresponding to a kiloton
 of scintillator shows that timing on the order of 0.1~ns is required
-to separate the directional Cerenkov light from the more abundant
+to separate the directional Cherenkov light from the more abundant
 scintillation light. This separation can be improved using
@@
 wavelengths. Furthermore, simple reconstruction algorithms adapted
-from those for water Cerenkov detectors are able to reliably
+from those for water Cherenkov detectors are able to reliably
 reconstruct the position and direction of 5~MeV electrons. More
@@
 \section{Acknowledgments}
-We thank Andrew Blake at University of Cambridge for his work
-authoring the WCSimAnalysis code used to perform the reconstructions
-shown in this paper. Special thanks are due to the neutrino
+The authors thank Andrew Blake at University of Cambridge for his work
+authoring the WCSimAnalysis code. The authors thank the neutrino
 reconstruction group at Iowa State, particularly Mayly Sanchez, Ioana
@@
 regarding issues related to Cherenkov reconstruction with
-fast-timing. Thanks also to Micheal Smy for his development of the
-quadruplet-based, vertex-finding method. L. Winslow would like to
-thank J.M. Conrad for useful discussion on the topic, and K. Arisaka
+fast-timing. The authors also thank Micheal Smy for his development of the
+quadruplet-based vertex-finding method. L. Winslow would like to
+thank J.M. Conrad for many useful discussion on the topic, and K. Arisaka
 for discussions on the possible reach of traditional PMTs and the
 characteristics of HPDs. C. Aberle and L. Winslow are supported by
-funds from University of California Los Angeles. The work at UC is
+funds from University of California Los Angeles. The work at the University of Chicago is
 partially supported by DOE DE-SC0008172, NSF PHY-1066014, and the

### 2013-07-21 Direction_v4.tex
Version v4 of paper draft. Relative to v3 there are changes in the text, changes in Figures 1 and 4 and addition to references. Work done on July 20, 2013. --Christoph

--- 2013-07-20 Direction_v2.tex
@@
 \newcommand{\GEANT}{GEANT4}
+\newcommand{\RCS}{$R_{C/S}$}
 \hyphenation{KamLAND}
 
-%\linenumbers
+\linenumbers
 
@@
 \newcommand{\ucla}{\affiliation{University of California Los Angeles, Los Angeles, CA 90095, USA}}
-\newcommand{\chicago}{\affiliation{University of Chicago, Chicago, IL, 60637, USA}}
+\newcommand{\chicago}{\affiliation{University of Chicago, Chicago, IL 60637, USA}}
 
@@
 especially for searches for neutrino-less double-beta decay. In this
-paper, we develop a technique for extracting particle direction and
-evaluate different detector advances that could be used to make
-direction reconstruction a reality in a kiloton-scale detector.
+paper, we develop a technique for extracting particle direction using
+the difference in arrival times for Cherenkov and scintillation light,
+and evaluate several detector advances in timing, photodetector
+spectral response, and scintillator emission spectra that could be
+used to make direction reconstruction a reality in a kiloton-scale
+detector.
 \end{abstract}
@@
 decay searches in which the final state consists of a pair of
-electrons with energies in the few MeV range.  Currently one of the
-best limits for the $0\nu\beta\beta$ half-life comes from the scintillating detector
+electrons with energies in the few MeV range.  The observation of this
+rare decay would prove that the neutrino is a Majorana particle, which
+has profound consequences to our understanding of the generation of
+mass and may provide a possible explanation of the matter-antimatter
+asymmetry in the universe.  Currently one of the best limits for the
+$0\nu\beta\beta$ half-life comes from the scintillating detector
 KamLAND-Zen\cite{KZ0nu}.
@@
 primary particle.  In contrast, the direction of the particle can be
-reconstructed from the Cherenkov cones in water-based detectors,
+reconstructed from the Cherenkov cone in water-based detectors,
 although the energy resolution rapidly degrades below $\sim$5~MeV. For
@@
 separating the Cherenkov and scintillation light using the photon arrival
-times and evaluate different detector technologies that would allow
+times and evaluate several detector technologies that would allow
 the realization of direction reconstruction in kilo-ton scale
@@
 
-\section{Light Production in Liquid Scintillators}
-Liquid scintillators are cocktails of aromatic hydrocarbons. When
+\section{Liquid Scintillator Detectors}
+Liquid scintillators are `cocktails' of aromatic hydrocarbons. When
 charged particles move through a scintillator, the molecules are
@@
 
-The molecules in liquid scintillators are not isolated. Radiation-less
-processes transfer energy between the molecules. The probability of
-energy transfer between molecules increases as the overlap between the
-molecule's absorption and emission spectra increase. The absorption
-and emission spectra overlap at some level for all
-molecules. Consequently if there is only one type of molecule in the
-scintillator cocktail, the light output would be reduced due to
+The absorption and emission spectra overlap at some level
+in all molecules. Consequently, if there is only one type of molecule in
+the scintillator cocktail the light output is reduced due to
 inefficiencies in the energy transfer through multiple absorption and
@@
 isotropically per MeV of deposited energy. Although less abundant,
-Cherenkov light will be produced as well if a particle is moving faster
-than the speed of light in the medium.  This light is emitted in a
-cone centered on the direction of the particle trajectory, and with a
-continuous spectrum weighted toward shorter wavelengths
-but extending well into the red. The spectrum is described by~\cite{Cherenkov34}:
+Cherenkov light will be produced as well if a particle is moving
+faster than the speed of light in the medium.  This light is emitted
+in a cone centered on the direction of the particle trajectory, and
+with a continuous spectrum weighted toward shorter wavelengths but
+extending well into the red. The spectrum is described
+by~\cite{Cherenkov34}:
 \begin{equation}
@@
 where $n(\lambda)$ is the wavelength-dependent index of refraction and
-$\beta$ is the velocity of the incoming particle. The Cherenkov light
-produced at wavelengths shorter than the absorption cutoff of the
-scintillator will be absorbed and re-emitted as isotropic light, but
-wavelengths longer than this cutoff will propagate across the
-detector, retaining their directional information. The yield is
-roughly 60 photons per MeV, assuming a 400~nm absorption
-cutoff\cite{qdot}. These undisturbed Cherenkov photons will have timing
-determined by the group velocity \cite{group_velocity_article} in the liquid,
+$\beta$ is the velocity of the incoming particle. In a large detector,
+the Cherenkov light produced at wavelengths shorter than the
+absorption cutoff of the scintillator will be absorbed and re-emitted
+as isotropic light, but wavelengths longer than this cutoff will
+propagate across the detector, retaining their directional
+information. The yield is roughly 60 photons per MeV, assuming a
+400~nm absorption cutoff\cite{qdot}. These undisturbed Cherenkov
+photons will have timing determined by the group velocity
+\cite{group_velocity_article} in the liquid,
 \begin{equation}
@@
 \end{equation}
-The longer wavelength Cherenkov photons typically arrive before
-the scintillation light, which is slowed by both the scintillation
+The longer wavelength Cherenkov photons typically arrive before the
+scintillation light, which is slowed by both the scintillation
 processes and the shorter wavelengths involved. Thus, with sufficient
@@
 
+In $0\nu\beta\beta$, the electrons emerge with a combined energy equal
+to the Q-value of the particular isotope. The individual
+electrons come out following distributions of energies and angular correlations, 
+a probable case being equal division of energy between back-to-back
+electrons\cite{phasespace}. This case is shown in FIG.~\ref{detector_view}
+for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
+half-life is inversely proportional to the phase-space factor, isotopes with
+higher Q-values are preferred and due to backgrounds from the daughters
+of the $^{238}$U and $^{232}$Th decay chains those with Q-values at or
+above 2.6~MeV are most often considered for $0\nu\beta\beta$ searches.
+There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb},
+but only a handful with large Q-values.  Most of the high Q-value candidates
+have been considered as a dopant for a liquid scintillator:
+$^{150}$Nd(Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr(Q=3.350~MeV)\cite{zr1},
+$^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=3.350~MeV)\cite{qdot},
+$^{116}$Cd(Q=2.802~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller},
+$^{136}$Xe(Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn(Q=2.29~MeV)\cite{sn1}.
+Xenon gas readily dissolves into liquid scintillator. For the other isotopes,
+suitable organometallic compound needs to be found that produces a stable
+scintillator with a long attenuation length in the wavelength region of
+interest. Recently, nanocrystals called quantum dots have been explored
+as an alternative method for doping scintillator\cite{qdot,qdot2}. Quantum dots
+have interesting optical properties which are discussed further
+in Section~\ref{scintillator_emission_sec}.
+
+The reconstruction of the direction of the two simultaneous
+emitted $\sim$1.5~MeV electrons from the $0\nu\beta\beta$ decay is
+quite challenging. In this paper, we decided to start our directional
+studies with 5~MeV single electrons since reconstruction of direction
+in this energy range had never been achieved. This is a logical
+energy to use since this is where historically scintillating detectors
+become more practical than water Cherenkov detectors. This is also
+an energy that is relevant for studies of $^{8}$B solar neutrinos
+and neutrino-electron elastic scattering measurement.
+
 \section{Geant4 Simulation}
-In order to study the effects relevant to directional
-reconstruction in liquid scintillators, a 
+In order to study the effects relevant to directional reconstruction
+in liquid scintillators, a
 \GEANT~\cite{geant4one,geant4two} simulation has been constructed. The 
@@
         \begin{center}
-        \includegraphics[scale=0.25]{graphs/detector_view.png}
+        \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
         \caption[]{Two back-to-back electrons with 1.41~MeV each
@@
         the center of the sphere with initial directions along the x
-        and -x-axis. Only Cherenkov photons are drawn to illustrate the
+        and -x-axis. Only Cherenkov photons (green lines) are drawn to illustrate the
         directionality of the event. \label{detector_view}}
@@
 
-The detector geometry is a sphere of 6.5~m diameter filled with scintillator.
-FIG.~\ref{detector_view} shows the geometry and the Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event.
-The default
-scintillator properties have been chosen to match the KamLAND
-scintillator \cite{kamland2003}: 80~\% n-Dodecane, 20~\% Pseudocumene
+The detector geometry is a sphere of 6.5~m radius filled with
+scintillator. FIG.~\ref{detector_view} shows the geometry and the
+Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
+default scintillator properties have been chosen to match the KamLAND
+scintillator \cite{kamland2003}: 80\% n-Dodecane, 20\% Pseudocumene
 (1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole). The
@@
 refractive index \cite{OlegThesis}, the scintillation emission
-spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns) and emission decay
-time constants ($\tau_{d1}$ = 6.9~ns and $\tau_{d2}$ = 8.8~ns with
-relative weights of 0.87 and 0.13)
+spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns)
+and emission decay time constants ($\tau_{d1}$ = 6.9~ns and
+$\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 0.13)
 \cite{tajimaThesis}, scintillator light yield (LY, 9030 photons/MeV)
@@
 as fully absorbing (no reflections), with a photodetector coverage of
-100~\%. Two important photodetector properties have been varied: 1)
+100\%. Two important photodetector properties have been varied: 1)
 the transit time spread (TTS, default $\sigma$ = 0.1~ns) and 2) the
@@
 production. The default is the QE of a bialkali photocathode (Hamamatsu
-R7081 PMT), for which digitized values come from the Double Chooz \cite{dctwo}
+R7081 PMT)\cite{Hamamatsu_R7081}, for which digitized values come from the Double Chooz \cite{dctwo}
 Monte Carlo simulation. We note that the KamLAND 17-inch PMTs use the
@@
 Before the simulation results for different simulation settings are
-discussed, we note that  we concentrate on 5~MeV single electrons since this is the energy where scintillating detectors become more practical than water Cherenkov detectors. 
-We now highlight the effects which
+discussed, we note that we concentrate on 5~MeV single electrons since
+this is the energy at which scintillating detectors become more practical
+than water Cherenkov detectors.  We now highlight the effects which
 contribute to the timing of the scintillator detector system.  First,
@@
 effect of the scintillation rise time, which is the reason why there
-is a lack of accurate numbers. We assume a rise time of 1.0~ns, more
+is a lack of accurate numbers. We assume a rise time of 1.0~ns; more
 detailed studies are needed in the future. The two time constants used
@@
 scintillator. Third, chromatic dispersion turns out to be an important
-effect in a 6.5~m diameter detector at the level of precision needed
-for direction reconstruction. Due to the wavelength-dependence of the
-refractive index the speed of light in the scintillator (see Equation
-(\ref{eqGroup})) increases with increasing photon wavelengths for
-normal dispersion. In order to study the size of this effect, we
-extracted results from a simplified simulation of 5~MeV electrons at
+effect in a 6.5m-radius detector at the level of precision needed for
+direction reconstruction.
+
+Due to the wavelength-dependence of the refractive index the speed of
+light in the scintillator (see Equation (\ref{eqGroup})) increases
+with increasing photon wavelengths for normal dispersion, with red
+light traveling faster than blue light.  In order to study the
+measurability of the time differences, we simulated 5~MeV electrons at
 the center of the sphere where we used instantaneous scintillation
-emission with the quantum-efficiency applied, but not including
-the transit-time spread. The true hit time distributions of
-photoelectrons were analyzed for scintillation light and Cherenkov
-light separately. Photoelectrons coming from Cherenkov light are on
-average created about 0.5~ns earlier than PEs from scintillation
-light. The RMS values from PE time distributions for Cherenkov and
-scintillation light are both about 0.5~ns. Note that these numbers
-include the effect of the finite electron travel time.
+emission with the quantum-efficiency applied, but not including a
+transit-time spread. The true hit time distributions of photoelectrons
+were analyzed for scintillation light and Cherenkov light
+separately. Photoelectrons coming from Cherenkov light are on average
+created about 0.5~ns earlier than PEs from scintillation light. The
+RMS values from PE time distributions for Cherenkov and scintillation
+light are both about 0.5~ns. Note that these numbers include the
+effect of the finite electron travel time.
 
 The measurement of the arrival times of single photoelectrons is
-affected by the TTS of the photodetectors, a number which can be
-different by orders of magnitude depending on the detector type. The
-default TTS of 0.1~ns ($\sigma$) is a value which can be achieved with
-the large area picosecond photodetectors
+affected by the transit-time spread (TTS) of the photodetectors, a
+number which can be different by orders of magnitude depending on the
+detector type. The default TTS of 0.1~ns ($\sigma$) is a value which
+can be achieved with the large area picosecond photodetectors
 (LAPPDs)\cite{LAPPDSum,LAPPDTDR} and possibly hybrid photodetectors
 (HPDs)\cite{hpdThesis}; even significantly lower TTS numbers are
-realistic with the LAPPD. A similar effect to the smearing due to the TTS, is the effect of the uncertainties in the vertex
-reconstruction.
+realistic with the LAPPD~\cite{RSI_paper,PSEC4_paper,anode_paper}. We
+note that uncertainties in the vertex reconstruction.will produce a
+similar effect to the smearing due to the TTS.
 
@@
 \ref{scintillator_emission_sec}, we study the
-photoelectron timing for different detector configurations. We focus on
-the idea to increase discrimination between Cherenkov and scintillation
-light by using improved detector timing. The main relevant quantities
-provided by the \GEANT~simulation are the photoelectron hit positions
-and the detection times after the TTS resolution has been applied. In
-Section~\ref{reconstruction_sec} these quantities are then used for event
-reconstruction.
+photoelectron timing for different detector configurations. We focus
+on the idea to increase discrimination between Cherenkov and
+scintillation light by using improved detector timing. The primary
+quantities provided by the \GEANT~simulation are the photoelectron hit
+positions and the detection times after the TTS resolution has been
+applied. In Section~\ref{reconstruction_sec} these quantities are then
+used for event reconstruction.
 
@@
 in the previous section. FIG.~\ref{time_plots_comparison} (a) shows
-the TTS-smeared photoelectron detection times for 1000 simulated
+the TTS-smeared photoelectron (PE) detection times for 1000 simulated
 electrons with 5~MeV energy in the center of the detector with initial
@@
 induced by Cherenkov light arrive earlier, as expected due to the
-instantaneous emission and the higher average photon speed compared to scintillation light. There is however
-significant overlap of the two arrival time distributions.
+instantaneous emission and the higher average photon speed compared to
+scintillation light. There is however significant overlap of the two
+arrival time distributions.
 
-In order to compare simulations with different parameters, a fixed time cut $t
-\leq$ 34.0~ns  is applied using the truth information to isolate the Cherenkov light in this early time window. For the default
-simulation case, the number of PEs per event coming from Cherenkov
-light in the early time window (108) is 98~\% of the total number of
-PEs from Cherenkov light (110). For scintillation, the number of PEs (171) is only 3.2~\% of the total
+In order to compare simulations with different parameters, a fixed
+time cut of $t\leq$ 34.0~ns is applied using the truth information to
+isolate the Cherenkov light in this early time window. For the default
+simulation case, the average number of PEs per event coming from
+Cherenkov light in the early time window (108) is 98\% of the total
+average number of PEs from Cherenkov light (110). For scintillation
+light, the average number of PEs (171) is only 3.2\% of the average total
 scintillation-induced PEs (5445). This indicates the effectiveness of
-a time cut to separate Cherenkov light from scintillation light. An
-important figure of merit when comparing different simulation settings
-is the ratio of Cherenkov/scintillation induced PE in the early time window since a higher ratio means more directional information per
-PE. The ratio is 0.63 for the default settings.
+a time cut to separate Cherenkov light from scintillation light.
+
+The ratio of Cherenkov-induced to scintillation-induced photo-electrons
+in the early time window (\RCS) is a useful figure-of-merit when
+comparing different simulation settings, since a higher ratio means
+more directional information per PE. For the default simulation
+settings $R_{C/S}=0.63$.
 
@@
         simulation settings are used and both Cherenkov and
-        scintillation light are included. } 
+        scintillation light are included.} 
         \label{Cherenkov_cone}
@@
 reconstruction effects, we can use it to indicate the spatial
-distribution of hits in this early time window. The Cherenkov ring structure
-can be clearly seen in the peak at about 46\textdegree, demonstrating
+distribution of hits in the early time window. The Cherenkov ring structure
+can be clearly seen in the peak near 46\textdegree, demonstrating
 that the directional signal conveyed by the Cherenkov photons is not
@@
 
-When the 17 inch KamLAND PMTs \cite{tajimaMaster,kume_1983} (TTS =
+When the 17-inch KamLAND PMTs \cite{tajimaMaster,kume_1983} (TTS =
 1.28~ns) are used in the simulation, the broadening of the time
 distributions leads to a strongly decreased ratio of Cherenkov over
-scintillation light (0.25) for $t<34$~ns (see
+scintillation light ($R_{C/S}=0.25$) for $t<34$~ns (see
 FIG.~\ref{time_plots_comparison} (b)). This shows that a low
@@
 \label{detector_wavelength_response_sec} 
-In addition to improving the photodetector TTS, optimization of the
-wavelength-dependent QE is promising. Since Cherenkov photons which
-passed 6.5~m of scintillator have higher average wavelengths than
-scintillation photons, a photodetector which is more sensitive at high
-wavelengths increases not only the absolute number of PEs but also the
-ratio between Cherenkov- and scintillation-induced PEs.
+In addition to decreasing the photodetector TTS to enhance $R_{C/S}$,
+it is possible to optimize the wavelength-dependence of the
+photocathode. Since Cherenkov photons which passed through meters of
+scintillator have longer average wavelengths than scintillation
+photons, a photodetector which is more sensitive at long wavelengths
+increases not only the absolute number of PEs but also the ratio
+between Cherenkov- and scintillation-induced PEs.
 
 We have run the simulation with the QE of an extended red-sensitive
-GaAsP photocathode (Hamamatsu R3809U-63 QE data was used).  FIG.~\ref{time_plots_comparison} (c) shows the results for the modified
-simulation with high QE in the red spectral region. The higher
-absolute number of photo-electrons coming from Cherenkov light (factor
-of $\approx$ 2) and the increased Cherenkov/scintillation ratio (1.6) in the early time window would significantly improve the directionality
-reconstruction.
+GaAsP photocathode (Hamamatsu R3809U-63)~\cite{Hamamatsu_R3899U}.
+FIG.~\ref{time_plots_comparison}(c) shows the results for the
+modified simulation with high QE in the red spectral region. The
+higher absolute number of photo-electrons coming from Cherenkov light
+(factor of $\approx$ 2) and the increased Cherenkov/scintillation
+ratio (1.6) in the early time window would significantly improve the
+directionality reconstruction.
+% Can we quote a value of R for this? hjf
 
@@
 (QDs) in liquid scintillators has been studied as a possibility to
-improve future large scale neutrino experiments \cite{qdot2}. One
+improve future large scale neutrino experiments~\cite{qdot2}. One
 major motivation for quantum-dot-doped scintillator is control of the
 emission spectra by tuning the size or composition of the quantum
-dots. The emission spectrum of commercial alloyed core/shell
-CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured. This spectrum shows a
-symmetric peak centered around 461~nm with FWHM = 29~nm.
+dots.
 
-In order to isolate the effect of a different emission spectrum, the
-other simulation settings, including the KamLAND absorption spectrum,
-are kept unchanged. Compared to the default case shown in
-FIG.~\ref{time_plots_comparison} (a) the separation is worse (as
-expected) because the scintillation light wavelengths are higher than
-in the KamLAND emission spectrum. However, advances in the production
-of commercial quantum dot samples could yield quantum dots which have
-similar, single peak emission shapes at lower wavelengths. This case
-has been simulated using the same spectral shape of the measured
-core-shell quantum dot emission but shifted to lower wavelengths such
-that the emission peak is centered at 384~nm. This peak emission value
-has been measured for other types of QDs, however with a much more
-pronounced tail \cite{qdot2}. The resulting PE time distribution shows
-improved separation of Cherenkov and scintillation light compared to
-the default simulation. After the 34.0~ns cut on the TTS-smeared PE
-time we obtain a higher Cherenkov/scintillation ratio of 0.86 (107 PE
+The emission spectrum of commercial alloyed core/shell
+CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured in
+Ref.~\cite{qdot2}. This spectrum shows a symmetric peak centered
+around 461~nm with FWHM = 29~nm.  In order to isolate the effect of
+the different emission spectrum, the other simulation settings,
+including the KamLAND absorption spectrum, were kept unchanged; we
+find $R_{C/S}=0.17$ for the default 34-nsec timing cut.  Compared to
+the default case shown in FIG.~\ref{time_plots_comparison}(a) the
+separation is worse (as expected) because the scintillation light
+wavelengths are higher than in the KamLAND emission spectrum.
+
+However, advances in the production of commercial quantum dot samples
+could yield quantum dots which have similar, single peak emission
+shapes at lower wavelengths. This case has been simulated using the
+same spectral shape of the measured core-shell quantum dot emission
+but shifted to lower wavelengths such that the emission peak is
+centered at 384~nm. This peak emission value has been measured for
+other types of QDs, however with a much more pronounced tail
+\cite{qdot2}. The resulting PE time distribution shows improved
+separation of Cherenkov and scintillation light compared to the
+default simulation. After the 34.0~ns cut on the TTS-smeared PE time
+we obtain a Cherenkov/scintillation ratio of $R_{C/S}=0.86$ (107 PE
 from Cherenkov light and 124 PE from scintillation). The number of
@@
         \begin{center}
-        \subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \caption[]{\label{fig:reco} (Top) The reconstructed direction,
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
+        \subfigure{\includegraphics[scale=0.4]{graphs/hDir_34ns.pdf}}
+        \subfigure{\includegraphics[scale=0.4]{graphs/hVtx_34ns.pdf}}
+        \caption[]{\label{fig:reco} (Left) The reconstructed direction,
         $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|},
@@
         X-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
-        from the center of the 6.5m-diameter detector, $\vec{r}$ =
+        from the center of the 6.5m-radius detector, $\vec{r}$ =
         (0,0,0). Only photons with arrival time of $t<$ 34~ns are used
         in the reconstruction. The quantum efficiency of the bialkali
-        photocathode is taken into account. (Bottom) The reconstructed
+        photocathode is taken into account. (Right) The reconstructed
         vertex position, $(x,y,z)$, for the same simulation.}
@@
 
-The timing studies show that in the early time window, $t\leq$ 34.0~ns, Cherenkov light dominates.
-In this paper, we apply reconstruction tools for a water Cherenkov detector, WCSimAnalysis, with only small modifications to remove water specific properties, to the problem of reconstructing the position and direction of 5~MeV electrons from this early light.
-WCSimAnalysis is a Water Cherenkov reconstruction package developed for the Long Baseline Neutrino Experiment (LBNE
-collaboration)\cite{Blake}. It provides a framework for generic event
-cleaning, track reconstruction, and particle identification, and 
-comes equipped with variety of pre-built algorithms. It is continuing to be expanded using new
-track-fitting techniques for Water Cherenkov detectors\cite{Sanchez2012525} based on
+The timing studies show that in the early time window, $t\leq$
+34.0~ns, Cherenkov light dominates. In this paper, we apply
+reconstruction tools for a water Cherenkov detector, WCSimAnalysis,
+with only small modifications to remove water specific properties, to
+the problem of reconstructing the position and direction of 5~MeV
+electrons from this early light.  WCSimAnalysis is a Water Cherenkov
+reconstruction package developed for the Long Baseline Neutrino
+Experiment (LBNE collaboration)\cite{Blake}. It provides a framework
+for generic event cleaning, track reconstruction, and particle
+identification, and comes equipped with variety of pre-built
+algorithms. It is continuing to be expanded using new track-fitting
+techniques for Water Cherenkov detectors\cite{Sanchez2012525} based on
 advanced photosensors with sub-cm imaging capabilities and timing
@@
 these randomly chosen quadruplets will produce anomalous solutions due
-to real world effects such as delayed emission and deviations from the
+to `real world' effects such as delayed emission and deviations from the
 point-like geometry. Nonetheless, we found that any chosen subset of
@@
 
-Once a set of vertex candidates has been found, we test the
-goodness of each vertex and select the one that best fits the full
-ensemble of photon hits. The goodness of fit is determined based on
-the distribution of an observable known as the ``point time
+Once a set of vertex candidates has been found, we test the goodness
+of each vertex and select the one that best fits the full ensemble of
+photon hits. The goodness of fit is determined based on the
+distribution of an observable known as the ``point time
 residual''\cite{SuperKalgo}. The point time residual is calculated by
@@
 the predicted time of the hit, given its distance from the vertex
-hypothesis, a single effective speed of light in the scintillator, and the hypothesized $t_0$ of the event. The width
-of the time residual distribution over all hits is minimized when the
-hypothesized vertex is near the true vertex. Based on this figure of
-merit, we select the vertex with the narrowest time residual
-distribution from among the 400 candidates.
+hypothesis, a single effective speed of light in the scintillator, and
+the hypothesized $t_0$ of the event. The width of the time residual
+distribution over all hits is minimized when the hypothesized vertex
+is near the true vertex. Based on this figure of merit, we select the
+vertex with the narrowest time residual distribution from among the
+400 candidates.
 
@@
 
-For the purpose of testing the reconstruction algorithm described
-above we use 1000 simulated electrons with an energy of 5~MeV. The
-electrons are simulated at the center of the detector, $\vec{r}$ =
-(0,0,0), along the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). FIG.~\ref{fig:reco} shows the vertex reconstruction. The vertex is
-reasonably well reconstructed around the center of the detector, $\vec{r}$
-= (0,0,0), except along the x-axis. The RMS of the distributions for all three reconstructed
-coordinates are smaller than 3.5~cm. The shift along the x-axis is
-partially attributed to the untuned effective speed of light and partially to the fact that electron is moving towards that direction
-and has a finite track length while the reconstruction algorithm
-treats all photons as they were emitted from the same point.
+For the purpose of testing the reconstruction algorithm we use 1000
+simulated electrons with an energy of 5~MeV. The electrons are
+simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
+the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). FIG.~\ref{fig:reco}
+shows the vertex reconstruction. The vertex is reasonably well
+reconstructed around the center of the detector, $\vec{r}$ = (0,0,0),
+except along the x-axis. The RMS values of the distributions for all three
+reconstructed coordinates are smaller than 3.5~cm. The shift along the
+x-axis is partially attributed to the untuned effective speed of light
+and partially to the fact that electron is moving towards that
+direction and has a finite track length while the reconstruction
+algorithm treats all photons as they were emitted from the same point.
 The reconstruction of the direction also is shown in
-Fig.~\ref{fig:reco}. It shows that the majority of the events are well reconstructed in the correct direction. This is a very promising result given the simplicity of the algorithms used.
+Fig.~\ref{fig:reco}. It shows that for the majority of the events the
+initial electron direction is reconstructed well. This is a very promising
+result given the simplicity of the algorithms used.
 
@@
 regarding issues related to Cherenkov reconstruction with
-fast-timing. The authors also thank Micheal Smy for his development of the
-quadruplet-based vertex-finding method. L. Winslow would like to
-thank J.M. Conrad for many useful discussion on the topic, and K. Arisaka
-for discussions on the possible reach of traditional PMTs and the
-characteristics of HPDs. C. Aberle and L. Winslow are supported by
-funds from University of California Los Angeles. The work at the University of Chicago is
-partially supported by DOE DE-SC0008172, NSF PHY-1066014, and the
-Grainger Foundation.
+fast-timing. The authors also thank Micheal Smy for his development of
+the quadruplet-based vertex-finding method. L. Winslow would like to
+thank J.M. Conrad for many useful discussion on the topic, and
+K. Arisaka for discussions on the possible reach of traditional PMTs
+and the characteristics of HPDs. C. Aberle and L. Winslow are
+supported by funds from University of California Los Angeles. The work
+at the University of Chicago is partially supported by DOE
+contract DE-SC0008172 and NSF grant PHY-1066014. Matthew Wetstein gratefully
+acknowledges support by the Grainger Foundation.
 

### 2013-07-22 Direction_v5.tex
Commit of version 5. Version 5 is the final version for proofreading. --Christoph

--- 2013-07-21 Direction_v4.tex
@@
 \author{A.~Elagin}\chicago
+\author{H.~J.~Frisch}\chicago
 \author{M.~Wetstein}\chicago
-\author{H.~J.~Frisch}\chicago
 \author{L.~Winslow}\ucla
@@
 
-\pacs{23.40.$-$s, 21.10.Tg, 14.60.Pq, 27.60.$+$j}
+\pacs{23.40.$-$s, 14.60.Pq, 13.15.+g, 29.40.Mc}
 
@@
 \section{Introduction}
-Liquid scintillator based detectors are responsible for several of the
+Liquid scintillator-based detectors are responsible for several of the
 critical measurements that have determined our present understanding
@@
 rare decay would prove that the neutrino is a Majorana particle, which
-has profound consequences to our understanding of the generation of
+would have profound consequences to our understanding of the generation of
 mass and may provide a possible explanation of the matter-antimatter
@@
 possible to reconstruct the direction of the primary particle or, in
-the case of double beta decay, to determine the existence and topology
+the case of double-beta decay, to determine the existence and topology
 of the pair.  The addition of directionality is thus a powerful tool
@@
 separating the Cherenkov and scintillation light using the photon arrival
-times and evaluate several detector technologies that would allow
+times and evaluate several detector advances in timing, photodetector 
+spectral response, and scintillator emission spectra that would allow
 the realization of direction reconstruction in kilo-ton scale
@@
 excited, predominantly via the non-localized electrons in the
-$\pi$-bonds of the phenyl groups~\cite{birks_book}. Vibrational
+$\pi$-bonds of the phenyl groups\cite{birks_book}. Vibrational
 and rotational modes of the molecules are turned into heat within
@@
 from higher levels through radiation-less transitions. The first
-excited state de-excites through photon emission. There are two
+excited state can de-excite through photon emission. There are two
 characteristic times for this de-excitation, depending if the singlet
@@
 inefficiencies in the energy transfer through multiple absorption and
-re-emission processes. Aromatic solutes or fluors are added to the
-primary solvent to shift the wavelength of the photons to longer
-wavelengths where the scintillator is more transparent. This
+re-emission processes. Aromatic solutes or fluorophores are added to the
+primary solvent to shift the wavelengths of the photons to higher values 
+where the scintillator is more transparent. This
 wavelength-shifting is also used to match the quantum efficiency as a
 function of wavelength for the photodetectors being used. One typical
-scintillator mixture uses pseudo-cumene as the solvent with 1-5~g/L of
-PPO as the fluor. This mixture has a peak emission at about 400~nm
+scintillator mixture uses pseudocumene as the solvent with 1-5~g/L of
+PPO as the fluorophore. This mixture has a peak emission at about 400~nm
 where bialkali photomultiplier tubes (PMTs) are most sensitive and the
-pseudo-cumene is relatively transparent.
+pseudocumene is relatively transparent.
 
@@
 extending well into the red. The spectrum is described
-by~\cite{Cherenkov34}:
+by\cite{Cherenkov34}:
 \begin{equation}
 \label{eqCherenkov}
-\frac{dN}{d\lambda dx} = \frac{2 \pi \alpha Z^2}{\lambda^2} \left [ 1 - \frac{1}{\beta^2 n(\lambda)^2} \right ]
+\frac{d^2N}{d\lambda dx} = \frac{2 \pi \alpha Z^2}{\lambda^2} \left [ 1 - \frac{1}{\beta^2 n(\lambda)^2} \right ]
 \end{equation}
@@
 400~nm absorption cutoff\cite{qdot}. These undisturbed Cherenkov
-photons will have timing determined by the group velocity
-\cite{group_velocity_article} in the liquid,
+photons will have timing determined by the group 
+velocity\cite{group_velocity_article,pdg_review_2012,tamm1939} in the liquid,
 \begin{equation}
@@
 to the Q-value of the particular isotope. The individual
-electrons come out following distributions of energies and angular correlations, 
+electrons follow distributions of energies and angular correlations, 
 a probable case being equal division of energy between back-to-back
@@
 Xenon gas readily dissolves into liquid scintillator. For the other isotopes,
-suitable organometallic compound needs to be found that produces a stable
+a suitable organometallic compound needs to be found that produces a stable
 scintillator with a long attenuation length in the wavelength region of
-interest. Recently, nanocrystals called quantum dots have been explored
-as an alternative method for doping scintillator\cite{qdot,qdot2}. Quantum dots
-have interesting optical properties which are discussed further
-in Section~\ref{scintillator_emission_sec}.
-
-The reconstruction of the direction of the two simultaneous
-emitted $\sim$1.5~MeV electrons from the $0\nu\beta\beta$ decay is
-quite challenging. In this paper, we decided to start our directional
-studies with 5~MeV single electrons since reconstruction of direction
-in this energy range had never been achieved. This is a logical
-energy to use since this is where historically scintillating detectors
-become more practical than water Cherenkov detectors. This is also
-an energy that is relevant for studies of $^{8}$B solar neutrinos
-and neutrino-electron elastic scattering measurement.
+interest. As an alternative to single-atom-doping, recently nanocrystals 
+formed by candidate isotopes have been explored as a dopant\cite{qdot,qdot2}.
 
@@
 in liquid scintillators, a
-\GEANT~\cite{geant4one,geant4two} simulation has been constructed. The 
-simulation uses \GEANT~Version 4.9.6 with the default liquid scintillator optical model, in which optical photons are
+\GEANT\cite{geant4one,geant4two} simulation has been constructed. The
+simulation uses \GEANT~Version 4.9.6 with the default liquid scintillator
+optical model, in which optical photons are
 assigned the group velocity in the wavelength region of normal
@@
 default scintillator properties have been chosen to match the KamLAND
-scintillator \cite{kamland2003}: 80\% n-Dodecane, 20\% Pseudocumene
-(1,2,4-Trimethylbenzene) and 1.52~g/l PPO (2,5-Diphenyloxazole). The
+scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene
+(1,2,4-trimethylbenzene) and 1.52~g/l PPO (2,5-diphenyloxazole). The
 scintillator properties implemented in the simulation include the
 atomic composition and density ($\rho$ = 0.78~g/ml), the
-wavelength-dependent attenuation length \cite{tajimaMaster} and
-refractive index \cite{OlegThesis}, the scintillation emission
+wavelength-dependent attenuation length\cite{tajimaMaster} and
+refractive index\cite{OlegThesis}, the scintillation emission
 spectrum\cite{tajimaMaster}, emission rise time ($\tau_r$ = 1.0~ns)
 and emission decay time constants ($\tau_{d1}$ = 6.9~ns and
-$\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 0.13)
-\cite{tajimaThesis}, scintillator light yield (LY, 9030 photons/MeV)
-and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)
-\cite{ChrisThesis}. Variations from the baseline KamLAND case are
+$\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 
+0.13)\cite{tajimaThesis}, scintillator light yield (9030 photons/MeV)
+and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}. 
+Variations from the baseline KamLAND case are
 discussed below. Re-emission of absorbed photons in the scintillator
@@
 production. The default is the QE of a bialkali photocathode (Hamamatsu
-R7081 PMT)\cite{Hamamatsu_R7081}, for which digitized values come from the Double Chooz \cite{dctwo}
+R7081 PMT)\cite{Hamamatsu_R7081}, for which digitized values come from the Double Chooz\cite{dctwo}
 Monte Carlo simulation. We note that the KamLAND 17-inch PMTs use the
@@
         (red, dotted line) are
-        compared. The dashed vertical line illustrates a time cut at
+        compared. The dash-dotted vertical line illustrates a time cut at
         34~ns. (a) Default simulation: bialkali photocathode and TTS =
@@
 
-Before the simulation results for different simulation settings are
-discussed, we note that we concentrate on 5~MeV single electrons since
-this is the energy at which scintillating detectors become more practical
-than water Cherenkov detectors.  We now highlight the effects which
-contribute to the timing of the scintillator detector system.  First,
-the simulated travel time of the initial 5~MeV electron is between
+The initial studies described here are made with single 5~MeV electrons. 
+Future studies will work on increasing the complexity and lowering 
+the electron energies as is needed for $0\nu\beta\beta$ studies. Three 
+effects primarily contribute to the timing of the scintillator detector 
+system. First, the simulated travel time of the initial 5~MeV electron is between
 0.10 and 0.15~ns, while the travel distance is about 3~cm. Second, the
@@
 (HPDs)\cite{hpdThesis}; even significantly lower TTS numbers are
-realistic with the LAPPD~\cite{RSI_paper,PSEC4_paper,anode_paper}. We
+realistic with the LAPPD\cite{RSI_paper,PSEC4_paper,anode_paper}. We
 note that uncertainties in the vertex reconstruction.will produce a
@@
 photoelectron timing for different detector configurations. We focus
-on the idea to increase discrimination between Cherenkov and
+on the idea for increasing discrimination between Cherenkov and
 scintillation light by using improved detector timing. The primary
@@
 
-The ratio of Cherenkov-induced to scintillation-induced photo-electrons
-in the early time window (\RCS) is a useful figure-of-merit when
+The ratio of Cherenkov-induced to scintillation-induced photoelectrons
+in the early time window ($R_{C/S}$) is a useful figure-of-merit when
 comparing different simulation settings, since a higher ratio means
@@
         \caption[]{The angular distribution of photoelectron hits
-        relative to the original electron direction $\cos(\Psi) =
+        relative to the original electron direction, $\cos(\Psi) =
         x_{hit}/\vec{r}_{hit}$. The sample consists of 1000 events
@@
 
-When the 17-inch KamLAND PMTs \cite{tajimaMaster,kume_1983} (TTS =
+When the 17-inch KamLAND PMTs\cite{tajimaMaster,kume_1983} (TTS =
 1.28~ns) are used in the simulation, the broadening of the time
@@
 photocathode. Since Cherenkov photons which passed through meters of
-scintillator have longer average wavelengths than scintillation
+scintillator have on average longer wavelengths than scintillation
 photons, a photodetector which is more sensitive at long wavelengths
@@
 We have run the simulation with the QE of an extended red-sensitive
-GaAsP photocathode (Hamamatsu R3809U-63)~\cite{Hamamatsu_R3899U}.
+GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}.
 FIG.~\ref{time_plots_comparison}(c) shows the results for the
 modified simulation with high QE in the red spectral region. The
-higher absolute number of photo-electrons coming from Cherenkov light
+higher absolute number of photoelectrons coming from Cherenkov light
 (factor of $\approx$ 2) and the increased Cherenkov/scintillation
-ratio (1.6) in the early time window would significantly improve the
+ratio ($R_{C/S}=1.6$) in the early time window would significantly improve the
 directionality reconstruction.
-% Can we quote a value of R for this? hjf
 
@@
 (QDs) in liquid scintillators has been studied as a possibility to
-improve future large scale neutrino experiments~\cite{qdot2}. One
+improve future large scale neutrino experiments\cite{qdot,qdot2}. One
 major motivation for quantum-dot-doped scintillator is control of the
 emission spectra by tuning the size or composition of the quantum
-dots.
+dots; quantum dots can also provide a mechanism for introducing an 
+isotope for studying double-beta decay.
 
@@
 CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured in
-Ref.~\cite{qdot2}. This spectrum shows a symmetric peak centered
+Ref.\cite{qdot2}. This spectrum shows a symmetric peak centered
 around 461~nm with FWHM = 29~nm.  In order to isolate the effect of
@@
 including the KamLAND absorption spectrum, were kept unchanged; we
-find $R_{C/S}=0.17$ for the default 34-nsec timing cut.  Compared to
+find $R_{C/S}=0.17$ for the default 34~ns timing cut.  Compared to
 the default case shown in FIG.~\ref{time_plots_comparison}(a) the
@@
 centered at 384~nm. This peak emission value has been measured for
-other types of QDs, however with a much more pronounced tail
-\cite{qdot2}. The resulting PE time distribution shows improved
+other types of QDs, however with a much more pronounced 
+tail\cite{qdot2}. The resulting PE time distribution shows improved
 separation of Cherenkov and scintillation light compared to the
@@
         \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons
-        (5MeV). In the simulation the electrons are produced along the
-        X-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
+        (5~MeV). In the simulation the electrons are produced along the
+        x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
         from the center of the 6.5m-radius detector, $\vec{r}$ =
@@
 The timing studies show that in the early time window, $t\leq$
-34.0~ns, Cherenkov light dominates. In this paper, we apply
+34.0~ns, the ratio $R_{C/S}$ is high. In this paper, we apply
 reconstruction tools for a water Cherenkov detector, WCSimAnalysis,
-with only small modifications to remove water specific properties, to
-the problem of reconstructing the position and direction of 5~MeV
-electrons from this early light.  WCSimAnalysis is a Water Cherenkov
+to the problem of reconstructing the position and direction of 5~MeV
+electrons from this early light.  WCSimAnalysis is a water Cherenkov
 reconstruction package developed for the Long Baseline Neutrino
@@
 algorithms. It is continuing to be expanded using new track-fitting
-techniques for Water Cherenkov detectors\cite{Sanchez2012525} based on
+techniques for water Cherenkov detectors\cite{Sanchez2012525} based on
 advanced photosensors with sub-cm imaging capabilities and timing
@@
 except along the x-axis. The RMS values of the distributions for all three
-reconstructed coordinates are smaller than 3.5~cm. The shift along the
-x-axis is partially attributed to the untuned effective speed of light
-and partially to the fact that electron is moving towards that
-direction and has a finite track length while the reconstruction
-algorithm treats all photons as they were emitted from the same point.
-The reconstruction of the direction also is shown in
-Fig.~\ref{fig:reco}. It shows that for the majority of the events the
-initial electron direction is reconstructed well. This is a very promising
-result given the simplicity of the algorithms used.
+reconstructed coordinates are smaller than 3.5~cm. The shift along
+the x-axis is due to two effects for which the reconstruction has to
+use average values rather than the unknown true value for each
+event: the wavelength and hence the speed of the light in the medium,
+and the point of emission for each of the photons, reconstructed
+as coming from a common point. The reconstruction of the direction
+also is shown in Fig.~\ref{fig:reco}. It shows that for the majority
+of the events the initial electron direction is reconstructed well.
+This is a promising result given the simplicity of the algorithms
+used.
 
@@
 experiments, especially those also
-searching for $0\nu\beta\beta$. More generally, this technique could
+searching for neutrino-less double-beta decay. More generally, this technique could
 be applied wherever scintillation-based detectors are used. A \GEANT~simulation of a simple spherical detector corresponding to a kiloton
@@
 to be developed to move to lower energies and more complicated event
-topography, such as those in $0\nu\beta\beta$, but the technique already
-appears very promising.
+topography, such as those in neutrino-less double-beta decay, but the technique already
+appears promising.
 

### 2013-07-22 Direction_v6.tex
Version 6 of draft. Comments included from emails on July 22. --Christoph

--- 2013-07-22 Direction_v5.tex
@@
 picoseconds, the $\pi$-electrons de-excite to the first excited state
-from higher levels through radiation-less transitions. The first
+from higher levels through radiationless transitions. The first
 excited state can de-excite through photon emission. There are two
@@
 function of wavelength for the photodetectors being used. One typical
-scintillator mixture uses pseudocumene as the solvent with 1-5~g/L of
+scintillator mixture uses pseudocumene as the solvent with 1-5~g/l of
 PPO as the fluorophore. This mixture has a peak emission at about 400~nm
@@
 a probable case being equal division of energy between back-to-back
-electrons\cite{phasespace}. This case is shown in FIG.~\ref{detector_view}
+electrons\cite{phasespace}. This case is shown in Fig.~\ref{detector_view}
 for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
@@
 There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb},
-but only a handful with large Q-values.  Most of the high Q-value candidates
+but only a handful with large Q-values.
+
+Most of the high Q-value candidates
 have been considered as a dopant for a liquid scintillator:
@@
 \GEANT\cite{geant4one,geant4two} simulation has been constructed. The
-simulation uses \GEANT~Version 4.9.6 with the default liquid scintillator
+simulation uses \GEANT~version 4.9.6 with the default liquid scintillator
 optical model, in which optical photons are
@@
         \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
-        \caption[]{Two back-to-back electrons with 1.41~MeV each
-        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay) at
+        \caption[]{The detector geometry and coordinate system.
+        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.41~MeV each
+        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay). The electrons originate at
         the center of the sphere with initial directions along the x
-        and -x-axis. Only Cherenkov photons (green lines) are drawn to illustrate the
+        and -x-axis. Only Cherenkov photons are drawn to illustrate the
         directionality of the event. \label{detector_view}}
@@
 The detector geometry is a sphere of 6.5~m radius filled with
-scintillator. FIG.~\ref{detector_view} shows the geometry and the
+scintillator. Figure~\ref{detector_view} shows the geometry and the
 Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
@@
         compared. The dash-dotted vertical line illustrates a time cut at
-        34~ns. (a) Default simulation: bialkali photocathode and TTS =
+        34.0~ns. (a) Default simulation: bialkali photocathode and TTS =
         0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
@@
 We first discuss results for the default simulation settings described
-in the previous section. FIG.~\ref{time_plots_comparison} (a) shows
+in the previous section. Figure~\ref{time_plots_comparison} (a) shows
 the TTS-smeared photoelectron (PE) detection times for 1000 simulated
-electrons with 5~MeV energy in the center of the detector with initial
+electrons with 5~MeV energy in the center of the detector, with initial
 momentum directions coinciding with the x-axis. The photoelectrons
@@
 instantaneous emission and the higher average photon speed compared to
-scintillation light. There is however significant overlap of the two
+scintillation light. There is, however, significant overlap of the two
 arrival time distributions.
@@
 
-FIG.~\ref{Cherenkov_cone} displays the angular distribution of PE
+Figure~\ref{Cherenkov_cone} displays the angular distribution of PE
 hits after the time cut. Although this time cut is a simplification of actual time
@@
 distributions leads to a strongly decreased ratio of Cherenkov over
-scintillation light ($R_{C/S}=0.25$) for $t<34$~ns (see
-FIG.~\ref{time_plots_comparison} (b)). This shows that a low
+scintillation light ($R_{C/S}=0.25$) for $t<34.0$~ns (see
+Fig.~\ref{time_plots_comparison} (b)). This shows that a low
 photodetector TTS is critical for directionality reconstruction and
-motivates the use of novel detector types.
+motivates the use of novel photodetector types.
 
@@
 it is possible to optimize the wavelength-dependence of the
-photocathode. Since Cherenkov photons which passed through meters of
+photocathode. Since Cherenkov photons which pass through meters of
 scintillator have on average longer wavelengths than scintillation
@@
 GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}.
-FIG.~\ref{time_plots_comparison}(c) shows the results for the
+Figure~\ref{time_plots_comparison}(c) shows the results for the
 modified simulation with high QE in the red spectral region. The
@@
 including the KamLAND absorption spectrum, were kept unchanged; we
-find $R_{C/S}=0.17$ for the default 34~ns timing cut.  Compared to
-the default case shown in FIG.~\ref{time_plots_comparison}(a) the
+find $R_{C/S}=0.17$ for the default 34.0~ns timing cut.  Compared to
+the default case shown in Fig.~\ref{time_plots_comparison}(a) the
 separation is worse (as expected) because the scintillation light
-wavelengths are higher than in the KamLAND emission spectrum.
+wavelengths are longer than in the KamLAND emission spectrum.
 
@@
         from the center of the 6.5m-radius detector, $\vec{r}$ =
-        (0,0,0). Only photons with arrival time of $t<$ 34~ns are used
+        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
         in the reconstruction. The quantum efficiency of the bialkali
@@
 The timing studies show that in the early time window, $t\leq$
-34.0~ns, the ratio $R_{C/S}$ is high. In this paper, we apply
+34.0~ns, the ratio $R_{C/S}$ is high, improving the event selection. In this section, we apply
 reconstruction tools for a water Cherenkov detector, WCSimAnalysis,
@@
 simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
-the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). FIG.~\ref{fig:reco}
+the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). Figure~\ref{fig:reco}
 shows the vertex reconstruction. The vertex is reasonably well
@@
 to separate the directional Cherenkov light from the more abundant
-scintillation light. This separation can be improved using
+scintillation light. The separation can be improved using
 photodetectors with more red sensitivity and liquid scintillators with

### 2013-07-22 Direction_v6.tex
Version 6 of draft. Rounded Q-values for 116Cd. --Christoph

@@
 $^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=3.350~MeV)\cite{qdot},
-$^{116}$Cd(Q=2.802~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller},
+$^{116}$Cd(Q=2.81~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller},
 $^{136}$Xe(Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn(Q=2.29~MeV)\cite{sn1}.
@@
         \caption[]{The detector geometry and coordinate system.
-        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.41~MeV each
+        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.4~MeV each
         (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay). The electrons originate at

### 2013-07-22 Direction_v6.tex
Version 6 of draft. Changed event to hit in two sentences. --Christoph

@@
 The timing studies show that in the early time window, $t\leq$
-34.0~ns, the ratio $R_{C/S}$ is high, improving the event selection. In this section, we apply
+34.0~ns, the ratio $R_{C/S}$ is high, improving the photoelectron hit selection. In this section, we apply
 reconstruction tools for a water Cherenkov detector, WCSimAnalysis,
@@
 use average values rather than the unknown true value for each
-event: the wavelength and hence the speed of the light in the medium,
+hit: the wavelength and hence the speed of the light in the medium,
 and the point of emission for each of the photons, reconstructed

### 2013-07-22 Direction_v6_submit.tex
The files submitted to the arXiv. --Lindley

--- 2013-07-22 Direction_v6.tex
@@
 
-\title{Measuring Directionality in Double Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors}
+\title{Measuring Directionality in Double-Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors}
 
@@
         \begin{center}
-        \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
+        \includegraphics[scale=0.4]{geometry_plot_labels.pdf}
         \caption[]{The detector geometry and coordinate system.
@@
         \subfigure[ ~Default
-        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        simulation.]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
         \subfigure[ ~Increased TTS
-        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        (1.28~ns).]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
         \subfigure[ ~Red-sensitive
-        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        photocathode.]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
         \caption[]{Photo-electron (PE) arrival times after application
@@
         \begin{center}
-        \includegraphics[scale=0.40]{graphs/cos_psi_34_h.pdf}
+        \includegraphics[scale=0.40]{cos_psi_34_h.pdf}
         \caption[]{The angular distribution of photoelectron hits
@@
         \begin{center}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.4]{graphs/hDir_34ns.pdf}}
-        \subfigure{\includegraphics[scale=0.4]{graphs/hVtx_34ns.pdf}}
+        \subfigure{\includegraphics[scale=0.4]{hDir_34ns.pdf}}
+        \subfigure{\includegraphics[scale=0.4]{hVtx_34ns.pdf}}
         \caption[]{\label{fig:reco} (Left) The reconstructed direction,
@@
 the quadruplet-based vertex-finding method. L. Winslow would like to
-thank J.M. Conrad for many useful discussion on the topic, and
-K. Arisaka for discussions on the possible reach of traditional PMTs
+thank Janet Conrad for many useful discussion on the topic, and
+Katsushi Arisaka for discussions on the possible reach of traditional PMTs
 and the characteristics of HPDs. C. Aberle and L. Winslow are

### 2013-08-15 Direction_v7.tex
Christph edits and a line for John. --Lindley

--- 2013-07-22 Direction_v6.tex
@@
 scintillating neutrino detectors.
+This is a different technique and application than the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz}.
 
@@
 propagate across the detector, retaining their directional
-information. The yield is roughly 60 photons per MeV, assuming a
-400~nm absorption cutoff\cite{qdot}. These undisturbed Cherenkov
+information. These undisturbed Cherenkov
 photons will have timing determined by the group 
@@
 a probable case being equal division of energy between back-to-back
-electrons\cite{phasespace}. This case is shown in Fig.~\ref{detector_view}
+electrons\cite{phasespace}. This case is shown in FIG.~\ref{detector_view}
 for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
@@
 $^{150}$Nd(Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr(Q=3.350~MeV)\cite{zr1},
-$^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=3.350~MeV)\cite{qdot},
+$^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=2.995~MeV)\cite{qdot},
 $^{116}$Cd(Q=2.81~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller},
@@
 The detector geometry is a sphere of 6.5~m radius filled with
-scintillator. Figure~\ref{detector_view} shows the geometry and the
+scintillator. FIG.~\ref{detector_view} shows the geometry and the
 Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
@@
 100\%. Two important photodetector properties have been varied: 1)
-the transit time spread (TTS, default $\sigma$ = 0.1~ns) and 2) the
+the transit-time spread (TTS, default $\sigma$ = 0.1~ns) and 2) the
 wavelength-dependent quantum efficiency (QE) for photoelectron
@@
         photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
-        \caption[]{Photo-electron (PE) arrival times after application
+        \caption[]{Photoelectron (PE) arrival times after application
         of the transit-time spread (TTS) for the simulation
@@
 effects primarily contribute to the timing of the scintillator detector 
-system. First, the simulated travel time of the initial 5~MeV electron is between
+system. First, the simulated travel time of the initial 5~MeV electron is typically between
 0.10 and 0.15~ns, while the travel distance is about 3~cm. Second, the
@@
 realistic with the LAPPD\cite{RSI_paper,PSEC4_paper,anode_paper}. We
-note that uncertainties in the vertex reconstruction.will produce a
+note that uncertainties in the vertex reconstruction will produce a
 similar effect to the smearing due to the TTS.
@@
 photoelectron timing for different detector configurations. We focus
-on the idea for increasing discrimination between Cherenkov and
+on the idea of increasing the discrimination between Cherenkov and
 scintillation light by using improved detector timing. The primary
@@
 We first discuss results for the default simulation settings described
-in the previous section. Figure~\ref{time_plots_comparison} (a) shows
+in the previous section. FIG.~\ref{time_plots_comparison} (a) shows
 the TTS-smeared photoelectron (PE) detection times for 1000 simulated
@@
 average number of PEs from Cherenkov light (110). For scintillation
-light, the average number of PEs (171) is only 3.2\% of the average total
+light, the average number of PEs (171) is only 3.1\% of the average total
 scintillation-induced PEs (5445). This indicates the effectiveness of
@@
         relative to the original electron direction, $\cos(\Psi) =
-        x_{hit}/\vec{r}_{hit}$. The sample consists of 1000 events
+        x_{hit}/|\vec{r}_{hit}|$. The sample consists of 1000 events
         with a 5~MeV electron produced at the detector center. Default
         simulation settings are used and both Cherenkov and
-        scintillation light are included.} 
+        scintillation light are included. The $t\leq$ 34.0~ns cut is applied.} 
         \label{Cherenkov_cone}
@@
 
-Figure~\ref{Cherenkov_cone} displays the angular distribution of PE
+FIG.~\ref{Cherenkov_cone} displays the angular distribution of PE
 hits after the time cut. Although this time cut is a simplification of actual time
@@
 scintillation light ($R_{C/S}=0.25$) for $t<34.0$~ns (see
-Fig.~\ref{time_plots_comparison} (b)). This shows that a low
+FIG.~\ref{time_plots_comparison} (b)). This shows that a low
 photodetector TTS is critical for directionality reconstruction and
@@
 GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}.
-Figure~\ref{time_plots_comparison}(c) shows the results for the
+FIG.~\ref{time_plots_comparison}(c) shows the results for the
 modified simulation with high QE in the red spectral region. The
@@
 (factor of $\approx$ 2) and the increased Cherenkov/scintillation
-ratio ($R_{C/S}=1.6$) in the early time window would significantly improve the
+ratio ($R_{C/S}=1.01$) in the early time window would significantly improve the
 directionality reconstruction.
@@
 find $R_{C/S}=0.17$ for the default 34.0~ns timing cut.  Compared to
-the default case shown in Fig.~\ref{time_plots_comparison}(a) the
+the default case shown in FIG.~\ref{time_plots_comparison}(a) the
 separation is worse (as expected) because the scintillation light
@@
 simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
-the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). Figure~\ref{fig:reco}
+the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). FIG.~\ref{fig:reco}
 shows the vertex reconstruction. The vertex is reasonably well
@@
 as coming from a common point. The reconstruction of the direction
-also is shown in Fig.~\ref{fig:reco}. It shows that for the majority
+also is shown in FIG.~\ref{fig:reco}. It shows that for the majority
 of the events the initial electron direction is reconstructed well.

### 2013-08-15 Direction_v7.tex
Read through one more time, and add ref for Manfred.  --Lindley

@@
 scintillating neutrino detectors.
-This is a different technique and application than the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz}.
+This is a different technique and application than the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
 
@@
 Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
-default scintillator properties have been chosen to match the KamLAND
+default scintillator properties have been chosen to match a KamLAND-like
 scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene
@@
 light, the average number of PEs (171) is only 3.1\% of the average total
-scintillation-induced PEs (5445). This indicates the effectiveness of
+scintillation-induced PEs (5445). This demonstrates the effectiveness of
 a time cut to separate Cherenkov light from scintillation light.
@@
 discussed in this paper, the extent of this track is small (a few cm)
-compared to the scale of the detector (R=6.5 meters) and thus typical
+compared to the scale of the detector (R=6.5 meters) and therefore the typical
 photon transit distances.

### 2013-10-14 Direction_JINST_v1.tex
Direction_JINST_v1.tex has the same content as Direction_v7.tex but is in JINST format. The references have been adjusted to match the JINST journal style. We use JHEP.bst as recommended in the JINST author manual. Direction_JINST_v2.tex has an additional section with three additional plots containing results for 1.4 and 2.1 MeV. Direction_JINST_v3.tex is another iteration of Direction_JINST_v2.tex and Direction_JINST_v4.tex is the version which Andrey, Henry, Matt and Christoph propose for publication. Some more details can be found in README.md --Christoph

--- 2013-08-15 Direction_v7.tex
@@
-%
-%%%%%%%%%%%%%%%%%%%%%
-% Special Instructions
-%%%%%%%%%%%%%%%%%%%%%
-%
-% Please follow these instructions to use line numbering package (lineno.sty):
-%
-%   http://fourforces.wordpress.com/2008/04/24/add-line-numbers-using-linenosty-in-revtex-4/
-%Cherenkov
-
 \RequirePackage{lineno}
 
-\documentclass[aps,prc,twocolumn,groupedaddress,showpacs,amsmath,amssymb,floatfix,superscriptaddress]{revtex4}
-\usepackage{multirow}
+\documentclass[cits]{JINST}
+%%\usepackage{multirow}
+\usepackage{times}
+%%\usepackage{hyperref}
+\usepackage{lineno}
+\usepackage{amsmath, amsthm, amssymb}
+\usepackage{mathptmx}
+\usepackage{mdwlist}
+\usepackage{textcomp}
 \usepackage{graphicx,subfigure}
-\usepackage{times}
-\usepackage{textcomp}
-\usepackage{hyperref}
-\usepackage{lineno}
-
-\bibliographystyle{apsrev}
-
-\usepackage{color}
-
-\newcommand{\nuebar}{$\overline{\nu}_{e}$}
-\newcommand{\PerTonDay}{(ton$\cdot$day)$^{-1}$}
-\newcommand{\GEANT}{GEANT4}
-\newcommand{\RCS}{$R_{C/S}$}
-\hyphenation{KamLAND}
-
-%\linenumbers
-
-\begin{document}
-
-\bibliographystyle{h-physrev3}
+\usepackage{fontenc}
+%%\usepackage{color}
 
@@
 
-% All university affiliations addresses go here:
-\newcommand{\ucla}{\affiliation{University of California Los Angeles, Los Angeles, CA 90095, USA}}
-\newcommand{\chicago}{\affiliation{University of Chicago, Chicago, IL 60637, USA}}
+\author{C.~Aberle$^a$, A.~Elagin$^b$, H.J.~Frisch$^b$, M.~Wetstein$^b$, and L.~Winslow$^a$\setcounter{footnote}{0}\thanks{corresponding author}\\
+\llap{$^a$}University of California, Los Angeles, Los Angeles, CA 90095, USA\\
+\llap{$^b$}University of Chicago, Chicago, IL 60637, USA\\
+  E-mail: \email{lwinslow@physics.ucla.edu}}
 
-%
-\author{C.~Aberle}\ucla
-\author{A.~Elagin}\chicago
-\author{H.~J.~Frisch}\chicago
-\author{M.~Wetstein}\chicago
-\author{L.~Winslow}\ucla
+\linenumbers
 
-\date{\today}
-
-\begin{abstract}
-Large liquid-scintillator-based detectors have proven to be
+\abstract{Large liquid-scintillator-based detectors have proven to be
 exceptionally effective for low energy neutrino measurements due to
@@
 used to make direction reconstruction a reality in a kiloton-scale
-detector.
-\end{abstract}
+detector.}
 
-\pacs{23.40.$-$s, 14.60.Pq, 13.15.+g, 29.40.Mc}
+\keywords{Scintillators, scintillation and light emission processes (solid, gas and liquid scintillators); Cherenkov detectors; Large detector systems for particle and astroparticle physics; Neutrino detectors; Simulation methods and programs}
 
-\maketitle
+\newcommand{\nuebar}{$\overline{\nu}_{e}$}
+\newcommand{\PerTonDay}{(ton$\cdot$day)$^{-1}$}
+\newcommand{\GEANT}{GEANT4}
+\newcommand{\RCS}{$R_{C/S}$}
+\hyphenation{KamLAND}
+
+\begin{document}
 
@@
 
-\section{Acknowledgments}
+\acknowledgments
 The authors thank Andrew Blake at University of Cambridge for his work
@@
 
+\newpage
+
+\bibliographystyle{JHEP}
 \bibliography{DirectionBibliography}

### 2013-10-14 Direction_JINST_v2.tex
Direction_JINST_v1.tex has the same content as Direction_v7.tex but is in JINST format. The references have been adjusted to match the JINST journal style. We use JHEP.bst as recommended in the JINST author manual. Direction_JINST_v2.tex has an additional section with three additional plots containing results for 1.4 and 2.1 MeV. Direction_JINST_v3.tex is another iteration of Direction_JINST_v2.tex and Direction_JINST_v4.tex is the version which Andrey, Henry, Matt and Christoph propose for publication. Some more details can be found in README.md --Christoph

--- 2013-10-14 Direction_JINST_v1.tex
@@
 
-\title{Measuring Directionality in Double-Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors}
+\title{Measuring directionality in double-beta decay and neutrino interactions with kiloton-scale scintillation detectors}
 
@@
 
-\section{Liquid Scintillator Detectors}
+\section{Liquid scintillator detectors}
 Liquid scintillators are `cocktails' of aromatic hydrocarbons. When
@@
 a probable case being equal division of energy between back-to-back
-electrons\cite{phasespace}. This case is shown in FIG.~\ref{detector_view}
+electrons\cite{phasespace}. This case is shown in figure \ref{detector_view}
 for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
@@
 have been considered as a dopant for a liquid scintillator:
-$^{150}$Nd(Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr(Q=3.350~MeV)\cite{zr1},
-$^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=2.995~MeV)\cite{qdot},
-$^{116}$Cd(Q=2.81~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller},
-$^{136}$Xe(Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn(Q=2.29~MeV)\cite{sn1}.
+$^{150}$Nd (Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr (Q=3.350~MeV)\cite{zr1},
+$^{100}$Mo (Q=3.034~MeV)\cite{mo1}, $^{82}$Se (Q=2.995~MeV)\cite{qdot},
+$^{116}$Cd (Q=2.81~MeV)\cite{qdot, cd1}, $^{130}$Te (Q=2.533~MeV)\cite{qdot, biller},
+$^{136}$Xe (Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn (Q=2.29~MeV)\cite{sn1}.
 Xenon gas readily dissolves into liquid scintillator. For the other isotopes,
@@
 
-\section{Geant4 Simulation}
+\section{Geant4 simulation}
+\label{sim_section}
 In order to study the effects relevant to directional reconstruction
-in liquid scintillators, a
-\GEANT\cite{geant4one,geant4two} simulation has been constructed. The
-simulation uses \GEANT~version 4.9.6 with the default liquid scintillator
+in liquid scintillators, a \GEANT \newline 
+\cite{geant4one,geant4two} simulation
+has been constructed. The simulation uses \GEANT~version 4.9.6 with the default liquid scintillator
 optical model, in which optical photons are
@@
 The detector geometry is a sphere of 6.5~m radius filled with
-scintillator. FIG.~\ref{detector_view} shows the geometry and the
+scintillator. Figure \ref{detector_view} shows the geometry and the
 Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event. The
@@
 
-\begin{figure*}[tbh]
+\begin{figure}[tbh]
 \begin{center}
@@
         \end{center}
-\end{figure*}
+\end{figure}
 
-The initial studies described here are made with single 5~MeV electrons. 
-Future studies will work on increasing the complexity and lowering 
-the electron energies as is needed for $0\nu\beta\beta$ studies. Three 
+The studies in sections \ref{sim_section} to \ref{reconstruction_sec} are made with single 5~MeV electrons, lower energies are
+discussed in section \ref{edep_section}. Future studies will work on increasing the complexity
+of the event as is needed for $0\nu\beta\beta$ studies. Three
 effects primarily contribute to the timing of the scintillator detector
@@
 
-\section{Detector Timing}
+\section{Detector timing}
 \label{detector_timing_sec}
@@
 We first discuss results for the default simulation settings described
-in the previous section. FIG.~\ref{time_plots_comparison} (a) shows
+in the previous section. Figure \ref{time_plots_comparison} (a) shows
 the TTS-smeared photoelectron (PE) detection times for 1000 simulated
@@
 
-FIG.~\ref{Cherenkov_cone} displays the angular distribution of PE
+Figure \ref{Cherenkov_cone} displays the angular distribution of PE
 hits after the time cut. Although this time cut is a simplification of actual time
@@
 scintillation light ($R_{C/S}=0.25$) for $t<34.0$~ns (see
-FIG.~\ref{time_plots_comparison} (b)). This shows that a low
+figure \ref{time_plots_comparison} (b)). This shows that a low
 photodetector TTS is critical for directionality reconstruction and
@@
 
-\section{Detector Wavelength Response}
+\section{Detector wavelength response}
 \label{detector_wavelength_response_sec} 
@@
 GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}.
-FIG.~\ref{time_plots_comparison}(c) shows the results for the
+Figure \ref{time_plots_comparison}(c) shows the results for the
 modified simulation with high QE in the red spectral region. The
@@
 
-\section{Scintillator Emission Spectrum}
+\section{Scintillator emission spectrum}
 \label{scintillator_emission_sec}
@@
 find $R_{C/S}=0.17$ for the default 34.0~ns timing cut.  Compared to
-the default case shown in FIG.~\ref{time_plots_comparison}(a) the
+the default case shown in figure \ref{time_plots_comparison}(a) the
 separation is worse (as expected) because the scintillation light
@@
 average photon travel times.
-
-\section{Reconstruction}
-\label{reconstruction_sec}
-
-\begin{figure*}[tbh]
+\begin{figure}[tbh]
         \begin{center}
@@
         %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.4]{graphs/hDir_34ns.pdf}}
-        \subfigure{\includegraphics[scale=0.4]{graphs/hVtx_34ns.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs/hDir_34ns.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs/hVtx_34ns.pdf}}
         \caption[]{\label{fig:reco} (Left) The reconstructed direction,
-        $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|},
-        \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons
+        $(p_x/|\vec{p}|, p_y/|\vec{p}|,
+        p_z/|\vec{p}|)$, for the simulation of 1000 electrons
         (5~MeV). In the simulation the electrons are produced along the
-        x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
+        x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0), and originate
         from the center of the 6.5m-radius detector, $\vec{r}$ =
@@
         vertex position, $(x,y,z)$, for the same simulation.}
+\end{center}
+\end{figure}
 
-\end{center}
-\end{figure*}
+\section{Reconstruction}
+\label{reconstruction_sec}
 
@@
 For the purpose of testing the reconstruction algorithm we use 1000
-simulated electrons with an energy of 5~MeV. The electrons are
-simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
-the x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0). FIG.~\ref{fig:reco}
+simulated electrons with an energy of 5~MeV, lower energies are studied in the next section.
+The electrons are simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
+the x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0). Figure \ref{fig:reco}
 shows the vertex reconstruction. The vertex is reasonably well
@@
 as coming from a common point. The reconstruction of the direction
-also is shown in FIG.~\ref{fig:reco}. It shows that for the majority
+also is shown in figure \ref{fig:reco}. It shows that for the majority
 of the events the initial electron direction is reconstructed well.
@@
 used.
+
+\section{Energy dependence}
+\label{edep_section}
+In the previous sections we presented results on single 5~MeV electrons.
+In this section we study two lower energies, 1.4~MeV and 2.1~MeV. These
+energies correspond to Q/2 for the double beta decay of
+$^{116}$Cd and $^{48}$Ca, respectively\cite{cd1,biller}. The isotope
+$^{116}$Cd was chosen because of its potential use in quantum-dot-doped
+scintillators\cite{qdot,qdot2} and $^{48}$Ca was chosen
+to cover the Q-value range of $0\nu\beta\beta$ candidate isotopes.
+
+\begin{figure}
+        \begin{center}
+        \includegraphics[scale=0.5]{graphs/PEandR_vs_energy.pdf}
+        \caption[]{The energy dependence of the mean number of PEs after the 34.0~ns time cut is shown for Cherenkov-induced PEs (black circles) and scintillation-induced PEs (black squares).
+The ratio between the mean number of Cherenkov-induced and scintillation-induced PEs is shown as blue triangles, values are given on the right y-axis.
+The statistical errors are too small to be seen.
+\label{Edep_NPE}}
+        \end{center}
+\end{figure}
+
+The two additional simulation sets with 1.4 MeV and 2.1 MeV electrons were
+generated using the default simulation configuration described in section
+\ref{sim_section}. The PE time distribution for the default settings is shown
+in figure \ref{time_plots_comparison} (a) for 5~MeV electrons. The shape
+of the scintillation and Cherenkov spectra is similar for the lower energies (not
+shown here). In figure \ref{Edep_NPE}, the energy-dependent mean number of
+PEs per event after the 34.0~ns time cut is shown for Cherenkov-induced and
+scintillation-induced PEs, as well as their ratio $R_{C/S}$. The mean number
+of PEs from Cherenkov (scintillation) light is 21.8 (52.1), 38.4 (76.8)
+and 108 (171) for 1.4~MeV, 2.1~MeV and 5~MeV, respectively. This gives the
+ratios $R_{C/S}$ = 0.42, 0.50 and 0.63: The decrease in Cherenkov-induced
+PEs is stronger than the decrease in scintillation-induced PEs as the energy
+is lowered.
+
+\begin{figure}[tbh]
+        \begin{center}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
+        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs/hDir_pq_1p4.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs/hVtx_pq_1p4.pdf}}
+        \caption[]{(Left) The reconstructed direction,
+        $(p_x/|\vec{p}|, p_y/|\vec{p}|,
+        p_z/|\vec{p}|)$, for the simulation of 1000 electrons
+        (1.4~MeV). In the simulation the electrons are produced along the
+        x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0), and originate
+        from the center of the 6.5m-radius detector, $\vec{r}$ =
+        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
+        in the reconstruction. The quantum efficiency of the bialkali
+        photocathode is taken into account. (Right) The reconstructed
+        vertex position, $(x,y,z)$, for the same simulation. \label{low_energy_reco}}
+\end{center}
+\end{figure}
+\begin{figure}
+        \begin{center}
+        \includegraphics[scale=0.4]{graphs/hCos_vs_E.pdf}
+        \caption[]{Mean cosine of the angle between the true initial electron direction and the reconstructed direction, depending on the energy of the initial electrons.
+For each energy 1000 events have been simulated.
+Statistical errors are
+shown. \label{Edep_angle}}
+        \end{center}
+\end{figure}
+
+The reconstruction algorithms outlined in section \ref{reconstruction_sec} have also been applied
+to the simulations at lower energies. Figure \ref{low_energy_reco} shows the results for 
+the lowest simulated electron energy, 1.4~MeV. Most events are still reconstructed well, despite the lower
+number of PEs and the decreased $R_{C/S}$. For 1.4~MeV electrons, the distribution RMS values for all three
+reconstructed coordinates are smaller than 4.5~cm. The mean cosine of the angle between the true direction
+and the reconstructed direction for different energies of the initial electron is shown in figure
+\ref{Edep_angle}. The direction reconstruction performance is still promising for energies as low as 1.4~MeV.
 
@@
 from those for water Cherenkov detectors are able to reliably
-reconstruct the position and direction of 5~MeV electrons. More
-detailed simulation and advanced reconstruction algorithms will need
-to be developed to move to lower energies and more complicated event
+reconstruct the position and direction of electrons with energies in the few MeV range.
+More detailed simulation and advanced reconstruction algorithms will need
+to be developed to move to more complicated event
 topography, such as those in neutrino-less double-beta decay, but the technique already

### 2013-10-14 Direction_JINST_v3.tex
Direction_JINST_v1.tex has the same content as Direction_v7.tex but is in JINST format. The references have been adjusted to match the JINST journal style. We use JHEP.bst as recommended in the JINST author manual. Direction_JINST_v2.tex has an additional section with three additional plots containing results for 1.4 and 2.1 MeV. Direction_JINST_v3.tex is another iteration of Direction_JINST_v2.tex and Direction_JINST_v4.tex is the version which Andrey, Henry, Matt and Christoph propose for publication. Some more details can be found in README.md --Christoph

--- 2013-10-14 Direction_JINST_v2.tex
@@
         \begin{center}
-        \includegraphics[scale=0.5]{graphs/PEandR_vs_energy.pdf}
+        \includegraphics[scale=0.5]{graphs/PEandR_vs_energy_line.pdf}
         \caption[]{The energy dependence of the mean number of PEs after the 34.0~ns time cut is shown for Cherenkov-induced PEs (black circles) and scintillation-induced PEs (black squares).
-The ratio between the mean number of Cherenkov-induced and scintillation-induced PEs is shown as blue triangles, values are given on the right y-axis.
+The ratio between the mean number of Cherenkov-induced and scintillation-induced PEs is shown as blue triangles connected by a line, values are given on the right y-axis.
 The statistical errors are too small to be seen.
@@
         \includegraphics[scale=0.4]{graphs/hCos_vs_E.pdf}
-        \caption[]{Mean cosine of the angle between the true initial electron direction and the reconstructed direction, depending on the energy of the initial electrons.
+        \caption[]{Mean cosine of the angle between the true initial electron direction and the reconstructed direction, as a function of the electron energy.
 For each energy 1000 events have been simulated.

### 2013-10-14 Direction_JINST_v4.tex
Direction_JINST_v1.tex has the same content as Direction_v7.tex but is in JINST format. The references have been adjusted to match the JINST journal style. We use JHEP.bst as recommended in the JINST author manual. Direction_JINST_v2.tex has an additional section with three additional plots containing results for 1.4 and 2.1 MeV. Direction_JINST_v3.tex is another iteration of Direction_JINST_v2.tex and Direction_JINST_v4.tex is the version which Andrey, Henry, Matt and Christoph propose for publication. Some more details can be found in README.md --Christoph

--- 2013-10-14 Direction_JINST_v3.tex
@@
 
+\pdfoutput=1
 \documentclass[cits]{JINST}
+
 %%\usepackage{multirow}
@@
 
-\keywords{Scintillators, scintillation and light emission processes (solid, gas and liquid scintillators); Cherenkov detectors; Large detector systems for particle and astroparticle physics; Neutrino detectors; Simulation methods and programs}
+\keywords{Scintillators, Large detector systems for particle and astroparticle physics; Neutrino detectors; Simulation methods and programs}
 
@@
 effects primarily contribute to the timing of the scintillator detector
-system. First, the simulated travel time of the initial 5~MeV electron is typically between
+system. First, the simulated travel time of an initial 5~MeV electron is typically between
 0.10 and 0.15~ns, while the travel distance is about 3~cm. Second, the
@@
 In this section we study two lower energies, 1.4~MeV and 2.1~MeV. These
-energies correspond to Q/2 for the double beta decay of
+energies correspond to Q/2 for the double-beta decay of
 $^{116}$Cd and $^{48}$Ca, respectively\cite{cd1,biller}. The isotope
@@
         \begin{center}
-        \includegraphics[scale=0.5]{graphs/PEandR_vs_energy_line.pdf}
-        \caption[]{The energy dependence of the mean number of PEs after the 34.0~ns time cut is shown for Cherenkov-induced PEs (black circles) and scintillation-induced PEs (black squares).
-The ratio between the mean number of Cherenkov-induced and scintillation-induced PEs is shown as blue triangles connected by a line, values are given on the right y-axis.
+        \includegraphics[scale=0.5]{graphs/PEandR_vs_energy_final.pdf}
+        \caption[]{The energy dependence of the mean number of PEs after the 34.0~ns time cut is shown for Cherenkov-induced PEs (black open circles, dotted line) and scintillation-induced PEs (black open squares, dashed line).
+The ratio between the mean number of Cherenkov-induced and scintillation-induced PEs is shown as blue filled triangles, values are given on the right y-axis.
 The statistical errors are too small to be seen.
@@
         \begin{center}
-        \includegraphics[scale=0.4]{graphs/hCos_vs_E.pdf}
+        \includegraphics[scale=0.4]{graphs/hCos_vs_E_final.pdf}
         \caption[]{Mean cosine of the angle between the true initial electron direction and the reconstructed direction, as a function of the electron energy.
@@
 Anghel, and Tian Xin, for their continued work in developing the
-WCSimAnalaysis algorithms and for their insights and expertise
+WCSimAnalysis algorithms and for their insights and expertise
 regarding issues related to Cherenkov reconstruction with
-fast-timing. The authors also thank Micheal Smy for his development of
+fast-timing. The authors also thank Michael Smy for his development of
 the quadruplet-based vertex-finding method. L. Winslow would like to

### 2014-01-06 Direction_JINST_v6.tex
version with referee comments and one pass by Henry. --Lindley

--- 2013-10-14 Direction_JINST_v4.tex
@@
 decay searches in which the final state consists of a pair of
-electrons with energies in the few MeV range.  The observation of this
+electrons with energies in the $\sim$1-2~MeV range.  The observation of this
 rare decay would prove that the neutrino is a Majorana particle, which
@@
 mass and may provide a possible explanation of the matter-antimatter
-asymmetry in the universe.  Currently one of the best limits for the
+asymmetry in the universe\cite{leptogenesis}.  Currently one of the best limits for the
 $0\nu\beta\beta$ half-life comes from the scintillating detector
@@
 $\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
-providing energy resolutions of $\sim$5\%. This is roughly a factor of
+providing energy resolutions of $\sim$5\% \sigma_{e}$/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
 two better than water Cherenkov detectors, the other developed
 technology that can be economically scaled to these large
-masses. However, for scintillator-based detectors, while the energy
-resolution is good due to the abundance of light, the light is
-isotropic and does not retain the directional information of the
-primary particle.  In contrast, the direction of the particle can be
-reconstructed from the Cherenkov cone in water-based detectors,
-although the energy resolution rapidly degrades below $\sim$5~MeV. For
-double-beta decay in particular, but also for neutrino interactions,
-the directional information can be a strong suppressant of
-backgrounds.
+masses.
+However, this energy resolution is much poorer than other technologies being used for $0\nu\beta\beta$ searches: Ge detectors\cite{gerda2013}, Te bolometers\cite{Alessandria:2011rc}, tracking detectors\cite{SuperNEMO}, liquid Xe time projection chambers (TPCs)\cite{EXO2012} and high pressure gaseous Xe TPCs\cite{NEXTsipm}.
 
-In a liquid-scintillation-based detector, Cherenkov light is also
-produced, although most is absorbed and re-emitted as part of the
-scintillation processes.  However, some fraction retains its
+Scintillation light is isotropic.
+At these low energies, it does not contain sufficient to reconstruct the track of the outgoing particles, although at higher energies it may\cite{john}.
+Cherenkov light is also
+produced for electrons above threshold. Most is absorbed and re-emitted as part of the
+scintillation processes; however some fraction retains its
 directional information. If this directional Cherenkov light can be
 isolated from the copious isotropic scintillation light, it may be
-possible to reconstruct the direction of the primary particle or, in
-the case of double-beta decay, to determine the existence and topology
-of the pair.  The addition of directionality is thus a powerful tool
-for background rejection.  In this paper, we develop a technique for
+possible to reconstruct the direction of the primary particle. The addition of directionality is a powerful tool
+for background rejection, especially for $0\nu\beta\beta$ searches.
+In the high pressure TPCs, reduction factors on the order of $10^4$ have been achieved\cite{Gotthard}.
+It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO}.
+The addition of a directional signal would make large-scale liquid scintillator detectors more competitive for the next generation of $0\nu\beta\beta$ searches and merits investigation.
+
+This is the first in a series of papers exploring directionality in large-scale liquid scintillator detectors. In this paper, we develop a technique for
 separating the Cherenkov and scintillation light using the photon arrival
@@
 scintillating neutrino detectors.
-This is a different technique and application than the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
+This is a different technique and application from the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
+We then use these results as input into a traditional direction reconstruction developed for water Cherenkov detectors.
+Since the reconstruction of the direction of $\sim$1~MeV particles has not been achieved before, we start these studies with the most simple case of a single particle at the center of the detector.
+We also start with an easier test case of a 5~MeV electron like that from a $^{8}$B solar neutrino interaction.
+With the higher photon statistics at this energy, we verify the technique and the different detector parameters that affect it.
+We then lower the energy to that relevant to a single electron in a $0\nu\beta\beta$.
 
@@
 function of wavelength for the photodetectors being used. One typical
-scintillator mixture uses pseudocumene as the solvent with 1-5~g/l of
-PPO as the fluorophore. This mixture has a peak emission at about 400~nm
+scintillator mixture uses pseudocumene(1,2,4-trimethylbenzene) as the solvent with 1-5~g/l of
+PPO (2,5-diphenyloxazole) as the fluorophore. This mixture has a peak emission at about 400~nm
 where bialkali photomultiplier tubes (PMTs) are most sensitive and the
@@
 \end{equation}
-where $n(\lambda)$ is the wavelength-dependent index of refraction and
-$\beta$ is the velocity of the incoming particle. In a large detector,
-the Cherenkov light produced at wavelengths shorter than the
+where $N$ is the number of Cherenkov photons, $\lambda$ is the wavelength of the photon, x is the distance travelled by the particle, $Z$ is the charge of the particle, $\alpha$ is the fine structure constant, $n(\lambda)$ is the wavelength-dependent index of refraction and
+$\beta$ is the velocity of the particle.
+
+The Cherenkov light produced at wavelengths shorter than the
 absorption cutoff of the scintillator will be absorbed and re-emitted
@@
 propagate across the detector, retaining their directional
-information. These undisturbed Cherenkov
+information.
+For the example scintillator above, n(400nm)=1.459\cite{OlegThesis}.
+The absorption cutoff is between 370~nm and 360~nm where the attenuation length rapidly drops from 6.5~m to 0.65~m.
+For a 5~MeV electron, this yields 744 Cherenkov photons per event; dropping down to a 1~MeV electron yields 89 Cherenkov photons per event when integrating from the cutoff wavelength at 360~nm to 550~nm, the point above which a small number of photons is detected due to a standard photocathode's quantum efficiency.
+For reference at this cutoff wavelength, the index of refraction n(370nm)=1.466.
+This can be translated into an effective Cherenkov threshold for electrons at 0.188~MeV.
+
+All photons including these undisturbed Cherenkov
 photons will have timing determined by the group 
@@
 \end{equation}
-The longer wavelength Cherenkov photons typically arrive before the
+This means photons at the scintillation cutoff at 370~nm will travel at 0.191~m/ns while those at 600~nm travel at 0.203 m/ns.
+Since on average the undisturbed Cherenkov photons have longer wavelengths, they will arrive before the
 scintillation light, which is slowed by both the scintillation
@@
 to the Q-value of the particular isotope. The individual
-electrons follow distributions of energies and angular correlations, 
-a probable case being equal division of energy between back-to-back
-electrons\cite{phasespace}. This case is shown in figure \ref{detector_view}
-for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
+electrons follow distributions of energies and angular correlations which depend on the underlying $0\nu\beta\beta$ decay mechanism\cite{SuperNEMO,newphysics0nuBB,bandv}.
+For light Majorana neutrino exchange, a probable case is the equal division of energy between two
+electrons emitted at a large angle relative to each other\cite{phasespace, SuperNEMO}. An example case is shown in figure \ref{detector_view}
+for a $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
 half-life is inversely proportional to the phase-space factor, isotopes with
-higher Q-values are preferred and due to backgrounds from the daughters
-of the $^{238}$U and $^{232}$Th decay chains those with Q-values at or
-above 2.6~MeV are most often considered for $0\nu\beta\beta$ searches.
+higher Q-values are preferred. Due to backgrounds from the daughters
+of the $^{238}$U and $^{232}$Th decay chains, isotopes with Q-values above the 2.6~MeV gamma from $^{208}$Tl are preferred; however slightly lower Q-values can be tolerated if there is sufficient background rejection.
 There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb},
-but only a handful with large Q-values.
+but only a handful with sufficiently large Q-values.
 
@@
 scintillator with a long attenuation length in the wavelength region of
-interest. As an alternative to single-atom-doping, recently nanocrystals 
-formed by candidate isotopes have been explored as a dopant\cite{qdot,qdot2}.
+interest. Recently nanocrystals formed by candidate isotopes have been explored as an alternative to doping by single atoms,\cite{qdot,qdot2}.
 
@@
 default scintillator properties have been chosen to match a KamLAND-like
-scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene
-(1,2,4-trimethylbenzene) and 1.52~g/l PPO (2,5-diphenyloxazole). The
+scintillator\cite{kamland2003}: 80\% n-dodecane, 20\% pseudocumene and 1.52~g/l PPO. The
 scintillator properties implemented in the simulation include the
@@
 and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}.
-Variations from the baseline KamLAND case are
-discussed below. Re-emission of absorbed photons in the scintillator
-bulk volume and scattering have not yet been included, but are not
-expected to change the conclusions here.
+This is a standard scintillator.
+The most critical value is the attenuation length at 400~nm which is 25~m.
+The attenuation length drops precipitously at 370~nm from 6.5~m to 0.65~m at the cutoff wavelength at 360~nm.
+Variations from the baseline KamLAND case are discussed below.
+
+Re-emission of absorbed photons in the scintillator
+bulk volume and optical scattering, specifically Rayleigh scattering, have not yet been included by default.
+A test simulation shows that the effect of optical scattering is negligible.
+The reason is that the cutoff is very steep below 360~nm and almost no photons reach the sphere, so optical scattering makes no difference for short wavelengths.
+Above about 395~nm, the attenuation length is greater than 20~m so both scattering and absorption are not very likely and scattering is negligible.
+The intermediate region is rather small.
+A similar argument holds for re-emission.
+Scattering length measurements and discussions can be found in Ref.~\cite{Wurm:2010ad}.
 
@@
 as fully absorbing (no reflections), with a photodetector coverage of
-100\%. Two important photodetector properties have been varied: 1)
+100\%. Similar to optical scattering, reflections at the sphere are a small effect that would create a small tail at longer times. Two important photodetector properties have been varied: 1)
 the transit-time spread (TTS, default $\sigma$ = 0.1~ns) and 2) the
@@
 production. The default is the QE of a bialkali photocathode (Hamamatsu
-R7081 PMT)\cite{Hamamatsu_R7081}, for which digitized values come from the Double Chooz\cite{dctwo}
+R7081 PMT)\cite{Hamamatsu_R7081}. The QE values as a function of wavelength come from the Double Chooz\cite{dctwo}
 Monte Carlo simulation. We note that the KamLAND 17-inch PMTs use the
@@
         34.0~ns. (a) Default simulation: bialkali photocathode and TTS =
-        0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
-        from scintillation and 108~PEs from Cherenkov light. (b)
+        0.1~ns ($\sigma$). After the 34.0~ns time cut, 171~PEs
+        from scintillation and 108~PEs from Cherenkov light are detected. (b)
         Default simulation settings except for TTS = 1.28~ns (KamLAND
-        17 in. PMTs). After the 34.0~ns time cut we get 349~PEs from
-        scintillation and 88~PEs from Cherenkov light. (c) Default
+        17 in. PMTs). After the 34.0~ns time cut, 349~PEs from
+        scintillation and 88~PEs from Cherenkov light are detected. (c) Default
         simulation settings except for a GaAsP photocathode. After the
-        34.0~ns time cut we get 226~PEs from scintillation and 229~PEs
-        from Cherenkov light. \label{time_plots_comparison}}
+        34.0~ns time cut, 226~PEs from scintillation and 229~EPS
+        from Cherenkov light are detected. \label{time_plots_comparison}}
         \end{center}
@@
 
-The studies in sections \ref{sim_section} to \ref{reconstruction_sec} are made with single 5~MeV electrons, lower energies are
-discussed in section \ref{edep_section}. Future studies will work on increasing the complexity
-of the event as is needed for $0\nu\beta\beta$ studies. Three
-effects primarily contribute to the timing of the scintillator detector
-system. First, the simulated travel time of an initial 5~MeV electron is typically between
-0.10 and 0.15~ns, while the travel distance is about 3~cm. Second, the
-scintillation light emission follows a distribution characterized by
-scintillator-specific rise and decay times. Before the solutes in
-liquid scintillator can emit optical photons, the energy has to be
-transferred from the solvent to the solute. The time constant of this
+%The studies in sections \ref{sim_section} to \ref{reconstruction_sec} are made with single 5~MeV electrons, lower energies are
+%discussed in section \ref{edep_section}. Future studies will work on increasing the complexity
+%of the event as is needed for $0\nu\beta\beta$ studies.
+
+Four effects primarily contribute to the timing of the scintillator detector
+system: the travel time of the particle, the time constants of the scintillation process, chromatic dispersion and the timing of the photodetector.
+First, the simulated travel time of a 5~MeV electron is 0.108$\pm$0.015~ns.
+This corresponds to an average path length of 3.1~cm and a final distance from the origin of 2.6~cm.
+The time until the electron drops below Cherenkov threshold is 0.106$\pm$0.015~ns.
+We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction; however before the electron drops below Cherenkov threshold the scattering angle is small and therefore the Cherenkov light encodes the direction of the primary electron.
+The scattering physics is handled by Geant4's ``Multiple Scattering" process which is valid down to 1~keV, where atomic shell structure becomes important\cite{geant4scatt}.
+In the energy range important for $0\nu\beta\beta$, a 1.4~MeV electron travels a total path length of 0.8~cm, has a distance from the origin of 0.6~cm in 0.030$\pm$0.004~ns  and takes 0.028$\pm$0.004~ns to drop below Cherenkov threshold.
+The scattering follows the same pattern.
+
+The scintillation light emission follows a distribution characterized by
+scintillator-specific rise and decay times. The first step in the scintillation process is the transfer of energy from the solvent to the solute. The time constant of this
 energy transfer accounts for a rise time in scintillation light
@@
 distribution (quoted above) are values specific to the KamLAND
-scintillator. Third, chromatic dispersion turns out to be an important
-effect in a 6.5m-radius detector at the level of precision needed for
-direction reconstruction.
+scintillator.
 
@@
 light traveling faster than blue light.  In order to study the
-measurability of the time differences, we simulated 5~MeV electrons at
+measurability of the time differences due to this chromatic dispersion, we simulated 5~MeV electrons at
 the center of the sphere where we used instantaneous scintillation
 emission with the quantum-efficiency applied, but not including a
-transit-time spread. The true hit time distributions of photoelectrons
+transit-time spread.
+The higher energy provides larger photon statistics with which to evaluate the method, debug the simulation and is of interest to neutrino-electron scattering experiments.
+The true hit time distributions of photoelectrons
 were analyzed for scintillation light and Cherenkov light
@@
 number which can be different by orders of magnitude depending on the
-detector type. The default TTS of 0.1~ns ($\sigma$) is a value which
-can be achieved with the large area picosecond photodetectors
-(LAPPDs)\cite{LAPPDSum,LAPPDTDR} and possibly hybrid photodetectors
+detector type. The default TTS of 0.1~ns ($\sigma$) can be achieved with large area picosecond photodetectors
+(LAPPDs)\cite{Adams:2013nva,RSI_paper,PSEC4_paper,anode_paper} and possibly hybrid photodetectors
 (HPDs)\cite{hpdThesis}; even significantly lower TTS numbers are
-realistic with the LAPPD\cite{RSI_paper,PSEC4_paper,anode_paper}. We
-note that uncertainties in the vertex reconstruction will produce a
-similar effect to the smearing due to the TTS.
+realistic with the LAPPD\cite{RSI_paper,PSEC4_paper,anode_paper}.
 
@@
 \ref{scintillator_emission_sec}, we study the
-photoelectron timing for different detector configurations. We focus
+photoelectron timing for different detector configurations at 5~MeV. We focus
 on the idea of increasing the discrimination between Cherenkov and
@@
 used for event reconstruction.
+With the successful reconstruction at 5~MeV, we then lower the energy of the simulated electrons and show that it is possible to reconstruct electrons in the range interesting for $0\nu\beta\beta$.
 
@@
 time cut of $t\leq$ 34.0~ns is applied using the truth information to
-isolate the Cherenkov light in this early time window. For the default
+isolate the Cherenkov light in this early time window, see figure \ref{time_plots_comparison} (a). For the default
 simulation case, the average number of PEs per event coming from
@@
         \begin{center}
-        \includegraphics[scale=0.40]{graphs/cos_psi_34_h.pdf}
+        \includegraphics[scale=0.40]{graphs2/combination_cos_psi_34_h.pdf}
         \caption[]{The angular distribution of photoelectron hits
         relative to the original electron direction, $\cos(\Psi) =
-        x_{hit}/|\vec{r}_{hit}|$. The sample consists of 1000 events
-        with a 5~MeV electron produced at the detector center. Default
+        x_{hit}/|\vec{r}_{hit}|$.
+Three energies are shown: 5~MeV (Black - Top), 2.1~MeV (Red - Middle), 1.4~MeV (Blue - Bottom).
+Each sample consists of 1000 events produced at the detector center.
+Default
         simulation settings are used and both Cherenkov and
@@
 that the directional signal conveyed by the Cherenkov photons is not
-erased by scattering of the initial 5~MeV electrons.
+erased by scattering of the initial electrons even at 1.4~MeV.
 
@@
 it is possible to optimize the wavelength-dependence of the
-photocathode. Since Cherenkov photons which pass through meters of
+photocathode. Since Cherenkov photons that pass through meters of
 scintillator have on average longer wavelengths than scintillation
-photons, a photodetector which is more sensitive at long wavelengths
+photons, a photodetector that is more sensitive at long wavelengths
 increases not only the absolute number of PEs but also the ratio
@@
 major motivation for quantum-dot-doped scintillator is control of the
-emission spectra by tuning the size or composition of the quantum
-dots; quantum dots can also provide a mechanism for introducing an 
-isotope for studying double-beta decay.
+emission spectrum by tuning the size or composition of the quantum
+dots.
 
-The emission spectrum of commercial alloyed core/shell
+Quantum dots can also provide a mechanism for introducing an 
+isotope for studying double-beta decay. The emission spectrum of commercial alloyed core/shell
 CdS$_x$Se$_{1-x}$/ZnS quantum dots was measured in
@@
 find $R_{C/S}=0.17$ for the default 34.0~ns timing cut.  Compared to
-the default case shown in figure \ref{time_plots_comparison}(a) the
+the default case shown in figure \ref{time_plots_comparison}(a), the
 separation is worse (as expected) because the scintillation light
@@
 could yield quantum dots which have similar, single peak emission
-shapes at lower wavelengths. This case has been simulated using the
+shapes at shorter wavelengths. This case has been simulated using the
 same spectral shape of the measured core-shell quantum dot emission
-but shifted to lower wavelengths such that the emission peak is
+but shifted to shorter wavelengths such that the emission peak is
 centered at 384~nm. This peak emission value has been measured for
@@
 average photon travel times.
+
 \begin{figure}[tbh]
@@
         %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.375]{graphs/hDir_34ns.pdf}}
-        \subfigure{\includegraphics[scale=0.375]{graphs/hVtx_34ns.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_5p0.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_5p0.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_2p1.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_2p1.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_1p4.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_1p4.pdf}}
         \caption[]{\label{fig:reco} (Left) The reconstructed direction,
         $(p_x/|\vec{p}|, p_y/|\vec{p}|,
-        p_z/|\vec{p}|)$, for the simulation of 1000 electrons
-        (5~MeV). In the simulation the electrons are produced along the
+        p_z/|\vec{p}|)$, for the simulation of 1000 electrons. In the simulation the electrons are produced along the
         x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0), and originate
@@
         photocathode is taken into account. (Right) The reconstructed
-        vertex position, $(x,y,z)$, for the same simulation.}
+        vertex position, $(x,y,z)$, for the same simulation. From Top to Bottom, 5~MeV, 2.1~MeV and 1.4 Mev are shown.}
 \end{center}
@@
 to the problem of reconstructing the position and direction of 5~MeV
-electrons from this early light.  WCSimAnalysis is a water Cherenkov
-reconstruction package developed for the Long Baseline Neutrino
-Experiment (LBNE collaboration)\cite{Blake}. It provides a framework
+electrons from this early light.
+To start, we are neglecting the effects of position dependence.
+In future work the arrival times can be corrected by the time of flight from the reconstructed vertex and the position and direction fitted simultaneously.
+The vertex reconstruction uncertainty leads to an additional smearing of the arrival time distribution of 0.15~ns for every 3~cm of position reconstruction uncertainty.
+
+WCSimAnalysis is a water Cherenkov
+reconstruction package developed by the Long Baseline Neutrino
+Experiment (LBNE) Collaboration\cite{Blake}. It provides a framework
 for generic event cleaning, track reconstruction, and particle
@@
 advanced photosensors with sub-cm imaging capabilities and timing
-resolutions below 100 picoseconds\cite{LAPPDSum,LAPPDTDR}.
+resolutions below 100 picoseconds.
 
@@
 Cherenkov light is emitted from a single point in space-time
-$(x_0,y_0,z_0,t_0)$. In actuality, the light is emitted along an
-extended, multi-scattered electron track. However, at the energies
+$(x_0,y_0,z_0,t_0)$. In actuality, the light is emitted along a multi-scattered electron track. However, at the energies
 discussed in this paper, the extent of this track is small (a few cm)
-compared to the scale of the detector (R=6.5 meters) and therefore the typical
+compared to the scale of the detector (R=6.5 meters), which sets the typical
 photon transit distances.
@@
 The electrons are simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
-the x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0). Figure \ref{fig:reco}
+the x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0). Figure \ref{fig:reco} (Top)
 shows the vertex reconstruction. The vertex is reasonably well
@@
 as coming from a common point. The reconstruction of the direction
-also is shown in figure \ref{fig:reco}. It shows that for the majority
+also is shown in figure \ref{fig:reco} (Top). It shows that for the majority
 of the events the initial electron direction is reconstructed well.
@@
 \label{edep_section}
-In the previous sections we presented results on single 5~MeV electrons.
-In this section we study two lower energies, 1.4~MeV and 2.1~MeV. These
+In the previous sections we presented results on single 5~MeV electrons such as might be observed in neutrino-electron scattering.
+In this section we study two lower energies, 1.4~MeV and 2.1~MeV, appropriate for searches for $0\nu\beta\beta$. These
 energies correspond to Q/2 for the double-beta decay of
@@
 \ref{sim_section}. The PE time distribution for the default settings is shown
-in figure \ref{time_plots_comparison} (a) for 5~MeV electrons. The shape
-of the scintillation and Cherenkov spectra is similar for the lower energies (not
+in figure \ref{time_plots_comparison} (a) for 5~MeV electrons. The shapes
+of the scintillation and Cherenkov spectra are similar for the lower energies (not
 shown here). In figure \ref{Edep_NPE}, the energy-dependent mean number of
@@
 scintillation-induced PEs, as well as their ratio $R_{C/S}$. The mean number
-of PEs from Cherenkov (scintillation) light is 21.8 (52.1), 38.4 (76.8)
+of PEs from Cherenkov (scintillation) light for electron energies 1.4(2.1)~MeV is 21.8 (52.1), 38.4 (76.8)
 and 108 (171) for 1.4~MeV, 2.1~MeV and 5~MeV, respectively. This gives the
@@
 
-\begin{figure}[tbh]
-        \begin{center}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
-        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.375]{graphs/hDir_pq_1p4.pdf}}
-        \subfigure{\includegraphics[scale=0.375]{graphs/hVtx_pq_1p4.pdf}}
-        \caption[]{(Left) The reconstructed direction,
-        $(p_x/|\vec{p}|, p_y/|\vec{p}|,
-        p_z/|\vec{p}|)$, for the simulation of 1000 electrons
-        (1.4~MeV). In the simulation the electrons are produced along the
-        x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0), and originate
-        from the center of the 6.5m-radius detector, $\vec{r}$ =
-        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
-        in the reconstruction. The quantum efficiency of the bialkali
-        photocathode is taken into account. (Right) The reconstructed
-        vertex position, $(x,y,z)$, for the same simulation. \label{low_energy_reco}}
-\end{center}
-\end{figure}
 \begin{figure}
@@
 The reconstruction algorithms outlined in section \ref{reconstruction_sec} have also been applied
-to the simulations at lower energies. Figure \ref{low_energy_reco} shows the results for 
-the lowest simulated electron energy, 1.4~MeV. Most events are still reconstructed well, despite the lower
-number of PEs and the decreased $R_{C/S}$. For 1.4~MeV electrons, the distribution RMS values for all three
+to the simulations at lower energies.
+Figure \ref{fig:reco} (Middle) shows the results for 2.1~MeV and Figure \ref{fig:reco} (Bottom) shows the results for 1.4~MeV.
+Most events are still reconstructed well, despite the lower
+number of PEs and the decreased $R_{C/S}$. For 1.4~MeV electrons, the RMS values of the distributions for all three
 reconstructed coordinates are smaller than 4.5~cm. The mean cosine of the angle between the true direction
@@
 be applied wherever scintillation-based detectors are used. A \GEANT~simulation of a simple spherical detector corresponding to a kiloton
-of scintillator shows that timing on the order of 0.1~ns is required
-to separate the directional Cherenkov light from the more abundant
+of scintillator shows that with timing on the order of 0.1~ns the directional Cherenkov light can be separated from the more abundant
 scintillation light. The separation can be improved using
@@
 More detailed simulation and advanced reconstruction algorithms will need
-to be developed to move to more complicated event
-topography, such as those in neutrino-less double-beta decay, but the technique already
-appears promising.
+to be developed to move to more complicated event structure, such as those in neutrino-less double-beta decay. The technique already
+appears promising and justifies future work on the topic.
 
@@
 the quadruplet-based vertex-finding method. L. Winslow would like to
-thank Janet Conrad for many useful discussion on the topic, and
+thank Janet Conrad for many useful discussions on the topic, and
 Katsushi Arisaka for discussions on the possible reach of traditional PMTs
@@
 \bibliographystyle{JHEP}
-\bibliography{DirectionBibliography}
+\bibliography{DirectionBibliography_v1}
 

### 2014-01-06 Direction_JINST_v7.tex
version with referee comments and one pass by Henry. --Lindley

--- 2014-01-06 Direction_JINST_v6.tex
@@
 $\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
-providing energy resolutions of $\sim$5\% \sigma_{e}$/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
+providing energy resolutions of $\sigma(E)=\sim$5 \% $/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
 two better than water Cherenkov detectors, the other developed
@@
 A test simulation shows that the effect of optical scattering is negligible.
+As shown in figure~\ref{scattplot}, scattering causes a small tail at longer times.
 The reason is that the cutoff is very steep below 360~nm and almost no photons reach the sphere, so optical scattering makes no difference for short wavelengths.
@@
 Scattering length measurements and discussions can be found in Ref.~\cite{Wurm:2010ad}.
+
+\begin{figure}
+        \begin{center}
+        \includegraphics[scale=0.4]{graphs2/Cerenkov_PEs_scattering_comparison.pdf}
+        \caption[]{Photoelectron (PE) arrival times after application
+        of the transit-time spread (TTS) for the simulation
+        of 1000 electrons (5~MeV).
+The default simulation (black) is compared to the simulation with optical scattering turned on (red).
+Scattering causes a small tail to longer times as expected.
+\label{scattplot}}
+        \end{center}
+\end{figure}
 

### 2014-01-08 graphs2/PEtime_Cerenkov_scattered_vs_unscattered.C
Added changes from Henry from the weekend. --Lidnley

@@ -1,0 +1,514 @@
+{
+//=========Macro generated from canvas: c1/c1
+//=========  (Sat Jan  4 17:33:13 2014) by ROOT version5.34/03
+   TCanvas *c1 = new TCanvas("c1", "c1",13,35,700,502);
+   c1->SetLogy();
+   c1->Range(23.85278,-1.374102,51.34291,9.258293);
+   c1->SetFillColor(0);
+   c1->SetBorderMode(0);
+   c1->SetBorderSize(2);
+   c1->SetLeftMargin(0.1508621);
+   c1->SetRightMargin(0.04885057);
+   c1->SetTopMargin(0.0720339);
+   c1->SetBottomMargin(0.1292373);
+   c1->SetFrameBorderMode(0);
+   c1->SetFrameLineWidth(2);
+   c1->SetFrameBorderMode(0);
+   
+   TH1F *th1 = new TH1F("th1","",220,28,50);
+   th1->SetBinContent(33,0.002);
+   th1->SetBinContent(34,0.002);
+   th1->SetBinContent(35,0.001);
+   th1->SetBinContent(36,0.002);
+   th1->SetBinContent(37,0.01);
+   th1->SetBinContent(38,0.074);
+   th1->SetBinContent(39,0.437);
+   th1->SetBinContent(40,1.803);
+   th1->SetBinContent(41,3.903);
+   th1->SetBinContent(42,6.269);
+   th1->SetBinContent(43,7.58);
+   th1->SetBinContent(44,8.088);
+   th1->SetBinContent(45,7.778);
+   th1->SetBinContent(46,7.551);
+   th1->SetBinContent(47,7.223);
+   th1->SetBinContent(48,6.847);
+   th1->SetBinContent(49,6.396);
+   th1->SetBinContent(50,6.097);
+   th1->SetBinContent(51,5.874);
+   th1->SetBinContent(52,5.44);
+   th1->SetBinContent(53,5.25);
+   th1->SetBinContent(54,4.78);
+   th1->SetBinContent(55,4.443);
+   th1->SetBinContent(56,3.642);
+   th1->SetBinContent(57,3.278);
+   th1->SetBinContent(58,2.816);
+   th1->SetBinContent(59,2.446);
+   th1->SetBinContent(60,1.991);
+   th1->SetBinContent(61,1.482);
+   th1->SetBinContent(62,0.916);
+   th1->SetBinContent(63,0.427);
+   th1->SetBinContent(64,0.214);
[…]

### 2014-01-08 graphs2/PEtime_Scintillation_scattered_vs_unscattered.C

@@ -1,0 +1,660 @@
+{
+//=========Macro generated from canvas: c1/c1
+//=========  (Sat Jan  4 17:21:05 2014) by ROOT version5.34/03
+   TCanvas *c1 = new TCanvas("c1", "c1",125,95,700,590);
+   c1->Range(24.07421,-10.05662,51.12721,66.43466);
+   c1->SetFillColor(0);
+   c1->SetBorderMode(0);
+   c1->SetBorderSize(2);
+   c1->SetLeftMargin(0.1451149);
+   c1->SetRightMargin(0.04166667);
+   c1->SetTopMargin(0.06972112);
+   c1->SetBottomMargin(0.1314741);
+   c1->SetFrameLineWidth(2);
+   c1->SetFrameBorderMode(0);
+   c1->SetFrameLineWidth(2);
+   c1->SetFrameBorderMode(0);
+   
+   TH1F *th4 = new TH1F("th4","",220,28,50);
+   th4->SetBinContent(42,0.006);
+   th4->SetBinContent(43,0.024);
+   th4->SetBinContent(44,0.059);
+   th4->SetBinContent(45,0.119);
+   th4->SetBinContent(46,0.304);
+   th4->SetBinContent(47,0.662);
+   th4->SetBinContent(48,1.136);
+   th4->SetBinContent(49,1.901);
+   th4->SetBinContent(50,3.015);
+   th4->SetBinContent(51,4.386);
+   th4->SetBinContent(52,6.051);
+   th4->SetBinContent(53,8.345);
+   th4->SetBinContent(54,10.885);
+   th4->SetBinContent(55,14.081);
+   th4->SetBinContent(56,17.366);
+   th4->SetBinContent(57,20.94);
+   th4->SetBinContent(58,24.545);
+   th4->SetBinContent(59,27.923);
+   th4->SetBinContent(60,31.847);
+   th4->SetBinContent(61,35.976);
+   th4->SetBinContent(62,38.451);
+   th4->SetBinContent(63,42.304);
+   th4->SetBinContent(64,44.729);
+   th4->SetBinContent(65,47.479);
+   th4->SetBinContent(66,50.344);
+   th4->SetBinContent(67,51.208);
+   th4->SetBinContent(68,53.045);
+   th4->SetBinContent(69,54.116);
+   th4->SetBinContent(70,54.846);
+   th4->SetBinContent(71,56.053);
+   th4->SetBinContent(72,56.566);
[…]

### 2014-01-08 Direction_JINST_v7.tex
Added changes from Henry from the weekend. --Lidnley

@@
 Scintillation light is isotropic.
-At these low energies, it does not contain sufficient to reconstruct the track of the outgoing particles, although at higher energies it may\cite{john}.
+At these low energies, it does not contain sufficient information to reconstruct the track of the outgoing particles, although at higher energies it may\cite{john}.
 Cherenkov light is also
@@
 scintillating neutrino detectors.
-This is a different technique and application from the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
+This is a different from the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
 We then use these results as input into a traditional direction reconstruction developed for water Cherenkov detectors.
@@
 With the higher photon statistics at this energy, we verify the technique and the different detector parameters that affect it.
-We then lower the energy to that relevant to a single electron in a $0\nu\beta\beta$.
+We then study the technique for two lower energies, 1.4 and 2.1 MeV, that are more relevant to $0\nu\beta\beta$.
 
@@
 information.
-For the example scintillator above, n(400nm)=1.459\cite{OlegThesis}.
+The absorption cutoff for the example scintillator above is 370~nm.
+The index of refraction at  n(370nm)=1.466 and can be translated into an effective Cherenkov threshold for electrons at 0.188~MeV.
+For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 360~nm to 550~nm,  the wavelength above which a small number of photons is detected due to a standard bialkali photocathode's quantum efficiency.
+Lowering the energy to 1~MeV yields 82 Cherenkov photons between 360-550~nm.
+
+%For the example scintillator above, n(400nm)=1.459\cite{OlegThesis}.
 The absorption cutoff is between 370~nm and 360~nm where the attenuation length rapidly drops from 6.5~m to 0.65~m.
@@
 \end{equation}
-This means photons at the scintillation cutoff at 370~nm will travel at 0.191~m/ns while those at 600~nm travel at 0.203 m/ns.
+Photons at the scintillation cutoff of 370nm have a velocity of 0.191 m/ns, while photons with wavelengths of 600 nm are appreciably faster, with a velocity of 0.203 m/ns.
 Since on average the undisturbed Cherenkov photons have longer wavelengths, they will arrive before the
@@
 
-In $0\nu\beta\beta$, the electrons emerge with a combined energy equal
+In neutrino-electron scattering events, a single electron emerges with a distribution of energies with a maximum energy related to the incoming neutrino's energy.
+In comparison, $0\nu\beta\beta$ events are more complicated with two electron emerge with a combined energy equal
 to the Q-value of the particular isotope. The individual
 electrons follow distributions of energies and angular correlations which depend on the underlying $0\nu\beta\beta$ decay mechanism\cite{SuperNEMO,newphysics0nuBB,bandv}.
-For light Majorana neutrino exchange, a probable case is the equal division of energy between two
-electrons emitted at a large angle relative to each other\cite{phasespace, SuperNEMO}. An example case is shown in figure \ref{detector_view}
-for a $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
-half-life is inversely proportional to the phase-space factor, isotopes with
-higher Q-values are preferred. Due to backgrounds from the daughters
-of the $^{238}$U and $^{232}$Th decay chains, isotopes with Q-values above the 2.6~MeV gamma from $^{208}$Tl are preferred; however slightly lower Q-values can be tolerated if there is sufficient background rejection.
-There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb},
-but only a handful with sufficiently large Q-values.
+Figure 1 shows a simulated $^{116}$Cd $0\nu\beta\beta$ event in a model with light Majorana neutrino exchange, for which a probable case is the emission of two electrons with comparable energies at a large angle relative to each other.
 
-Most of the high Q-value candidates
+High Q-value candidates are preferred for $0\nu\beta\beta$  for two reasons.
+First, the measured half-life is inversely proportional to the phase-space factor, therefore the measured rate is expected to be higher from isotopes with higher Q-values.
+Second, the main background for these experiments come from the daughters of the $^{238}$U and $^{232}$Th decay chains.
+The Compton shoulder from gamma-rays is particularly problematic.
+The 2.6~MeV gamma-ray from $^{208}$Tl decay is the highest energy gamma from these decay chains and therefore isotopes above 2.6~MeV are preferred.
+Experiments using isotopes with Q-values below this energy must compensate with improved background rejection techniques.
+
+Most of the high Q-value candidates\cite{tabledbb}
 have been considered as a dopant for a liquid scintillator:
@@
 interest. Recently nanocrystals formed by candidate isotopes have been explored as an alternative to doping by single atoms,\cite{qdot,qdot2}.
-
-\section{Geant4 simulation}
-\label{sim_section}
-In order to study the effects relevant to directional reconstruction
-in liquid scintillators, a \GEANT \newline 
-\cite{geant4one,geant4two} simulation
-has been constructed. The simulation uses \GEANT~version 4.9.6 with the default liquid scintillator
-optical model, in which optical photons are
-assigned the group velocity in the wavelength region of normal
-dispersion.
 
@@
 \end{figure}
+
+\section{Geant4 simulation}
+\label{sim_section}
+In order to study the effects relevant to directional reconstruction
+in liquid scintillators, a \GEANT \newline 
+\cite{geant4one,geant4two} simulation
+has been constructed. The simulation uses \GEANT~version 4.9.6 with the default liquid scintillator
+optical model, in which optical photons are
+assigned the group velocity in the wavelength region of normal
+dispersion.
 
@@
         \begin{center}
-        \includegraphics[scale=0.4]{graphs2/Cerenkov_PEs_scattering_comparison.pdf}
+        \includegraphics[scale=0.4]{graphs2/PEtime_Cerenkov_scattered_vs_unscattered.pdf}
         \caption[]{Photoelectron (PE) arrival times after application
@@
         of 1000 electrons (5~MeV).
-The default simulation (black) is compared to the simulation with optical scattering turned on (red).
-Scattering causes a small tail to longer times as expected.
+The default simulation is shown with optical scattering turned on: all photons (black), un-scattered photons (red dotted), and scattered (blue dashed).
+Scattering causes a very small tail at longer times as expected.
 \label{scattplot}}
@@
 as fully absorbing (no reflections), with a photodetector coverage of
-100\%. Similar to optical scattering, reflections at the sphere are a small effect that would create a small tail at longer times. Two important photodetector properties have been varied: 1)
+100\%. As in the case of optical scattering, reflections at the sphere are a small effect that would create a small tail at longer times. Two important photodetector properties have been varied: 1)
 the transit-time spread (TTS, default $\sigma$ = 0.1~ns) and 2) the
@@
 
-\begin{figure}[tbh]
+\begin{figure}
 \begin{center}
@@
         Default simulation settings except for TTS = 1.28~ns (KamLAND
-        17 in. PMTs). After the 34.0~ns time cut, 349~PEs from
+        17-inch PMTs). After the 34.0~ns time cut, 349~PEs from
         scintillation and 88~PEs from Cherenkov light are detected. (c) Default
@@
 Four effects primarily contribute to the timing of the scintillator detector
-system: the travel time of the particle, the time constants of the scintillation process, chromatic dispersion and the timing of the photodetector.
+system: the travel time of the particle, the time constants of the scintillation process, chromatic dispersion, and the timing of the photodetector.
 First, the simulated travel time of a 5~MeV electron is 0.108$\pm$0.015~ns.
@@
 transit-time spread.
-The higher energy provides larger photon statistics with which to evaluate the method, debug the simulation and is of interest to neutrino-electron scattering experiments.
+The higher energy provides larger photon statistics with which to evaluate the method, debug the simulation, and is of interest to neutrino-electron scattering experiments.
 The true hit time distributions of photoelectrons
@@
 
-In Sections \ref{detector_timing_sec} to
+In sections \ref{detector_timing_sec} to
 \ref{scintillator_emission_sec}, we study the
@@
 positions and the detection times after the TTS resolution has been
-applied. In Section~\ref{reconstruction_sec} these quantities are then
+applied. In section~\ref{reconstruction_sec} these quantities are then
 used for event reconstruction.
@@
 time cut of $t\leq$ 34.0~ns is applied using the truth information to
-isolate the Cherenkov light in this early time window, see figure \ref{time_plots_comparison} (a). For the default
+isolate the Cherenkov light in this early time window, as shown in figure \ref{time_plots_comparison} (a). For the default
 simulation case, the average number of PEs per event coming from
@@
 scintillation light ($R_{C/S}=0.25$) for $t<34.0$~ns (see
-figure \ref{time_plots_comparison} (b)). This shows that a low
-photodetector TTS is critical for directionality reconstruction and
+figure \ref{time_plots_comparison} (b)). This shows that a
+photodetector with a low TTS is critical for directionality reconstruction and
 motivates the use of novel photodetector types.
@@
 \label{detector_wavelength_response_sec} 
-In addition to decreasing the photodetector TTS to enhance $R_{C/S}$,
-it is possible to optimize the wavelength-dependence of the
-photocathode. Since Cherenkov photons that pass through meters of
+Since Cherenkov photons that pass through meters of
 scintillator have on average longer wavelengths than scintillation
@@
 increases not only the absolute number of PEs but also the ratio
-between Cherenkov- and scintillation-induced PEs.
-
-We have run the simulation with the QE of an extended red-sensitive
-GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}.
+between Cherenkov- and scintillation-induced PEs. We have run the simulation with the QE of an extended red-sensitive
+GaAsP photocathode (Hamamatsu R3809U-63)\cite{Hamamatsu_R3899U}, but with the default TTS of 0.1~ns.
 Figure \ref{time_plots_comparison}(c) shows the results for the
@@
 to the problem of reconstructing the position and direction of 5~MeV
-electrons from this early light.
-To start, we are neglecting the effects of position dependence.
-In future work the arrival times can be corrected by the time of flight from the reconstructed vertex and the position and direction fitted simultaneously.
-The vertex reconstruction uncertainty leads to an additional smearing of the arrival time distribution of 0.15~ns for every 3~cm of position reconstruction uncertainty.
-
-WCSimAnalysis is a water Cherenkov
+electrons from this early light. WCSimAnalysis is a water Cherenkov
 reconstruction package developed by the Long Baseline Neutrino
@@
 
+To start, we are neglecting the effects of position dependence.
+In future work the arrival times can be corrected by the time of flight from the reconstructed vertex and the position and direction fitted simultaneously.
+For isotropic light, the vertex reconstruction uncertainty leads to an additional smearing of the arrival time distribution of 0.15~ns for every 3~cm of position reconstruction uncertainty.
+For directional light, the vertex reconstruction uncertainty leads to an effective shift of the arrival time distribution.
+This would change the ideal time cut by 0.15~ns  for every 3~cm of position reconstruction uncertainty.
+We have studied other time cuts from 33~ns to 34.5~ns and the reconstruction remained reasonable.
+
 The results presented in this paper rely on a simple vertex
@@
 discussed in this paper, the extent of this track is small (a few cm)
-compared to the scale of the detector (R=6.5 meters), which sets the typical
+compared to the scale of the detector (R=6.5 meters) that sets the typical
 photon transit distances.
@@
 distribution of an observable known as the ``point time
-residual''\cite{SuperKalgo}. The point time residual is calculated by
-taking the difference between the measured time of a photon hit, and
-the predicted time of the hit, given its distance from the vertex
-hypothesis, a single effective speed of light in the scintillator, and
-the hypothesized $t_0$ of the event. The width of the time residual
+residual''\cite{SuperKalgo}.
+The point time residual is calculated by first choosing a hypothesis for the vertex position and $t_0$ for the event.
+The goodness of fit for these values  is then calculated by taking the difference between the measured and predicted times of each photon hit, using a single effective speed of light in the scintillator.
+The width of the time residual
 distribution over all hits is minimized when the hypothesized vertex
@@
 For the purpose of testing the reconstruction algorithm we use 1000
-simulated electrons with an energy of 5~MeV, lower energies are studied in the next section.
+simulated electrons with an energy of 5~MeV; lower energies are studied in the next section.
 The electrons are simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
 the x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0). Figure \ref{fig:reco} (Top)
-shows the vertex reconstruction. The vertex is reasonably well
-reconstructed around the center of the detector, $\vec{r}$ = (0,0,0),
+shows the vertex reconstruction. The vertex is reasonably well-reconstructed around the center of the detector, $\vec{r}$ = (0,0,0),
 except along the x-axis. The RMS values of the distributions for all three
@@
 to the simulations at lower energies.
-Figure \ref{fig:reco} (Middle) shows the results for 2.1~MeV and Figure \ref{fig:reco} (Bottom) shows the results for 1.4~MeV.
+Figure \ref{fig:reco} (Middle) shows the results for 2.1~MeV and figure \ref{fig:reco} (Bottom) shows the results for 1.4~MeV.
 Most events are still reconstructed well, despite the lower

### 2014-01-10 RefereeResponse/RefereeResponse_v2.tex
Added Christophs edits and committing Henrys referee response. --Lindley

\documentclass[11pt]{article}

\usepackage[usenames,dvips]{pstcol}

\usepackage{color}

\usepackage{amssymb}

\usepackage{amsmath}

%

\input{symbols.tex}

%

\newcounter{question}

\newcommand{\noteq}{{\vskip0.1in\bf \large \color{red} [Q:{\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\noteQ}{{\bf \LARGE \color{blue} [Q: {\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\notea}{{\bf \large \color{red} [A]~}}

\newcommand{\noteA}{{\bf \LARGE \color{blue} [A]~}}

\newcommand{\noteqq}[1]{{\bf \large \color{red} [Q: {\color{black}#1}]~}}

\newcommand{\noteaa}[1]{{\bf \large \color{red} [A: {\color{black}#1}]~}}

%

\textwidth 6.5in

\textheight 9.4in

\topmargin -0.8in

\oddsidemargin 0.0in

\evensidemargin 0.0in

\parindent 0.0in

\def\baselinestretch{1.0}

\def \Russ {Translated from Russian. Fixed. Thanks.}

\def \F    {Fixed.}

\def \FT   {Fixed. Thanks.}

\def \D    {Done.}

\def \AP   {Basically Stylistic- Author's Prerogative.}

\begin{document}

\begin{flushright}

\today

\end{flushright}

%----------------------------------------------------------------

Dear JINST Editors,\\

\parindent 1.0in

We thank the referees for their very helpful and knowledgeable

comments. We have substantially rewritten the manuscript in response

to the comments as detailed in the list below.

There are several areas, however, in which the referees requested

substantially expanding the scope of the paper.  We envisioned this

paper as a proof of  the principle that the use of the fast timing

photodetectors presently being developed would enable the separation

of Cherenkov and scintillation light, as the reconstruction of the

direction of low energy ($\sim$1 MeV) electrons in a liquid

scintillator detector has never been shown before.  In this paper, we

want to: 1) show scintillation/Cherenkov separation is possible; 2)

determine the parameters of a large scintillator detector that are

important to this separation; and 3) demonstrate that reconstruction

techniques based on Cherenkov light would work in idealized

situations.

We are now working on two follow-up papers that are substantially

beyond the scope of the present paper, moving beyond

proof-of-principle into the laborious and time-intensive area of

reconstruction algorithms, areas for which the difficulty is known

well by some of us from our experience on KamLand. The first is a

detailed study of reconstruction in more realistic scenarios. The

second is a proposal for a $0\nu\beta\beta$ experiment  where the detailed

background reduction factors and sensitivity to new physics are

presented, SuperNEMO has a similar paper, but for scintillator

detectors nothing similar exists since a directional signal has not

been considered for them in the past.  Finally, we will be having a

turnover of young personnel so this paper provides an accessible summary

of their work.

\parindent 3.0in

    Sincerely,\\

\parindent 4.0in

		         Lindley Winslow, for the Authors

\parindent 0.0in

%\begin{verbatim}

%\end{verbatim}

\noteq [Referee 1 Intro] ``I like  very much the clean, transparent, simple

and well-motivated methodology (with one exception..). The structure

of the paper is nice''

\notea{\bf Many thanks.}\\

\noteq [Referee 1 Intro] ``However the presentation of the results is a

different thing... I am particularly concerned with the fact...,

instead of showing the resulting CosTheta distribution that would seem

like the obvious result one wants to see. .... Why not to disclose

this openly to allow the reader to evaluate the results in detail?''

\notea{\bf  This is our fault- the desired plot was (is) there

(Figure 3), but was labeled as $p_x/|\vec{p}|$ rather than the obvious

CosTheta that it is. We have made this clear on the figure and in the

text. Thanks.}\\

\noteq [Referee 1: Page 1] ``Please specify FWHM or sigma and give a

reference for both cases.''

\notea{\bf  \D }\\

DOES ANYBODY KNOW WHAT HSE IS ASKING FOR WRT THE REFERENCE?

\noteq [Referee 1: Page 1] ``...but please avoid misplacing the adjective good in

this context. ...''

\notea{\bf  \FT}\\

\noteq [Referee 1: Page 2] ``..the authors should..check and refer to

the work of NEXT and Super-Nemo collaborations:''

\notea{\bf  \D}\\

I DON'T SEE THE REFERENCE TO NEXT IN THE TOPOLOGY PARAGRAPH- DID I MISS

IT? IT SHOULD GO AFTER ``..ESPECIALLY FOR 0NUBETABETA SEARCHES'''?

\noteq [Referee 1: Page 2] ``Define PPO here, not later''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3] ``Please define all magnitudes in eq. 2.1 and 2.2''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3]  ``Indicating the direction of the

electrons should not harm''

\notea{\bf  \D}\\

WAS ANYTHING CHANGED IN THE FIGURE? THE CAPTION MAKES IT CLEAR, AND

THE BRIGHT WHITE AXIS ARROW IS CLEAR... IF WE DON'T CHANGE THE FIG, WE

SHOULD SAY SO AND WHY NOT (EASY)

\noteq [Referee 1: Page 3] ``None of the best limits or claims in bb0

sues an isotope that satisfies this condition..''

\notea{\bf \FT}\\

IS THIS IS FIXED? I COULDN'T FIND IT IN THE CURRENT DRAFT AROUND

THE 2.6 MEV DISCUSSION. .I'M CONFUSED BY THE COMMENT, FRANKLY. DOES

THE PRESENT TEXT NEED MORE WORK? (I.E. I NEED EDUCATING...)

\noteq [Referee 1: Page 3]

\notea{\bf \FT}\\

THIS IS FIXED, RIGHT? I COULDN'T FIND IT IN THE CURRENT DRAFT AROUND

THE 2.6 MEV DISCUSSION. .I'M CONFUSED BY THE COMMENT, FRANKLY. DOES

THE TEXT NEED WORK? (I.E. I NEED EDUCATING...)

\noteq [Referee 1: Page 4] ``could you perhaps include the

attenuation @400nm and n@400..?''

\notea{\bf \D}\\

I SEE THE ATTENUATION LENGTH, BUT NOT N@400

\noteq [Referee 1: Page 4] Mentioning the figure in the citation.

\notea{\bf ?}\\

I DIDN'T UNDERSTAND THIS COMMENT- THE PAPER MENTIONS FIG 4C..

\noteq [Referee 1: Page 4] ``Could you please include also the cutoff

wavelength previously referred to?''

\notea{\bf ?}\\

I DIDN'T UNDERSTAND THIS ONE EITHER

\noteq [Referee 1: Page 4] `` Scattering have not been included..''

\notea{\bf We have added a figure and accompanying text to show that

  the scattering only adds a late tail. Thanks.}\\

\noteq [Referee 1: Page 4] ``For which digitized values come from ??''

\notea{\bf \F}\\

\noteq [Referee 1: Page 4] ``5 MeV electrons.. It is very disturbing

to benchmark the simulation with electrons having an energy doubling

the Qbb of the most common isotopes..''

\notea{\bf We have clarified in several places

that the 5 MeV case is for neutrino-electron experiments. For dbb we

have made it explicit that we consider 1.4 and 2.1 MeV}\\

\noteq [Referee 1: Page 4] ``Please specify also the track length in

the direction of emission''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``Before the solutes in liquid... Please rephrase''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``I hope that [43] is not the final version

but a draft.

\notea{\bf It was a draft; we have changed the reference. Thanks!}\\

\noteq [Referee 1: Page 6] ``So I would be very much convinced ..if

the authors would report Fig. 3 during the electron history

(equivalently, energy or distance, including several histograms for

each case. It would be nice to see at which energy scale or length the

electron isotropizes and the Cherenkov cone smears out.''

\notea{\bf We have added studies of 1.4 MeV and 2.1 MeV electrons,

  which show that the separation and directionality based on timing

  hold as the electron energy is lowered, as shown in the newly-added

  Figures 5 and 6. We have also added the path

  lengths and total distance and time from the vertex. We believe that

  additional plots beyond the ones we have added won't add any useful

  information.}\\

 HSE'S CLEARLY STUCK ON THIS- MISSED THE POINT. HOWEVER, WE NEED TO

  ANSWER THIS DIRECTLY - IS IT POSSIBLE TO MAKE ONE PLOT OF

\noteq [Referee 1: Pages 7-8] ``This is interesting, but specially

section 6 seems a bit off-topic to me. I am not against keeping it...''

\notea{\bf We'd like to keep it. }\\

\noteq [Referee 1: Page 9] ``Please add a histogram of the reconstructed

angle for all energies...''

\notea{\bf }\\

NEEDS ATTENTION...

\noteq [Referee 1: Page 9] ``..the vertex reconstruction has to be

obtained from all photons and the direction from the Cherenkov photons..''

\notea{\bf }\\

I CAN'T WRITE AN ANSWER TO THIS- CHRISTOPHE,ANDREY, MATT, LINDLEY?

(there's lots of email that contains the answer, it seems...)

\noteq [Referee 1: Fig. 7] ``the authors should show the spread of the

cos(theta) distribution for each of the energies.

\notea{\bf \D See Figure 3.}\\

I DON'T HAVE A FIGURE 7. ???

\noteq [Referee 1: Fig. 7 ``It would be nice to show/indicate..also the

Cherenkov threshold for high wave-length (n about 1.4)

\notea{\bf We have added this to the text.}\\

I DON'T HAVE A FIGURE 7. ???

\noteq [Referee 1: Conclusions] ``Should be rewritten after the

previous objections are addressed.''

\notea{\bf \D Thanks!}\\

WE SHOULD REWRITE THE CONCLUSIONS TO BE MORE SPECIFIC WITH MORE

FACTS. AS THEY STAND THEY'RE SORT OF GENERIC- IF THAT'S ALL ONE READS

OF THE PAPER ONE'S NOT MUCH WISER ...

\noteq [Referee 1: Page 7-8] ``I see no convincing proof of it with

the information here presented. Hope the answers help clarifying the issue.''

\notea{\bf We have made the sentence more specific. Thanks.}\\

WE SHOULD BE MORE SPECIFIC- CAN GIVE NUMBERS RATHER THAN OPINIONS

HERE..

%\newpage

\noteq [Referee 2: General Comments] ``The paper introduces an

original technical development which could increase the interest of

large scale scintillation detectors for neutrino physics in general

and neutrinoless double beta decay in particular''

\notea{\bf Warms the cockles of our hearts...}\\

\noteq [Referee 2: General comments] `` ..however the discussion

should deserve...a deeper analysis.''

\notea{\bf Yes. We have extended the analysis to lower energies and

  have extracted much more detail from the simulation, and have

  extensively rewritten the manuscript. It is still at the

  `proof-of-principle level, as to go deeper yet will require vertex

  reconstruction, which is a much bigger effort (we are working on it).}\\

\noteq [Referee 2: General Comments] ``The physics case is included in

the title and mentioned in the text but never discussed in some

detail...A description of why the method is powerful for double beta

decay, and..to what extent it can improve the experimental sensitivity

would complete the discussion

\notea{\bf }\\

NYGREN'S RIGHT- CAN WE ADD A SENTENCE OR TWO TO THE INTRO AND

(PERHAPS) THE CONCLUSIONS?

\noteq [Referee 2: General Comments] ``The chosen example (5 MeV,

head-to-head single electrons) is of little interest for double beta decay.''

\notea{\bf We have added plots and numbers for 1.4 and 2.1 MeV, and

  substantially rewritten the text to differentiate between the cases

  of neutrino-electron scattering and double-beta decay.}\\

\noteq [Referee 2: General Comments] ``The dependence on the decay

position.. is not discussed.''

\notea{\bf To do a credible and useful job on the position dependence

  will require a fairly-detailed vertex reconstruction. This is beyond

  the scope of this present `proof-of-principle' paper. We have tried

  to clarify this in the text. }\\

\noteq [Referee 2: General Comments] Threshold effects as well as the

effect of statistical fluctuations for the low energy tails.. are not

discussed. The same holds for position and direction reconstruction

capabilities''

\notea{\bf }\\

I DON'T KNOW WHAT TO SAY HERE- SOMEBODY MORE EXPERT?

\noteq [Referee 2: Abstract] ``While appropriate for neutrino physics,

energy resolution is definitely worse with respect to other techniques..''

\notea{\bf We have fixed the text. Thanks.}\\

\noteq [Referee 2: Introduction] ``A reference ... would help to understand

the connection with Majorana neutrinos.''

\notea{\bf \D}\\

\noteq [Referee 2: Page 1] Is this 1 sigma or FWHM?

\notea{\bf It's sigma- we have fixed the text.}\\

\noteq [Referee 2: Page 2] ``better than water Cherenkov detectors''

\notea{\bf \FT}\\

PLEASE ADD `` for neutrino detectors'' after ``the other developed

technology'' AT THE TOP OF PAGE 2.

\noteq [Referee 2: Page 2 ] ``strong suppressant of backgrounds''..this

statement should deserve a deeper quantitative discussion ..''

\notea{\bf We agree. However, a useful ``quantitative discussion''

  will require a vertex reconstruction. We will get there, but at present

  it's beyond the scope of `proof-of-principle, looking to see if we

  can see the signal. We have tried to make this clearer in the text. }\\

\noteq [Referee 2: Equ. 2.1] ``..the meaning of the parameter `x' is

not described.''

\notea{\bf \F}\\

\noteq [Referee 2: Page 3] What is the typical cutoff value and the

corresponding useful fraction of Cherenkov light?

\notea{\bf We have added the cutoff values.}\\

DO WE HAVE THE USEFUL FRACTIONS OF CHERENKOV LIGHT? (I DON'T EVEN KNOW

IF THAT'S WELL-DEFINED?)

\noteq [Referee 2: Page 3] `` `Probable case being...' Events with

asymmetric electron energies are far more abundant.''

\notea{\bf \FT}\\

WE SHOULD REWRITE THIS SENTENCE ..E.G. ``We consider the case of

equal..'' OR SOMETHING.. (?) (MAYBE CHANGE 'PROBABLE' TO 'SIMPLE'?)

\noteq [Referee 2: Page 4] `` `Fully absorbing' ''. How are results

influenced by deviations from this hypothesis?''

\notea{\bf This is a level of detail well beyond what we've done so

  far; once we have a real detector design and a full simulation it

  can be addressed, but at this point devoting effort to it

doesn't seem productive.}\\

\noteq [Referee 2: Page 4] ``Single 5 MeV electrons''.

\notea{\bf \F}\\

\noteq [Referee 2: Page 4] ``I believe the `realistic' analysis

mentioned here is indispensable ..for this process''

\notea{\bf We agree, and are working on it. It isn't easy, or,

  unfortunately, fast compared to a post-doc tenure.}\\

\noteq [Referee 2: Page 6] ``The dependence on the position and the

corresponding transit time variations cannot in my opinion be neglected.''

\notea{\bf }\\

THIS NEEDS AN EXPERT'S ANSWER...

\noteq [Referee 2: Figure 3] `` This figure is very interesting

to understand the capacity of the method to maintain directionality

information... The comparison with a single event (or equivalently the

distribution of the variances) and the distribution for 1 MeV

electrons would improve to understand also the limits/power of the method''

\notea{\bf }\\

THIS NEEDS AN EXPERT'S ANSWER...

\noteq [Referee 2: reconstruction] ``It is not clear if limiting the

reconstruction to Cherenkov photons improve reconstruction''

\notea{\bf Yes, we agree. We are not proposing throwing away the

scintillation light, but in addition using the selected Cherenkov light to

determine the electron direction, i.e., not `limiting the

reconstruction to Cherenkov photons''}\\

CAN THIS ANSWER BE IMPROVED? (YES)

\end{document}

\noteq [Referee 1]

\notea{\bf }\\

### 2014-01-10 Direction_JINST_v5.tex
Added Christophs edits and committing Henrys referee response. --Lindley

--- 2013-10-14 Direction_JINST_v4.tex
@@
-\RequirePackage{lineno}
+%\RequirePackage{lineno}
 
@@
 $\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
-providing energy resolutions of $\sim$5\%. This is roughly a factor of
+providing energy resolutions of $\sim$5\%$/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
 two better than water Cherenkov detectors, the other developed
 technology that can be economically scaled to these large
-masses. However, for scintillator-based detectors, while the energy
-resolution is good due to the abundance of light, the light is
-isotropic and does not retain the directional information of the
-primary particle.  In contrast, the direction of the particle can be
-reconstructed from the Cherenkov cone in water-based detectors,
-although the energy resolution rapidly degrades below $\sim$5~MeV. For
-double-beta decay in particular, but also for neutrino interactions,
-the directional information can be a strong suppressant of
-backgrounds.
+masses.
+However, this energy resolution is much poorer than other technologies being used for $0\nu\beta\beta$ searches: Ge detectors\cite{gerda2013}, Te bolometers, liquid Xe time projection chambers (TPCs)\cite{EXO2012} and high pressure gaseous Xe TPCs\cite{NEXTsipm}.
 
-In a liquid-scintillation-based detector, Cherenkov light is also
-produced, although most is absorbed and re-emitted as part of the
-scintillation processes.  However, some fraction retains its
+Scintillation light is isotropic and any information on the primary particle's direction is lost.  Cherenkov light is also
+produced. Most is absorbed and re-emitted as part of the
+scintillation processes, however some fraction retains its
 directional information. If this directional Cherenkov light can be
 isolated from the copious isotropic scintillation light, it may be
-possible to reconstruct the direction of the primary particle or, in
-the case of double-beta decay, to determine the existence and topology
-of the pair.  The addition of directionality is thus a powerful tool
-for background rejection.  In this paper, we develop a technique for
+possible to reconstruct the direction of the primary particle. The addition of directionality is a powerful tool
+for background rejection especially for $0\nu\beta\beta$ searches.
+In the high pressure TPCs, reduction factors on the order of $10^4$ have been achieved\cite{Gotthard}.
+It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO}.
+The addition of a directional signal would make large-scale liquid scintillator detectors more competitive for the next generation of $0\nu\beta\beta$ searches and merits investigation.
+
+This is the first in a series of papers exploring directionality in large-scale liquid scintillator detectors. In this paper, we develop a technique for
 separating the Cherenkov and scintillation light using the photon arrival
@@
 This is a different technique and application than the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
+We then use these results as input into a traditional direction reconstruction developed for water Cherenkov detectors.
+Since the reconstruction of the direction of $\sim$1~MeV particles has not been achieved before, we will start these studies with the most simple case of a single particle at the center of the detector.
+We will also start with an easier test case of a 5~MeV electron like that from a $^{8}$B solar neutrino interaction.
+With the higher photon statistics at this energy, we will verify the technique and the different detector parameters that effect it.
+We will then lower the energy to that relevant to a single electron in a $0\nu\beta\beta$.
+Since these simpler scenarios are successful, this justifies future work, specifically on the reconstruction of the two simultaneous electrons throughout the detector volume and the study of the sensitivity of a detector instrumented to take advantage a directional signal to both $0\nu\beta\beta$ and to new physics.
 
@@
 \end{equation}
-where $n(\lambda)$ is the wavelength-dependent index of refraction and
-$\beta$ is the velocity of the incoming particle. In a large detector,
-the Cherenkov light produced at wavelengths shorter than the
+where x is the distance travelled by the particle, $n(\lambda)$ is the wavelength-dependent index of refraction and
+$\beta$ is the velocity of the incoming particle.
+
+The Cherenkov light produced at wavelengths shorter than the
 absorption cutoff of the scintillator will be absorbed and re-emitted
@@
 propagate across the detector, retaining their directional
-information. These undisturbed Cherenkov
+information.
+For the example scintillator above, n(400nm) is 1.459\cite{OlegThesis}.
+The absorption cutoff is between 360~nm and 370~nm where the attenuation length rapidly drops from 6.5~m to 0.65~m.
+For a 5~MeV electron, this yields 744 Cherenkov photons per event, and dropping down to a 1~MeV electron yields 89.3 Cherenkov photons per event when integrating from the cutoff wavelength at 360~nm to 550~nm, the point above which a negligible number of photons is produced.
+
+All photons including these undisturbed Cherenkov
 photons will have timing determined by the group 
@@
 \end{equation}
-The longer wavelength Cherenkov photons typically arrive before the
+This means photons right above the scintillation cutoff at 370~nm will travel at 0.1914~m/ns while those at longer wavelengths for example 600~nm travel at 0.2031 m/ns.
+Since on average the undisturbed Cherenkov photons have longer wavelengths, they will arrive before the
 scintillation light, which is slowed by both the scintillation
@@
 to the Q-value of the particular isotope. The individual
-electrons follow distributions of energies and angular correlations, 
-a probable case being equal division of energy between back-to-back
-electrons\cite{phasespace}. This case is shown in figure \ref{detector_view}
+electrons follow distributions of energies and angular correlations which depend on the underlying decay $0\nu\beta\beta$ mechanism\cite{newphysics0nuBB, SuperNEMO}.
+For the basic mechanism of light Majorana neutrino exchange, a probable case is the equal division of energy between back-to-back
+electrons\cite{phasespace, SuperNEMO}. This case is shown in figure \ref{detector_view}
 for an example $^{116}$Cd $0\nu\beta\beta$ event. Since the decay
 half-life is inversely proportional to the phase-space factor, isotopes with
-higher Q-values are preferred and due to backgrounds from the daughters
-of the $^{238}$U and $^{232}$Th decay chains those with Q-values at or
-above 2.6~MeV are most often considered for $0\nu\beta\beta$ searches.
+higher Q-values are preferred. Due to backgrounds from the daughters
+of the $^{238}$U and $^{232}$Th decay chains, isotopes with Q-values above the $^{208}$Tl  2.6~MeV gamma are preferred, however slightly lower Q-values can be tolerated if there is sufficient background rejection.
 There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb},
-but only a handful with large Q-values.
+but only a handful with sufficiently large Q-values.
 
@@
 and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}.
-Variations from the baseline KamLAND case are
-discussed below. Re-emission of absorbed photons in the scintillator
-bulk volume and scattering have not yet been included, but are not
-expected to change the conclusions here.
+This is a standard scintillator.
+The most critical value is the attenuation length at 400~nm which is 25~m.
+The attenuation length drops precipitously at 370~nm from 6.5~m to 0.65~m at the cutoff wavelength at 360~nm.
+Variations from the baseline KamLAND case are discussed below.
+
+Re-emission of absorbed photons in the scintillator
+bulk volume and optical scattering, specifically Rayleigh scattering, have not yet been included by default.
+A test simulation shows that their effects are negligible.
+The reason is that the cutoff is very steep below 360~nm and almost no photons reach the sphere, so optical scattering makes no difference.
+Above about 395~nm, the attenuation length is greater than 20~m so both scattering and absorption are not very likely and scattering is negligible.
+The intermediate region is rather small.
+This is the conclusion of a more detailed study found in Ref.~\cite{Wurm:2010ad}.
 
@@
 
-The studies in sections \ref{sim_section} to \ref{reconstruction_sec} are made with single 5~MeV electrons, lower energies are
-discussed in section \ref{edep_section}. Future studies will work on increasing the complexity
-of the event as is needed for $0\nu\beta\beta$ studies. Three
-effects primarily contribute to the timing of the scintillator detector
-system. First, the simulated travel time of an initial 5~MeV electron is typically between
-0.10 and 0.15~ns, while the travel distance is about 3~cm. Second, the
-scintillation light emission follows a distribution characterized by
-scintillator-specific rise and decay times. Before the solutes in
-liquid scintillator can emit optical photons, the energy has to be
-transferred from the solvent to the solute. The time constant of this
+%The studies in sections \ref{sim_section} to \ref{reconstruction_sec} are made with single 5~MeV electrons, lower energies are
+%discussed in section \ref{edep_section}. Future studies will work on increasing the complexity
+%of the event as is needed for $0\nu\beta\beta$ studies.
+
+Four effects primarily contribute to the timing of the scintillator detector
+system: the travel time of the particle, the time constants of the scintillation process, chromatic dispersion and the timing of the photodetector.
+First, the simulated travel time of an initial 5~MeV electron is 0.108$\pm$0.015~ns.
+This corresponds to an average travel distance of 3.1~cm and a distance from the origin of 2.6~cm.
+We can define an effective Cherenkov production threshold and the time until the electron drops below threshold is 0.106$\pm$0.015~ns.
+We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction, however in the steps before the electron drops below Cherenkov threshold the scattering angle is small and therefore the Cherenkov light encodes the direction of the primary electron.
+The scattering physics is handled by Geant4's ``Multiple Scattering" process which is valid down to 1~keV where atomic shell structure becomes important\cite{geant4scatt}.
+In the energy range important for $0\nu\beta\beta$, a 1.4~MeV electron travels a total distance of 0.8~cm and has a distance from the origin of 0.6~cm in 0.030$\pm$0.004~ns  and takes 0.028$\pm$0.004~ns to drop below Cherenkov threshold.
+The scattering follows the same pattern.
+
+The scintillation light emission follows a distribution characterized by
+scintillator-specific rise and decay times. The first step in the scintillation process is the transfer of energy from the solvent to the solute. The time constant of this
 energy transfer accounts for a rise time in scintillation light
@@
 light traveling faster than blue light.  In order to study the
-measurability of the time differences, we simulated 5~MeV electrons at
+measurability of the time differences due to this chromatic dispersion, we simulated 5~MeV electrons at
 the center of the sphere where we used instantaneous scintillation
 emission with the quantum-efficiency applied, but not including a
-transit-time spread. The true hit time distributions of photoelectrons
+transit-time spread. The higher energy provides larger photon statistics with which to evaluate the method and debug the simulation. The true hit time distributions of photoelectrons
 were analyzed for scintillation light and Cherenkov light
@@
 can be achieved with the large area picosecond photodetectors
-(LAPPDs)\cite{LAPPDSum,LAPPDTDR} and possibly hybrid photodetectors
+(LAPPDs)\cite{Adams:2013nva,RSI_paper,PSEC4_paper,anode_paper} and possibly hybrid photodetectors
 (HPDs)\cite{hpdThesis}; even significantly lower TTS numbers are
@@
 to the problem of reconstructing the position and direction of 5~MeV
-electrons from this early light.  WCSimAnalysis is a water Cherenkov
+electrons from this early light.
+To start, we are neglecting the effects of position dependence.
+In future work the arrival times can be corrected by the time of flight from the reconstructed vertex and the position and reconstruction fitted simultaneously.
+The vertex reconstruction uncertainty leads to an additional smearing of the arrival time distribution of 0.15~ns for every 3~cm of position reconstruction uncertainty.
+
+WCSimAnalysis is a water Cherenkov
 reconstruction package developed for the Long Baseline Neutrino
@@
 The reconstruction algorithms outlined in section \ref{reconstruction_sec} have also been applied
-to the simulations at lower energies. Figure \ref{low_energy_reco} shows the results for 
-the lowest simulated electron energy, 1.4~MeV. Most events are still reconstructed well, despite the lower
+to the simulations at lower energies.
+Figure \ref{fig:reco} (Middle) shows the results for 2.1~MeV and Figure \ref{fig:reco} (Bottom) shows the results for 1.4~MeV.
+Most events are still reconstructed well, despite the lower
 number of PEs and the decreased $R_{C/S}$. For 1.4~MeV electrons, the distribution RMS values for all three

### 2014-01-10 Direction_JINST_v8.tex
Added Christophs edits and committing Henrys referee response. --Lindley

--- 2014-01-08 Direction_JINST_v7.tex
@@
 
-The scintillation light emission follows a distribution characterized by
-scintillator-specific rise and decay times. The first step in the scintillation process is the transfer of energy from the solvent to the solute. The time constant of this
+The scintillator-specific rise and decay times are the second effect that determines the timing in a scintillator detector.
+The first step in the scintillation process is the transfer of energy from the solvent to the solute.
+The time constant of this
 energy transfer accounts for a rise time in scintillation light
@@
 
-Due to the wavelength-dependence of the refractive index the speed of
+Chromatic dispersion is the third effect that determines the timing in a scintillator detector. Due to the wavelength-dependence of the refractive index the speed of
 light in the scintillator (see Equation (\ref{eqGroup})) increases
 with increasing photon wavelengths for normal dispersion, with red
-light traveling faster than blue light.  In order to study the
-measurability of the time differences due to this chromatic dispersion, we simulated 5~MeV electrons at
+light traveling faster than blue light.  In order to study the time differences due to this chromatic dispersion, we used a simplified simulation of 5~MeV electrons at
 the center of the sphere where we used instantaneous scintillation
@@
 transit-time spread.
-The higher energy provides larger photon statistics with which to evaluate the method, debug the simulation, and is of interest to neutrino-electron scattering experiments.
+The higher energy 5~MeV electron provides larger photon statistics with which to evaluate the method, debug the simulation, and is of interest to neutrino-electron scattering experiments.
 The true hit time distributions of photoelectrons
@@
 
-The measurement of the arrival times of single photoelectrons is
+The fourth effect determining the timing in a scintillator detector is the timing of the photodetectors. The measurement of the arrival times of single photoelectrons is
 affected by the transit-time spread (TTS) of the photodetectors, a
@@
 \section{Conclusions}
-The ability to reconstruct direction in kiloton-scale scintillation
-detectors would be a major technological advance for neutrino
-experiments, especially those also
-searching for neutrino-less double-beta decay. More generally, this technique could
-be applied wherever scintillation-based detectors are used. A \GEANT~simulation of a simple spherical detector corresponding to a kiloton
-of scintillator shows that with timing on the order of 0.1~ns the directional Cherenkov light can be separated from the more abundant
-scintillation light. The separation can be improved using
-photodetectors with more red sensitivity and liquid scintillators with
-a more narrow emission spectrum shifted to shorter
-wavelengths. Furthermore, simple reconstruction algorithms adapted
-from those for water Cherenkov detectors are able to reliably
-reconstruct the position and direction of electrons with energies in the few MeV range.
-More detailed simulation and advanced reconstruction algorithms will need
-to be developed to move to more complicated event structure, such as those in neutrino-less double-beta decay. The technique already
-appears promising and justifies future work on the topic.
+
+We developed a technique to separate scintillation and Cherenkov light to reconstruct direction of electrons with energies 5~MeV, 2.1~MeV and 1.4~MeV in a liquid scintillator detector.
+These energies were chosen to represent typical neutrino-electron elastic scattering energies and
+$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.2~MeV
+
+While scintillation light is isotropic, Cherenkov light with wavelengths above $\sim$370~nm carries the information about the direction of the electrons.
+All light shorter than $\sim$370~nm get absorbed and re-emitted isotropicaly as a part of the scintillation process.
+On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite rise time of the scintillation.
+Therefore early light contains directional information.
+
+Using a GEANT~simulation of a spherical detector with radius of 6.5~m, we showed that for electrons originated at the center, Cherenkov light arrives on average $\sim$1~ns before the scintillation light.
+Chromatic dispersion alone accounts for $\sim$0.5~ns delay between scintillation and Cherenkov light.
+We then defined an early time window and found for a 5~MeV electron that the ratio of Cherenkov to scintillation light is 0.63 for photodetectors with TTS of 0.1~ns.
+This was degraded by a factor of 2.5 if current photodetectors with $\sim$1~ns resolution were used.
+This can be improved by 1.6 or 1.4  if more red sensitive photodetectors or scintillators with narrower emission spectra are used
+
+We applied reconstruction algorithms from water Cherenkov detectors. QUANTATIVE STATEMENT ABOUT VERTEX DIRECTION PERFORMANCE.
+
+We find the technique promising and plan to continue work on the topic.
+The ability to reconstruct direction in kiloton-scale scintillation detectors would expand capabilities of neutrino experiments, especially those also searching for neutrino-less double-beta decay.
+More generally, this technique could be applied wherever scintillation-based detectors are used.
+
+%This work should include a more detailed simulation including second order effects such as reemission and a more detailed photodetector model including electronics thresholds.
+More advanced reconstruction methods are also needed to simultaneously fit position and direction and to handle the more complicated event structure found in neutrino-less double-beta decay.
+
+%The separation can be improved using photodetectors with more red sensitivity and liquid scintillators with a more narrow emission spectrum shifted to shorter wavelengths.
+Furthermore, simple reconstruction algorithms adapted from those for water Cherenkov detectors are able to reliably reconstruct the position and direction of electrons with energies in the few MeV range.
+More detailed simulation and advanced reconstruction algorithms will need to be developed to move to more complicated event structure, such as those in neutrino-less double-beta decay.
 

### 2014-01-10 Direction_JINST_v9.tex
Added Christophs edits and committing Henrys referee response. --Lindley

--- 2014-01-10 Direction_JINST_v8.tex
@@
 for background rejection, especially for $0\nu\beta\beta$ searches.
-In the high pressure TPCs, reduction factors on the order of $10^4$ have been achieved\cite{Gotthard}.
+In the high pressure TPCs, reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}.
 It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO}.
@@
 This is a standard scintillator.
-The most critical value is the attenuation length at 400~nm which is 25~m.
+The attenuation length at 400~nm, the position of the peak standard bialkali photocathod efficiency, is 25~m.
 The attenuation length drops precipitously at 370~nm from 6.5~m to 0.65~m at the cutoff wavelength at 360~nm.
@@
         \includegraphics[scale=0.4]{graphs2/PEtime_Cerenkov_scattered_vs_unscattered.pdf}
-        \caption[]{Photoelectron (PE) arrival times after application
+        \caption[]{Cherenkov photoelectron (PE) arrival times after application
         of the transit-time spread (TTS) for the simulation
@@
 The time until the electron drops below Cherenkov threshold is 0.106$\pm$0.015~ns.
-We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction; however before the electron drops below Cherenkov threshold the scattering angle is small and therefore the Cherenkov light encodes the direction of the primary electron.
+We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction; however the scattering angle is small while the majority of Cherenkov light is produced.
+The Cherenkov light thus encodes the direction of the primary electron.
 The scattering physics is handled by Geant4's ``Multiple Scattering" process which is valid down to 1~keV, where atomic shell structure becomes important\cite{geant4scatt}.
@@
         photocathode is taken into account. (Right) The reconstructed
-        vertex position, $(x,y,z)$, for the same simulation. From Top to Bottom, 5~MeV, 2.1~MeV and 1.4 Mev are shown.}
+        vertex position, $(x,y,z)$, for the same simulation. From Top to Bottom, 5~MeV, 2.1~MeV and 1.4 MeV are shown.}
 \end{center}
@@
 
-We developed a technique to separate scintillation and Cherenkov light to reconstruct direction of electrons with energies 5~MeV, 2.1~MeV and 1.4~MeV in a liquid scintillator detector.
-These energies were chosen to represent typical neutrino-electron elastic scattering energies and
-$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.2~MeV
+We have developed a technique to separate scintillation and Cherenkov light to reconstruct direction of electrons with energies 5~MeV, 2.1~MeV and 1.4~MeV in a liquid scintillator detector.
+These energies have been chosen to represent typical neutrino-electron elastic scattering energies and
+$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.188~MeV
 
-While scintillation light is isotropic, Cherenkov light with wavelengths above $\sim$370~nm carries the information about the direction of the electrons.
-All light shorter than $\sim$370~nm get absorbed and re-emitted isotropicaly as a part of the scintillation process.
-On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite rise time of the scintillation.
-Therefore early light contains directional information.
+While scintillation light is isotropic, Cherenkov light with wavelengths above the absorption cutoff $\sim$370~nm carries the information about the direction of the electrons.
+All light shorter than this gets absorbed and re-emitted isotropically as a part of the scintillation process.
+On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and and the finite time of the scintillation processes, therefore early light contains directional information.
 
-Using a GEANT~simulation of a spherical detector with radius of 6.5~m, we showed that for electrons originated at the center, Cherenkov light arrives on average $\sim$1~ns before the scintillation light.
-Chromatic dispersion alone accounts for $\sim$0.5~ns delay between scintillation and Cherenkov light.
-We then defined an early time window and found for a 5~MeV electron that the ratio of Cherenkov to scintillation light is 0.63 for photodetectors with TTS of 0.1~ns.
-This was degraded by a factor of 2.5 if current photodetectors with $\sim$1~ns resolution were used.
-This can be improved by 1.6 or 1.4  if more red sensitive photodetectors or scintillators with narrower emission spectra are used
+Using a Geant4~simulation of a spherical detector with radius of 6.5~m and photodetectors with TTS of 0.1~ns, we have showed that for electrons originated at the center a time cut on the early light was effective at isolating the directional light, improving the ratio of Cherenkov light to scintillation light from 0.02 to 0.63.
+This ratio is degraded by a factor of 2.5 if current photodetectors with $\sim$1~ns resolution were used.
+This can be improved by factors of 1.6 or 1.4 if more red sensitive photodetectors or scintillators with narrower emission spectra are used.
 
-We applied reconstruction algorithms from water Cherenkov detectors. QUANTATIVE STATEMENT ABOUT VERTEX DIRECTION PERFORMANCE.
+Reconstruction algorithms developed for water Cherenkov detectors have been applied to this early light.
+The algorithms are able to converge on reasonable reconstructed vertices and directions for all simulated energies: 5~MeV, 2.1~MeV, 1.4~MeV.
+As expected, the reduction in photon statistics with lower energies leads to a broadening of the reconstructed vertex by 1.5~cm.
+Similarly, the mean reconstructed direction drops from XX to XX as the $\cos(\theta)$ distribution broadens.
 
-We find the technique promising and plan to continue work on the topic.
+We have shown that this technique promising, and we plan to continue work on the topic.
 The ability to reconstruct direction in kiloton-scale scintillation detectors would expand capabilities of neutrino experiments, especially those also searching for neutrino-less double-beta decay.

### 2014-01-14 Direction_JINST_v10.tex
the last few versions. --Lindley

--- 2014-01-10 Direction_JINST_v9.tex
@@
 two better than water Cherenkov detectors, the other developed
-technology that can be economically scaled to these large
+technology for neutrino detectors that can be economically scaled to these large
 masses.
@@
 In the high pressure TPCs, reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}.
-It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO}.
+It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO,NEXTsipm}.
 The addition of a directional signal would make large-scale liquid scintillator detectors more competitive for the next generation of $0\nu\beta\beta$ searches and merits investigation.
@@
 The index of refraction at  n(370nm)=1.466 and can be translated into an effective Cherenkov threshold for electrons at 0.188~MeV.
-For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 360~nm to 550~nm,  the wavelength above which a small number of photons is detected due to a standard bialkali photocathode's quantum efficiency.
-Lowering the energy to 1~MeV yields 82 Cherenkov photons between 360-550~nm.
+For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 370~nm to 550~nm,  the wavelength above which a small number of photons is detected due to a standard bialkali photocathode's quantum efficiency.
+Lowering the energy to 1~MeV yields 82 Cherenkov photons between 370-550~nm.
 
@@
 In order to study the effects relevant to directional reconstruction
-in liquid scintillators, a \GEANT \newline 
-\cite{geant4one,geant4two} simulation
-has been constructed. The simulation uses \GEANT~version 4.9.6 with the default liquid scintillator
+in liquid scintillators, a Geant4\cite{geant4one,geant4two} simulation
+has been constructed. The simulation uses Geant4~version 4.9.6 with the default liquid scintillator
 optical model, in which optical photons are
@@
 This is a standard scintillator.
-The attenuation length at 400~nm, the position of the peak standard bialkali photocathod efficiency, is 25~m.
-The attenuation length drops precipitously at 370~nm from 6.5~m to 0.65~m at the cutoff wavelength at 360~nm.
+The attenuation length at 400~nm, the position of the peak standard bialkali photocathode efficiency, is 25~m.
+The attenuation length drops precipitously between 370~nm and 360~nm from 6.5~m to 0.65~m.
+We use this drop to define the cutoff wavelength at 370~nm.
 Variations from the baseline KamLAND case are discussed below.
@@
 Monte Carlo simulation. We note that the KamLAND 17-inch PMTs use the
-same photocathode type with similar quantum efficiency.
+same photocathode type with similar quantum efficiency. We are neglecting any threshold effects in the photodetector readout electronics.
 
@@
 scintillation light by using improved detector timing. The primary
-quantities provided by the \GEANT~simulation are the photoelectron hit
+quantities provided by the Geant4~simulation are the photoelectron hit
 positions and the detection times after the TTS resolution has been
@@
 of the events the initial electron direction is reconstructed well.
-This is a promising result given the simplicity of the algorithms
-used.
+This is a promising result given the simplicity of the algorithms.
 
@@
 These energies have been chosen to represent typical neutrino-electron elastic scattering energies and
-$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.188~MeV
+$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.188~MeV.
 
@@
 All light shorter than this gets absorbed and re-emitted isotropically as a part of the scintillation process.
-On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and and the finite time of the scintillation processes, therefore early light contains directional information.
+On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite time of the scintillation processes, therefore early light contains directional information.
 
@@
 The algorithms are able to converge on reasonable reconstructed vertices and directions for all simulated energies: 5~MeV, 2.1~MeV, 1.4~MeV.
-As expected, the reduction in photon statistics with lower energies leads to a broadening of the reconstructed vertex by 1.5~cm.
-Similarly, the mean reconstructed direction drops from XX to XX as the $\cos(\theta)$ distribution broadens.
+As expected, the reduction in photon statistics with lower energies leads to a broadening of the reconstructed vertex by $\sim$1~cm.
+Similarly, the mean reconstructed direction drops from 0.88 to 0.74 as the $\cos(\theta)$ distribution broadens.
 
-We have shown that this technique promising, and we plan to continue work on the topic.
+We have shown that this technique is promising, and we plan to continue work on the topic.
 The ability to reconstruct direction in kiloton-scale scintillation detectors would expand capabilities of neutrino experiments, especially those also searching for neutrino-less double-beta decay.

### 2014-01-14 Direction_JINST_v11.tex
the last few versions. --Lindley

--- 2014-01-14 Direction_JINST_v10.tex
@@
 providing energy resolutions of $\sigma(E)=\sim$5 \% $/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
-two better than water Cherenkov detectors, the other developed
+two better than for water Cherenkov detectors, the other developed
 technology for neutrino detectors that can be economically scaled to these large
@@
 for background rejection, especially for $0\nu\beta\beta$ searches.
-In the high pressure TPCs, reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}.
+In high pressure TPCs, reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}.
 It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO,NEXTsipm}.
-The addition of a directional signal would make large-scale liquid scintillator detectors more competitive for the next generation of $0\nu\beta\beta$ searches and merits investigation.
+The addition of a directional signal would make large-scale liquid scintillator detectors more competitive for the next generation of $0\nu\beta\beta$ searches.
+%and merits investigation.
 
@@
 scintillating neutrino detectors.
-This is a different from the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
+This is different from the direction reconstruction described for high-energy neutrino interactions\cite{john} or that for neutrons from inverse beta decay\cite{chooz,dcDirection}.
 We then use these results as input into a traditional direction reconstruction developed for water Cherenkov detectors.
-Since the reconstruction of the direction of $\sim$1~MeV particles has not been achieved before, we start these studies with the most simple case of a single particle at the center of the detector.
+Since the reconstruction of the direction of $\sim$1~MeV particles has not been achieved before, we start these studies with the %most
+simple case of a single particle at the center of the detector.
 We also start with an easier test case of a 5~MeV electron like that from a $^{8}$B solar neutrino interaction.
@@
 propagate across the detector, retaining their directional
-information.
-The absorption cutoff for the example scintillator above is 370~nm.
-The index of refraction at  n(370nm)=1.466 and can be translated into an effective Cherenkov threshold for electrons at 0.188~MeV.
+information. The absorption cutoff for the example scintillator above is 370~nm. The index of refraction at $\lambda$=370~nm is n=1.466, which translates
+%n(370nm)=1.466 and can be translated 
+into an effective Cherenkov threshold for electrons at 0.188~MeV.
 For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 370~nm to 550~nm,  the wavelength above which a small number of photons is detected due to a standard bialkali photocathode's quantum efficiency.
@@
 In neutrino-electron scattering events, a single electron emerges with a distribution of energies with a maximum energy related to the incoming neutrino's energy.
-In comparison, $0\nu\beta\beta$ events are more complicated with two electron emerge with a combined energy equal
+In comparison, $0\nu\beta\beta$ events are more complicated, with two electrons emerging with a combined energy equal
 to the Q-value of the particular isotope. The individual
@@
 High Q-value candidates are preferred for $0\nu\beta\beta$  for two reasons.
-First, the measured half-life is inversely proportional to the phase-space factor, therefore the measured rate is expected to be higher from isotopes with higher Q-values.
+First, as the measured half-life is inversely proportional to the phase-space factor, the measured rate is expected to be higher from isotopes with higher Q-values.
 Second, the main background for these experiments come from the daughters of the $^{238}$U and $^{232}$Th decay chains.
 The Compton shoulder from gamma-rays is particularly problematic.
-The 2.6~MeV gamma-ray from $^{208}$Tl decay is the highest energy gamma from these decay chains and therefore isotopes above 2.6~MeV are preferred.
+The 2.6~MeV gamma-ray from $^{208}$Tl decay is the highest energy gamma from these decay chains and so isotopes above 2.6~MeV are preferred.
 Experiments using isotopes with Q-values below this energy must compensate with improved background rejection techniques.
@@
 scintillator with a long attenuation length in the wavelength region of
-interest. Recently nanocrystals formed by candidate isotopes have been explored as an alternative to doping by single atoms,\cite{qdot,qdot2}.
+interest. Recently nanocrystals formed by candidate isotopes have been explored as an alternative to doping by single atoms\cite{qdot,qdot2}.
 
@@
 $\tau_{d2}$ = 8.8~ns with relative weights of 0.87 and 
-0.13)\cite{tajimaThesis}, scintillator light yield (9030 photons/MeV)
+0.13)\cite{tajimaThesis}, scintillator light yield (9030 photons/MeV),
 and the Birks constant ($kB$ $\approx$ 0.1~mm/MeV)\cite{ChrisThesis}.
@@
         of 1000 electrons (5~MeV).
-The default simulation is shown with optical scattering turned on: all photons (black), un-scattered photons (red dotted), and scattered (blue dashed).
+The default simulation is shown with optical scattering turned on: all photons (black solid), un-scattered photons (red dotted), and scattered (blue dashed).
 Scattering causes a very small tail at longer times as expected.
@@
 However, advances in the production of commercial quantum dot samples
-could yield quantum dots which have similar, single peak emission
+could yield quantum dots that have similar, single peak emission
 shapes at shorter wavelengths. This case has been simulated using the
@@
         x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0), and originate
-        from the center of the 6.5m-radius detector, $\vec{r}$ =
+        at the center of the 6.5m-radius detector, $\vec{r}$ =
         (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
@@
 This would change the ideal time cut by 0.15~ns  for every 3~cm of position reconstruction uncertainty.
-We have studied other time cuts from 33~ns to 34.5~ns and the reconstruction remained reasonable.
+We have studied other time cuts from 33~ns to 34.5~ns and the reconstruction remains reasonable.
 
@@
 scintillation-induced PEs, as well as their ratio $R_{C/S}$. The mean number
-of PEs from Cherenkov (scintillation) light for electron energies 1.4(2.1)~MeV is 21.8 (52.1), 38.4 (76.8)
+of PEs from Cherenkov (scintillation) light for electron energies of 1.4, 2.1, and 5~MeV is 21.8 (52.1), 38.4 (76.8)
 and 108 (171) for 1.4~MeV, 2.1~MeV and 5~MeV, respectively. This gives the
@@
 These energies have been chosen to represent typical neutrino-electron elastic scattering energies and
-$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.188~MeV.
+$0\nu\beta\beta$ energies. The Cherenkov threshold for an electron in a typical liquid scintillator is $\sim$0.2~MeV.
 
@@
 All light shorter than this gets absorbed and re-emitted isotropically as a part of the scintillation process.
-On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite time of the scintillation processes, therefore early light contains directional information.
+On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite time of the scintillation processes; the early light thus contains directional information.
 
-Using a Geant4~simulation of a spherical detector with radius of 6.5~m and photodetectors with TTS of 0.1~ns, we have showed that for electrons originated at the center a time cut on the early light was effective at isolating the directional light, improving the ratio of Cherenkov light to scintillation light from 0.02 to 0.63.
+Using a Geant4~simulation of a spherical detector with radius of 6.5~m and photodetectors with TTS of 0.1~ns, we have shown that for electrons originated at the center a time cut on the early light is effective at isolating the directional light, improving the ratio of Cherenkov light to scintillation light from 0.02 to 0.63.
 This ratio is degraded by a factor of 2.5 if current photodetectors with $\sim$1~ns resolution were used.
-This can be improved by factors of 1.6 or 1.4 if more red sensitive photodetectors or scintillators with narrower emission spectra are used.
+This can be improved by factors of 1.6 or 1.4 if more red-sensitive photodetectors or scintillators with narrower emission spectra are used.
 
@@
 As expected, the reduction in photon statistics with lower energies leads to a broadening of the reconstructed vertex by $\sim$1~cm.
-Similarly, the mean reconstructed direction drops from 0.88 to 0.74 as the $\cos(\theta)$ distribution broadens.
+Similarly, the mean $\cos(\theta)$ between the reconstructed and initial directions drops from 0.88 to 0.74.
+% reconstructed direction drops from 0.88 to 0.74 as the $\cos(\theta)$ distribution broadens.
 
-We have shown that this technique is promising, and we plan to continue work on the topic.
+This technique is promising, and we plan to continue work on the topic.
 The ability to reconstruct direction in kiloton-scale scintillation detectors would expand capabilities of neutrino experiments, especially those also searching for neutrino-less double-beta decay.

### 2014-04-21 Direction_18Jul.tex
Add old versions of paper. --Lindley

--- 2013-07-18 Direction.tex
@@
 \section{Conclusion}
-This is going to work.
+The ability to reconstruct direction in kiloton-scale scintillating detectors would be a major technological advance.
+It would find immediate application in neutrino experiments especially those searching for $0\nu\beta\beta$.
+More generally, this technique could be applied wherever scintillation-based detectors are used.
+A Geant4 simulation of a simple spherical detector corresponding to a kiloton of scintillator shows that  timing on the order of  0.1~ns is required to separate the directional Cerenkov light from the more abundant scintillation light.
+This separation can be improved using photodetectors with more red sensitivity and liquid scintillators with a more narrow emission spectrum shifted to shorter wavelengths.
+Furthermore, simple reconstruction algorithms adapted from those for water Cerenkov detectors are able to reliable reconstruct the position and direction of 5~MeV electrons.
+More detailed simulation and advanced reconstruction algorithms will need to be developed to move to lower energies and more complicated event topography, like those in $0\nu\beta\beta$, but the technique already appears very promising.
 

### 2014-04-21 Direction_JINST_v11.1.tex
Add old versions of paper. --Lindley

--- 2014-01-14 Direction_JINST_v11.tex
@@
 into an effective Cherenkov threshold for electrons at 0.188~MeV.
-For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 370~nm to 550~nm,  the wavelength above which a small number of photons is detected due to a standard bialkali photocathode's quantum efficiency.
+For a 5~MeV electron, this yields 685 Cherenkov photons per event when integrating from the cutoff wavelength at 370~nm to 550~nm,  the wavelength above which a small number of photons is detected due to the quantum efficiency of a standard bialkali photocathode.
 Lowering the energy to 1~MeV yields 82 Cherenkov photons between 370-550~nm.
@@
 scintillation-induced PEs, as well as their ratio $R_{C/S}$. The mean number
-of PEs from Cherenkov (scintillation) light for electron energies of 1.4, 2.1, and 5~MeV is 21.8 (52.1), 38.4 (76.8)
-and 108 (171) for 1.4~MeV, 2.1~MeV and 5~MeV, respectively. This gives the
+of PEs from Cherenkov (scintillation) light is 21.8 (52.1), 38.4 (76.8)
+and 108 (171) for electron energies of 1.4~MeV, 2.1~MeV and 5~MeV, respectively. This gives the
 ratios $R_{C/S}$ = 0.42, 0.50 and 0.63: The decrease in Cherenkov-induced

### 2014-04-21 Direction_JINST_v12.tex
Add old versions of paper. --Lindley

--- 2014-01-14 Direction_JINST_v11.tex
@@
 especially for searches for neutrino-less double-beta decay. In this
-paper, we develop a technique for extracting particle direction using
+paper, we propose a technique for extracting particle direction using
 the difference in arrival times for Cherenkov and scintillation light,
@@
 for background rejection, especially for $0\nu\beta\beta$ searches.
-In high pressure TPCs, reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}.
+High pressure TPCs are another technology that can record the direction of particles.
+In this type of experiment reduction factors on the order of $\sim10^4$ have been achieved\cite{Gotthard}.
 It is also possible to look for new physics in the angular correlation of the emerging electrons\cite{newphysics0nuBB}, as has been proposed for the tracking-based detectors\cite{SuperNEMO,NEXTsipm}.
@@
 
+\begin{figure}
+        \begin{center}
+        \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
+        \caption[]{The detector geometry and coordinate system.
+        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.4~MeV each
+        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay). The electrons originate at
+        the center of the sphere with initial directions along the x
+        and -x-axis. Only Cherenkov photons are drawn to illustrate the
+        directionality of the event. \label{detector_view}}
+        \end{center}
+\end{figure}
+
 All photons including these undisturbed Cherenkov
@@
 Figure 1 shows a simulated $^{116}$Cd $0\nu\beta\beta$ event in a model with light Majorana neutrino exchange, for which a probable case is the emission of two electrons with comparable energies at a large angle relative to each other.
+In the $0\nu\beta\beta$ decay of $^{116}$Cd, 52.1\% of the electrons have energies above 1.4~MeV and 97.4\% have energies above the Cherenkov threshold of 0.2~MeV\cite{phasespace}.
 
@@
 
-\begin{figure}
-        \begin{center}
-        \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
-        \caption[]{The detector geometry and coordinate system.
-        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.4~MeV each
-        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay). The electrons originate at
-        the center of the sphere with initial directions along the x
-        and -x-axis. Only Cherenkov photons are drawn to illustrate the
-        directionality of the event. \label{detector_view}}
-        \end{center}
-\end{figure}
-
 \section{Geant4 simulation}
@@
 In order to study the effects relevant to directional reconstruction
-in liquid scintillators, a Geant4\cite{geant4one,geant4two} simulation
+in liquid scintillators, a Geant4\cite{geant4one,geant4two} Monte Carlo (MC)  simulation
 has been constructed. The simulation uses Geant4~version 4.9.6 with the default liquid scintillator
@@
 We use this drop to define the cutoff wavelength at 370~nm.
+The baseline simulation is summarized in table~\ref{summary_table}.
 Variations from the baseline KamLAND case are discussed below.
+
+\begin{table}
+\caption{The basic parameters of the default simulation.
+The photodetector transit time spread (TTS), photodetector peak sensitivity, and scintillator emission spectrum are varied as part of the study.
+}
+  \begin{center}
+    \begin{tabular}{ll}
+      Parameter & Value\\
+      \hline\hline\\[-5px]
+      Radius & 6.5~m \\
+      Scintillator density & 0.78 g/ml \\
+      Scintillator rise time & 1.0~ns\\
+      Scintillator decay time & 6.9~ns and 8.8~ns \\
+      Scintillator light yield & 9030 photons/MeV \\
+      Scintillator peak emission & 400~nm \\
+      Photodetector TTS & 0.1~ns \\
+      Photodetector peak sensitivity & 400~nm \\
+      \hline
+    \label{summary_table}
+    \end{tabular}
+  \end{center}
+\end{table}
 
@@
 A similar argument holds for re-emission.
+These are second order effects that will be studied in future work on uniformly distributed and lower energy events.
 Scattering length measurements and discussions can be found in Ref.~\cite{Wurm:2010ad}.
@@
         \caption[]{Cherenkov photoelectron (PE) arrival times after application
-        of the transit-time spread (TTS) for the simulation
+        of the photodetector transit-time spread (TTS) for the simulation
         of 1000 electrons (5~MeV).
@@
 as fully absorbing (no reflections), with a photodetector coverage of
-100\%. As in the case of optical scattering, reflections at the sphere are a small effect that would create a small tail at longer times. Two important photodetector properties have been varied: 1)
+100\%.
+A uniform reduction in photodetector coverage would lower the number of measured photons.
+As in the case of optical scattering, reflections at the sphere are a small effect that would create a small tail at longer times.
+Two important photodetector properties have been varied: 1)
 the transit-time spread (TTS, default $\sigma$ = 0.1~ns) and 2) the
@@
 R7081 PMT)\cite{Hamamatsu_R7081}. The QE values as a function of wavelength come from the Double Chooz\cite{dctwo}
-Monte Carlo simulation. We note that the KamLAND 17-inch PMTs use the
+MC simulation. We note that the KamLAND 17-inch PMTs use the
 same photocathode type with similar quantum efficiency. We are neglecting any threshold effects in the photodetector readout electronics.
-
-\begin{figure}
-\begin{center}
-        \subfigure[ ~Default
-        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
-        \subfigure[ ~Increased TTS
-        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
-        \subfigure[ ~Red-sensitive
-        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
-        \caption[]{Photoelectron (PE) arrival times after application
-        of the transit-time spread (TTS) for the simulation
-        of 1000 electrons (5~MeV) with different values of the TTS and
-        wavelength response. PEs from
-        Cherenkov light (black, solid line) and scintillation light
-        (red, dotted line) are
-        compared. The dash-dotted vertical line illustrates a time cut at
-        34.0~ns. (a) Default simulation: bialkali photocathode and TTS =
-        0.1~ns ($\sigma$). After the 34.0~ns time cut, 171~PEs
-        from scintillation and 108~PEs from Cherenkov light are detected. (b)
-        Default simulation settings except for TTS = 1.28~ns (KamLAND
-        17-inch PMTs). After the 34.0~ns time cut, 349~PEs from
-        scintillation and 88~PEs from Cherenkov light are detected. (c) Default
-        simulation settings except for a GaAsP photocathode. After the
-        34.0~ns time cut, 226~PEs from scintillation and 229~EPS
-        from Cherenkov light are detected. \label{time_plots_comparison}}
-        \end{center}
-\end{figure}
 
@@
 The time until the electron drops below Cherenkov threshold is 0.106$\pm$0.015~ns.
-We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction; however the scattering angle is small while the majority of Cherenkov light is produced.
+We note that due to Coulomb scattering the final direction of the electron before it stops does not correspond to the initial direction; however the scattering angle is small while the majority of Cherenkov light is produced.
 The Cherenkov light thus encodes the direction of the primary electron.
 The scattering physics is handled by Geant4's ``Multiple Scattering" process which is valid down to 1~keV, where atomic shell structure becomes important\cite{geant4scatt}.
-In the energy range important for $0\nu\beta\beta$, a 1.4~MeV electron travels a total path length of 0.8~cm, has a distance from the origin of 0.6~cm in 0.030$\pm$0.004~ns  and takes 0.028$\pm$0.004~ns to drop below Cherenkov threshold.
+In the energy range important for $0\nu\beta\beta$, a 1.4~MeV electron travels a total path length of 0.8~cm, travels a distance from the origin of 0.6~cm in 0.030$\pm$0.004~ns  and takes 0.028$\pm$0.004~ns to drop below Cherenkov threshold.
 The scattering follows the same pattern.
@@
 quantities provided by the Geant4~simulation are the photoelectron hit
-positions and the detection times after the TTS resolution has been
-applied. In section~\ref{reconstruction_sec} these quantities are then
+positions and the detection times. The hit times are then smeared according to the TTS resolution. In section~\ref{reconstruction_sec} these quantities are then
 used for event reconstruction.
@@
 
-\section{Detector timing}
+\section{Instrumentation Requirements}
+\subsection{Photodetector timing}
 \label{detector_timing_sec}
@@
 
+\begin{figure}
+\begin{center}
+        \subfigure[ ~Default
+        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        \subfigure[ ~Increased TTS
+        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        \subfigure[ ~Red-sensitive
+        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        \caption[]{Photoelectron (PE) arrival times after application
+        of the photodetector transit-time spread (TTS) for the simulation
+        of 1000 single electrons (5~MeV) moving along the x-axis with different values of the TTS and
+        wavelength response. PEs from
+        Cherenkov light (black, solid line) and scintillation light
+        (red, dotted line) are
+        compared. The dash-dotted vertical line illustrates a time cut at
+        34.0~ns. (a) Default simulation: bialkali photocathode and TTS =
+        0.1~ns ($\sigma$). After the 34.0~ns time cut, 171~PEs
+        from scintillation and 108~PEs from Cherenkov light are detected. (b)
+        Default simulation settings except for TTS = 1.28~ns (KamLAND
+        17-inch PMTs). After the 34.0~ns time cut, 349~PEs from
+        scintillation and 88~PEs from Cherenkov light are detected. (c) Default
+        simulation settings except for a GaAsP photocathode. After the
+        34.0~ns time cut, 226~PEs from scintillation and 229~PEs
+        from Cherenkov light are detected. \label{time_plots_comparison}}
+        \end{center}
+\end{figure}
+
 In order to compare simulations with different parameters, a fixed
-time cut of $t\leq$ 34.0~ns is applied using the truth information to
+time cut of $t\leq$ 34.0~ns is applied using the MC truth information to
 isolate the Cherenkov light in this early time window, as shown in figure \ref{time_plots_comparison} (a). For the default
@@
 erased by scattering of the initial electrons even at 1.4~MeV.
+At this lower energy, photon statistics are reduced, making reconstruction more difficult as we move to lower energies that will occur in $0\nu\beta\beta$ and neutrino scattering events.
 
@@
 
-\section{Detector wavelength response}
+\subsection{Photodetector wavelength response}
 \label{detector_wavelength_response_sec} 
@@
 
-\section{Scintillator emission spectrum}
+\subsection{Scintillator emission spectrum}
 \label{scintillator_emission_sec}
@@
         %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
-        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_5p0.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq3_5p0.pdf}}
         \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_5p0.pdf}}
-        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_2p1.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq3_2p1.pdf}}
         \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_2p1.pdf}}
-        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq2_1p4.pdf}}
+        \subfigure{\includegraphics[scale=0.375]{graphs2/hDir_pq3_1p4.pdf}}
         \subfigure{\includegraphics[scale=0.375]{graphs2/hVtx_pq2_1p4.pdf}}
@@
 
-\section{Reconstruction}
+\section{Event Reconstruction}
 \label{reconstruction_sec}
 
+\subsection{Description of algorithm}
 The timing studies show that in the early time window, $t\leq$
@@
 
+\subsection{Reconstruction Results}
+
 For the purpose of testing the reconstruction algorithm we use 1000
@@
 The electrons are simulated at the center of the detector, $\vec{r}$ = (0,0,0), along
-the x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0). Figure \ref{fig:reco} (Top)
+the x-axis, $\vec{p}/|\vec{p}|$ = (1,0,0). Figure \ref{fig:reco} (Left)
 shows the vertex reconstruction. The vertex is reasonably well-reconstructed around the center of the detector, $\vec{r}$ = (0,0,0),
@@
 as coming from a common point. The reconstruction of the direction
-also is shown in figure \ref{fig:reco} (Top). It shows that for the majority
+also is shown in figure \ref{fig:reco} (Right). It shows that for the majority
 of the events the initial electron direction is reconstructed well.
@@
 
-\section{Energy dependence}
+\subsection{Energy dependence}
 \label{edep_section}
@@
 
-\begin{figure}
-        \begin{center}
-        \includegraphics[scale=0.4]{graphs/hCos_vs_E_final.pdf}
-        \caption[]{Mean cosine of the angle between the true initial electron direction and the reconstructed direction, as a function of the electron energy.
+%\begin{figure}
+%        \begin{center}
+%        \includegraphics[scale=0.4]{graphs/hCos_vs_E_final.pdf}
+%        \caption[]{Mean cosine of the angle between the true initial electron direction and the reconstructed direction, as a function of the electron energy.
 For each energy 1000 events have been simulated.
 Statistical errors are
-shown. \label{Edep_angle}}
-        \end{center}
-\end{figure}
+%shown. \label{Edep_angle}}
+%        \end{center}
+%\end{figure}
 
@@
 number of PEs and the decreased $R_{C/S}$. For 1.4~MeV electrons, the RMS values of the distributions for all three
-reconstructed coordinates are smaller than 4.5~cm. The mean cosine of the angle between the true direction
-and the reconstructed direction for different energies of the initial electron is shown in figure
-\ref{Edep_angle}. The direction reconstruction performance is still promising for energies as low as 1.4~MeV.
+reconstructed coordinates are smaller than 4.5~cm. The direction reconstruction performance is still promising for energies as low as 1.4~MeV.
 
@@
 
-We have developed a technique to separate scintillation and Cherenkov light to reconstruct direction of electrons with energies 5~MeV, 2.1~MeV and 1.4~MeV in a liquid scintillator detector.
+We have proposed a technique to separate scintillation and Cherenkov light to reconstruct direction of electrons with energies 5~MeV, 2.1~MeV and 1.4~MeV in a liquid scintillator detector.
 These energies have been chosen to represent typical neutrino-electron elastic scattering energies and
@@
 
-While scintillation light is isotropic, Cherenkov light with wavelengths above the absorption cutoff $\sim$370~nm carries the information about the direction of the electrons.
-All light shorter than this gets absorbed and re-emitted isotropically as a part of the scintillation process.
+While scintillation light is isotropic, Cherenkov light with wavelengths above the absorption cutoff, $\sim$370~nm for a typical scintillator, carries the information about the direction of the electrons.
+All light with wavelength shorter than this gets absorbed and re-emitted isotropically as part of the scintillation process.
 On average scintillation light is delayed with respect to the direct Cherenkov light due to chromatic dispersion and the finite time of the scintillation processes; the early light thus contains directional information.
 
-Using a Geant4~simulation of a spherical detector with radius of 6.5~m and photodetectors with TTS of 0.1~ns, we have shown that for electrons originated at the center a time cut on the early light is effective at isolating the directional light, improving the ratio of Cherenkov light to scintillation light from 0.02 to 0.63.
+Using a Geant4~simulation of a spherical detector with radius of 6.5~m and photodetectors with transit time spread of 0.1~ns, we have shown that for electrons originated at the center a time cut on the early light is effective at isolating the directional light, improving the ratio of Cherenkov light to scintillation light from 0.02 to 0.63.
 This ratio is degraded by a factor of 2.5 if current photodetectors with $\sim$1~ns resolution were used.
@@
 As expected, the reduction in photon statistics with lower energies leads to a broadening of the reconstructed vertex by $\sim$1~cm.
-Similarly, the mean $\cos(\theta)$ between the reconstructed and initial directions drops from 0.88 to 0.74.
+%Similarly, the mean $\cos(\theta)$ between the reconstructed and initial directions drops from 0.88 to 0.74.
 % reconstructed direction drops from 0.88 to 0.74 as the $\cos(\theta)$ distribution broadens.
@@
 This technique is promising, and we plan to continue work on the topic.
+The next step will be the extension of the algorithm to multi-particle events distributed throughout the volume.
+This will be followed by a study of the background discrimination capability of the technique given realistic backgrounds and the detailed kinematics of double-beta decay events and other neutrino interactions.
 The ability to reconstruct direction in kiloton-scale scintillation detectors would expand capabilities of neutrino experiments, especially those also searching for neutrino-less double-beta decay.
-More generally, this technique could be applied wherever scintillation-based detectors are used.
+More generally, this technique could be applied whenever scintillation-based detectors are used.
 

### 2014-04-21 Direction_v3_withLindley.tex
Add old versions of paper. --Lindley

--- 2013-07-20 Direction_v2.tex
@@
 
-%\linenumbers
+\linenumbers
 
@@
 
+In $0\nu\beta\beta$, the electrons emerge with a combined energy equal to the endpoint of the particular isotope's Q-value.
+The individual electrons come out with a distribution of energies and angular correlations.
+The most probable being equal division of energy between the back-to-back electrons\cite{phasespace} as is shown in FIG.~\ref{detector_view} for an example $^{116}$Cd $0\nu\beta\beta$ event.
+Since the decay half-life is proportional to the phase-space factor, isotope's with higher Q-values are preferred and due to backgrounds from the daughters of the $^{238}$U and $^{232}$Th decay chains those with Q-values at or above 2.6~MeV at most often considered for $0\nu\beta\beta$ searches.
+There are hundreds of candidate $0\nu\beta\beta$ isotopes\cite{tabledbb}, but only a handful with large Q-values.
+Most of the high Q-value candidates have been considered as a dopant for a liquid scintillator: $^{150}$Nd(Q=3.367~MeV)\cite{minfang,nd1}, $^{96}$Zr(Q=3.350~MeV)\cite{zr1},  $^{100}$Mo(Q=3.034~MeV)\cite{mo1}, $^{82}$Se(Q=3.350~MeV)\cite{qdot}, $^{116}$Cd(Q=2.802~MeV)\cite{qdot, cd1}, $^{130}$Te(Q=2.533~MeV)\cite{qdot, biller}, $^{136}$Xe(Q=2.479~MeV)\cite{KZ0nu} and $^{124}$Sn(Q=2.29~MeV)\cite{sn1}.
+Xenon gas readily dissolves into liquid scintillator.
+For the other isotopes, the correct organometallic compound needs to be found that produces a stable scintillator with a long attenuation length in the wavelength region of interest.
+Recently, nanocrystals called quantum dots have been explored as an alternative method for doping scintillator\cite{qdot}.
+Quantum dots have interesting optical properties which are discussed further in Section~\ref{scintillator_emission_sec}.
+
+The reconstruction of the direction of the two simultaneous emitted $\sim$1.5~MeV electrons from the $0\nu\beta\beta$ decay is quite challenging.
+In this paper, we decided to start our directional studies with 5~MeV single electrons since reconstruction of direction in this energy range had never been achieved.
+This is a logical energy to use since this is where historically scintillating detectors become more practical than water Cherenkov detectors.
+This is also an energy that is relevant for studies of $^{8}$B solar neutrinos and neutrino-electron elastic scattering measurement.
+
 \section{Geant4 Simulation}
@@
 
-The detector geometry is a sphere of 6.5~m diameter filled with scintillator.
-FIG.~\ref{detector_view} shows the geometry and the Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event.
+The detector geometry is a sphere of 6.5~m diameter filled with scintillator, see FIG.~\ref{detector_view} shows the geometry and the Cherenkov light from an example $^{116}$Cd $0\nu\beta\beta$ event.
 The default
@@
 
-\bibliography{DirectionBibliography}
+\bibliography{DirectionBibliography_withLindley}
 

### 2014-04-21 Direction_v7_submit.tex
Add old versions of paper. --Lindley

--- 2013-08-15 Direction_v7.tex
@@
         \subfigure[ ~Default
-        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
+        simulation.]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
         \subfigure[ ~Increased TTS
-        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
+        (1.28~ns).]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
         \subfigure[ ~Red-sensitive
-        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
+        photocathode.]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
         \caption[]{Photoelectron (PE) arrival times after application

### 2014-04-21 RefereeResponse/RefereeResponseTwo_v0.tex
Referee response. --Lindley

\documentclass[11pt]{article}

\usepackage[usenames,dvips]{pstcol}

\usepackage{color}

\usepackage{amssymb}

\usepackage{amsmath}

\usepackage{graphicx,color}

%

%\input{symbols.tex}

%

\newcounter{question}

\newcommand{\noteq}{{\vskip0.1in\bf \large \color{red} [Q:{\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\noteQ}{{\bf \LARGE \color{blue} [Q: {\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\notea}{{\bf \large \color{red} [A]~}}

\newcommand{\noteA}{{\bf \LARGE \color{blue} [A]~}}

\newcommand{\noteqq}[1]{{\bf \large \color{red} [Q: {\color{black}#1}]~}}

\newcommand{\noteaa}[1]{{\bf \large \color{red} [A: {\color{black}#1}]~}}

%

\textwidth 6.5in

\textheight 9.4in

\topmargin -0.8in

\oddsidemargin 0.0in

\evensidemargin 0.0in

\parindent 0.0in

\def\baselinestretch{1.0}

\def \Russ {Translated from Russian. Fixed. Thanks.}

\def \F    {Fixed.}

\def \FT   {Fixed. Thanks.}

\def \D    {Done.}

\def \AP   {Basically Stylistic- Author's Prerogative.}

\begin{document}

\begin{flushright}

\today

\end{flushright}

%----------------------------------------------------------------

Dear JINST Editors,\\

\parindent 1.0in

We thank the referees again for their very helpful and knowledgeable
comments. We have made all the minor changes that were suggested by referee 1 and referee 2.

The remaining point to discuss is the title of the paper, \textit{Measuring directionality in double-beta decay and neutrino interactions with kiloton-scale scintillation detectors}.  We feel that the title indicates that we will be developing a method to measure direction and  the title guides the reader to what type of event we will be proposing to reconstruct, events with $\sim$MeV energies. It is important to note that reconstruction of $\sim$GeV events in scintillating detectors has been tackled by other groups and we reference their work in this publication.

Referee 2 says ``I believe the title is sending a different message to the reader, indicating the measurement of (low energy) particle directionality as a powerful tool to improve the performance of experiments on neutrino interactions or double-beta decay."  In order to keep the title, Referee 2 requests that we address in detail the kinematics of the interactions, the Cherenkov threshold, and the spatial distribution of the decays.  We have addressed part of this request by including the number of electrons above the Cherenkov threshold and above 1.4~MeV in $^{116}$Cd neutrinoless double beta decay events. However, to address these issues further we would need to do a significant amount of work on the reconstruction algorithm. We feel goes beyond the proof of principle paper we were intending. If the editors feel strongly that we should change the title, we can, but we believe it is the correct title for the paper.

Below we list the referee requests and changes in detail.\\ \\

\parindent 3.0in

    Sincerely,\\

\parindent 4.0in

		         Lindley Winslow, for the Authors

\parindent 0.0in

%\begin{verbatim}

%\end{verbatim}

{\bf Referee 1: }

\noteq [Referee 1 1) Organization] ``Please consider organizing the paper along these items/subsections be-
low (or others equivalent), in order to avoid a pure sequential enumeration
of (disconnected) sections''

\notea{\bf We have changed the organization.}\\

\noteq [Referee 1: 2) Figures] ``In general, the figures are not really easy to find''

\notea{\bf  \F~We have worked on the figure placement}\\

\noteq [Referee 1: 3) ] ``it is very difficult for an outsider to understand why the authors are so attached to cos(theta)
in Fig.7.''

\notea{\bf We have removed Fig. 7. We were not attached to the figure.}\\

{\bf Referee 1: Specific Comments}

\noteq [Referee 1:  Abstract and Conclusions] ``Develop a technique probably too strong: ”we propose a technique” I would say. Developing it will take some more time, to my understanding...''

\notea{\bf \FT}\\

\noteq [Referee 1:  Abstract and Conclusions] ``Generally, I suggest not to assume that everybody understands what TTS (or even what transit-time spread) is, specially outside a context. In the figures, one could add *photo-detector* transit-time spread, for instance.''

\notea{\bf \FT}\\

\noteq [Referee 1:  Abstract and Conclusions] ``Fig. 3 EPS$\rightarrow$PEs''

\notea{\bf \FT}\\

\noteq [Referee 1:  Page 7, Top] ``We note that due to scattering the final... $\rightarrow$one could
add Coulomb scattering since the word scattering was used before in other
context.'' and "A bit later: has a distance from... $\rightarrow$travels a distance from''

\notea{\bf \FT}\\

\noteq [Referee 1:  Page 7, Bottom] ``after the TTS resolution has been applied $\rightarrow$probably
rephrase.''

\notea{\bf \D}\\

\noteq [Referee 1:  section 4:] ``truth information $\rightarrow$’MC truth’ information.''

\notea{\bf \FT}\\

\noteq [Referee 1:  section 4:] ``For the default simulation case... $\rightarrow$I am fine, but it is not easy to re-
member all details and a table may help.''

\notea{\bf We have added a table.}\\

\noteq [Referee 1:  Fig. 5(Left)]  ``Some strange points appear in between the p x(y,z)/p lines...''

\notea{\bf  \FT}\\

\noteq [Referee 1:  Fig. 5(Top)]  ``Fig. 5(Top) $\rightarrow$Fig. 5(Left).''

\notea{\bf \FT}\\

\noteq [Referee 1:  Conclusions]  ``We have developed a technique $\rightarrow$not quite''

\notea{\bf Changed to $\rightarrow$ We have proposed..}\\

\noteq [Referee 1:  Conclusions]  ``370nm $\rightarrow$for pseudocumene''

\notea{\bf Added for a typical scintillator.}\\

\noteq [Referee 1:  Conclusions]  ``All light shorter $\rightarrow$with wavelength shorter''

\notea{\bf \D}\\

\noteq [Referee 1:  Conclusions]  ``as a part $\rightarrow$as part?''

\notea{\bf \D}\\

\noteq [Referee 1:  Conclusions]  ``TTS $\rightarrow$Do not use acronym here, for sure, at least not only.''

\notea{\bf \D}\\

\noteq [Referee 1:  Conclusions]  ``Similarly the mean cos(theta)... please see my comment above on this re-
gard.''

\notea{\bf \F~We have removed this line as it does not add any information.}\\

\noteq [Referee 1:  Conclusions]  ``We plan to continue work on the topic $\rightarrow$please address what you plan to do
very shortly and what future steps are.''

\notea{\bf \F~We have added a sentence about our plans.}\\

\noteq [Referee 1:  Conclusions]  ``wherever$\rightarrow$whenever?'

\notea{\bf \D}\\

{\bf Referee 2: }

\noteq [Referee 2:  Physics Case]  ``The reference to the Gotthard experiment (in which the identification of the two electron tracks
was based on a different principle) is probably not straightforward and
should be made more transparent.''

\notea{\bf \F~We are now more explicit that this is a different technology.}\\

\noteq [Referee 2:  Physics Case]  ``5 MeV, head to head single electrons: the authors have added 2 more
examples at lower energies. These are more appropriate for double-beta
decay and provide useful information.''

\notea{\bf  We are glad that they are useful.}\\

\noteq [Referee 2:  Physics Case]  ``Cherenkov thresholds: a mention to the inefficiency in the identification
of both electron tracks would be worth."

\notea{\bf \D~We have calculated the efficiency for $^{116}$Cd and put that with the description of the first figure. A more detailed study is part of our future work.}\\

\noteq [Referee 2:  Page 5, line 3 from bottom] `` ``greater than 20 m'' ...The conclusion is
reasonable for an event originating in the center of the detector. The
situation could be substantially different for (a large fraction of) events
originating far from the center (close to the balloon surface) or larger
scale detectors.''

\notea{\bf \D~We now note that these effects will become important with uniformly distributed and lower energy events.}\\

\noteq [Referee 2:  Caption to Figure 3.]  ``I understand this is probably not very relevant
but I assume that the 1000 electrons generation is carried out in the
same conditions described for Figure 1. If this is the case maybe it
would be worth to specifiy it.''

\notea{\bf \D~We now specify that these are single electrons moving in the x-direction.}\\

\noteq [Referee 2:  Page 5, line 2.] ``What is the main effect of incomplete coverage? Simply
a lowering of the measured number of photons?"

\notea{\bf \D~We now specify that a uniform decrease in coverage would simply lower the measured number of photons.}\\

\noteq [Referee 2: Page 8, line 4 from bottom] ``even at 1.4 MeV” looks a quite strong
statement taken into account that this is still a favorable condition in
double-beta decay (one of the two electrons has always an energy in
the interval 0-Q/2.)''

\notea{\bf \D~We now note that the photon statistics will continue to drop as we head to lower energies.}\\

\end{document}

### 2014-04-21 RefereeResponse/RefereeResponse_v1.tex
Referee response. --Lindley

%This is Janet's for the NSF.
\documentclass[11pt]{article}
\usepackage{graphicx,color}
\usepackage{rotating}
\usepackage{epsfig,graphics,rotate,color}
\usepackage{wrapfig}
\topmargin 0pt
\advance \topmargin by -\headheight
\advance \topmargin by -\headsep
\textheight 8.9in     
\oddsidemargin 0pt
\evensidemargin \oddsidemargin
\marginparwidth 0.5in
\textwidth 6.5in

%This is mine for changing the font.
\usepackage[T1]{fontenc}
\usepackage{times}
 \usepackage{hyperref}

%This is for NSF style
\bibliographystyle{utphys}

\begin{document} 
%Do this myself
%\begin{center}
%\Large{Next-Generation Liquid-Scintillator-Based Detectors: \\ Quantums Dots and Picosecond Timing}\\[0.25cm]
%\large{Lindley Winslow,  \today}\\[0.25cm]
%\end{center}
\noindent
Dear JINST Editors, \\

In the following draft we have attempted to address the excellent comments of the referees and below we provide some additional information. In some areas we would like to limit the scope of the work proposed by the referees. We envisioned this paper as a proof-of-principle. The reconstruction of the direction of low energy ($\sim$1 MeV) electrons in a liquid scintillator detector has never been shown before. In this paper, we wanted to show the following: the scintillation/Cherenkov separation was possible, the parameters of a detector that were important to this separation, and finally that Cherenkov reconstruction techniques would work in idealized situations. We are now working on at least two follow-up papers. The first is a detailed study of  reconstruction in more realistic scenarios. The second is a $0\nu\beta\beta$  experiment proposal where the detailed background reduction factors and sensitivity to new physics are presented, SuperNEMO has a similar paper but nothing similar exists for scintillator detectors since a directional signal has not been considered for them in the past.  Finally, we will be having a turnover of personnel so this paper provides an excellent summary of their work.\\

\noindent
The following is a more detailed response to the referees' comments.\\

\noindent
Thank you,\\ 
Lindley Winslow for the authors
\\
\\
\noindent
{\bf Response to Referee 1}\\
Page 1: \\
Liquid scintillators detectors quote their resolution in terms of  $\sim$5\%$/\sqrt{E(MeV)}$. We have updated the text to make that more clear. \\ \\
Rewrote introduction to include NEXT/SuperNEMO and to be more clear about the energy resolution issues and our approach to this paper. \\ \\
Defined PPO earlier.\\

\noindent
Page 3: \\
Added some example values for the equations.\\ \\
There is one referee comment here that we did not know how to interpret, {\it This case is shown in Fig.1.
->Indicating the direction of the electrons should not harm.}\\ \\
Changed the wording about the Q-value and the background considerations. \\ \\

\noindent
Page 4: \\
Also added here some reference values for the cutoff wavelength etc. as requested. \\ \\
We have added a more detailed explanation on why optical scattering can be neglected and a detailed reference. In Fig.~\ref{scatt} below we also show the simulation results with and without optical scattering for reference. In the simulation without scattering a ``pure" attenuation length is used so photons are simply absorbed and disappear from the simulation, therefore when scattering is turned on more photons reach the sphere. The timing distribution gains a small tail at longer times as is expected.\\ \\
We fixed the wording for the explanation of the QE values.\\ \\
We added some text explaining why 5~MeV is good for evaluating the technique.\\ \\
We added the distance between the start and stop points. As can be seen in Fig.~\ref{truthscatt}, the scattering angles are small until the electron falls below Cherenkov threshold, therefore the light emitted encodes the emission direction of the primary electron.\\ \\

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{Cerenkov_PEs_scattering_comparison.pdf} 
         \caption{The arrival time of Cherenkov light with (red) and without scattering (black).\label{scatt}}
        \end{center}
\end{figure}

\noindent
Page 5: \\
We rephrased the discussion of the energy transfer between solvent and solute.\\ \\
We updated the citations for the LAPPDs to more recent ones.\\ \\

\noindent
Page 6: \\
We have added a reference to the Geant4 standard electron physics list.  The main request of the referee is to see Fig. 3 as a function of the electron path. We are not setup to easily extract that information. Instead we point them to the data in Fig.~\ref{truthscatt} and we note again that the electron scattering is small during the time the electron is above Cherenkov threshold. In addition, we have remade Fig. 3 with all three energies and the Cherenkov structure is clear down to 1.4~MeV.

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.50]{1p4_MeV_Direction_to_Stop.pdf} 
        \includegraphics[scale=0.50]{1p4_MeV_Direction_to_c_threshold_point.pdf} 
        \includegraphics[scale=0.50]{1p4MeV_direction_of_last_step_h.pdf} 
         \caption[]{Truth scattering data.\label{truthscatt}}
        \end{center}
\end{figure}

\noindent
Page 7-8: \\
We would like to leave the sections on scintillator in for completeness.\\ \\

\noindent
Page 9: \\
Added reconstruction plots for all energies and now note that the p$_x$ distribution is the $cos\theta$ distribution. \\ \\
The referee correctly notes that the vertex resolution is a sizable effect, so our use of the truth timing information is a non-negligible simplification. The effect is 0.15~s for the 3cm resolution. Eventually, we should simultaneously reconstruct the direction and vertex. This also addresses the issue of direction reconstruction in other sections of the detector. {\bf Should I add more here about our logic in neglecting this effect for this paper.  Since we are putting in the TTS smearing by hand,  we could simply double the transit time spread smearing we are using. We were interested in whether timing on the order of 0.1~ns, compared to the standard 1~ns, would allow us to reconstruct direction.}\\ \\
We now show the $cos\theta$ distribution for all energies. Fig.~7  is nice as you can see the performance of the reconstruction and extrapolate to higher or lower energies. For reference,  the Cherenkov threshold for  n(370nm)=1.465892 is 0.1879~MeV. This a logical choice of wavelength for defining the Cherenkov threshold since light produced below 370 nm will be absorbed before reaching the edge of the sphere since once again this is the absorption cutoff of the scintillator. \\ \\

\noindent
Conclusions:\\
We rewrote the conclusion to emphasize more that this is the first attempt to reconstruct direction in a liquid scintillator. \\ \\ \\ \\

\noindent
{\bf Response to Referee 2}\\

\begin{itemize}
\item The referee asked for more discussion on the possible sensitivity improvement. This is what we would like to work on next. It is obvious that it will but the magnitude requires a detailed sensitivity analysis and a better understanding of the final reconstruction performance. This is why we wanted to show it was possible first and then work on these more detailed and time intensive analyses. We have added some more discussion on the experience of other experiments as was requested by referee 1.
\item We agree 5~MeV is not interesting for $0\nu\beta\beta$, however it is interesting to neutrino-electron scattering interactions (also referred to in the title) and this allowed us to verify our simulation and reconstruction and then push to lower energies where lower photon statistics make the verification more difficult. In the reconstruction section we do present the results of lower energy.
\item  We agree that position dependence is going to be important. Our next step is to simultaneously reconstruct position and direction. A preliminary attempt  using all the photons to do the position reconstruction at 5~MeV showed no bias and a decrease in the resolution to 8.5~cm. The position reconstruction uncertainty is equivalent to a larger TTS smearing so our direction reconstruction results will degrade or improve depending on the vertex reconstruction. This is the content of a more detailed paper to follow.
\item The statistical fluctuations are a good point. We added discussion on this to accompany what was Fig.~7, the average $cos\theta$ plot.  {\bf I still need to address this, any suggestions?}
\end{itemize}

Most of the more detailed requests were addressed in the response to referee 1.  These are the ones unique to referee 2:

\begin{itemize}
\item Added reference for matter-antimatter asymmetry.
\item Added definition of x.
\item Clarified wording of a ``probable case", now dependent on model. We also added the reference to Boehm and Vogel.
\item Added discussion of how scattering at the sphere would affect the timing discussion.
\end{itemize}

{\bf \it This completes the comments of the referees.}

%Here is the bibliography
%\bibliography{QuantumDotWhitepaper} 
\end{document}

### 2014-04-21 RefereeResponse/RefereeResponse_v3.tex
Referee response. --Lindley

--- 2014-01-10 RefereeResponse/RefereeResponse_v2.tex
@@
 
+\usepackage{graphicx,color}
+
 %
 
-\input{symbols.tex}
+%\input{symbols.tex}
 
@@
 
-(Figure 3), but was labeled as $p_x/|\vec{p}|$ rather than the obvious
+(now Figure 5), but was labeled as $p_x/|\vec{p}|$ rather than the obvious
 
-CosTheta that it is. We have made this clear on the figure and in the
+$\cos\theta$ that it is. We have made this clear on the figure and in the
 
-text. Thanks.}\\
+text. We have also added the distribution for the 2.1~MeV point. Thanks.}\\
 
@@
 
-\notea{\bf  \D }\\
-
-DOES ANYBODY KNOW WHAT HSE IS ASKING FOR WRT THE REFERENCE?
+\notea{\bf  \D  ~We have added references to KamLAND and Borexino.}\\
 
@@
 
-\notea{\bf  \D}\\
-
-I DON'T SEE THE REFERENCE TO NEXT IN THE TOPOLOGY PARAGRAPH- DID I MISS
-
-IT? IT SHOULD GO AFTER ``..ESPECIALLY FOR 0NUBETABETA SEARCHES'''?
+\notea{\bf  \D We have added a discussion of previous work on directionality.}\\
 
@@
 
-\notea{\bf  \D}\\
-
-WAS ANYTHING CHANGED IN THE FIGURE? THE CAPTION MAKES IT CLEAR, AND
-
-THE BRIGHT WHITE AXIS ARROW IS CLEAR... IF WE DON'T CHANGE THE FIG, WE
-
-SHOULD SAY SO AND WHY NOT (EASY)
+\notea{\bf  We were not sure what the referee was requesting.
+If this is a request for figure 1, we have tried adding the arrows and it becomes too busy.
+We think the caption description works well.}\\
 
@@
 
-sues an isotope that satisfies this condition..''
+uses an isotope that satisfies this condition..''
 
-\notea{\bf \FT}\\
-
-IS THIS IS FIXED? I COULDN'T FIND IT IN THE CURRENT DRAFT AROUND
-
-THE 2.6 MEV DISCUSSION. .I'M CONFUSED BY THE COMMENT, FRANKLY. DOES
-
-THE PRESENT TEXT NEED MORE WORK? (I.E. I NEED EDUCATING...)
-
-\noteq [Referee 1: Page 3]
-
-\notea{\bf \FT}\\
-
-THIS IS FIXED, RIGHT? I COULDN'T FIND IT IN THE CURRENT DRAFT AROUND
-
-THE 2.6 MEV DISCUSSION. .I'M CONFUSED BY THE COMMENT, FRANKLY. DOES
-
-THE TEXT NEED WORK? (I.E. I NEED EDUCATING...)
+\notea{\bf \FT We have rewritten this section to address this comment.}\\
 
@@
 
-\notea{\bf \D}\\
-
-I SEE THE ATTENUATION LENGTH, BUT NOT N@400
-
-\noteq [Referee 1: Page 4] Mentioning the figure in the citation.
-
-\notea{\bf ?}\\
-
-I DIDN'T UNDERSTAND THIS COMMENT- THE PAPER MENTIONS FIG 4C..
+\notea{\bf \D We have added the attenuation length at 400nm.
+We have also added the n at other wavelengths in other locations.
+Unfortunately we cannot cite one figure in the references as the index of refraction is taken from the KamLAND simulation which is based on several measurements that are in the cited references.}\\
 
@@
 
-\notea{\bf ?}\\
-
-I DIDN'T UNDERSTAND THIS ONE EITHER
+\notea{\bf This has been added.}\\
 
@@
 
-that the 5 MeV case is for neutrino-electron experiments. For dbb we
+that the 5 MeV case is for neutrino-electron scattering experiments. For $0\nu\beta\beta$ we
 
@@
 
-\notea{\bf We have added studies of 1.4 MeV and 2.1 MeV electrons,
+\notea{\bf  In the updated figure 4 (angular PE hit distribution after time cut), we added the histograms for 1.4~MeV and 2.1~MeV.
+This shows that also for 2.1 and 1.4~MeV the PE hits are not isotropized, including the full history of the electron tracks.
+We have also added the path lengths and total distance and time from the vertex to the text.
+We also present in Fig.~\ref{truthscatt} of this report  more raw data for the referees which supports the statement that the scattering angle is small while the majority of the Cherenkov light is produced.}\\
 
-  which show that the separation and directionality based on timing
-
-  hold as the electron energy is lowered, as shown in the newly-added
-
-  Figures 5 and 6. We have also added the path
-
-  lengths and total distance and time from the vertex. We believe that
-
-  additional plots beyond the ones we have added won't add any useful
-
-  information.}\\
-
- HSE'S CLEARLY STUCK ON THIS- MISSED THE POINT. HOWEVER, WE NEED TO
-
-  ANSWER THIS DIRECTLY - IS IT POSSIBLE TO MAKE ONE PLOT OF
+\begin{figure}
+        \begin{center}
+        \includegraphics[scale=0.50]{Figure1_1.pdf} 
+        \includegraphics[scale=0.50]{Figure1_2.pdf} 
+         \caption[]{"For electrons with and initial energy of 1.4~MeV, we show (Top) the angle between the initial electron direction and the direction at 0.7 MeV (=half of the initial electron energy) and (Bottom) the angle between the initial electron direction and the direction of the last Geant4 step before the electron stops.
+We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction; however the scattering angle is small while the majority of Cherenkov light is produced.}
+        \end{center}
+\end{figure}
 
@@
 
-\notea{\bf }\\
-
-NEEDS ATTENTION...
+\notea{\bf  \D}\\
 
@@
 
-\notea{\bf }\\
-
-I CAN'T WRITE AN ANSWER TO THIS- CHRISTOPHE,ANDREY, MATT, LINDLEY?
-
-(there's lots of email that contains the answer, it seems...)
+\notea{\bf We agree with the referee and plan to address this in future papers.
+To do a credible and useful job on the position dependence will require a fairly-detailed vertex reconstruction.
+This is beyond the scope of this present ‘proof-of-principle’ paper.
+We have tried to clarify this in the text }\\
 
@@
 
-\notea{\bf \D See Figure 3.}\\
-
-I DON'T HAVE A FIGURE 7. ???
+\notea{\bf \D See Figure 5.}\\
 
@@
 
-I DON'T HAVE A FIGURE 7. ???
-
 \noteq [Referee 1: Conclusions] ``Should be rewritten after the
@@
 
-WE SHOULD REWRITE THE CONCLUSIONS TO BE MORE SPECIFIC WITH MORE
-
-FACTS. AS THEY STAND THEY'RE SORT OF GENERIC- IF THAT'S ALL ONE READS
-
-OF THE PAPER ONE'S NOT MUCH WISER ...
-
 \noteq [Referee 1: Page 7-8] ``I see no convincing proof of it with
@@
 \notea{\bf We have made the sentence more specific. Thanks.}\\
-
-WE SHOULD BE MORE SPECIFIC- CAN GIVE NUMBERS RATHER THAN OPINIONS
-
-HERE..
 
@@
 
-\notea{\bf Warms the cockles of our hearts...}\\
+\notea{\bf  Thank you, this warms our hearts.}\\
 
@@
 
-\notea{\bf Yes. We have extended the analysis to lower energies and
+\notea{\bf Yes. We included more data from the lower energy studies,
 
@@
 
-  `proof-of-principle level, as to go deeper yet will require vertex
-
-  reconstruction, which is a much bigger effort (we are working on it).}\\
+  `proof-of-principle level.
+Our next work will include a full sensitivity analysis including \a deeper understanding of the performance of the reconstruction algorithms with a more detailed simulation.
+This is a big job and we are working on it!
+}\\
 
@@
 
-\notea{\bf }\\
-
-NYGREN'S RIGHT- CAN WE ADD A SENTENCE OR TWO TO THE INTRO AND
-
-(PERHAPS) THE CONCLUSIONS?
+\notea{\bf  \D}\\
 
@@
 
-\notea{\bf }\\
-
-I DON'T KNOW WHAT TO SAY HERE- SOMEBODY MORE EXPERT?
+\notea{\bf Statistical fluctuations are addressed by the studies at lower energies and we mention explicitly in the conclusions the widening of the distributions due to decreased statistics.
+Threshold effect are outside of the scope of this paper.
+In general in these detectors, the largest loss of photon statistics is a the photocathode.
+The electronics for PMT readout are very efficient.
+}\\
 
@@
 \notea{\bf \FT}\\
-
-PLEASE ADD `` for neutrino detectors'' after ``the other developed
-
-technology'' AT THE TOP OF PAGE 2.
 
@@
 
-DO WE HAVE THE USEFUL FRACTIONS OF CHERENKOV LIGHT? (I DON'T EVEN KNOW
-
-IF THAT'S WELL-DEFINED?)
-
 \noteq [Referee 2: Page 3] `` `Probable case being...' Events with
@@
 \notea{\bf \FT}\\
-
-WE SHOULD REWRITE THIS SENTENCE ..E.G. ``We consider the case of
-
-equal..'' OR SOMETHING.. (?) (MAYBE CHANGE 'PROBABLE' TO 'SIMPLE'?)
 
@@
 
-\notea{\bf }\\
-
-THIS NEEDS AN EXPERT'S ANSWER...
+\notea{\bf This part of the paper describes a simplified test simulation (where the photodetector TTS was switched off) to extract the time scales of various effects.
+We made changes in the text to make this clear.}\\
 
@@
 
-\notea{\bf }\\
-
-THIS NEEDS AN EXPERT'S ANSWER...
+\notea{\bf  We have added to what is now figure 4 the distributions for the two lower energies. We think this is a nice illustration of the limits/power of the method.}\\
 
@@
 
-\notea{\bf Yes, we agree. We are not proposing throwing away the
-
-scintillation light, but in addition using the selected Cherenkov light to
-
-determine the electron direction, i.e., not `limiting the
-
-reconstruction to Cherenkov photons''}\\
-
-CAN THIS ANSWER BE IMPROVED? (YES)
+\notea{\bf Yes, we agree.
+In this paper we were focussing on the Cherenkov photons as a simple way to apply direction reconstruction algorithms.
+In the future we will work to reconstruct position and direction simultaneously from the scintillation and Cherenkov light.}\\
 

### 2014-04-21 RefereeResponse/RefereeResponse_v4.1.tex
Referee response. --Lindley

\documentclass[11pt]{article}

\usepackage[usenames,dvips]{pstcol}

\usepackage{color}

\usepackage{amssymb}

\usepackage{amsmath}

\usepackage{graphicx,color}

%

%\input{symbols.tex}

%

\newcounter{question}

\newcommand{\noteq}{{\vskip0.1in\bf \large \color{red} [Q:{\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\noteQ}{{\bf \LARGE \color{blue} [Q: {\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\notea}{{\bf \large \color{red} [A]~}}

\newcommand{\noteA}{{\bf \LARGE \color{blue} [A]~}}

\newcommand{\noteqq}[1]{{\bf \large \color{red} [Q: {\color{black}#1}]~}}

\newcommand{\noteaa}[1]{{\bf \large \color{red} [A: {\color{black}#1}]~}}

%

\textwidth 6.5in

\textheight 9.4in

\topmargin -0.8in

\oddsidemargin 0.0in

\evensidemargin 0.0in

\parindent 0.0in

\def\baselinestretch{1.0}

\def \Russ {Translated from Russian. Fixed. Thanks.}

\def \F    {Fixed.}

\def \FT   {Fixed. Thanks.}

\def \D    {Done.}

\def \AP   {Basically Stylistic- Author's Prerogative.}

\begin{document}

\begin{flushright}

\today

\end{flushright}

%----------------------------------------------------------------

Dear JINST Editors,\\

\parindent 1.0in

We thank the referees for their very helpful and knowledgeable
comments. We have substantially rewritten the manuscript in response 
to the comments as detailed in the list below.

There are several areas in which the referees requested
substantially expanding the scope of the paper.  We envisioned this
paper as a proof of  the principle that the use of the fast timing
photodetectors presently being developed would enable the separation
of Cherenkov and scintillation light, as the reconstruction of the
direction of low energy ($\sim$1 MeV) electrons in a liquid
scintillator detector has never been shown before.  In this paper, we
want to: 1) show scintillation/Cherenkov separation is possible; 2)
determine the parameters of a large scintillator detector that are
important to this separation; and 3) demonstrate that reconstruction
techniques based on Cherenkov light would work in idealized
situations.

We are now working on two follow-up papers that are substantially
beyond the scope of the present paper, moving beyond
proof-of-principle into the laborious and time-intensive area of
reconstruction algorithms, areas for which the difficulty is known
well by some of us from our experience on KamLand. The first is a
detailed study of reconstruction in more realistic scenarios. The
second is a proposal for a $0\nu\beta\beta$ experiment  where the detailed
background reduction factors and sensitivity to new physics are
presented. SuperNEMO has a similar paper, but for scintillator
detectors nothing similar exists since a directional signal has not
been considered for them in the past.  Finally, we will be having a
turnover of young personnel so this paper provides an accessible summary
of their work.

\parindent 3.0in

    Sincerely,\\

\parindent 4.0in

		         Lindley Winslow, for the Authors

\parindent 0.0in

%\begin{verbatim}

%\end{verbatim}

\noteq [Referee 1 Intro] ``I like  very much the clean, transparent, simple
and well-motivated methodology (with one exception..). The structure
of the paper is nice''

\notea{\bf Many thanks.}\\

\noteq [Referee 1 Intro] ``However the presentation of the results is a
different thing... I am particularly concerned with the fact...,
instead of showing the resulting CosTheta distribution that would seem
like the obvious result one wants to see. .... Why not to disclose
this openly to allow the reader to evaluate the results in detail?''

\notea{\bf  This is our fault- the desired plot was (is) there
(now Figure 5), but was labeled as $p_x/|\vec{p}|$ rather than 
$\cos\theta$. We have made this clear on the figure and in the
text. We have also added the distribution for the 2.1~MeV point. Thanks.}\\

\noteq [Referee 1: Page 1] ``Please specify FWHM or sigma and give a
reference for both cases.''

\notea{\bf  \D  ~We have added references to KamLAND and Borexino.}\\

\noteq [Referee 1: Page 1] ``...but please avoid misplacing the adjective good in
this context. ...''

\notea{\bf  \FT}\\

\noteq [Referee 1: Page 2] ``..the authors should..check and refer to
the work of NEXT and Super-Nemo collaborations:''

\notea{\bf  \D We have added a discussion of previous work on directionality.}\\

\noteq [Referee 1: Page 2] ``Define PPO here, not later''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3] ``Please define all magnitudes in eq. 2.1 and 2.2''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3]  ``Indicating the direction of the
electrons should not harm''

\notea{\bf  We were not sure what the referee was requesting. If this is a request for figure 1, we have tried adding the arrows and it becomes too busy. We think the caption description works well.}\\

\noteq [Referee 1: Page 3] ``None of the best limits or claims in bb0
uses an isotope that satisfies this condition..''

\notea{\bf \FT We have rewritten this section to address this comment.}\\

\noteq [Referee 1: Page 4] ``could you perhaps include the
attenuation @400nm and n@400..?''

\notea{\bf \D We have added the attenuation length at 400nm. We have also added the n at other wavelengths in other locations. Unfortunately we cannot cite one figure in the references as the index of refraction is taken from the KamLAND simulation which is based on several measurements that are in the cited references.}\\

\noteq [Referee 1: Page 4] ``Could you please include also the cutoff
wavelength previously referred to?''

\notea{\bf This has been added.}\\

\noteq [Referee 1: Page 4] `` Scattering have not been included..''

\notea{\bf We have added a figure and accompanying text to show that
the scattering only adds a late tail. Thanks.}\\

\noteq [Referee 1: Page 4] ``For which digitized values come from ??''

\notea{\bf \F}\\

\noteq [Referee 1: Page 4] ``5 MeV electrons.. It is very disturbing
to benchmark the simulation with electrons having an energy doubling
the Qbb of the most common isotopes..''

\notea{\bf We have clarified in several places 
that the 5 MeV case is for neutrino-electron scattering experiments. For $0\nu\beta\beta$ we
have made it explicit that we consider 1.4 and 2.1 MeV.}\\

\noteq [Referee 1: Page 4] ``Please specify also the track length in
the direction of emission''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``Before the solutes in liquid... Please rephrase''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``I hope that [43] is not the final version
but a draft.

\notea{\bf It was a draft; we have changed the reference. Thanks!}\\

\noteq [Referee 1: Page 6] ``So I would be very much convinced ..if
the authors would report Fig. 3 during the electron history
(equivalently, energy or distance, including several histograms for
each case. It would be nice to see at which energy scale or length the
electron isotropizes and the Cherenkov cone smears out.''

\notea{\bf  In the updated figure 4 (angular PE hit distribution after time cut), we added the histograms for 1.4~MeV and 2.1~MeV. These show that also for 2.1 and 1.4~MeV the PE hits are not `isotropized`.%, including the full history of the electron tracks. 
We have also added the path lengths and total distance and time from the vertex to the text. We also present in Fig.~\ref{truthscatt} of this report  more raw data for the referees which support the statement that the scattering angle is small while the majority of the Cherenkov light is produced.}\\

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.50]{Figure1_1.pdf} 
        \includegraphics[scale=0.50]{Figure1_2.pdf} 
	   \caption[]{For electrons with an initial energy of 1.4~MeV, we show (Top) the angle between the initial electron direction and the direction at 0.7 MeV (=half of the initial electron energy) and (Bottom) the angle between the initial electron direction and the direction of the last Geant4 step before the electron stops. We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction; however the scattering angle is small while the majority of Cherenkov light is produced.}
	  \label{truthscatt}
        \end{center}
\end{figure}

\noteq [Referee 1: Pages 7-8] ``This is interesting, but specially
section 6 seems a bit off-topic to me. I am not against keeping it...''

\notea{\bf We'd like to keep it. }\\

\noteq [Referee 1: Page 9] ``Please add a histogram of the reconstructed
angle for all energies...''

\notea{\bf  \D}\\

\noteq [Referee 1: Page 9] ``..the vertex reconstruction has to be
obtained from all photons and the direction from the Cherenkov photons..''

\notea{\bf We agree with the referee and plan to address this in future papers. To do a credible and useful job on the position dependence will require a fairly-detailed vertex reconstruction. This is beyond the scope of this present 'proof-of-principle' paper. We have tried to clarify this in the text }\\

\noteq [Referee 1: Fig. 7] ``the authors should show the spread of the
cos(theta) distribution for each of the energies.

\notea{\bf \D\ See Figure 5.}\\

\noteq [Referee 1: Fig. 7 ``It would be nice to show/indicate..also the
Cherenkov threshold for high wave-length (n about 1.4)

\notea{\bf We have added this to the text.}\\

\noteq [Referee 1: Conclusions] ``Should be rewritten after the
previous objections are addressed.''

\notea{\bf \D\ Thanks!}\\

\noteq [Referee 1: Page 7-8] ``I see no convincing proof of it with
the information here presented. Hope the answers help clarifying the issue.''

\notea{\bf We have made the sentence more specific. Thanks.}\\

%\newpage

\noteq [Referee 2: General Comments] ``The paper introduces an
original technical development which could increase the interest of
large scale scintillation detectors for neutrino physics in general
and neutrinoless double beta decay in particular''

\notea{\bf  Thank you.}\\ %, this warms our hearts.}\\

\noteq [Referee 2: General comments] `` ..however the discussion
should deserve...a deeper analysis.''

\notea{\bf Yes. We have included more data from the lower energy studies,
have extracted much more detail from the simulation, and have
extensively rewritten the manuscript. It is still at the
proof-of-principle level. Our next work will include a full sensitivity analysis including a deeper understanding of the performance of the reconstruction algorithms with a more detailed simulation. This is a big job and we are working on it! }\\

\noteq [Referee 2: General Comments] ``The physics case is included in
the title and mentioned in the text but never discussed in some
detail...A description of why the method is powerful for double beta
decay, and..to what extent it can improve the experimental sensitivity
would complete the discussion

\notea{\bf  \D}\\

\noteq [Referee 2: General Comments] ``The chosen example (5 MeV,
head-to-head single electrons) is of little interest for double beta decay.''

\notea{\bf We have added plots and numbers for 1.4 and 2.1 MeV, and
substantially rewritten the text to differentiate between the cases
of neutrino-electron scattering and double-beta decay.}\\

\noteq [Referee 2: General Comments] ``The dependence on the decay
position.. is not discussed.''

\notea{\bf To do a credible and useful job on the position dependence
will require a fairly-detailed vertex reconstruction. This is beyond
the scope of this present `proof-of-principle' paper. We have tried
to clarify this in the text. }\\

\noteq [Referee 2: General Comments] Threshold effects as well as the
effect of statistical fluctuations for the low energy tails.. are not
discussed. The same holds for position and direction reconstruction
capabilities''

\notea{\bf Statistical fluctuations are addressed by the studies at lower energies. We mention explicitly in the conclusions the widening of the distributions due to decreased statistics. Threshold effects are outside of the scope of this paper. In general in these detectors, the largest loss of photon statistics is at the photocathode. The electronics for PMT readout are very efficient. }\\

\noteq [Referee 2: Abstract] ``While appropriate for neutrino physics,
energy resolution is definitely worse with respect to other techniques..''

\notea{\bf We have fixed the text. Thanks.}\\

\noteq [Referee 2: Introduction] ``A reference ... would help to understand
the connection with Majorana neutrinos.''

\notea{\bf \D}\\

\noteq [Referee 2: Page 1] Is this 1 sigma or FWHM?

\notea{\bf It's sigma- we have fixed the text.}\\

\noteq [Referee 2: Page 2] ``better than water Cherenkov detectors''

\notea{\bf \FT}\\

\noteq [Referee 2: Page 2 ] ``strong suppressant of backgrounds''..this
statement should deserve a deeper quantitative discussion ..''

\notea{\bf We agree. However, a useful ``quantitative discussion''
will require a vertex reconstruction. We will get there, but at present
it's beyond the scope of `proof-of-principle, looking to see if we
can see the signal. We have tried to make this clearer in the text. }\\

\noteq [Referee 2: Equ. 2.1] ``..the meaning of the parameter `x' is
not described.''

\notea{\bf \F}\\

\noteq [Referee 2: Page 3] What is the typical cutoff value and the
corresponding useful fraction of Cherenkov light?

\notea{\bf We have added the cutoff values.}\\

\noteq [Referee 2: Page 3] `` `Probable case being...' Events with
asymmetric electron energies are far more abundant.''

\notea{\bf \FT}\\

\noteq [Referee 2: Page 4] `` `Fully absorbing' ''. How are results
influenced by deviations from this hypothesis?''

\notea{\bf This is a level of detail well beyond what we've done so
far; once we have a real detector design and a full simulation it
can be addressed, but at this point devoting effort to it 
doesn't seem productive.}\\

\noteq [Referee 2: Page 4] ``Single 5 MeV electrons''.

\notea{\bf \F}\\

\noteq [Referee 2: Page 4] ``I believe the `realistic' analysis
mentioned here is indispensable ..for this process''

\notea{\bf We agree, and are working on it. It isn't easy, or,
unfortunately, fast compared to a post-doc tenure.}\\

\noteq [Referee 2: Page 6] ``The dependence on the position and the
corresponding transit time variations cannot in my opinion be neglected.''

\notea{\bf This part of the paper describes a simplified test simulation (where the photo-detector TTS was switched off) to extract the time scales of various effects. We made changes in the text to make this clear.}\\

\noteq [Referee 2: Figure 3] `` This figure is very interesting
to understand the capacity of the method to maintain directionality
information... The comparison with a single event (or equivalently the
distribution of the variances) and the distribution for 1 MeV
electrons would improve to understand also the limits/power of the method''

\notea{\bf  We have added to what is now figure 4 the distributions for the two lower energies. We think this is a nice illustration of the limits/power of the method.}\\

\noteq [Referee 2: reconstruction] ``It is not clear if limiting the
reconstruction to Cherenkov photons improve reconstruction''

\notea{\bf We agree. In this paper we were focussing on the Cherenkov photons as a simple way to apply direction reconstruction algorithms. In the future we will work to reconstruct position and direction simultaneously from the scintillation and Cherenkov light.}\\

\end{document}

\noteq [Referee 1]

\notea{\bf }\\
